### 2022-09-20 study manuscript.qmd
ATACH import

---
title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
author: "Nikhil Patel, MD"
format: docx
editor: visual
---

## Introduction

Intracerebral hemorrhage encompassess 10-15% of all strokes but is associated with a disproportionate level of death and disability. While recent attention has been paid to avoiding a self-fulfilling prophecy when prognosticating functional outcomes, there is a lingering fear that injuries in certain locations and hemispheres may portend a poorer prognosis.

There is limited research on the outcomes in these situations.

We aimed to conduct a pooled analysis of randomized trials in ICH to examine the following:

-   Is there a difference in functional outcome or HRQOL in patients with left vs. right hemispheric injuries?

## Methods

### 2022-09-20 study protocol.qmd
ATACH import

---
title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
subtitle: "Study Protocol and Statistical Analysis Plan"
author: "Nikhil Patel, MD"
format: docx
editor: visual
---

## Introduction

Patients with neurological injury experience different symptoms between the left and right hemispheres, and this may have.

## Methods

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tableone)
library(data.table)
```

### Pooling RCTs

Randomized trials and observational studies included in the analysis:

- ATACH2
- ERICH
- MISTIE
- MISTIE III
- CLEAR-IVH

```{r setup, include=FALSE}

# ATACH2

# Create function to paste together string of FORMxx.txt
atach.import <- function(y) {
  dataset <- paste0("atach2",".",y)
  filepath <- paste0("./ATACH2/FORM",y,".txt")
  dataset <- read.table(filepath, sep = "\t", header = TRUE)
}

atach00 <- atach.import("00")
atach01 <- atach.import("01")
atach02 <- atach.import("02")
atach03 <- atach.import("03")
atach04 <- atach.import("04")
atach05 <- atach.import("05")
# atach05c <- atach.import("05c") Detailed Blood Pressure Data, Unsure if Needed
atach06 <- atach.import("06")
atach07 <- atach.import("07")
atach08 <- atach.import("08")
atach09 <- atach.import("09")
atach10 <- atach.import("10")
atach11 <- atach.import("11")
atach13 <- atach.import("13")
atach14 <- atach.import("14")
atach15 <- atach.import("15")
atach16 <- atach.import("16")
atach17 <- atach.import("17")
atach18 <- atach.import("18")
atach20 <- atach.import("20")
atach21 <- atach.import("21")
atach23 <- atach.import("23")
atach33 <- atach.import("33")

# Merging all tidy data
atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach33)

atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")

# To Do's: write function to quickly import separate files, merge all datasets into one, and extract variables to describe how WLST was done and how often

# https://www.statology.org/merge-multiple-data-frames-in-r/

```

### 2022-09-21 study protocol.qmd
combined atach2 data

@@
 format: docx
-editor: visual
 ---
@@
 atach33 <- atach.import("33")
+atachct <- read.table("./ATACH2/CTCR.txt", sep = "\t", header = TRUE)
 
 # Merging all tidy data
-atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach33)
+atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach21, atach33)
+atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")
+atach.whole <-atach.whole %>% mutate(study = "ATACH2")
 
-atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")
+# Importing CT Data
+atachct <- atachct %>%
+  mutate(ctnumber = case_when(
+    CTCRMIN < 0 ~ "baseline",
+    CTCRMIN > 0 ~ "24hourpost"
+  ))
 
-# To Do's: write function to quickly import separate files, merge all datasets into one, and extract variables to describe how WLST was done and how often
+# Extracting First CT Data
+atachctbaseline <- atachct %>% 
+  filter(ctnumber == "baseline") %>%
+  select(SUBJECT_ID, CTCRMIN, CTCRQ05, CTCRQ07, CTCRQ08, CTCRQ09, CTCRQ12, CTCRQ13, CTCRQ11, CTCRQ14) %>%
+  rename(CTCRMIN.baseline = CTCRMIN) %>%
+  rename(ich.location.baseline = CTCRQ05) %>%
+  rename(ivh.baseline = CTCRQ07) %>%
+  rename(hydrocephalus.baseline = CTCRQ08) %>%
+  rename(pineal.shift.baseline = CTCRQ09) %>%
+  rename(ich.volume.baseline = CTCRQ12) %>%
+  rename(ivh.volume.baseline = CTCRQ13) %>%
+  rename(phe.volume.baseline = CTCRQ11) %>%
+  rename(sp.shift.baseline = CTCRQ14)
 
-# https://www.statology.org/merge-multiple-data-frames-in-r/
+atachct24hrpost <- atachct %>% # Extracting Second CT Data
+  filter(ctnumber == "24hourpost") %>%
+    select(SUBJECT_ID, CTCRMIN, CTCRQ05, CTCRQ07, CTCRQ08, CTCRQ09, CTCRQ12, CTCRQ13, CTCRQ11, CTCRQ14) %>%
+  rename(CTCRMIN.24hourpost = CTCRMIN) %>%
+  rename(ich.location.24hourpost = CTCRQ05) %>%
+  rename(ivh.24hourpost = CTCRQ07) %>%
+  rename(hydrocephalus.24hourpost = CTCRQ08) %>%
+  rename(pineal.shift.24hourpost = CTCRQ09) %>%
+  rename(ich.volume.24hourpost = CTCRQ12) %>%
+  rename(ivh.volume.24hourpost = CTCRQ13) %>%
+  rename(phe.volume.24hourpost = CTCRQ11) %>%
+  rename(sp.shift.24hourpost = CTCRQ14)
+
+atach.whole <- left_join(atach.whole, atachctbaseline, by = "SUBJECT_ID")
+atach.whole <- left_join(atach.whole, atachct24hrpost, by = "SUBJECT_ID")
+
+# Importing mRS Data
+
+atachmrs30 <- atach09 %>%
+  filter(mrsday == 30) %>%
+  select(SUBJECT_ID, F09Q01) %>%
+  rename(mrs30 = F09Q01)
+
+atachmrs90 <- atach09 %>%
+  filter(mrsday == 90) %>%
+  select(SUBJECT_ID, F09Q01) %>%
+  rename(mrs90 = F09Q01)
+
+atach.whole <-left_join(atach.whole, atachmrs30, by = "SUBJECT_ID")
+atach.whole <-left_join(atach.whole, atachmrs90, by = "SUBJECT_ID")
+
+# Table 1
+
+atachVarsTableOne <- c("F33O02", "F01Q01", "F01Q02", "F01Q03M1", "F01Q03M2", "F01Q03M3", "F01Q03M5", "F01Q03M6", "F01Q03M7", "SITE_ID", "COUNTRY", "F03Q01", "F03Q04", "F03Q05", "F03Q06", "F03Q07", "F03Q09", "F03Q12", "F03Q13", "F04Q01", "F04Q03", "F00Q02MIN", "F05Q07a", "F05Q07b", "F33Q04", "F10Q22", "F15Q11", "F15Q04", "F15Q07", "F15Q09", "F15Q12", "F15Q17", "CTCRQ05", "CTCRQ07
+", "CTCRQ08", "CTCRQ12", "CTCRQ13", "CTCRQ11", "CTCRQ14", "F07Q01", "F07Q02", "F07Q11", "F07Q04DAY", "F07Q05", "F07Q07", "F07Q08", "F07Q09DAY")
 
 ```
+
+Variables included:
+
+-   Age
+
+-   Gender
+
+-   Country Randomized
+
+-   RCT enrolled in
+
+-   Intracerebral hemorrhage
+
+-   Ischemic stroke
+
+-   Coronary artery disease
+
+-   Congestive heart failure
+
+-   Chronic kidney disease
+
+-   Obesity
+
+-   Hypertension
+
+-   Diabetes
+
+-   Antihypertensive therapy
+
+-   Anticoagluation therapy
+
+-   Antiplatelet therapy
+
+-   Lipid lowering therapy
+
+-   Time from onset to presentation
+
+-   Systolic BP
+
+-   Diastolic BP
+
+-   Median GCS
+
+-   Median NIHSS
+
+-   Glucose
+
+-   Hematoma volume
+
+-   Hematoma location
+
+-   IVH
+
+-   ICH Score
+
+-   Extraventricular drain
+
+-   Craniotomy
+
+-   WLST
+
+-   Median time to WLST
+
+-   mRS
+
+-   EuroQOL
+
+-   BI
+
+-   Living at home

### 2022-09-22 study protocol.qmd
Combined ATACH and ERICH

@@
 library(data.table)
+library(haven) # Importing SAS Files
+
 ```
 
-### Pooling RCTs
-
-Randomized trials and observational studies included in the analysis:
-
--   ATACH2
--   ERICH
--   MISTIE
--   MISTIE III
--   CLEAR-IVH
-
-```{r setup, include=FALSE}
-
-# ATACH2
+```{r ATACH2, include=FALSE}
 
@@
   select(SUBJECT_ID, F09Q01) %>%
-  rename(mrs30 = F09Q01)
+  rename(mrs.30 = F09Q01)
 
@@
   select(SUBJECT_ID, F09Q01) %>%
-  rename(mrs90 = F09Q01)
+  rename(mrs.90 = F09Q01)
 
@@
 
+# Categorizing and Recoding Variables
+
+# Race
+atach.whole <- atach.whole %>%
+  mutate(race = case_when(
+    F01Q03M1 == 1 ~ "American Indian or Alaskan Native",
+    F01Q03M2 == 1 ~ "Asian",
+    F01Q03M3 == 1 ~ "Black or African-American",
+    F01Q03M5 == 1 ~ "White",
+    F01Q03M6 == 1 ~ "Other/Not Reported",
+    F01Q03M7 == 1 ~ "Other/Not Reported"
+  ))
+
+# Ethnicity
+atach.whole <- atach.whole %>% 
+  mutate(ethnicity = recode(F01Q02,
+                             "1" = "Hispanic or Latino",
+                             "2" = "Not Hispanic or Latino",
+                             .default = "Unknown/Not Reported"
+                             ))
+
+# Sex
+atach.whole <- atach.whole %>% 
+  mutate(sex = dplyr::recode(F01Q01,
+                      "1" = "Male",
+                      "2" = "Female"))
+
+# Country of Enrolling Site
+atach.whole <- atach.whole %>% 
+  mutate(enrolling.country = recode(COUNTRY,
+                                    "1" = "United States",
+                                    "2" = "Japan",
+                                    "3" = "China",
+                                    "4" = "South Korea",
+                                    "5" = "Taiwan",
+                                    "6" = "Germany"))
+
+# Past Medical History
+atach.whole <- atach.whole %>%
+  mutate(stroke = recode(F03Q01,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(chf = recode(F03Q03,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(afib = recode(F03Q04,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(cad = case_when(
+    F03Q05 == 0 ~ "No",
+    F03Q05 == 1 ~ "Yes",
+    F03Q05 == 98 ~ "NA",
+    F03Q06 == 0 ~ "No",
+    F03Q06 == 1 ~ "Yes",
+    F03Q06 == 98 ~ "NA"
+  ))
+
+atach.whole <- atach.whole %>%
+  mutate(htn = recode(F03Q07,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(pvd = recode(F03Q08,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(hld = recode(F03Q09,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(dm2 = recode(F03Q12,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(tobacco = recode(F03Q13,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+# ICH Laterality
+atach.whole <- atach.whole %>%
+  mutate(ich.laterality = recode(ich.location.baseline,
+                                 "1" = "Right",
+                                 "2" = "Right",
+                                 "3" = "Right",
+                                 "4" = "Left",
+                                 "5" = "Left",
+                                 "6" = "Left",
+                                 .default = "NA"))
+
+# ICH Location
+atach.whole <- atach.whole %>%
+  mutate(ich.location.baseline = recode(ich.location.baseline,
+                                        "1" = "Thalamus",
+                                        "2" = "Basal Ganglia",
+                                        "3" = "Lobar",
+                                        "4" = "Thalamus",
+                                        "5" = "Basal Ganglia",
+                                        "6" = "Lobar",
+                                        "7" = "Pons",
+                                        "8" = "Cerebellum"))
+
+atach.whole <- atach.whole %>%
+  mutate(ivh = recode(ivh.baseline,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(hydrocephalus = recode(hydrocephalus.baseline,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(hydrocephalus = recode(hydrocephalus.baseline,
+                         "0" = "No",
+                         "1" = "Yes",
+                         "98" = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(ich.volume.cc = ich.volume.baseline/1000)
+
+atach.whole <- atach.whole %>%
+  mutate(phe.volume.cc = phe.volume.baseline/1000)
+
+atach.whole <- atach.whole %>%
+  mutate(ivh.volume.cc = ivh.volume.baseline/1000)
+
+atach.whole <- atach.whole %>%
+  mutate(mech.ventilation = recode(F07Q01,
+                                   "0" = "No",
+                                   "1" = "Yes",
+                                   .default = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(evd = recode(F07Q05,
+                                   "0" = "No",
+                                   "1" = "Yes",
+                                   .default = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(nsgy.evacuation = recode(F07Q08,
+                                   "0" = "No",
+                                   "1" = "Yes",
+                                   .default = "NA"))
+
+atach.whole <- atach.whole %>%
+  mutate(trach = recode(F07Q11,
+                                   "0" = "No",
+                                   "1" = "Yes",
+                                   .default = "NA"))
+
+# NIHSS
+atach.nihss.baseline <- atach10 %>%
+  filter(nihss.assessment == "baseline") %>%
+  select(SUBJECT_ID, F10Q22) %>%
+  rename(nihss.baseline = F10Q22)
+
+atach.nihss.24hourpost <- atach10 %>%
+  filter(nihss.assessment == "24hourpost") %>%
+  select(SUBJECT_ID, F10Q22) %>%
+  rename(nihss.24hourpost = F10Q22)
+
+atach.whole <-left_join(atach.whole, atach.nihss.baseline, by = "SUBJECT_ID")
+atach.whole <-left_join(atach.whole, atach.nihss.24hourpost, by = "SUBJECT_ID")
+
+# EuroQol
+
+atach.whole <- atach.whole %>%
+  mutate(euro.mobility.90 = factor(F16Q02,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
+
+atach.whole <- atach.whole %>%
+  mutate(euro.selfcare.90 = factor(F16Q03,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
+
+atach.whole <- atach.whole %>%
+  mutate(euro.usual.90 = factor(F16Q04,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
+
+atach.whole <- atach.whole %>%
+  mutate(euro.pain.90 = factor(F16Q05,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
+
+atach.whole <- atach.whole %>%
+  mutate(euro.anxiety.90 = factor(F16Q06,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
+
+# Filtering out cases without ICH laterality
+atach.whole <- atach.whole %>%
+  filter(ich.laterality != "NA")
+
 # Table 1
 
-atachVarsTableOne <- c("F33O02", "F01Q01", "F01Q02", "F01Q03M1", "F01Q03M2", "F01Q03M3", "F01Q03M5", "F01Q03M6", "F01Q03M7", "SITE_ID", "COUNTRY", "F03Q01", "F03Q04", "F03Q05", "F03Q06", "F03Q07", "F03Q09", "F03Q12", "F03Q13", "F04Q01", "F04Q03", "F00Q02MIN", "F05Q07a", "F05Q07b", "F33Q04", "F10Q22", "F15Q11", "F15Q04", "F15Q07", "F15Q09", "F15Q12", "F15Q17", "CTCRQ05", "CTCRQ07
-", "CTCRQ08", "CTCRQ12", "CTCRQ13", "CTCRQ11", "CTCRQ14", "F07Q01", "F07Q02", "F07Q11", "F07Q04DAY", "F07Q05", "F07Q07", "F07Q08", "F07Q09DAY")
+atach.whole <- atach.whole %>%
+  rename(age = F33Q02) %>%
+  rename(symptom.onset = F00Q02MIN) %>%
+  rename(sbp.initial = F05Q07A) %>%
+  rename(dbp.initial = F05Q07B) %>%
+  rename(gcs.baseline = F33Q04) %>%
+  rename(intubation.days = F07Q02) %>%
+  rename(extubation.date = F07Q03DAY) %>%
+  rename(trach.date = F07Q04DAY) %>%
+  rename(evd.days = F07Q07) %>%
+  rename(nsgy.evacuation.date = F07Q09DAY) %>%
+  rename(euro.vas.90 = F16Q07) %>%
+  rename(EQ_INDEX.90 = EQ_INDEX)
+
+VarsTableOne <- c("study",
+                  "age", # Age
+                       "sex", # Sex
+                       "ethnicity", # Ethnicity
+                       "race", # Race
+                       "enrolling.country", # Country of Enrolling Site
+                       "stroke", 
+                       "cad", 
+                       "chf", 
+                       "afib", 
+                       "pvd",
+                       "hld",
+                       "dm2",
+                       "tobacco",
+                       #"symptom.onset", # Time of symptom onset
+                       "sbp.initial", # Initial SBP
+                       "dbp.initial", # Initial DBP
+                       "gcs.baseline", # Baseline GCS
+                       #"nihss.baseline", # NIHSS
+                       "ich.location.baseline", 
+                       "ivh",
+                      # "hydrocephalus",
+                       "ich.volume.cc",
+                       "ivh.volume.cc",
+                       "phe.volume.cc",
+                       "sp.shift.baseline",
+                       "mech.ventilation",
+                       "intubation.days", 
+                       "extubation.date", # Date of extubation
+                       "trach",
+                       "trach.date", # Date of Tracheostomy
+                       "evd",
+                       "evd.days", # Total EVD days
+                       "nsgy.evacuation",
+                       "nsgy.evacuation.date",  # Date of Evacuation
+                       "mrs.30",
+                       "mrs.90",
+                      "mrs.180",
+                      "mrs.365",
+                      "euro.mobility.90",
+                      "euro.mobility.180",
+                      "euro.mobility.365",
+                       "euro.selfcare.90",
+                      "euro.selfcare.180",
+                      "euro.selfcare.365",
+                       "euro.usual.90",
+                      "euro.usual.180",
+                      "euro.usual.365",
+                       "euro.pain.90",
+                      "euro.pain.180",
+                      "euro.pain.365",
+                       "euro.anxiety.90",
+                      "euro.anxiety.180",
+                      "euro.anxiety.365",
+                       "euro.vas.90", # EuroQOL VAS
+                       "euro.vas.180",
+                       "euro.vas.365",
+                       "EQ_INDEX.90", # EQ-5D Utility Index,,
+                      "current.location.90",
+                      "current.location.180",
+                      "current.location.365",
+                       "ich.laterality"
+                  )
+
+# Maybe import lab data, but this may be superfluous
 
@@
 
-Variables included:
+## Table 1 ATACH-2
 
--   Age
+```{r ATACH table1, echo=FALSE}
 
--   Gender
+atach.table1.data <- atach.whole %>%
+  select(atachVarsTableOne)
 
--   Country Randomized
+CreateTableOne(vars = atachVarsTableOne,
+               data = atach.whole,
+               strata = "ich.laterality")
 
--   RCT enrolled in
+```
 
--   Intracerebral hemorrhage
+```{r ERICH, include=FALSE}
 
--   Ischemic stroke
+erich.case <- read_sas("ERICH/erich_case_100217.sas7bdat")
 
--   Coronary artery disease
+erich.whole <- erich.case %>%
+  select(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30, O3, O3A, O5, O20, AU8, AU8_OTH, RANKIN_FU1, RANKIN_FU2, RANKIN_FU3, MOBILITY_FU1, MOBILITY_FU2, MOBILITY_FU3, SELFCARE_FU1, SELFCARE_FU2, SELFCARE_FU3, USUALACT_FU1, USUALACT_FU2, USUALACT_FU3, PAIN_FU1, PAIN_FU2, PAIN_FU3, ANXIETY_FU1, ANXIETY_FU2, ANXIETY_FU3, HLTHSTAT_FU1, HLTHSTAT_FU2, HLTHSTAT_FU3, FEED_FU1, FEED_FU2, FEED_FU3, BATHE_FU1, BATHE_FU2, BATHE_FU3, GROOM_FU1, GROOM_FU2, GROOM_FU3, DRESS_FU1, DRESS_FU2, DRESS_FU3, BOWELS_FU1, BOWELS_FU2, BOWELS_FU3, BLADDER_FU1, BLADDER_FU2, BLADDER_FU3, TOILET_FU1, TOILET_FU2, TOILET_FU3, TRANSFER_FU1, TRANSFER_FU2, TRANSFER_FU3, MOBLEVEL_FU1, MOBLEVEL_FU2, MOBLEVEL_FU3, STAIRS_FU1, STAIRS_FU2, STAIRS_FU3, BART_TOT_FU1, BART_TOT_FU2, BART_TOT_FU3, EMS33, ED44, CURR_LOC_FU1, CURR_LOC_FU2, CURR_LOC_FU3)
 
--   Congestive heart failure
+erich.whole <- erich.whole %>%
+  mutate(study = "ERICH")
 
--   Chronic kidney disease
+erich.whole <- erich.whole %>%
+  mutate(enrolling.country = "United States")
 
--   Obesity
+# Renaming and Recoding Variables
 
--   Hypertension
+erich.whole <- erich.whole %>%
+  mutate(gender = recode(gender,
+                         "F" = "Female",
+                         "M" = "Male")) %>%
+  rename(sex = gender)
 
--   Diabetes
+erich.whole <- erich.whole %>%
+  mutate(Race = case_when(
+    D21 == "W" ~ "White",
+    D21 == "B" ~ "Black or African-American",
+    D21 == "A" ~ "Asian",
+    D21 == "N" ~ "Native Hawaiian/Pacific Islander",
+    D21 == "I" ~ "American Indian or Alaskan Native",
+    D21 == "O" ~ "Other/Not Reported",
+    D21 == "U" ~ "Other/Not Reported"
+    )) %>%
+  rename(race = Race)
 
--   Antihypertensive therapy
+erich.whole <- erich.whole %>%
+  mutate(Ethnicity = case_when(
+    D21A == "H" ~ "Hispanic or Latino",
+    D21A == "N" ~ "Not Hispanic or Latino",
+    D21A == "U" ~ "Not Hispanic or Latino"
+  ))
 
--   Anticoagluation therapy
+erich.whole <- erich.whole %>%
+  mutate(stroke = recode(SH1,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Antiplatelet therapy
+erich.whole <- erich.whole %>%
+  mutate(cad = recode(MHX8,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Lipid lowering therapy
+erich.whole <- erich.whole %>%
+  mutate(chf = recode(MHX13,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Time from onset to presentation
+erich.whole <- erich.whole %>%
+  mutate(afib = recode(MHX11,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Systolic BP
+erich.whole <- erich.whole %>%
+  mutate(pvd = recode(MHX27C,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Diastolic BP
+erich.whole <- erich.whole %>%
+  mutate(hld = recode(MHX6,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
 
--   Median GCS
+erich.whole <- erich.whole %>%
+  mutate(dm2 = case_when(
+    MHX3 == "2" ~ "No",
+    MHX3 == "8" ~ "NA",
+    MHX3A == "1" ~ "No",
+    MHX3A == "2" ~ "Yes",
+    MHX3A == "8" ~ "NA"))
 
--   Median NIHSS
+erich.whole <- erich.whole %>%
+  mutate(tobacco = recode(SMK1A,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "5" = "NA",
+                         "8" = "NA"))
 
--   Glucose
+erich.whole <- erich.whole %>%
+  rename(sbp.initial = ED47_SYS)
 
--   Hematoma volume
+erich.whole <- erich.whole %>%
+  rename(dbp.initial = ED47_DIAS)
 
--   Hematoma location
+erich.whole <- erich.whole %>%
+  rename(gcs.baseline = ED39)
 
--   IVH
+erich.whole <- erich.whole %>%
+  mutate(ich.laterality = recode(PHY6A,
+                                 "L" = "Left",
+                                 "R" = "Right",
+                                 "M" = "NA",
+                                 .default = "NA"))
 
--   ICH Score
+erich.whole <- erich.whole %>%
+  mutate(ich.location.baseline = recode(erich.case$detail,
+                                        "L Thalamic" = "Thalamus",
+                                        "R Thalamic" = "Thalamus",
+                                        "L Caudate Nucleus" = "Basal Ganglia",
+                                        "L Cerebellar" = "Cerebellum",
+                                        "L Corona Radiata" = "Lobar",
+                                        "L Frontal" = "Lobar",
+                                        "L Frontal Temporal" = "Lobar",
+                                        "L Frontoparietal" = "Lobar",
+                                        "L Occipital" = "Lobar",
+                                        "L Parietal" = "Lobar",
+                                        "L Parieto-Occipital" = "Lobar",
+                                        "L Periventricular" = "Basal Ganglia",
+                                        "L Putaminal" = "Basal Ganglia",
+                                        "L Temporal" = "Lobar",
+                                        "L Temporal-Parietal-Occipital (L Hemisphere)" = "Lobar",
+                                        "L Temporal Occipital" = "Lobar",
+                                        "L Temporal parietal" = "Lobar",
+                                        "Multiple Hemorrhages - Deep" = "Basal Ganglia",
+                                        "Multiple Hemorrhages - Lobar" = "Lobar",
+                                        "Not applicable" = "NA",
+                                        "Primary IVH" = "Primary IVH",
+                                        "R Basal Ganglia" = "Basal Ganglia",
+                                        "R Caudate Nucleus" = "Basal Ganglia",
+                                        "R Cerebellar" = "Cerebellum",
+                                        "R Frontal" = "Lobar",
+                                        "R Frontal Temporal" = "Lobar",
+                                        "R Frontoparietal" = "Lobar",
+                                        "R Occipital" = "Lobar",
+                                        "R Parietal" = "Lobar",
+                                        "R Parieto-Occipital" = "Lobar",
+                                        "R Periventricular" = "Basal Ganglia",
+                                        "R Putaminal" = "Basal Ganglia",
+                                        "R Temporal" = "Lobar",
+                                        "R Temporal-Parietal-Occipital (R Hemisphere)" = "Lobar",
+                                        "R Temporal Occipital" = "Lobar",
+                                        "R Temporal Perietal" = "Lobar",
+                                        "Unknown" = "NA",
+                                        "Brainstem" = "Brainstem",
+                                        "Cerebellar" = "Cerebellum",
+                                        "R Temporal parietal" = "Lobar",
+                                        "L Basal Ganglia" = "Basal Ganglia",                                        ))
 
--   Extraventricular drain
+erich.whole <- erich.whole %>%
+  mutate(ivh = case_when(erich.case$IVH_Vol_CT1 > 0 ~ "Yes",
+                         erich.case$IVH_Vol_CT1 < 0.1 ~ "No"))
 
--   Craniotomy
+erich.whole <- erich.whole %>%
+  rename(ich.volume.cc = ICH_Vol_CT1)
 
--   WLST
+erich.whole <- erich.whole %>%
+  rename(ivh.volume.cc = IVH_Vol_CT1)
 
--   Median time to WLST
+erich.whole <- erich.whole %>%
+  mutate(mech.ventilation = case_when(
+    EMS33 == "1" ~ "Yes",
+    EMS33 == "2" ~ "No",
+    ED44 == "1" ~ "Yes",
+    ED44 == "2" ~ "No",
+    IT7 == "1" ~ "Yes",
+    IT7 == "2" ~ "No"))
 
--   mRS
+erich.whole <- erich.whole %>%
+  mutate(evd = case_when(
+    IT8 == "1" ~ "Yes",
+    IT8 == "2" ~ "No",
+  ))
 
--   EuroQOL
+erich.whole <- erich.whole %>%
+  mutate(nsgy.evacuation = case_when(
+    IT5D == "1" ~ "Yes",
+    IT5D == "2" ~ "No",
+    IT5H == "1" ~ "Yes",
+    IT5H == "2" ~ "No"
+  ))
 
--   BI
+erich.whole <- erich.whole %>%
+  rename(mrs.90 = RANKIN_FU1) %>%
+  rename(mrs.180 = RANKIN_FU2) %>%
+  rename(mrs.365 = RANKIN_FU3) %>%
+  rename(euro.vas.90 = HLTHSTAT_FU1) %>%
+  rename(euro.vas.180 = HLTHSTAT_FU2) %>%
+  rename(euro.vas.365 = HLTHSTAT_FU3) %>%
+  rename(barthel.feeding.90 = FEED_FU1) %>%
+  rename(barthel.bathing.90 = BATHE_FU1) %>%
+  rename(barthel.grooming.90 = GROOM_FU1) %>%
+  rename(barthel.dressing.90 = DRESS_FU1) %>%
+  rename(barthel.bowels.90 = BOWELS_FU1) %>%
+  rename(barthel.bladder.90 = BLADDER_FU1) %>%
+  rename(barthel.toilet.90 = TOILET_FU1) %>%
+  rename(barthel.transfer.90 = TRANSFER_FU1) %>%
+  rename(barthel.mobility.90 = MOBLEVEL_FU1) %>%
+  rename(barthel.stairs.90 = STAIRS_FU1) %>%
+  rename(barthel.total.90 = BART_TOT_FU1) %>%
+  rename(barthel.feeding.180 = FEED_FU2) %>%
+  rename(barthel.bathing.180 = BATHE_FU2) %>%
+  rename(barthel.grooming.180 = GROOM_FU2) %>%
+  rename(barthel.dressing.180 = DRESS_FU2) %>%
+  rename(barthel.bowels.180 = BOWELS_FU2) %>%
+  rename(barthel.bladder.180 = BLADDER_FU2) %>%
+  rename(barthel.toilet.180 = TOILET_FU2) %>%
+  rename(barthel.transfer.180 = TRANSFER_FU2) %>%
+  rename(barthel.mobility.180 = MOBLEVEL_FU2) %>%
+  rename(barthel.stairs.180 = STAIRS_FU2) %>%
+  rename(barthel.total.180 = BART_TOT_FU2) %>%
+  rename(barthel.feeding.365 = FEED_FU3) %>%
+  rename(barthel.bathing.365 = BATHE_FU3) %>%
+  rename(barthel.grooming.365 = GROOM_FU3) %>%
+  rename(barthel.dressing.365 = DRESS_FU3) %>%
+  rename(barthel.bowels.365 = BOWELS_FU3) %>%
+  rename(barthel.bladder.365 = BLADDER_FU3) %>%
+  rename(barthel.toilet.365 = TOILET_FU3) %>%
+  rename(barthel.transfer.365 = TRANSFER_FU3) %>%
+  rename(barthel.mobility.365 = MOBLEVEL_FU3) %>%
+  rename(barthel.stairs.365 = STAIRS_FU3) %>%
+  rename(barthel.total.365 = BART_TOT_FU3)
 
--   Living at home
+erich.whole <- erich.whole %>%
+  mutate(current.location.90 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      )) %>%
+  mutate(current.location.180 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      )) %>%
+  mutate(current.location.365 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      ))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.90 = factor(MOBILITY_FU1,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.90 = factor(SELFCARE_FU1,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.90 = factor(USUALACT_FU1,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.90 = factor(PAIN_FU1,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.90 = factor(ANXIETY_FU1,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.180 = factor(MOBILITY_FU2,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.180 = factor(SELFCARE_FU2,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.180 = factor(USUALACT_FU2,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.180 = factor(PAIN_FU2,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.180 = factor(ANXIETY_FU2,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.365 = factor(MOBILITY_FU3,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.365 = factor(SELFCARE_FU3,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.365 = factor(USUALACT_FU3,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.365 = factor(PAIN_FU3,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.365 = factor(ANXIETY_FU3,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
+
+# Variables left: symptom.onset
+
+# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days, tracheostomy, nsgy.evacuation.date,
+
+# Filtering out ICH Laterality and Infratentorial Location
+erich.whole <- erich.whole %>%
+  filter(ich.laterality != "NA") %>%
+  filter(ich.location.baseline == "Basal Ganglia" |
+           ich.location.baseline == "Thalamus" |
+           ich.location.baseline == "Lobar")
+
+```
+
+```{r combining ATACH2 and ERICH, include=FALSE}
+
+atach.erich <- union_all(atach.whole, erich.whole)
+
+atach.erich <- atach.erich %>%
+  select(VarsTableOne) %>%
+  mutate(ich.score.age = case_when(
+    age < 80 ~ 0,
+    age >= 80 ~ 1
+  )) %>%
+  mutate(ich.score.gcs = case_when(
+    gcs.baseline >= 13 ~ 0,
+    gcs.baseline <= 4 ~ 2,
+    TRUE ~ 1)) %>%
+  mutate(ich.score.volume = if_else(ich.volume.cc >= 30,1,0)) %>%
+  mutate(ich.score.ivh = if_else(ivh == "Yes",1,0)) %>%
+  mutate(ich.score.total = ich.score.age + 
+                           ich.score.gcs + 
+                           ich.score.volume +
+                           ich.score.ivh)
+
+```

### 2022-09-23 study protocol.qmd
Update study protocol.qmd

@@
 library(haven) # Importing SAS Files
+library(janitor)
 
@@
 
-## Table 1 ATACH-2
-
-```{r ATACH table1, echo=FALSE}
-
-atach.table1.data <- atach.whole %>%
-  select(atachVarsTableOne)
-
-CreateTableOne(vars = atachVarsTableOne,
-               data = atach.whole,
-               strata = "ich.laterality")
-
-```
-
 ```{r ERICH, include=FALSE}
@@
 
+atach.erich %>%
+  as_tibble()
+
 ```

### 2022-10-06 study protocol.qmd
Updated stuff

@@
 library(janitor)
+library(lubridate)
 
@@
 erich.whole <- erich.case %>%
-  select(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30, O3, O3A, O5, O20, AU8, AU8_OTH, RANKIN_FU1, RANKIN_FU2, RANKIN_FU3, MOBILITY_FU1, MOBILITY_FU2, MOBILITY_FU3, SELFCARE_FU1, SELFCARE_FU2, SELFCARE_FU3, USUALACT_FU1, USUALACT_FU2, USUALACT_FU3, PAIN_FU1, PAIN_FU2, PAIN_FU3, ANXIETY_FU1, ANXIETY_FU2, ANXIETY_FU3, HLTHSTAT_FU1, HLTHSTAT_FU2, HLTHSTAT_FU3, FEED_FU1, FEED_FU2, FEED_FU3, BATHE_FU1, BATHE_FU2, BATHE_FU3, GROOM_FU1, GROOM_FU2, GROOM_FU3, DRESS_FU1, DRESS_FU2, DRESS_FU3, BOWELS_FU1, BOWELS_FU2, BOWELS_FU3, BLADDER_FU1, BLADDER_FU2, BLADDER_FU3, TOILET_FU1, TOILET_FU2, TOILET_FU3, TRANSFER_FU1, TRANSFER_FU2, TRANSFER_FU3, MOBLEVEL_FU1, MOBLEVEL_FU2, MOBLEVEL_FU3, STAIRS_FU1, STAIRS_FU2, STAIRS_FU3, BART_TOT_FU1, BART_TOT_FU2, BART_TOT_FU3, EMS33, ED44, CURR_LOC_FU1, CURR_LOC_FU2, CURR_LOC_FU3)
+  select(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30, O3, O3A, O5, O20, AU8, AU8_OTH, RANKIN_FU1, RANKIN_FU2, RANKIN_FU3, MOBILITY_FU1, MOBILITY_FU2, MOBILITY_FU3, SELFCARE_FU1, SELFCARE_FU2, SELFCARE_FU3, USUALACT_FU1, USUALACT_FU2, USUALACT_FU3, PAIN_FU1, PAIN_FU2, PAIN_FU3, ANXIETY_FU1, ANXIETY_FU2, ANXIETY_FU3, HLTHSTAT_FU1, HLTHSTAT_FU2, HLTHSTAT_FU3, FEED_FU1, FEED_FU2, FEED_FU3, BATHE_FU1, BATHE_FU2, BATHE_FU3, GROOM_FU1, GROOM_FU2, GROOM_FU3, DRESS_FU1, DRESS_FU2, DRESS_FU3, BOWELS_FU1, BOWELS_FU2, BOWELS_FU3, BLADDER_FU1, BLADDER_FU2, BLADDER_FU3, TOILET_FU1, TOILET_FU2, TOILET_FU3, TRANSFER_FU1, TRANSFER_FU2, TRANSFER_FU3, MOBLEVEL_FU1, MOBLEVEL_FU2, MOBLEVEL_FU3, STAIRS_FU1, STAIRS_FU2, STAIRS_FU3, BART_TOT_FU1, BART_TOT_FU2, BART_TOT_FU3, EMS33, ED44, CURR_LOC_FU1, CURR_LOC_FU2, CURR_LOC_FU3, IT2, O1, O2_DATE, O2_TIME, O3, O3_OTH, DOD)
 
@@
     IT5D == "1" ~ "Yes",
-    IT5D == "2" ~ "No",
     IT5H == "1" ~ "Yes",
-    IT5H == "2" ~ "No"
+    TRUE ~ "No"
       ))
@@
 
-# Variables left: symptom.onset
+# Parsing Dates and Times in ERICH
 
-# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days, tracheostomy, nsgy.evacuation.date,
+# symptom.onset
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$SE2)))
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.estimated.time = case_when( # Converting estimated times to numbers
+    SE3 == "1" ~ "06:00:00",
+    SE3 == "2" ~ "NA",
+    SE3 == "3" ~ "03:00:00",
+    SE3 == "4" ~ "09:00:00",
+    SE3 == "5" ~ "15:00:00",
+    SE3 == "6" ~ "21:00:00"
+  ))
+
+erich.whole$symptom.onset.estimated.time <- hms(erich.whole$symptom.onset.estimated.time)
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.estimated.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$symptom.onset.estimated.time)))
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset = case_when(
+    is.na(symptom.onset.date.time) ~ symptom.onset.estimated.date.time,
+    TRUE ~ symptom.onset.date.time
+  ))
+
+# Time of ED presentation
+
+erich.whole <- erich.whole %>%
+  mutate(ed.presentation.date.time = ymd_hms(paste(erich.whole$ED7, erich.whole$ED8)))
+
+# Neurosurgery Dates
+
+erich.whole <- erich.whole %>%
+  mutate(nsgy.evacuation.date = case_when(
+    IT5D == "1" ~ ymd_hms(paste(IT5D_DATE, IT5D_TIME)),
+    IT5H == "1" ~ ymd_hms(paste(IT5H_DATE, IT5H_TIME)),
+    IT5F == "1" ~ ymd_hms(paste(IT5F_DATE, IT5F_TIME))
+  ))
+
+# DNR and WLST Status
+
+erich.whole <- erich.whole %>%
+  mutate(dnr.binary = case_when(
+    CC16 == "1" ~ "Yes",
+    CC16 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(dni.binary = case_when(
+    CC17 == "1" ~ "Yes",
+    CC17 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(comfort.binary = case_when(
+    CC18 == "1" ~ "Yes",
+    CC18 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(dnr.date.time = case_when(
+    CC16 == "1" ~ ymd_hms(paste(CC16_DATE, CC16_TIME))
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(dni.date.time = case_when(
+    CC17 == "1" ~ ymd_hms(paste(CC17_DATE, CC17_TIME))
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(comfort.date.time = case_when(
+    CC18 == "1" ~ ymd_hms(paste(CC18_DATE, CC18_TIME))
+  ))
+
+# Mortality Status
+
+erich.whole$death.date <- ymd(erich.whole$DOD)
+
+erich.whole <- erich.whole %>%
+  mutate(death.interval = as_date(death.date) - as_date(symptom.onset)) %>%
+  as.duration(death.interval)
+
+erich.whole$death.interval <- parse_integer(erich.whole$death.interval)
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.90 = case_when(
+    death.interval > 90 ~ "6",
+    TRUE ~ mrs.90
+  ))
+
+# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days, tracheostomy
 

### 2022-10-07 study protocol.qmd
Update study protocol.qmd

@@
 erich.whole <- erich.whole %>%
-  mutate(Ethnicity = case_when(
+  mutate(ethnicity = case_when(
     D21A == "H" ~ "Hispanic or Latino",
@@
 erich.whole <- erich.whole %>%
-  mutate(symptom.onset = case_when(
+  mutate(symptom.onset.combined = case_when(
     is.na(symptom.onset.date.time) ~ symptom.onset.estimated.date.time,
@@
 erich.whole <- erich.whole %>%
-  mutate(nsgy.evacuation.date = case_when(
+  mutate(nsgy.evacuation.date.erich = case_when(
     IT5D == "1" ~ ymd_hms(paste(IT5D_DATE, IT5D_TIME)),
@@
   ))
+
+erich.whole$nsgy.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$nsgy.evacuation.date.erich)
+erich.whole$nsgy.interval <- time_length(erich.whole$nsgy.interval, unit = "days")
+erich.whole$nsgy.evacuation.date <- round(erich.whole$nsgy.interval, digits = 0) # Makes it compatabible to merge with ATACH
 
@@
 
-# Mortality Status
+# Mortality Status and Adding mRS 6
 
@@
 
-erich.whole <- erich.whole %>%
-  mutate(death.interval = as_date(death.date) - as_date(symptom.onset)) %>%
-  as.duration(death.interval)
-
-erich.whole$death.interval <- parse_integer(erich.whole$death.interval)
+erich.whole$death.interval <- erich.whole$symptom.onset.combined %--% erich.whole$death.date
+erich.whole$death.interval <- as.period(erich.whole$death.interval)
+erich.whole$death.interval <- time_length(erich.whole$death.interval, unit = "days")
 
@@
   mutate(mrs.90 = case_when(
-    death.interval > 90 ~ "6",
+    death.interval < 90 ~ 6,
     TRUE ~ mrs.90
+    ))
+
+erich.whole$mrs.90 <- na_if(erich.whole$mrs.90, 8)
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.180 = case_when(
+    death.interval < 180 ~ 6,
+    TRUE ~ mrs.180
+    ))
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.365 = case_when(
+    death.interval < 365 ~ 6,
+    TRUE ~ mrs.365
     ))

### 2022-11-04 study protocol.qmd
Updated Quarto Markdown

@@
 title: "Hemispheric Lateralization and Outcomes After ICH: A Pooled Analysis of Clinical Trials"
-subtitle: "Study Protocol and Statistical Analysis Plan"
+subtitle: "Study Protocol"
 author: "Nikhil Patel, MD"
-format: docx
+format: 
+  html:
+    toc: true
+    toc-location: left
+    toc-title: ""
+    css: styles.css
+    anchor-sections: true
+    code-fold: true
+    code-overflow: wrap
+    code-summary: "Show the code"
+    output: false
 ---
@@
 
-Patients with neurological injury experience different symptoms between the left and right hemispheres, and this may have.
+Patients with neurological injury experience different symptoms between the left and right hemispheres.
+This may affect prognostic estimates, and how frequently patients are offered aggressive care.
+In this study, we aimed to investigate any differential treatment patterns based on hemispheric lateralization, as well as differences in outcome with a pooled analysis of NINDS randomized trials enrolling patients with ICH.
 
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+**Loading packages and data and define helper functions**
+
+```{r Initial Setup}
+
 library(tidyverse)
@@
 
-```{r ATACH2, include=FALSE}
+**Importing and wrangling ATACH2 data
+
+```{r ATACH2}
 
@@
 
-```{r ERICH, include=FALSE}
+Importing and wrangling ERICH data
+
+```{r ERICH}
 
@@
 
-```{r combining ATACH2 and ERICH, include=FALSE}
+Combining Datasets
+
+```{r combining ATACH2 and ERICH}
 
@@
 
+table1 <- CreateTableOne(data = atach.erich,
+                         strata = "ich.laterality",)
+
+print(table1, showAllLevels = FALSE)
+
 ```

### 2022-11-16 study protocol.qmd
Update
More updates

@@
 library(lubridate)
+library(readxl)
 
@@
 
-**Importing and wrangling ATACH2 data
+**Importing and wrangling ATACH2 data**
 
@@
 atach.whole <- atach.whole %>%
-  mutate(trach = recode(F07Q11,
+  mutate(tracheostomy = recode(F07Q11,
                                    "0" = "No",
@@
                        "extubation.date", # Date of extubation
-                       "trach",
+                       "tracheostomy",
                        "trach.date", # Date of Tracheostomy
@@
 
-# Maybe import lab data, but this may be superfluous
-
 ```
 
-Importing and wrangling ERICH data
+**Importing and wrangling ERICH data**
 
@@
     IT8 == "2" ~ "No",
+    str_detect(erich.whole$IT5F_SPEC, "VENTR") ~ "Yes",
   ))
@@
     IT5H == "1" ~ "Yes",
+    str_detect(erich.whole$IT5F_SPEC, "CRANI") ~ "Yes",
+    str_detect(erich.whole$IT5F_SPEC, "EVAC") ~ "Yes",
     TRUE ~ "No"
@@
 
-# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days, tracheostomy
+# Adding tracheostomy information
+
+erich.whole <- erich.whole %>%
+  mutate(
+    tracheostomy = case_when(
+      str_detect(erich.whole$IT5F_SPEC, "TRACHEO") ~ "Yes",
+      ))
+
+# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days
 
@@
 
-Combining Datasets
+**Importing and wrangling iDEF data**
+
+```{r iDEF}
+
+# Create function to paste together string of formXX.sas7bdat
+idef.import <- function(y) {
+  dataset <- paste0("idef",".",y)
+  filepath <- paste0("./iDEF/form",y,".sas7bdat")
+  dataset <- read_sas(filepath)
+}
+
+idefdictionary <- read_sas("iDEF/datadictionary.sas7bdat")
+idef01 <- idef.import("01")
+idef02 <- idef.import("02")
+idef03 <- idef.import("03")
+idef03c <- idef.import("03c")
+idef04 <- idef.import("04")
+idef05 <- idef.import("05")
+idef06 <- idef.import("06")
+idef07 <- idef.import("07")
+idef08 <- idef.import("08")
+idef10c <- idef.import("10c")
+idef11 <- idef.import("11")
+idef15 <- idef.import("15")
+idef18 <- idef.import("18")
+idef19 <- idef.import("19")
+idef19c <- idef.import("19c")
+idef22 <- idef.import("22")
+idef23 <- idef.import("23")
+idef24 <- idef.import("24")
+idef25 <- idef.import("25")
+idef25c <- idef.import("25c")
+idef26 <- idef.import("26")
+idef27 <- idef.import("27")
+idef29 <- idef.import("29")
+idef30 <- idef.import("30")
+idef32 <- idef.import("32")
+idef32c <- idef.import("32c")
+idef43 <- idef.import("43")
+
+```
+
+**Importing and wrangling CLEAR-III data**
+
+```{r CLEAR III}
+
+# Importing Excel files
+clear3.demographics <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 1)
+clear3.outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 2)
+clear3.180outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_clinicaltrials.gov results_2018.1.30.xlsx", sheet = 1)
+
+# Combining all dataframes into one
+clear3.whole <- left_join(clear3.demographics, clear3.outcomes)
+clear3.whole <- left_join(clear3.whole, clear3.180outcomes)
+
+# Adding category to indicate study identity
+clear3.whole <- clear3.whole %>% mutate(study = "CLEAR III")
+
+# Categorizing and Recoding Variables
+
+clear3.whole <- clear3.whole %>%
+  mutate(race = case_when(
+    race == "African-American" ~ "Black or African-American",
+    race == "Asian" ~ "Asian",
+    race == "Hawaiian or Pacific" ~ "Native Hawaiian/Pacific Islander",
+    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
+    race == "Indian or Alaskan;White" ~ "American Indian or Alaskan Native",
+    race == "unknown/missing" ~ "Other/Not Reported",
+    race == "White" ~ "White"
+  ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(ethnicity = recode(hispanic_latino,
+                            "0" = "Not Hispanic or Latino",
+                            "1" = "Hispanic or Latino"
+  ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(sex = recode(gender,
+                         "female" = "Female",
+                         "male" = "Male"))
+
+clear3.whole <- clear3.whole %>%
+  mutate(ich.laterality = recode(hemisphere,
+                                 "left" = "Left",
+                                 "right" = "Right",
+                                 "unknown/missing" = "NA" ))
+
+clear3.whole <- clear3.whole %>% mutate(ivh = "Yes")
+
+clear3.whole <- clear3.whole %>%
+  rename(age = age_at_registration) %>%
+  rename(tobacco = tobacco_use) %>%
+  rename(sbp.initial = er_present_systolic) %>%
+  rename(dbp.initial = er_present_diastolic) %>%
+  rename(ich.location.baseline = adjudicated_ich_location) %>%
+  rename(gcs.baseline = er_present_gcs_total) %>%
+  rename(nihss.baseline = er_present_nihss_total) %>%
+  rename(ich.volume.cc = dct_ich_volume_rc) %>%
+  rename(ivh.volume.cc = dct_ivh_volume_rc) %>%
+  rename(mrs.30 = glasgow_rankin_30) %>%
+  rename(mrs.180 = glasgow_rankin_180) %>%
+  rename(gose.30 = gose_30) %>%
+  rename(gose.180 = gose_180) %>%
+  rename(comfort.binary = withdrawal_contribution) %>%
+  rename(comfort.date.time = ictus_to_withdrawal_days) %>%
+  rename(symptom.onset.to.presentation = ictus_to_presentation)
+
+```
+
+**Importing and Wrangling MISTIE-II data**
+
+```{r MISTIE II}
+
+# Importing MISTIE II data
+
+mistie2.demographics <- readxl::read_excel("./MISTIE2/MISTIE II_Publication results_2016.4.14_RT_RD updates 2017.10.2.xlsx", sheet = 1)
+
+mistie2.demographics2 <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_Demographics data_2017.10.16.xlsx", sheet = 1)
+
+mistie2.results <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_ctgov results_2017.10.17.xlsx", sheet = 3)
+
+mistie2.demographics3 <- readxl::read_excel("./MISTIE2/ICES_Publication results_2015.12.23_RT_RD updates 2017.10.16.xlsx", sheet = 1)
+
+mistie2.combined.demographics <- full_join(mistie2.demographics, mistie2.demographics3)
+mistie2.combined.demographics <- full_join(mistie2.combined.demographics, mistie2.demographics2)
+
+mistie2.wide.results <- mistie2.results %>%
+  pivot_wider(names_from = Follow_up_Visit, values_from = c(GOS, eGOS_Score, rankin_score, SIS_Total, SIS_16_Total))
+
+mistie2.whole <- full_join(mistie2.wide.results, mistie2.combined.demographics)
+
+mistie2.whole <- mistie2.whole %>% mutate(study = "MISTIE II")
+
+# Renaming and Recoding Variables
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(ethnicity = case_when(
+    Race == "African American not Hispanic" ~ "Not Hispanic or Latino",
+    Race == "Hispanic" ~ "Hispanic or Latino",
+    Race == "Asian or Pacific Islander" ~ "Not Hispanic or Latino",
+    Race == "Caucasian not Hispanic" ~ "Not Hispanic or Latino",
+    Race == "Other or unknown" ~ "Not Hispanic or Latino"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(race = recode(Race,
+                       "African American not Hispanic" = "Black or African American",
+                       "Asian or Pacific Islander" = "Asian",
+                       "Caucasian not Hispanic" = "White",
+                       "Hispanic" = "Other/Not Reported",
+                       "Other or unknown" = "Other/Not Reported"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  rename(sbp.initial = er_sbp) %>%
+  rename(dbp.initial = er_dbp) %>%
+  rename(gcs.baseline = enrollment_gcs) %>%
+  rename(nihss.baseline = enrollment_nihss_total) %>%
+  rename(ich.volume.cc = pre_rand_ichvol) %>%
+  rename(ivh.volume.cc = pre_rand_ivhvol) %>%
+  rename(gos.365 = GOS_365) %>%
+  rename(gos.270 = GOS_270) %>%
+  rename(gos.180 = GOS_180) %>%
+  rename(gos.90 = GOS_90) %>%
+  rename(gos.30 = GOS_30) %>%
+  rename(egos.365 = eGOS_Score_365) %>%
+  rename(egos.270 = eGOS_Score_270) %>%
+  rename(egos.180 = eGOS_Score_180) %>%
+  rename(egos.90 = eGOS_Score_90) %>%
+  rename(egos.30 = eGOS_Score_30) %>%
+  rename(mrs.365 = rankin_score_365) %>%
+  rename(mrs.270 = rankin_score_270) %>%
+  rename(mrs.180 = rankin_score_180) %>%
+  rename(mrs.90 = rankin_score_90) %>%
+  rename(mrs.30 = rankin_score_30)
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(ivh = recode(any_ivh_prerand,
+                      "0" = "No",
+                      "1" = "Yes"
+  ))
+
+```
+
+**Importing and Wrangling MISTIE-III data**
+
+```{r MISTIE-III}
+
+# Importing MISTIE III data
+mistie3 <- read.table("./MISTIE3/Deidentified_Mistie_III_Wide_Analytic_2019_07_11 (1).csv", sep = ",", header = TRUE)
+
+mistie3 <- mistie3 %>% mutate(study = "MISTIE III")
+
+mistie3 <- mistie3 %>%
+  rename(age = age_at_consent) %>%
+  rename(race = race_ninds) %>%
+  rename(sbp.initial = er_present_systolic) %>%
+  rename(dbp.initial = er_present_diastolic) %>%
+  rename(gcs.baseline = gcs_total) %>%
+  rename(nihss.baseline = nihss_total) %>%
+  rename(ich.volume.cc = diagct_ich_volume) %>%
+  rename(ivh.volume.cc = diagct_ivh_volume) %>%
+  rename(ich.laterality = ich_hemisphere_adj) %>%
+  rename(mrs.30 = glasgow_rankin_30) %>%
+  rename(mrs.180 = glasgow_rankin_180) %>%
+  rename(mrs.365 = glasgow_rankin_365) %>%
+  rename(barthel.total.365 = barthel_index_365) %>%
+  rename(eq.vas.365 = eq_vas_365) %>%
+  rename(time.to.presentation = ictus_to_site_arrival_hours)
+
+mistie3 <- mistie3 %>%
+  mutate(race = case_when(
+    race == "African-American" ~ "Black of African-American",
+    race == "Asian Hawaiian or Pacific" ~ "Asian",
+    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
+    race == "More than one race" ~ "Other/Not Reported",
+    race == "White" ~ "White"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(ethnicity = case_when(
+         hispanic == "0" ~ "Not Hispanic or Latino",
+         hispanic == "1" ~ "Hispanic or Latino"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(sex = case_when(
+    male_gender == "0" ~ "Woman",
+    male_gender == "1" ~ "Man"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(tobacco = case_when(
+    current_smoker == "0" ~ "No",
+    current_smoker == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(dm2 = case_when(
+    diabetes_pmh == "0" ~ "No",
+    diabetes_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(htn = case_when(
+    htn_pmh == "0" ~ "No",
+    htn_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(cad = case_when(
+    cvd_pmh == "0" ~ "No",
+    cvd_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(comfort = case_when(
+    withdrawal_contribution == "0" ~ "No",
+    withdrawal_contribution == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.mobility.365 = case_when(
+    eq5d_mobility_365 == "1" ~ "No Problems",
+    eq5d_mobility_365 == "2" ~ "Some Problems",
+    eq5d_mobility_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.selfcare.365 = case_when(
+    eq5d_selfcare_365 == "1" ~ "No Problems",
+    eq5d_selfcare_365 == "2" ~ "Some Problems",
+    eq5d_selfcare_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.usual.365 = case_when(
+    eq5d_activities_365 == "1" ~ "No Problems",
+    eq5d_activities_365 == "2" ~ "Some Problems",
+    eq5d_activities_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.pain.365 = case_when(
+    eq5d_pain_365 == "1" ~ "No Problems",
+    eq5d_pain_365 == "2" ~ "Some Problems",
+    eq5d_pain_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.anxiety.365 = case_when(
+    eq5d_anxiety_365 == "1" ~ "No Problems",
+    eq5d_anxiety_365 == "2" ~ "Some Problems",
+    eq5d_anxiety_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.30 = case_when(
+    location_d30 == "Acute - Hospital" ~ "Hospital",
+    location_d30 == "Acute - ICU" ~ "Hospital",
+    location_d30 == "Dead" ~ "Dead",
+    location_d30 == "Home" ~ "Home",
+    location_d30 == "Rehab" ~ "Rehabilitation",
+    location_d30 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d30 == "Hospice" ~ "Hospice",
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.180 = case_when(
+    location_d180 == "Acute - Hospital" ~ "Hospital",
+    location_d180 == "Acute - ICU" ~ "Hospital",
+    location_d180 == "Dead" ~ "Dead",
+    location_d180 == "Home" ~ "Home",
+    location_d180 == "Rehab" ~ "Rehabilitation",
+    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d180 == "Hospice" ~ "Hospice",
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.365 = case_when(
+    location_d365 == "Acute - Hospital" ~ "Hospital",
+    location_d365 == "Acute - ICU" ~ "Hospital",
+    location_d365 == "Dead" ~ "Dead",
+    location_d365 == "Home" ~ "Home",
+    location_d365 == "Rehab" ~ "Rehabilitation",
+    location_d365 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d365 == "Hospice" ~ "Hospice",
+  ))
+
+```
+
+**Combining Datasets**
 

### 2022-11-16 study protocol.qmd
Combined all datasets

@@
 
-# Table 1
+# Renaming variables
 
@@
   rename(age = F33Q02) %>%
-  rename(symptom.onset = F00Q02MIN) %>%
+  rename(symptom.onset = F00Q02MIN) %>% # All times in ATACH2 dataset are expressed as minutes from randomization
+  rename(arrivaltime.initial = F02Q03MIN) %>% 
+  rename(arrivaltime.strokecenter = F02Q05MIN) %>%
   rename(sbp.initial = F05Q07A) %>%
@@
 
-VarsTableOne <- c("study",
-                  "age", # Age
-                       "sex", # Sex
-                       "ethnicity", # Ethnicity
+# Calculating Time to ED Presentation as expressed by minutes and hours from symptom onset
+
+atach.whole <- atach.whole %>%
+  mutate(time.to.ed.minutes = abs(arrivaltime.strokecenter - symptom.onset)) %>%
+  mutate(time.to.ed.hours = time.to.ed.minutes / 60)
+
+atach.whole$time.to.ed.hours <- round(atach.whole$time.to.ed.hours, digits = 2)
+
+atach.times <- atach.whole %>%
+  select(c(symptom.onset, arrivaltime.initial, arrivaltime.strokecenter, time.to.ed.minutes, time.to.ed.hours))
+
+# DNR and WLST status
+
+atach.whole <- atach.whole %>%
+  mutate(dnr.binary = case_when(
+    F21Q08 == "0" ~ "No",
+    F21Q08 == "1" ~ "Yes"
+  ))
+
+atach.whole <- atach.whole %>%
+  mutate(comfort.binary = case_when(
+    F21Q06 == "0" ~ "No",
+    F21Q06 == "1" ~ "Yes"
+  ))
+
+atach.whole <- atach.whole %>%
+  rename(comfort.day = F21Q07DAY) %>%
+  rename(dnr.day = F21Q09DAY)
+
+```
+
+**Importing and wrangling ERICH data**
+
+```{r ERICH}
+
+erich.case <- read_sas("ERICH/erich_case_100217.sas7bdat")
+
+erich.whole <- erich.case %>%
+  select(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30, O3, O3A, O5, O20, AU8, AU8_OTH, RANKIN_FU1, RANKIN_FU2, RANKIN_FU3, MOBILITY_FU1, MOBILITY_FU2, MOBILITY_FU3, SELFCARE_FU1, SELFCARE_FU2, SELFCARE_FU3, USUALACT_FU1, USUALACT_FU2, USUALACT_FU3, PAIN_FU1, PAIN_FU2, PAIN_FU3, ANXIETY_FU1, ANXIETY_FU2, ANXIETY_FU3, HLTHSTAT_FU1, HLTHSTAT_FU2, HLTHSTAT_FU3, FEED_FU1, FEED_FU2, FEED_FU3, BATHE_FU1, BATHE_FU2, BATHE_FU3, GROOM_FU1, GROOM_FU2, GROOM_FU3, DRESS_FU1, DRESS_FU2, DRESS_FU3, BOWELS_FU1, BOWELS_FU2, BOWELS_FU3, BLADDER_FU1, BLADDER_FU2, BLADDER_FU3, TOILET_FU1, TOILET_FU2, TOILET_FU3, TRANSFER_FU1, TRANSFER_FU2, TRANSFER_FU3, MOBLEVEL_FU1, MOBLEVEL_FU2, MOBLEVEL_FU3, STAIRS_FU1, STAIRS_FU2, STAIRS_FU3, BART_TOT_FU1, BART_TOT_FU2, BART_TOT_FU3, EMS33, ED44, CURR_LOC_FU1, CURR_LOC_FU2, CURR_LOC_FU3, IT2, O1, O2_DATE, O2_TIME, O3, O3_OTH, DOD)
+
+erich.whole <- erich.whole %>%
+  mutate(study = "ERICH")
+
+erich.whole <- erich.whole %>%
+  mutate(enrolling.country = "United States")
+
+# Renaming and Recoding Variables
+
+erich.whole <- erich.whole %>%
+  mutate(gender = recode(gender,
+                         "F" = "Female",
+                         "M" = "Male")) %>%
+  rename(sex = gender)
+
+erich.whole <- erich.whole %>%
+  mutate(Race = case_when(
+    D21 == "W" ~ "White",
+    D21 == "B" ~ "Black or African-American",
+    D21 == "A" ~ "Asian",
+    D21 == "N" ~ "Native Hawaiian/Pacific Islander",
+    D21 == "I" ~ "American Indian or Alaskan Native",
+    D21 == "O" ~ "Other/Not Reported",
+    D21 == "U" ~ "Other/Not Reported"
+    )) %>%
+  rename(race = Race)
+
+erich.whole <- erich.whole %>%
+  mutate(ethnicity = case_when(
+    D21A == "H" ~ "Hispanic or Latino",
+    D21A == "N" ~ "Not Hispanic or Latino",
+    D21A == "U" ~ "Not Hispanic or Latino"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(stroke = recode(SH1,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(cad = recode(MHX8,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(chf = recode(MHX13,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(afib = recode(MHX11,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(pvd = recode(MHX27C,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(hld = recode(MHX6,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(dm2 = case_when(
+    MHX3 == "2" ~ "No",
+    MHX3 == "8" ~ "NA",
+    MHX3A == "1" ~ "No",
+    MHX3A == "2" ~ "Yes",
+    MHX3A == "8" ~ "NA"))
+
+erich.whole <- erich.whole %>%
+  mutate(tobacco = recode(SMK1A,
+                         "2" = "No",
+                         "1" = "Yes",
+                         "5" = "NA",
+                         "8" = "NA"))
+
+erich.whole <- erich.whole %>%
+  rename(sbp.initial = ED47_SYS)
+
+erich.whole <- erich.whole %>%
+  rename(dbp.initial = ED47_DIAS)
+
+erich.whole <- erich.whole %>%
+  rename(gcs.baseline = ED39)
+
+erich.whole <- erich.whole %>%
+  mutate(ich.laterality = recode(PHY6A,
+                                 "L" = "Left",
+                                 "R" = "Right",
+                                 "M" = "NA",
+                                 .default = "NA"))
+
+# ICH Location
+erich.whole <- erich.whole %>%
+  mutate(
+    ich.location.baseline = case_when(
+      str_detect(erich.whole$detail, "Ganglia") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Brainstem") ~ "Brainstem",
+      str_detect(erich.whole$detail, "Cerebell") ~ "Cerebellum",
+      str_detect(erich.whole$detail, "Front") ~ "Lobar",
+      str_detect(erich.whole$detail, "Tempor") ~ "Lobar",
+      str_detect(erich.whole$detail, "Occi") ~ "Lobar",
+      str_detect(erich.whole$detail, "Perivent") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Caudate") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Putam") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Deep") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Lobar") ~ "Lobar",
+      str_detect(erich.whole$detail, "IVH") ~ "Primary IVH",
+      str_detect(erich.whole$detail, "Hemis") ~ "Lobar",
+      str_detect(erich.whole$detail, "Thal") ~ "Thalamus",
+      str_detect(erich.whole$detail, "Ganglia") ~ "Basal Ganglia",
+      str_detect(erich.whole$detail, "Corona") ~ "Lobar",
+      str_detect(erich.whole$detail, "applic") ~ "NA",
+      detail == "Unknown" ~ "NA"
+      ))
+
+erich.whole <- erich.whole %>%
+  mutate(ivh = case_when(erich.case$IVH_Vol_CT1 > 0 ~ "Yes",
+                         erich.case$IVH_Vol_CT1 < 0.1 ~ "No"))
+
+erich.whole <- erich.whole %>%
+  rename(ich.volume.cc = ICH_Vol_CT1)
+
+erich.whole <- erich.whole %>%
+  rename(ivh.volume.cc = IVH_Vol_CT1)
+
+erich.whole <- erich.whole %>%
+  mutate(mech.ventilation = case_when(
+    EMS33 == "1" ~ "Yes",
+    EMS33 == "2" ~ "No",
+    ED44 == "1" ~ "Yes",
+    ED44 == "2" ~ "No",
+    IT7 == "1" ~ "Yes",
+    IT7 == "2" ~ "No"))
+
+erich.whole <- erich.whole %>%
+  mutate(evd = case_when(
+    IT8 == "1" ~ "Yes",
+    IT8 == "2" ~ "No",
+    str_detect(erich.whole$IT5F_SPEC, "VENTR") ~ "Yes",
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(nsgy.evacuation = case_when(
+    IT5D == "1" ~ "Yes",
+    IT5H == "1" ~ "Yes",
+    str_detect(erich.whole$IT5F_SPEC, "CRANI") ~ "Yes",
+    str_detect(erich.whole$IT5F_SPEC, "EVAC") ~ "Yes",
+    TRUE ~ "No"
+      ))
+
+erich.whole <- erich.whole %>%
+  rename(mrs.90 = RANKIN_FU1) %>%
+  rename(mrs.180 = RANKIN_FU2) %>%
+  rename(mrs.365 = RANKIN_FU3) %>%
+  rename(euro.vas.90 = HLTHSTAT_FU1) %>%
+  rename(euro.vas.180 = HLTHSTAT_FU2) %>%
+  rename(euro.vas.365 = HLTHSTAT_FU3) %>%
+  rename(barthel.feeding.90 = FEED_FU1) %>%
+  rename(barthel.bathing.90 = BATHE_FU1) %>%
+  rename(barthel.grooming.90 = GROOM_FU1) %>%
+  rename(barthel.dressing.90 = DRESS_FU1) %>%
+  rename(barthel.bowels.90 = BOWELS_FU1) %>%
+  rename(barthel.bladder.90 = BLADDER_FU1) %>%
+  rename(barthel.toilet.90 = TOILET_FU1) %>%
+  rename(barthel.transfer.90 = TRANSFER_FU1) %>%
+  rename(barthel.mobility.90 = MOBLEVEL_FU1) %>%
+  rename(barthel.stairs.90 = STAIRS_FU1) %>%
+  rename(barthel.total.90 = BART_TOT_FU1) %>%
+  rename(barthel.feeding.180 = FEED_FU2) %>%
+  rename(barthel.bathing.180 = BATHE_FU2) %>%
+  rename(barthel.grooming.180 = GROOM_FU2) %>%
+  rename(barthel.dressing.180 = DRESS_FU2) %>%
+  rename(barthel.bowels.180 = BOWELS_FU2) %>%
+  rename(barthel.bladder.180 = BLADDER_FU2) %>%
+  rename(barthel.toilet.180 = TOILET_FU2) %>%
+  rename(barthel.transfer.180 = TRANSFER_FU2) %>%
+  rename(barthel.mobility.180 = MOBLEVEL_FU2) %>%
+  rename(barthel.stairs.180 = STAIRS_FU2) %>%
+  rename(barthel.total.180 = BART_TOT_FU2) %>%
+  rename(barthel.feeding.365 = FEED_FU3) %>%
+  rename(barthel.bathing.365 = BATHE_FU3) %>%
+  rename(barthel.grooming.365 = GROOM_FU3) %>%
+  rename(barthel.dressing.365 = DRESS_FU3) %>%
+  rename(barthel.bowels.365 = BOWELS_FU3) %>%
+  rename(barthel.bladder.365 = BLADDER_FU3) %>%
+  rename(barthel.toilet.365 = TOILET_FU3) %>%
+  rename(barthel.transfer.365 = TRANSFER_FU3) %>%
+  rename(barthel.mobility.365 = MOBLEVEL_FU3) %>%
+  rename(barthel.stairs.365 = STAIRS_FU3) %>%
+  rename(barthel.total.365 = BART_TOT_FU3)
+
+erich.whole <- erich.whole %>%
+  mutate(current.location.90 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      )) %>%
+  mutate(current.location.180 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      )) %>%
+  mutate(current.location.365 = recode(CURR_LOC_FU1,
+                      "1" = "Home",
+                      "2" = "Rehabilitation",
+                      "3" = "Skilled Nursing Facility",
+                      "4" = "Hospital",
+                      "7" = "Other"
+                      ))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.90 = factor(MOBILITY_FU1,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.90 = factor(SELFCARE_FU1,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.90 = factor(USUALACT_FU1,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.90 = factor(PAIN_FU1,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.90 = factor(ANXIETY_FU1,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed", "NA")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.180 = factor(MOBILITY_FU2,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.180 = factor(SELFCARE_FU2,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.180 = factor(USUALACT_FU2,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.180 = factor(PAIN_FU2,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.180 = factor(ANXIETY_FU2,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.mobility.365 = factor(MOBILITY_FU3,
+                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.selfcare.365 = factor(SELFCARE_FU3,
+                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.usual.365 = factor(USUALACT_FU3,
+                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.pain.365 = factor(PAIN_FU3,
+                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
+
+erich.whole <- erich.whole %>%
+  mutate(euro.anxiety.365 = factor(ANXIETY_FU3,
+                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
+
+# Parsing Dates and Times in ERICH
+
+# symptom.onset
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$SE2)))
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.estimated.time = case_when( # Converting estimated times to numbers
+    SE3 == "1" ~ "06:00:00",
+    SE3 == "2" ~ "NA",
+    SE3 == "3" ~ "03:00:00",
+    SE3 == "4" ~ "09:00:00",
+    SE3 == "5" ~ "15:00:00",
+    SE3 == "6" ~ "21:00:00"
+  ))
+
+erich.whole$symptom.onset.estimated.time <- hms(erich.whole$symptom.onset.estimated.time)
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.estimated.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$symptom.onset.estimated.time)))
+
+erich.whole <- erich.whole %>%
+  mutate(symptom.onset.combined = case_when(
+    is.na(symptom.onset.date.time) ~ symptom.onset.estimated.date.time,
+    TRUE ~ symptom.onset.date.time
+  ))
+
+# Time of ED presentation
+
+erich.whole <- erich.whole %>%
+  mutate(ed.presentation.date.time = ymd_hms(paste(erich.whole$ED7, erich.whole$ED8)))
+
+# Calculating Time to ED Presentation as expressed by minutes and hours from symptom onset
+
+erich.whole$symptom.onset.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$ed.presentation.date.time)
+erich.whole$time.to.ed.minutes <- time_length(erich.whole$symptom.onset.interval, unit = "minutes")
+erich.whole$time.to.ed.hours <- time_length(erich.whole$symptom.onset.interval, unit = "hours")
+
+erich.whole$time.to.ed.minutes[erich.whole$time.to.ed.minutes < 0] = NA
+erich.whole$time.to.ed.hours[erich.whole$time.to.ed.hours < 0] = NA
+
+# Neurosurgery Dates
+
+erich.whole <- erich.whole %>%
+  mutate(nsgy.evacuation.date.erich = case_when(
+    IT5D == "1" ~ ymd_hms(paste(IT5D_DATE, IT5D_TIME)),
+    IT5H == "1" ~ ymd_hms(paste(IT5H_DATE, IT5H_TIME)),
+    IT5F == "1" ~ ymd_hms(paste(IT5F_DATE, IT5F_TIME))
+  ))
+
+erich.whole$nsgy.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$nsgy.evacuation.date.erich)
+erich.whole$nsgy.interval <- time_length(erich.whole$nsgy.interval, unit = "days")
+erich.whole$nsgy.evacuation.date <- round(erich.whole$nsgy.interval, digits = 0) # Makes it compatible to merge with ATACH
+
+# DNR and WLST Status
+
+erich.whole <- erich.whole %>%
+  mutate(dnr.binary = case_when(
+    CC16 == "1" ~ "Yes",
+    CC16 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(dni.binary = case_when(
+    CC17 == "1" ~ "Yes",
+    CC17 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(comfort.binary = case_when(
+    CC18 == "1" ~ "Yes",
+    CC18 == "2" ~ "No",
+    TRUE ~ "NA"
+  ))
+
+erich.whole <- erich.whole %>%
+  mutate(dnr.date.time = case_when(
+    CC16 == "1" ~ ymd_hms(paste(CC16_DATE, CC16_TIME))
+  ))
+
+erich.whole$dnr.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$dnr.date.time)
+erich.whole$dnr.interval <- time_length(erich.whole$dnr.interval, unit = "days")
+erich.whole$dnr.day <- round(erich.whole$dnr.interval, digits = 0)
+
+erich.whole <- erich.whole %>%
+  mutate(dni.date.time = case_when(
+    CC17 == "1" ~ ymd_hms(paste(CC17_DATE, CC17_TIME))
+  ))
+
+erich.whole$dni.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$dni.date.time)
+erich.whole$dni.interval <- time_length(erich.whole$dni.interval, unit = "days")
+erich.whole$dni.day <- round(erich.whole$dni.interval, digits = 0)
+
+erich.whole <- erich.whole %>%
+  mutate(comfort.date.time = case_when(
+    CC18 == "1" ~ ymd_hms(paste(CC18_DATE, CC18_TIME))
+  ))
+
+erich.whole$comfort.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$comfort.date.time)
+erich.whole$comfort.interval <- time_length(erich.whole$comfort.interval, unit = "days")
+erich.whole$comfort.day <- round(erich.whole$comfort.interval, digits = 0)
+
+# Mortality Status and Adding mRS 6
+
+erich.whole$death.date <- ymd(erich.whole$DOD)
+
+erich.whole$death.interval <- erich.whole$symptom.onset.combined %--% erich.whole$death.date
+erich.whole$death.interval <- as.period(erich.whole$death.interval)
+erich.whole$death.interval <- time_length(erich.whole$death.interval, unit = "days")
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.90 = case_when(
+    death.interval < 90 ~ 6,
+    TRUE ~ mrs.90
+    ))
+
+erich.whole$mrs.90 <- na_if(erich.whole$mrs.90, 8)
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.180 = case_when(
+    death.interval < 180 ~ 6,
+    TRUE ~ mrs.180
+    ))
+
+erich.whole <- erich.whole %>%
+  mutate(mrs.365 = case_when(
+    death.interval < 365 ~ 6,
+    TRUE ~ mrs.365
+    ))
+
+# Adding tracheostomy information
+
+erich.whole <- erich.whole %>%
+  mutate(
+    tracheostomy = case_when(
+      str_detect(erich.whole$IT5F_SPEC, "TRACHEO") ~ "Yes",
+      ))
+
+# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days
+
+```
+
+**Importing and wrangling iDEF data**
+
+```{r iDEF}
+
+# Create function to paste together string of formXX.sas7bdat
+idef.import <- function(y) {
+  dataset <- paste0("idef",".",y)
+  filepath <- paste0("./iDEF/form",y,".sas7bdat")
+  dataset <- read_sas(filepath)
+}
+
+idefdictionary <- read_sas("iDEF/datadictionary.sas7bdat")
+idef01 <- idef.import("01")
+idef02 <- idef.import("02")
+idef03 <- idef.import("03")
+idef03c <- idef.import("03c")
+idef04 <- idef.import("04")
+idef05 <- idef.import("05")
+idef06 <- idef.import("06")
+idef07 <- idef.import("07")
+idef08 <- idef.import("08")
+idef10c <- idef.import("10c")
+idef11 <- idef.import("11")
+idef15 <- idef.import("15")
+idef18 <- idef.import("18")
+idef19 <- idef.import("19")
+idef19c <- idef.import("19c")
+idef22 <- idef.import("22")
+idef23 <- idef.import("23")
+idef24 <- idef.import("24")
+idef25 <- idef.import("25")
+idef25c <- idef.import("25c")
+idef26 <- idef.import("26")
+idef27 <- idef.import("27")
+idef29 <- idef.import("29")
+idef30 <- idef.import("30")
+idef32 <- idef.import("32")
+idef32c <- idef.import("32c")
+idef43 <- idef.import("43")
+
+```
+
+**Importing and wrangling CLEAR-III data**
+
+```{r CLEAR III}
+
+# Importing Excel files
+clear3.demographics <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 1)
+clear3.outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 2)
+clear3.180outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_clinicaltrials.gov results_2018.1.30.xlsx", sheet = 1)
+
+# Combining all dataframes into one
+clear3.whole <- left_join(clear3.demographics, clear3.outcomes)
+clear3.whole <- left_join(clear3.whole, clear3.180outcomes)
+
+# Adding category to indicate study identity
+clear3.whole <- clear3.whole %>% mutate(study = "CLEAR III")
+
+# Renaming Variables
+
+clear3.whole <- clear3.whole %>%
+  rename(age = age_at_registration) %>%
+  rename(sbp.initial = er_present_systolic) %>%
+  rename(dbp.initial = er_present_diastolic) %>%
+  rename(ich.location.baseline = adjudicated_ich_location) %>%
+  rename(gcs.baseline = er_present_gcs_total) %>%
+  rename(nihss.baseline = er_present_nihss_total) %>%
+  rename(ich.volume.cc = dct_ich_volume_rc) %>%
+  rename(ivh.volume.cc = dct_ivh_volume_rc) %>%
+  rename(mrs.30 = glasgow_rankin_30) %>%
+  rename(mrs.180 = glasgow_rankin_180) %>%
+  rename(gose.30 = gose_30) %>%
+  rename(gose.180 = gose_180) %>%
+  rename(comfort.day = ictus_to_withdrawal_days)
+
+# Categorizing and Recoding Variables
+
+clear3.whole <- clear3.whole %>%
+  mutate(ich.location.baseline = case_when(
+    ich.location.baseline == "Caudate" ~ "Basal Ganglia",
+    ich.location.baseline == "Frontal" ~ "Lobar",
+    ich.location.baseline == "Globus Pallidus" ~ "Basal Ganglia",
+    ich.location.baseline == "Occipital" ~ "Lobar",
+    ich.location.baseline == "Other" ~ "NA",
+    ich.location.baseline == "Probable Primary IVH" ~ "Primary IVH",
+    ich.location.baseline == "Putamen" ~ "Basal Ganglia",
+    ich.location.baseline == "Temporal" ~ "Lobar",
+    ich.location.baseline == "Thalamus" ~ "Thalamus"
+  ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(race = case_when(
+    race == "African-American" ~ "Black or African-American",
+    race == "Asian" ~ "Asian",
+    race == "Hawaiian or Pacific" ~ "Native Hawaiian/Pacific Islander",
+    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
+    race == "Indian or Alaskan;White" ~ "American Indian or Alaskan Native",
+    race == "unknown/missing" ~ "Other/Not Reported",
+    race == "White" ~ "White"
+  ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(ethnicity = recode(hispanic_latino,
+                            "0" = "Not Hispanic or Latino",
+                            "1" = "Hispanic or Latino"
+  ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(sex = recode(gender,
+                         "female" = "Female",
+                         "male" = "Male"))
+
+clear3.whole <- clear3.whole %>%
+  mutate(ich.laterality = recode(hemisphere,
+                                 "left" = "Left",
+                                 "right" = "Right",
+                                 "unknown/missing" = "NA" ))
+
+clear3.whole <- clear3.whole %>%
+  mutate(tobacco = recode(tobacco_use,
+                          "1" = "Yes",
+                          "NA" = "No"))
+
+clear3.whole <- clear3.whole %>% mutate(ivh = "Yes")
+
+clear3.whole <- clear3.whole %>%
+  mutate(comfort.binary = case_when(
+    withdrawal_contribution == "on" ~ "Yes",
+    TRUE ~ "No"
+  ))
+
+# Recalculating Time to ED Presentation
+
+clear3.whole <- clear3.whole %>%
+  mutate(time.to.ed.minutes = ictus_to_presentation*24*60) %>%
+  mutate(time.to.ed.hours = ictus_to_presentation*24)
+
+clear3.whole$time.to.ed.minutes <- round(clear3.whole$time.to.ed.minutes, digits = 0)
+clear3.whole$time.to.ed.hours <- round(clear3.whole$time.to.ed.hours, digits = 2)
+
+```
+
+**Importing and Wrangling MISTIE-II data**
+
+```{r MISTIE II}
+
+# Importing MISTIE II data
+
+mistie2.demographics <- readxl::read_excel("./MISTIE2/MISTIE II_Publication results_2016.4.14_RT_RD updates 2017.10.2.xlsx", sheet = 1)
+
+mistie2.demographics2 <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_Demographics data_2017.10.16.xlsx", sheet = 1)
+
+mistie2.results <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_ctgov results_2017.10.17.xlsx", sheet = 3)
+
+mistie2.demographics3 <- readxl::read_excel("./MISTIE2/ICES_Publication results_2015.12.23_RT_RD updates 2017.10.16.xlsx", sheet = 1)
+
+mistie2.combined.demographics <- left_join(mistie2.demographics3, mistie2.demographics)
+mistie2.combined.demographics <- left_join(mistie2.demographics2, mistie2.combined.demographics)
+
+mistie2.wide.results <- mistie2.results %>%
+  pivot_wider(names_from = Follow_up_Visit, values_from = c(GOS, eGOS_Score, rankin_score, SIS_Total, SIS_16_Total))
+
+mistie2.whole <- full_join(mistie2.wide.results, mistie2.combined.demographics)
+
+mistie2.whole <- mistie2.whole %>% mutate(study = "MISTIE II")
+
+# Renaming and Recoding Variables
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(ethnicity = case_when(
+    Race == "African American not Hispanic" ~ "Not Hispanic or Latino",
+    Race == "Hispanic" ~ "Hispanic or Latino",
+    Race == "Asian or Pacific Islander" ~ "Not Hispanic or Latino",
+    Race == "Caucasian not Hispanic" ~ "Not Hispanic or Latino",
+    Race == "Other or unknown" ~ "Not Hispanic or Latino"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(race = recode(Race,
+                       "African American not Hispanic" = "Black or African-American",
+                       "Asian or Pacific Islander" = "Asian",
+                       "Caucasian not Hispanic" = "White",
+                       "Hispanic" = "Other/Not Reported",
+                       "Other or unknown" = "Other/Not Reported"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(ich.location.baseline = case_when(
+    clot_location_rc == "Globus Palidus" ~ "Basal Ganglia",
+    clot_location_rc == "Lobar" ~ "Lobar",
+    clot_location_rc == "Putamen" ~ "Basal Ganglia",
+    clot_location_rc == "Thalamus" ~ "Thalamus"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  mutate(ivh = recode(any_ivh_prerand,
+                      "1" = "Yes",
+                      "0" = "No",
+                      .default = "NA"
+  ))
+
+mistie2.whole <- mistie2.whole %>%
+  rename(sbp.initial = er_sbp) %>%
+  rename(dbp.initial = er_dbp) %>%
+  rename(gcs.baseline = enrollment_gcs) %>%
+  rename(nihss.baseline = enrollment_nihss_total) %>%
+  rename(ich.volume.cc = pre_rand_ichvol) %>%
+  rename(ivh.volume.cc = pre_rand_ivhvol) %>%
+  rename(htn = Hypertension) %>%
+  rename(hld = Hyperlipidemia) %>%
+  rename(dm2 = Diabetes) %>%
+  rename(tobacco = Tobacco_Use) %>%
+  rename(cad = Other_CV) %>%
+  rename(gos.365 = GOS_365) %>%
+  rename(gos.270 = GOS_270) %>%
+  rename(gos.180 = GOS_180) %>%
+  rename(gos.90 = GOS_90) %>%
+  rename(gos.30 = GOS_30) %>%
+  rename(egos.365 = eGOS_Score_365) %>%
+  rename(egos.270 = eGOS_Score_270) %>%
+  rename(egos.180 = eGOS_Score_180) %>%
+  rename(egos.90 = eGOS_Score_90) %>%
+  rename(egos.30 = eGOS_Score_30) %>%
+  rename(mrs.365 = rankin_score_365) %>%
+  rename(mrs.270 = rankin_score_270) %>%
+  rename(mrs.180 = rankin_score_180) %>%
+  rename(mrs.90 = rankin_score_90) %>%
+  rename(mrs.30 = rankin_score_30) %>%
+  rename(ich.laterality = hemisphere)
+
+```
+
+**Importing and Wrangling MISTIE-III data**
+
+```{r MISTIE-III}
+
+# Importing MISTIE III data
+mistie3 <- read.table("./MISTIE3/Deidentified_Mistie_III_Wide_Analytic_2019_07_11 (1).csv", sep = ",", header = TRUE)
+
+mistie3 <- mistie3 %>% mutate(study = "MISTIE III")
+
+mistie3 <- mistie3 %>%
+  rename(age = age_at_consent) %>%
+  rename(race = race_ninds) %>%
+  rename(sbp.initial = er_present_systolic) %>%
+  rename(dbp.initial = er_present_diastolic) %>%
+  rename(gcs.baseline = gcs_total) %>%
+  rename(nihss.baseline = nihss_total) %>%
+  rename(ich.volume.cc = diagct_ich_volume) %>%
+  rename(ivh.volume.cc = diagct_ivh_volume) %>%
+  rename(ich.laterality = ich_hemisphere_adj) %>%
+  rename(mrs.30 = glasgow_rankin_30) %>%
+  rename(mrs.180 = glasgow_rankin_180) %>%
+  rename(mrs.365 = glasgow_rankin_365) %>%
+  rename(barthel.total.365 = barthel_index_365) %>%
+  rename(eq.vas.365 = eq_vas_365) %>%
+  rename(time.to.ed.hours = ictus_to_diagct_hours)
+
+mistie3$time.to.ed.hours = round(mistie3$time.to.ed.hours, digits = 2)
+mistie3$time.to.ed.minutes = mistie3$time.to.ed.hours * 60
+mistie3$time.to.ed.minutes = round(mistie3$time.to.ed.minutes, digits = 0)
+
+mistie3 <- mistie3 %>%
+  mutate(race = case_when(
+    race == "African-American" ~ "Black or African-American",
+    race == "Asian Hawaiian or Pacific" ~ "Asian",
+    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
+    race == "More than one race" ~ "Other/Not Reported",
+    race == "White" ~ "White"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(ethnicity = case_when(
+         hispanic == "0" ~ "Not Hispanic or Latino",
+         hispanic == "1" ~ "Hispanic or Latino"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(sex = case_when(
+    male_gender == "0" ~ "Woman",
+    male_gender == "1" ~ "Man"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(tobacco = case_when(
+    current_smoker == "0" ~ "No",
+    current_smoker == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(dm2 = case_when(
+    diabetes_pmh == "0" ~ "No",
+    diabetes_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(htn = case_when(
+    htn_pmh == "0" ~ "No",
+    htn_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(cad = case_when(
+    cvd_pmh == "0" ~ "No",
+    cvd_pmh == "1" ~ "Yes"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(ich.location.baseline = case_when(
+    ich_location_specify == "BG" ~ "Basal Ganglia",
+    ich_location_specify == "Frontal" ~ "Lobar",
+    ich_location_specify == "Occipital" ~ "Lobar",
+    ich_location_specify == "Parietal" ~ "Lobar",
+    ich_location_specify == "Temporal" ~ "Lobar",
+    ich_location_specify == "Thalamus" ~ "Thalamus",
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(comfort.binary = case_when(
+    withdrawal_contribution == "0" ~ "No",
+    withdrawal_contribution == "1" ~ "Yes"
+  ))
+
+mistie3 <- subset(mistie3, select = -withdrawal_contribution)
+
+mistie3 <- mistie3 %>%
+  mutate(euro.mobility.365 = case_when(
+    eq5d_mobility_365 == "1" ~ "No Problems",
+    eq5d_mobility_365 == "2" ~ "Some Problems",
+    eq5d_mobility_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.selfcare.365 = case_when(
+    eq5d_selfcare_365 == "1" ~ "No Problems",
+    eq5d_selfcare_365 == "2" ~ "Some Problems",
+    eq5d_selfcare_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.usual.365 = case_when(
+    eq5d_activities_365 == "1" ~ "No Problems",
+    eq5d_activities_365 == "2" ~ "Some Problems",
+    eq5d_activities_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.pain.365 = case_when(
+    eq5d_pain_365 == "1" ~ "No Problems",
+    eq5d_pain_365 == "2" ~ "Some Problems",
+    eq5d_pain_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(euro.anxiety.365 = case_when(
+    eq5d_anxiety_365 == "1" ~ "No Problems",
+    eq5d_anxiety_365 == "2" ~ "Some Problems",
+    eq5d_anxiety_365 == "3" ~ "Extreme Problems"
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.30 = case_when(
+    location_d30 == "Acute - Hospital" ~ "Hospital",
+    location_d30 == "Acute - ICU" ~ "Hospital",
+    location_d30 == "Dead" ~ "Dead",
+    location_d30 == "Home" ~ "Home",
+    location_d30 == "Rehab" ~ "Rehabilitation",
+    location_d30 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d30 == "Hospice" ~ "Hospice",
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.180 = case_when(
+    location_d180 == "Acute - Hospital" ~ "Hospital",
+    location_d180 == "Acute - ICU" ~ "Hospital",
+    location_d180 == "Dead" ~ "Dead",
+    location_d180 == "Home" ~ "Home",
+    location_d180 == "Rehab" ~ "Rehabilitation",
+    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d180 == "Hospice" ~ "Hospice",
+  ))
+
+mistie3 <- mistie3 %>%
+  mutate(current.location.365 = case_when(
+    location_d365 == "Acute - Hospital" ~ "Hospital",
+    location_d365 == "Acute - ICU" ~ "Hospital",
+    location_d365 == "Dead" ~ "Dead",
+    location_d365 == "Home" ~ "Home",
+    location_d365 == "Rehab" ~ "Rehabilitation",
+    location_d365 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d365 == "Hospice" ~ "Hospice",
+  ))
+
+# Calculating WLST Day
+
+mistie3 <- mistie3 %>%
+  mutate(symptom.onset.date = mdy(ninds_symptom_onset_date)) %>%
+  mutate(withdrawal.date = mdy(ninds_withdrawal_date))
+
+mistie3$withdrawal.interval <- as.period(mistie3$symptom.onset.date %--% mistie3$withdrawal.date)
+mistie3$withdrawal.interval <- time_length(mistie3$withdrawal.interval, unit = "days")
+mistie3$comfort.day <- round(mistie3$withdrawal.interval, digits = 0)
+
+```
+
+**Combining Datasets**
+
+```{r Combining Everything}
+
+# Combining Data from All Studies
+combined.dataset.list <- list(atach.whole, erich.whole, clear3.whole, mistie2.whole, mistie3)
+
+atach.erich <- full_join(atach.whole, erich.whole)
+atach.erich.clear <- full_join(atach.erich, clear3.whole)
+atach.erich.clear.mistie2 <- full_join(atach.erich.clear, mistie2.whole)
+atach.erich.clear.mistie2.mistie3 <- full_join(atach.erich.clear.mistie2, mistie3)
+
+VarsTableOne <- c("study","age", "sex", "ethnicity", # Ethnicity
                        "race", # Race
@@
                        "tobacco",
-                       #"symptom.onset", # Time of symptom onset
+                       "time.to.ed.minutes", # Time from symptom onset to Ed presentation
                        "sbp.initial", # Initial SBP
@@
                        "gcs.baseline", # Baseline GCS
-                       #"nihss.baseline", # NIHSS
+                       "nihss.baseline", # NIHSS
                        "ich.location.baseline", 
@@
                       "current.location.365",
-                       "ich.laterality"
+                       "ich.laterality",
+                  "dnr.binary",
+                  "comfort.binary",
+                  "dnr.day",
+                  "comfort.day"
                   )
 
-```
-
-**Importing and wrangling ERICH data**
-
-```{r ERICH}
-
-erich.case <- read_sas("ERICH/erich_case_100217.sas7bdat")
-
-erich.whole <- erich.case %>%
-  select(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30, O3, O3A, O5, O20, AU8, AU8_OTH, RANKIN_FU1, RANKIN_FU2, RANKIN_FU3, MOBILITY_FU1, MOBILITY_FU2, MOBILITY_FU3, SELFCARE_FU1, SELFCARE_FU2, SELFCARE_FU3, USUALACT_FU1, USUALACT_FU2, USUALACT_FU3, PAIN_FU1, PAIN_FU2, PAIN_FU3, ANXIETY_FU1, ANXIETY_FU2, ANXIETY_FU3, HLTHSTAT_FU1, HLTHSTAT_FU2, HLTHSTAT_FU3, FEED_FU1, FEED_FU2, FEED_FU3, BATHE_FU1, BATHE_FU2, BATHE_FU3, GROOM_FU1, GROOM_FU2, GROOM_FU3, DRESS_FU1, DRESS_FU2, DRESS_FU3, BOWELS_FU1, BOWELS_FU2, BOWELS_FU3, BLADDER_FU1, BLADDER_FU2, BLADDER_FU3, TOILET_FU1, TOILET_FU2, TOILET_FU3, TRANSFER_FU1, TRANSFER_FU2, TRANSFER_FU3, MOBLEVEL_FU1, MOBLEVEL_FU2, MOBLEVEL_FU3, STAIRS_FU1, STAIRS_FU2, STAIRS_FU3, BART_TOT_FU1, BART_TOT_FU2, BART_TOT_FU3, EMS33, ED44, CURR_LOC_FU1, CURR_LOC_FU2, CURR_LOC_FU3, IT2, O1, O2_DATE, O2_TIME, O3, O3_OTH, DOD)
-
-erich.whole <- erich.whole %>%
-  mutate(study = "ERICH")
-
-erich.whole <- erich.whole %>%
-  mutate(enrolling.country = "United States")
-
-# Renaming and Recoding Variables
-
-erich.whole <- erich.whole %>%
-  mutate(gender = recode(gender,
-                         "F" = "Female",
-                         "M" = "Male")) %>%
-  rename(sex = gender)
-
-erich.whole <- erich.whole %>%
-  mutate(Race = case_when(
-    D21 == "W" ~ "White",
-    D21 == "B" ~ "Black or African-American",
-    D21 == "A" ~ "Asian",
-    D21 == "N" ~ "Native Hawaiian/Pacific Islander",
-    D21 == "I" ~ "American Indian or Alaskan Native",
-    D21 == "O" ~ "Other/Not Reported",
-    D21 == "U" ~ "Other/Not Reported"
-    )) %>%
-  rename(race = Race)
-
-erich.whole <- erich.whole %>%
-  mutate(ethnicity = case_when(
-    D21A == "H" ~ "Hispanic or Latino",
-    D21A == "N" ~ "Not Hispanic or Latino",
-    D21A == "U" ~ "Not Hispanic or Latino"
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(stroke = recode(SH1,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(cad = recode(MHX8,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(chf = recode(MHX13,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(afib = recode(MHX11,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(pvd = recode(MHX27C,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(hld = recode(MHX6,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(dm2 = case_when(
-    MHX3 == "2" ~ "No",
-    MHX3 == "8" ~ "NA",
-    MHX3A == "1" ~ "No",
-    MHX3A == "2" ~ "Yes",
-    MHX3A == "8" ~ "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(tobacco = recode(SMK1A,
-                         "2" = "No",
-                         "1" = "Yes",
-                         "5" = "NA",
-                         "8" = "NA"))
-
-erich.whole <- erich.whole %>%
-  rename(sbp.initial = ED47_SYS)
-
-erich.whole <- erich.whole %>%
-  rename(dbp.initial = ED47_DIAS)
-
-erich.whole <- erich.whole %>%
-  rename(gcs.baseline = ED39)
-
-erich.whole <- erich.whole %>%
-  mutate(ich.laterality = recode(PHY6A,
-                                 "L" = "Left",
-                                 "R" = "Right",
-                                 "M" = "NA",
-                                 .default = "NA"))
-
-erich.whole <- erich.whole %>%
-  mutate(ich.location.baseline = recode(erich.case$detail,
-                                        "L Thalamic" = "Thalamus",
-                                        "R Thalamic" = "Thalamus",
-                                        "L Caudate Nucleus" = "Basal Ganglia",
-                                        "L Cerebellar" = "Cerebellum",
-                                        "L Corona Radiata" = "Lobar",
-                                        "L Frontal" = "Lobar",
-                                        "L Frontal Temporal" = "Lobar",
-                                        "L Frontoparietal" = "Lobar",
-                                        "L Occipital" = "Lobar",
-                                        "L Parietal" = "Lobar",
-                                        "L Parieto-Occipital" = "Lobar",
-                                        "L Periventricular" = "Basal Ganglia",
-                                        "L Putaminal" = "Basal Ganglia",
-                                        "L Temporal" = "Lobar",
-                                        "L Temporal-Parietal-Occipital (L Hemisphere)" = "Lobar",
-                                        "L Temporal Occipital" = "Lobar",
-                                        "L Temporal parietal" = "Lobar",
-                                        "Multiple Hemorrhages - Deep" = "Basal Ganglia",
-                                        "Multiple Hemorrhages - Lobar" = "Lobar",
-                                        "Not applicable" = "NA",
-                                        "Primary IVH" = "Primary IVH",
-                                        "R Basal Ganglia" = "Basal Ganglia",
-                                        "R Caudate Nucleus" = "Basal Ganglia",
-                                        "R Cerebellar" = "Cerebellum",
-                                        "R Frontal" = "Lobar",
-                                        "R Frontal Temporal" = "Lobar",
-                                        "R Frontoparietal" = "Lobar",
-                                        "R Occipital" = "Lobar",
-                                        "R Parietal" = "Lobar",
-                                        "R Parieto-Occipital" = "Lobar",
-                                        "R Periventricular" = "Basal Ganglia",
-                                        "R Putaminal" = "Basal Ganglia",
-                                        "R Temporal" = "Lobar",
-                                        "R Temporal-Parietal-Occipital (R Hemisphere)" = "Lobar",
-                                        "R Temporal Occipital" = "Lobar",
-                                        "R Temporal Perietal" = "Lobar",
-                                        "Unknown" = "NA",
-                                        "Brainstem" = "Brainstem",
-                                        "Cerebellar" = "Cerebellum",
-                                        "R Temporal parietal" = "Lobar",
-                                        "L Basal Ganglia" = "Basal Ganglia",                                        ))
-
-erich.whole <- erich.whole %>%
-  mutate(ivh = case_when(erich.case$IVH_Vol_CT1 > 0 ~ "Yes",
-                         erich.case$IVH_Vol_CT1 < 0.1 ~ "No"))
-
-erich.whole <- erich.whole %>%
-  rename(ich.volume.cc = ICH_Vol_CT1)
-
-erich.whole <- erich.whole %>%
-  rename(ivh.volume.cc = IVH_Vol_CT1)
-
-erich.whole <- erich.whole %>%
-  mutate(mech.ventilation = case_when(
-    EMS33 == "1" ~ "Yes",
-    EMS33 == "2" ~ "No",
-    ED44 == "1" ~ "Yes",
-    ED44 == "2" ~ "No",
-    IT7 == "1" ~ "Yes",
-    IT7 == "2" ~ "No"))
-
-erich.whole <- erich.whole %>%
-  mutate(evd = case_when(
-    IT8 == "1" ~ "Yes",
-    IT8 == "2" ~ "No",
-    str_detect(erich.whole$IT5F_SPEC, "VENTR") ~ "Yes",
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(nsgy.evacuation = case_when(
-    IT5D == "1" ~ "Yes",
-    IT5H == "1" ~ "Yes",
-    str_detect(erich.whole$IT5F_SPEC, "CRANI") ~ "Yes",
-    str_detect(erich.whole$IT5F_SPEC, "EVAC") ~ "Yes",
-    TRUE ~ "No"
-      ))
-
-erich.whole <- erich.whole %>%
-  rename(mrs.90 = RANKIN_FU1) %>%
-  rename(mrs.180 = RANKIN_FU2) %>%
-  rename(mrs.365 = RANKIN_FU3) %>%
-  rename(euro.vas.90 = HLTHSTAT_FU1) %>%
-  rename(euro.vas.180 = HLTHSTAT_FU2) %>%
-  rename(euro.vas.365 = HLTHSTAT_FU3) %>%
-  rename(barthel.feeding.90 = FEED_FU1) %>%
-  rename(barthel.bathing.90 = BATHE_FU1) %>%
-  rename(barthel.grooming.90 = GROOM_FU1) %>%
-  rename(barthel.dressing.90 = DRESS_FU1) %>%
-  rename(barthel.bowels.90 = BOWELS_FU1) %>%
-  rename(barthel.bladder.90 = BLADDER_FU1) %>%
-  rename(barthel.toilet.90 = TOILET_FU1) %>%
-  rename(barthel.transfer.90 = TRANSFER_FU1) %>%
-  rename(barthel.mobility.90 = MOBLEVEL_FU1) %>%
-  rename(barthel.stairs.90 = STAIRS_FU1) %>%
-  rename(barthel.total.90 = BART_TOT_FU1) %>%
-  rename(barthel.feeding.180 = FEED_FU2) %>%
-  rename(barthel.bathing.180 = BATHE_FU2) %>%
-  rename(barthel.grooming.180 = GROOM_FU2) %>%
-  rename(barthel.dressing.180 = DRESS_FU2) %>%
-  rename(barthel.bowels.180 = BOWELS_FU2) %>%
-  rename(barthel.bladder.180 = BLADDER_FU2) %>%
-  rename(barthel.toilet.180 = TOILET_FU2) %>%
-  rename(barthel.transfer.180 = TRANSFER_FU2) %>%
-  rename(barthel.mobility.180 = MOBLEVEL_FU2) %>%
-  rename(barthel.stairs.180 = STAIRS_FU2) %>%
-  rename(barthel.total.180 = BART_TOT_FU2) %>%
-  rename(barthel.feeding.365 = FEED_FU3) %>%
-  rename(barthel.bathing.365 = BATHE_FU3) %>%
-  rename(barthel.grooming.365 = GROOM_FU3) %>%
-  rename(barthel.dressing.365 = DRESS_FU3) %>%
-  rename(barthel.bowels.365 = BOWELS_FU3) %>%
-  rename(barthel.bladder.365 = BLADDER_FU3) %>%
-  rename(barthel.toilet.365 = TOILET_FU3) %>%
-  rename(barthel.transfer.365 = TRANSFER_FU3) %>%
-  rename(barthel.mobility.365 = MOBLEVEL_FU3) %>%
-  rename(barthel.stairs.365 = STAIRS_FU3) %>%
-  rename(barthel.total.365 = BART_TOT_FU3)
-
-erich.whole <- erich.whole %>%
-  mutate(current.location.90 = recode(CURR_LOC_FU1,
-                      "1" = "Home",
-                      "2" = "Rehabilitation",
-                      "3" = "Skilled Nursing Facility",
-                      "4" = "Hospital",
-                      "7" = "Other"
-                      )) %>%
-  mutate(current.location.180 = recode(CURR_LOC_FU1,
-                      "1" = "Home",
-                      "2" = "Rehabilitation",
-                      "3" = "Skilled Nursing Facility",
-                      "4" = "Hospital",
-                      "7" = "Other"
-                      )) %>%
-  mutate(current.location.365 = recode(CURR_LOC_FU1,
-                      "1" = "Home",
-                      "2" = "Rehabilitation",
-                      "3" = "Skilled Nursing Facility",
-                      "4" = "Hospital",
-                      "7" = "Other"
-                      ))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.mobility.90 = factor(MOBILITY_FU1,
-                                labels = c("No Problems", "Some Problems", "Confined to Bed", "NA")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.selfcare.90 = factor(SELFCARE_FU1,
-                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself", "NA")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.usual.90 = factor(USUALACT_FU1,
-                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities", "NA")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.pain.90 = factor(PAIN_FU1,
-                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort", "NA")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.anxiety.90 = factor(ANXIETY_FU1,
-                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed", "NA")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.mobility.180 = factor(MOBILITY_FU2,
-                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.selfcare.180 = factor(SELFCARE_FU2,
-                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.usual.180 = factor(USUALACT_FU2,
-                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.pain.180 = factor(PAIN_FU2,
-                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.anxiety.180 = factor(ANXIETY_FU2,
-                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.mobility.365 = factor(MOBILITY_FU3,
-                                labels = c("No Problems", "Some Problems", "Confined to Bed")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.selfcare.365 = factor(SELFCARE_FU3,
-                                labels = c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.usual.365 = factor(USUALACT_FU3,
-                                labels = c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.pain.365 = factor(PAIN_FU3,
-                                labels = c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")))
-
-erich.whole <- erich.whole %>%
-  mutate(euro.anxiety.365 = factor(ANXIETY_FU3,
-                                labels = c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")))
-
-# Parsing Dates and Times in ERICH
-
-# symptom.onset
-
-erich.whole <- erich.whole %>%
-  mutate(symptom.onset.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$SE2)))
-
-erich.whole <- erich.whole %>%
-  mutate(symptom.onset.estimated.time = case_when( # Converting estimated times to numbers
-    SE3 == "1" ~ "06:00:00",
-    SE3 == "2" ~ "NA",
-    SE3 == "3" ~ "03:00:00",
-    SE3 == "4" ~ "09:00:00",
-    SE3 == "5" ~ "15:00:00",
-    SE3 == "6" ~ "21:00:00"
-  ))
-
-erich.whole$symptom.onset.estimated.time <- hms(erich.whole$symptom.onset.estimated.time)
-
-erich.whole <- erich.whole %>%
-  mutate(symptom.onset.estimated.date.time = ymd_hms(paste(erich.whole$SE1, erich.whole$symptom.onset.estimated.time)))
-
-erich.whole <- erich.whole %>%
-  mutate(symptom.onset.combined = case_when(
-    is.na(symptom.onset.date.time) ~ symptom.onset.estimated.date.time,
-    TRUE ~ symptom.onset.date.time
-  ))
-
-# Time of ED presentation
-
-erich.whole <- erich.whole %>%
-  mutate(ed.presentation.date.time = ymd_hms(paste(erich.whole$ED7, erich.whole$ED8)))
-
-# Neurosurgery Dates
-
-erich.whole <- erich.whole %>%
-  mutate(nsgy.evacuation.date.erich = case_when(
-    IT5D == "1" ~ ymd_hms(paste(IT5D_DATE, IT5D_TIME)),
-    IT5H == "1" ~ ymd_hms(paste(IT5H_DATE, IT5H_TIME)),
-    IT5F == "1" ~ ymd_hms(paste(IT5F_DATE, IT5F_TIME))
-  ))
-
-erich.whole$nsgy.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$nsgy.evacuation.date.erich)
-erich.whole$nsgy.interval <- time_length(erich.whole$nsgy.interval, unit = "days")
-erich.whole$nsgy.evacuation.date <- round(erich.whole$nsgy.interval, digits = 0) # Makes it compatabible to merge with ATACH
-
-# DNR and WLST Status
-
-erich.whole <- erich.whole %>%
-  mutate(dnr.binary = case_when(
-    CC16 == "1" ~ "Yes",
-    CC16 == "2" ~ "No",
-    TRUE ~ "NA"
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(dni.binary = case_when(
-    CC17 == "1" ~ "Yes",
-    CC17 == "2" ~ "No",
-    TRUE ~ "NA"
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(comfort.binary = case_when(
-    CC18 == "1" ~ "Yes",
-    CC18 == "2" ~ "No",
-    TRUE ~ "NA"
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(dnr.date.time = case_when(
-    CC16 == "1" ~ ymd_hms(paste(CC16_DATE, CC16_TIME))
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(dni.date.time = case_when(
-    CC17 == "1" ~ ymd_hms(paste(CC17_DATE, CC17_TIME))
-  ))
-
-erich.whole <- erich.whole %>%
-  mutate(comfort.date.time = case_when(
-    CC18 == "1" ~ ymd_hms(paste(CC18_DATE, CC18_TIME))
-  ))
-
-# Mortality Status and Adding mRS 6
-
-erich.whole$death.date <- ymd(erich.whole$DOD)
-
-erich.whole$death.interval <- erich.whole$symptom.onset.combined %--% erich.whole$death.date
-erich.whole$death.interval <- as.period(erich.whole$death.interval)
-erich.whole$death.interval <- time_length(erich.whole$death.interval, unit = "days")
-
-erich.whole <- erich.whole %>%
-  mutate(mrs.90 = case_when(
-    death.interval < 90 ~ 6,
-    TRUE ~ mrs.90
-    ))
-
-erich.whole$mrs.90 <- na_if(erich.whole$mrs.90, 8)
-
-erich.whole <- erich.whole %>%
-  mutate(mrs.180 = case_when(
-    death.interval < 180 ~ 6,
-    TRUE ~ mrs.180
-    ))
-
-erich.whole <- erich.whole %>%
-  mutate(mrs.365 = case_when(
-    death.interval < 365 ~ 6,
-    TRUE ~ mrs.365
-    ))
-
-# Adding tracheostomy information
-
-erich.whole <- erich.whole %>%
-  mutate(
-    tracheostomy = case_when(
-      str_detect(erich.whole$IT5F_SPEC, "TRACHEO") ~ "Yes",
-      ))
-
-# Variables missing: nihss.baseline, hydrocephalus, phe.volume.cc, sp.shift.baseline, intubation.days
-
-# Filtering out ICH Laterality and Infratentorial Location
-erich.whole <- erich.whole %>%
-  filter(ich.laterality != "NA") %>%
-  filter(ich.location.baseline == "Basal Ganglia" |
-           ich.location.baseline == "Thalamus" |
-           ich.location.baseline == "Lobar")
-
-```
-
-**Importing and wrangling iDEF data**
-
-```{r iDEF}
-
-# Create function to paste together string of formXX.sas7bdat
-idef.import <- function(y) {
-  dataset <- paste0("idef",".",y)
-  filepath <- paste0("./iDEF/form",y,".sas7bdat")
-  dataset <- read_sas(filepath)
-}
-
-idefdictionary <- read_sas("iDEF/datadictionary.sas7bdat")
-idef01 <- idef.import("01")
-idef02 <- idef.import("02")
-idef03 <- idef.import("03")
-idef03c <- idef.import("03c")
-idef04 <- idef.import("04")
-idef05 <- idef.import("05")
-idef06 <- idef.import("06")
-idef07 <- idef.import("07")
-idef08 <- idef.import("08")
-idef10c <- idef.import("10c")
-idef11 <- idef.import("11")
-idef15 <- idef.import("15")
-idef18 <- idef.import("18")
-idef19 <- idef.import("19")
-idef19c <- idef.import("19c")
-idef22 <- idef.import("22")
-idef23 <- idef.import("23")
-idef24 <- idef.import("24")
-idef25 <- idef.import("25")
-idef25c <- idef.import("25c")
-idef26 <- idef.import("26")
-idef27 <- idef.import("27")
-idef29 <- idef.import("29")
-idef30 <- idef.import("30")
-idef32 <- idef.import("32")
-idef32c <- idef.import("32c")
-idef43 <- idef.import("43")
-
-```
-
-**Importing and wrangling CLEAR-III data**
-
-```{r CLEAR III}
-
-# Importing Excel files
-clear3.demographics <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 1)
-clear3.outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_Publication data_2018.1.30.xlsx", sheet = 2)
-clear3.180outcomes <- readxl::read_excel("./CLEAR3/CLEAR III_clinicaltrials.gov results_2018.1.30.xlsx", sheet = 1)
-
-# Combining all dataframes into one
-clear3.whole <- left_join(clear3.demographics, clear3.outcomes)
-clear3.whole <- left_join(clear3.whole, clear3.180outcomes)
-
-# Adding category to indicate study identity
-clear3.whole <- clear3.whole %>% mutate(study = "CLEAR III")
-
-# Categorizing and Recoding Variables
-
-clear3.whole <- clear3.whole %>%
-  mutate(race = case_when(
-    race == "African-American" ~ "Black or African-American",
-    race == "Asian" ~ "Asian",
-    race == "Hawaiian or Pacific" ~ "Native Hawaiian/Pacific Islander",
-    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
-    race == "Indian or Alaskan;White" ~ "American Indian or Alaskan Native",
-    race == "unknown/missing" ~ "Other/Not Reported",
-    race == "White" ~ "White"
-  ))
-
-clear3.whole <- clear3.whole %>%
-  mutate(ethnicity = recode(hispanic_latino,
-                            "0" = "Not Hispanic or Latino",
-                            "1" = "Hispanic or Latino"
-  ))
-
-clear3.whole <- clear3.whole %>%
-  mutate(sex = recode(gender,
-                         "female" = "Female",
-                         "male" = "Male"))
-
-clear3.whole <- clear3.whole %>%
-  mutate(ich.laterality = recode(hemisphere,
-                                 "left" = "Left",
-                                 "right" = "Right",
-                                 "unknown/missing" = "NA" ))
-
-clear3.whole <- clear3.whole %>% mutate(ivh = "Yes")
-
-clear3.whole <- clear3.whole %>%
-  rename(age = age_at_registration) %>%
-  rename(tobacco = tobacco_use) %>%
-  rename(sbp.initial = er_present_systolic) %>%
-  rename(dbp.initial = er_present_diastolic) %>%
-  rename(ich.location.baseline = adjudicated_ich_location) %>%
-  rename(gcs.baseline = er_present_gcs_total) %>%
-  rename(nihss.baseline = er_present_nihss_total) %>%
-  rename(ich.volume.cc = dct_ich_volume_rc) %>%
-  rename(ivh.volume.cc = dct_ivh_volume_rc) %>%
-  rename(mrs.30 = glasgow_rankin_30) %>%
-  rename(mrs.180 = glasgow_rankin_180) %>%
-  rename(gose.30 = gose_30) %>%
-  rename(gose.180 = gose_180) %>%
-  rename(comfort.binary = withdrawal_contribution) %>%
-  rename(comfort.date.time = ictus_to_withdrawal_days) %>%
-  rename(symptom.onset.to.presentation = ictus_to_presentation)
-
-```
-
-**Importing and Wrangling MISTIE-II data**
-
-```{r MISTIE II}
-
-# Importing MISTIE II data
-
-mistie2.demographics <- readxl::read_excel("./MISTIE2/MISTIE II_Publication results_2016.4.14_RT_RD updates 2017.10.2.xlsx", sheet = 1)
-
-mistie2.demographics2 <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_Demographics data_2017.10.16.xlsx", sheet = 1)
-
-mistie2.results <- readxl::read_excel("./MISTIE2/MISTIE II and ICES_ctgov results_2017.10.17.xlsx", sheet = 3)
-
-mistie2.demographics3 <- readxl::read_excel("./MISTIE2/ICES_Publication results_2015.12.23_RT_RD updates 2017.10.16.xlsx", sheet = 1)
-
-mistie2.combined.demographics <- full_join(mistie2.demographics, mistie2.demographics3)
-mistie2.combined.demographics <- full_join(mistie2.combined.demographics, mistie2.demographics2)
-
-mistie2.wide.results <- mistie2.results %>%
-  pivot_wider(names_from = Follow_up_Visit, values_from = c(GOS, eGOS_Score, rankin_score, SIS_Total, SIS_16_Total))
-
-mistie2.whole <- full_join(mistie2.wide.results, mistie2.combined.demographics)
-
-mistie2.whole <- mistie2.whole %>% mutate(study = "MISTIE II")
-
-# Renaming and Recoding Variables
-
-mistie2.whole <- mistie2.whole %>%
-  mutate(ethnicity = case_when(
-    Race == "African American not Hispanic" ~ "Not Hispanic or Latino",
-    Race == "Hispanic" ~ "Hispanic or Latino",
-    Race == "Asian or Pacific Islander" ~ "Not Hispanic or Latino",
-    Race == "Caucasian not Hispanic" ~ "Not Hispanic or Latino",
-    Race == "Other or unknown" ~ "Not Hispanic or Latino"
-  ))
-
-mistie2.whole <- mistie2.whole %>%
-  mutate(race = recode(Race,
-                       "African American not Hispanic" = "Black or African American",
-                       "Asian or Pacific Islander" = "Asian",
-                       "Caucasian not Hispanic" = "White",
-                       "Hispanic" = "Other/Not Reported",
-                       "Other or unknown" = "Other/Not Reported"
-  ))
-
-mistie2.whole <- mistie2.whole %>%
-  rename(sbp.initial = er_sbp) %>%
-  rename(dbp.initial = er_dbp) %>%
-  rename(gcs.baseline = enrollment_gcs) %>%
-  rename(nihss.baseline = enrollment_nihss_total) %>%
-  rename(ich.volume.cc = pre_rand_ichvol) %>%
-  rename(ivh.volume.cc = pre_rand_ivhvol) %>%
-  rename(gos.365 = GOS_365) %>%
-  rename(gos.270 = GOS_270) %>%
-  rename(gos.180 = GOS_180) %>%
-  rename(gos.90 = GOS_90) %>%
-  rename(gos.30 = GOS_30) %>%
-  rename(egos.365 = eGOS_Score_365) %>%
-  rename(egos.270 = eGOS_Score_270) %>%
-  rename(egos.180 = eGOS_Score_180) %>%
-  rename(egos.90 = eGOS_Score_90) %>%
-  rename(egos.30 = eGOS_Score_30) %>%
-  rename(mrs.365 = rankin_score_365) %>%
-  rename(mrs.270 = rankin_score_270) %>%
-  rename(mrs.180 = rankin_score_180) %>%
-  rename(mrs.90 = rankin_score_90) %>%
-  rename(mrs.30 = rankin_score_30)
-
-mistie2.whole <- mistie2.whole %>%
-  mutate(ivh = recode(any_ivh_prerand,
-                      "0" = "No",
-                      "1" = "Yes"
-  ))
-
-```
-
-**Importing and Wrangling MISTIE-III data**
-
-```{r MISTIE-III}
-
-# Importing MISTIE III data
-mistie3 <- read.table("./MISTIE3/Deidentified_Mistie_III_Wide_Analytic_2019_07_11 (1).csv", sep = ",", header = TRUE)
-
-mistie3 <- mistie3 %>% mutate(study = "MISTIE III")
-
-mistie3 <- mistie3 %>%
-  rename(age = age_at_consent) %>%
-  rename(race = race_ninds) %>%
-  rename(sbp.initial = er_present_systolic) %>%
-  rename(dbp.initial = er_present_diastolic) %>%
-  rename(gcs.baseline = gcs_total) %>%
-  rename(nihss.baseline = nihss_total) %>%
-  rename(ich.volume.cc = diagct_ich_volume) %>%
-  rename(ivh.volume.cc = diagct_ivh_volume) %>%
-  rename(ich.laterality = ich_hemisphere_adj) %>%
-  rename(mrs.30 = glasgow_rankin_30) %>%
-  rename(mrs.180 = glasgow_rankin_180) %>%
-  rename(mrs.365 = glasgow_rankin_365) %>%
-  rename(barthel.total.365 = barthel_index_365) %>%
-  rename(eq.vas.365 = eq_vas_365) %>%
-  rename(time.to.presentation = ictus_to_site_arrival_hours)
-
-mistie3 <- mistie3 %>%
-  mutate(race = case_when(
-    race == "African-American" ~ "Black of African-American",
-    race == "Asian Hawaiian or Pacific" ~ "Asian",
-    race == "Indian or Alaskan" ~ "American Indian or Alaskan Native",
-    race == "More than one race" ~ "Other/Not Reported",
-    race == "White" ~ "White"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(ethnicity = case_when(
-         hispanic == "0" ~ "Not Hispanic or Latino",
-         hispanic == "1" ~ "Hispanic or Latino"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(sex = case_when(
-    male_gender == "0" ~ "Woman",
-    male_gender == "1" ~ "Man"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(tobacco = case_when(
-    current_smoker == "0" ~ "No",
-    current_smoker == "1" ~ "Yes"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(dm2 = case_when(
-    diabetes_pmh == "0" ~ "No",
-    diabetes_pmh == "1" ~ "Yes"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(htn = case_when(
-    htn_pmh == "0" ~ "No",
-    htn_pmh == "1" ~ "Yes"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(cad = case_when(
-    cvd_pmh == "0" ~ "No",
-    cvd_pmh == "1" ~ "Yes"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(comfort = case_when(
-    withdrawal_contribution == "0" ~ "No",
-    withdrawal_contribution == "1" ~ "Yes"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(euro.mobility.365 = case_when(
-    eq5d_mobility_365 == "1" ~ "No Problems",
-    eq5d_mobility_365 == "2" ~ "Some Problems",
-    eq5d_mobility_365 == "3" ~ "Extreme Problems"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(euro.selfcare.365 = case_when(
-    eq5d_selfcare_365 == "1" ~ "No Problems",
-    eq5d_selfcare_365 == "2" ~ "Some Problems",
-    eq5d_selfcare_365 == "3" ~ "Extreme Problems"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(euro.usual.365 = case_when(
-    eq5d_activities_365 == "1" ~ "No Problems",
-    eq5d_activities_365 == "2" ~ "Some Problems",
-    eq5d_activities_365 == "3" ~ "Extreme Problems"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(euro.pain.365 = case_when(
-    eq5d_pain_365 == "1" ~ "No Problems",
-    eq5d_pain_365 == "2" ~ "Some Problems",
-    eq5d_pain_365 == "3" ~ "Extreme Problems"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(euro.anxiety.365 = case_when(
-    eq5d_anxiety_365 == "1" ~ "No Problems",
-    eq5d_anxiety_365 == "2" ~ "Some Problems",
-    eq5d_anxiety_365 == "3" ~ "Extreme Problems"
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(current.location.30 = case_when(
-    location_d30 == "Acute - Hospital" ~ "Hospital",
-    location_d30 == "Acute - ICU" ~ "Hospital",
-    location_d30 == "Dead" ~ "Dead",
-    location_d30 == "Home" ~ "Home",
-    location_d30 == "Rehab" ~ "Rehabilitation",
-    location_d30 == "LTCF" ~ "Skilled Nursing Facility",
-    location_d30 == "Hospice" ~ "Hospice",
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(current.location.180 = case_when(
-    location_d180 == "Acute - Hospital" ~ "Hospital",
-    location_d180 == "Acute - ICU" ~ "Hospital",
-    location_d180 == "Dead" ~ "Dead",
-    location_d180 == "Home" ~ "Home",
-    location_d180 == "Rehab" ~ "Rehabilitation",
-    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
-    location_d180 == "Hospice" ~ "Hospice",
-  ))
-
-mistie3 <- mistie3 %>%
-  mutate(current.location.365 = case_when(
-    location_d365 == "Acute - Hospital" ~ "Hospital",
-    location_d365 == "Acute - ICU" ~ "Hospital",
-    location_d365 == "Dead" ~ "Dead",
-    location_d365 == "Home" ~ "Home",
-    location_d365 == "Rehab" ~ "Rehabilitation",
-    location_d365 == "LTCF" ~ "Skilled Nursing Facility",
-    location_d365 == "Hospice" ~ "Hospice",
-  ))
-
-```
-
-**Combining Datasets**
-
-```{r combining ATACH2 and ERICH}
-
-atach.erich <- union_all(atach.whole, erich.whole)
-
-atach.erich <- atach.erich %>%
-  select(VarsTableOne) %>%
+ninds.combined <- atach.erich.clear.mistie2.mistie3 %>%
+  select(VarsTableOne)
+
+ninds.combined <- ninds.combined %>%
   mutate(ich.score.age = case_when(
@@
 
-atach.erich %>%
-  as_tibble()
-
-table1 <- CreateTableOne(data = atach.erich,
-                         strata = "ich.laterality",)
-
+# Cleaning up combined dataset
+ninds.combined <- ninds.combined %>%
+  mutate(sex = case_when(
+    sex == "Man" ~ "Male",
+    sex == "Male" ~ "Male",
+    sex == "Woman" ~ "Female",
+    sex == "Female" ~ "Female")) %>%
+  mutate(ethnicity = case_when(
+    ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
+    ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino",
+    ethnicity == "Unknown/Not Reported" ~ "NA")) %>%
+  mutate(ich.location.baseline = case_when(
+    ich.location.baseline == "Basal Ganglia" ~ "Basal Ganglia",
+    ich.location.baseline == "Brainstem" ~ "Brainstem",
+    ich.location.baseline == "Caudate" ~ "Basal Ganglia",
+    ich.location.baseline == "Cerebellum" ~ "Cerebellum",
+    ich.location.baseline == "Frontal" ~ "Lobar",
+    ich.location.baseline == "Globus Pallidus" ~ "Basal Ganglia",
+    ich.location.baseline == "Lobar" ~ "Lobar",
+    ich.location.baseline == "Occipital" ~ "Lobar",
+    ich.location.baseline == "Other" ~ "Other",
+    ich.location.baseline == "Parietal" ~ "Lobar",
+    ich.location.baseline == "Primary IVH" ~ "Primary IVH",
+    ich.location.baseline == "Probable Primary IVH" ~ "Primary IVH",
+    ich.location.baseline == "Putamen" ~ "Basal Ganglia",
+    ich.location.baseline == "Temporal" ~ "Lobar",
+    ich.location.baseline == "Thalamus" ~ "Thalamus"
+  ))
+
+ninds.combined <- ninds.combined %>%
+  filter(ich.location.baseline == "Lobar" | 
+           ich.location.baseline == "Thalamus" |
+           ich.location.baseline == "Basal Ganglia") %>%
+  filter(!is.na(ich.laterality))
+
+table1 <- CreateTableOne(data = ninds.combined,
+                         strata = "ich.laterality")
+
+```
+
+```{r output = TRUE}
 print(table1, showAllLevels = FALSE)

### 2023-01-06 manuscript.qmd
manuscript

---
title: "Hemispheric Lateralization and Outcomes After ICH"
subtitle: "A Pooled Analysis of NINDS Studies"
author: "Nikhil Patel, MD"
format: 
  html:
    toc: true
    toc-location: left
    toc-title: ""
    css: styles.css
    anchor-sections: true
    code-fold: true
    code-overflow: wrap
    code-summary: "Show the code"
    output: false
---

# Introduction

Brain injured patients may have different symptoms based on the lesion's laterality

This may affect patient's quality of life or disability down the road

This may also affect clinicians' attitude toward the need for aggressive treatment

We sought to figure out if there was differences in processes among clinicians with left hemispheric versus right hemispheric lesions

We also sought to determine if there are differences in functional outcomes and health-related quality of life

This has never been done in ICH

# Methods

We used data from NINDS clinical trials and 1 observational study: ATACH2, MISTIEII, MISTIEIII, CLEARIII, and ERICH.

### 2023-01-18 data-wrangling.qmd
updates

--- 2022-11-16 study protocol.qmd
@@
 
-mistie2.combined.demographics <- left_join(mistie2.demographics3, mistie2.demographics)
-mistie2.combined.demographics <- left_join(mistie2.demographics2, mistie2.combined.demographics)
+mistie2.combined.demographics <- left_join(mistie2.demographics, mistie2.demographics3)
+mistie2.combined.demographics <- left_join(mistie2.combined.demographics, mistie2.demographics2)
 
@@
 
+mistie2.whole <- mistie2.whole %>%
+  mutate(ich.laterality = case_when(
+    ich.laterality == "Discrepancy" ~ "NA",
+    ich.laterality == "Left" ~ "Left",
+    ich.laterality == "Right" ~ "Right",
+  ))
+
 ```
@@
                        "afib", 
-                       "pvd",
+                       "htn",
                        "hld",
@@
     ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino",
-    ethnicity == "Unknown/Not Reported" ~ "NA")) %>%
-  mutate(ich.location.baseline = case_when(
-    ich.location.baseline == "Basal Ganglia" ~ "Basal Ganglia",
-    ich.location.baseline == "Brainstem" ~ "Brainstem",
-    ich.location.baseline == "Caudate" ~ "Basal Ganglia",
-    ich.location.baseline == "Cerebellum" ~ "Cerebellum",
-    ich.location.baseline == "Frontal" ~ "Lobar",
-    ich.location.baseline == "Globus Pallidus" ~ "Basal Ganglia",
-    ich.location.baseline == "Lobar" ~ "Lobar",
-    ich.location.baseline == "Occipital" ~ "Lobar",
-    ich.location.baseline == "Other" ~ "Other",
-    ich.location.baseline == "Parietal" ~ "Lobar",
-    ich.location.baseline == "Primary IVH" ~ "Primary IVH",
-    ich.location.baseline == "Probable Primary IVH" ~ "Primary IVH",
-    ich.location.baseline == "Putamen" ~ "Basal Ganglia",
-    ich.location.baseline == "Temporal" ~ "Lobar",
-    ich.location.baseline == "Thalamus" ~ "Thalamus"
-  ))
-
+    ethnicity == "Unknown/Not Reported" ~ "NA")) 
 ninds.combined <- ninds.combined %>%
@@
            ich.location.baseline == "Basal Ganglia") %>%
+  mutate(ich.laterality = na_if(ich.laterality, "NA")) %>%
+  mutate(ethnicity = na_if(ethnicity, "NA")) %>%
+  mutate(stroke = na_if(stroke, "NA")) %>%
+  mutate(cad = na_if(cad, "NA")) %>%
+  mutate(chf = na_if(chf, "NA")) %>%
+  mutate(afib = na_if(afib, "NA")) %>%
+  mutate(htn = na_if(htn, "NA")) %>%
+  mutate(hld = na_if(hld, "NA")) %>%
+  mutate(dm2 = na_if(dm2, "NA")) %>%
+  mutate(tobacco = na_if(tobacco, "NA")) %>%
+  mutate(euro.mobility.90 = na_if(euro.mobility.90, "NA")) %>%
+  mutate(euro.selfcare.90 = na_if(euro.selfcare.90, "NA")) %>%
+  mutate(euro.usual.90 = na_if(euro.usual.90, "NA")) %>%
+  mutate(euro.pain.90 = na_if(euro.pain.90, "NA")) %>%
+  mutate(euro.anxiety.90 = na_if(euro.anxiety.90, "NA")) %>%
+  mutate(dnr.binary = na_if(dnr.binary, "NA")) %>%
+  mutate(comfort.binary = na_if(comfort.binary, "NA")) %>%
   filter(!is.na(ich.laterality))

### 2023-01-18 data-wrangling.qmd
writing to csv
html updates
updates

@@
 library(readxl)
+library(Hmisc)
 
@@
 
+ninds.combined$comfort.day <- round(ninds.combined$comfort.day)
+
 table1 <- CreateTableOne(data = ninds.combined,
@@
 ```
+
+```{r output = TRUE}
+
+write.csv(ninds.combined, file = "nindsCombined.csv")
+
+```

### 2023-01-19 code/data-wrangling.qmd
rearranged files

--- 2023-01-18 data-wrangling.qmd
@@
 library(Hmisc)
+library(kableExtra)
 
@@
     location_d180 == "Rehab" ~ "Rehabilitation",
-    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
+    location_d180 == "LTCF" ~ "Skilled Nursing Facility",library(tidyverse)
+library(tableone)
+library(data.table)
+library(haven) # Importing SAS Files
+library(janitor)
+library(lubridate)
+library(readxl)
+library(Hmisc)
     location_d180 == "Hospice" ~ "Hospice",
@@
 ```{r output = TRUE}
-print(table1, showAllLevels = FALSE)
-
+kbl(table1)
 ```

### 2023-01-19 code/importWrangling.r

@@ -1,0 +1,173 @@
+# loading libraries----
+
+library(tidyverse)
+library(tableone)
+library(data.table)
+library(haven) # Importing SAS Files
+library(janitor)
+library(lubridate)
+library(readxl)
+library(Hmisc)
+
+# loading raw data----
+
+# variables needed from each dataset-----
+# demographics: age, gender, site of enrollment
+# pmh: hypertension, diabetes, stroke, ICH, CAD, CKD
+# meds: anticoagulation, antiplatelet, lipids, bp
+# presentation features: symptom onset, SBP/DBP,GCS, NIHSS
+# presentation features: intubated
+# imaging: ICH volume, location, IVH, midline shift?
+# processes: EVD, neurosurgery, trach, PEG, WLST
+# outcomes: mRS,
+
+
+# atach-2----
+
+## import raw data----
+
+# Create function to paste together string of FORMxx.txt
+atachImportFunc <- function(y) { # nolint: object_name_linter.
+  filepath <- paste0("./data/ATACH2/FORM", y, ".txt")
+  dataset <- read.table(filepath, sep = "\t", header = TRUE) # nolint
+}
+
+## import data files that are tidy-structured
+atachImportVector <- c("00", "01", "02", "03", "04", "05", "07", "08", "11", "16", "17", "21", "33") # nolint # nolint
+
+atachList <- lapply(atachImportVector, FUN = atachImportFunc) # nolint
+
+atachWhole <- as.data.table(reduce(atachList, full_join, by = "SUBJECT_ID")) # nolint
+
+atachWhole[, study := "ATACH2"]
+
+## renaming variables----
+renameFunc <- function(x, n) setnames(x, names(n), n)
+
+renameFunc(
+  atachWhole,
+  .q(
+    SUBJECT_ID = id,
+    F33Q02 = age,
+    F00Q02MIN = symptomOnset,
+    F02Q03MIN = arrivalTimeInitial,
+    F02Q05MIN = arrivalTimeStrokeCenter,
+    F05Q07A = sbpInitial,
+    F05Q07B = dbpInitial,
+    F33Q04 = gcsBaseline,
[…]

### 2023-01-19 code/olddatawrangling.r

@@ -1,0 +1,361 @@
+
+library(tidyverse)
+library(tableone)
+library(data.table)
+library(haven) # Importing SAS Files
+library(janitor)
+library(lubridate)
+library(readxl)
+library(Hmisc)
+library(kableExtra)
+
+# Create function to paste together string of FORMxx.txt
+atach.import <- function(y) {
+  dataset <- paste0("atach2",".",y)
+  filepath <- paste0("./data/ATACH2/FORM",y,".txt")
+  dataset <- read.table(filepath, sep = "\t", header = TRUE)
+}
+
+atach00 <- atach.import("00")
+atach01 <- atach.import("01")
+atach02 <- atach.import("02")
+atach03 <- atach.import("03")
+atach04 <- atach.import("04")
+atach05 <- atach.import("05")
+# atach05c <- atach.import("05c") Detailed Blood Pressure Data, Unsure if Needed
+atach06 <- atach.import("06")
+atach07 <- atach.import("07")
+atach08 <- atach.import("08")
+atach09 <- atach.import("09")
+atach10 <- atach.import("10")
+atach11 <- atach.import("11")
+atach13 <- atach.import("13")
+atach14 <- atach.import("14")
+atach15 <- atach.import("15")
+atach16 <- atach.import("16")
+atach17 <- atach.import("17")
+atach18 <- atach.import("18")
+atach20 <- atach.import("20")
+atach21 <- atach.import("21")
+atach23 <- atach.import("23")
+atach33 <- atach.import("33")
+atachct <- read.table("./data/ATACH2/CTCR.txt", sep = "\t", header = TRUE)
+
+# Merging all tidy data
+atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach21, atach33)
+atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")
+atach.whole <-atach.whole %>% mutate(study = "ATACH2")
+
+# Importing CT Data
+atachct <- atachct %>%
+  mutate(ctnumber = case_when(
+    CTCRMIN < 0 ~ "baseline",
+    CTCRMIN > 0 ~ "24hourpost"
+  ))
+
+# Extracting First CT Data
+atachctbaseline <- atachct %>% 
+  filter(ctnumber == "baseline") %>%
[…]

### 2023-01-20 code/data-wrangling.qmd
completed ATACH2, working on ERICH

@@
   rename(dnr.day = F21Q09DAY)
+
+# variables needed from each dataset-----
+# demographics: age, gender, site of enrollment
+# pmh: hypertension, diabetes, stroke, ICH, CAD, CKD
+# meds: anticoagulation, antiplatelet, lipids, bp
+# presentation features: symptom onset, SBP/DBP,GCS, NIHSS
+# presentation features: intubated
+# imaging: ICH volume, location, IVH, midline shift?
+# processes: EVD, neurosurgery, trach, PEG, WLST
+# outcomes: mRS, euroQOL
 

### 2023-01-20 code/importWrangling.r

@@ -55,16 +55,25 @@ renameFunc(
     F05Q07A = sbpInitial,
     F05Q07B = dbpInitial,
     F33Q04 = gcsBaseline,
-    F07Q02 = intubationDays,
-    F07Q03DAY = extubationDays,
+    F07Q02 = daysVentilated,
+    F07Q03DAY = extubationDay,
     F07Q04DAY = trachDate,
+    F07Q06DAY = evdDate,
     F07Q07 = evdDays,
     F07Q09DAY = nsgyEvacDate,
     F16Q07 = euroVAS90,
-    EQ_INDEX = eqIndex90
+    EQ_INDEX = eqIndex90,
+    F16Q02 = euroMobility90,
+    F16Q03 = euroSelfCare90,
+    F16Q04 = euroUsual90,
+    F16Q05 = euroPain90,
+    F16Q06 = euroAnxiety90,
+    F21Q07DAY = comfortCareDate,
+    F21Q09DAY = dnrDate,
+    F11Q02 = icuLOS,
+    F11Q01DAY = hospitalLOS
   )
 )
-
 ## categorizing and recoding Variables----
 
 ### demographics----
@@ -102,6 +111,8 @@ atachWhole[, enrollingCountry := factor(fcase(
 
 # Having trouble using this function
 atachPMHFunc <- function(x, n) {
+  x <- deparse(substitute(x))
+  n <- deparse(substitute(n))
   atachWhole[, x := factor(fcase(
     n == 0, "No",
     n == 1, "Yes",
@@ -167,6 +178,251 @@ atachWhole[, cad := factor(fcase(
   F03Q06 == 98, "NA"
 ))]
 
+### medications-----
+
+atachWhole[, htnMeds := factor(fcase(
+  F04Q01 == 0, "No",
+  F04Q01 == 1, "Yes"
+))]
+
+atachWhole[, diabetesMeds := factor(fcase(
+  F04Q03 == 0, "No",
+  F04Q03 == 1, "Yes"
+))]
+
+### presentation features-----
+
+#### time from symptom onset to ED presentation | everything is relative to randomization but making it relative to symptom onset
+atachWhole[, symptomsToED := round(abs(arrivalTimeStrokeCenter - symptomOnset), digits = 2)]
+
+# mechanical ventilation
+
+atachWhole[, mechVentilation := factor(fcase(
[…]

### 2023-01-21 code/data-wrangling.qmd
erich wrangling

@@
 erich.whole$symptom.onset.interval <- as.period(erich.whole$symptom.onset.combined %--% erich.whole$ed.presentation.date.time)
+
 erich.whole$time.to.ed.minutes <- time_length(erich.whole$symptom.onset.interval, unit = "minutes")
+
 erich.whole$time.to.ed.hours <- time_length(erich.whole$symptom.onset.interval, unit = "hours")

### 2023-01-21 code/importWrangling.r

@@ -354,10 +354,7 @@ atachWhole[, hospDischargeDisp := factor(fcase(
   default = "NA"
 ))]
 
-
-
-## filtering only variables need in final analysis-----
-
+## filtering only variables need in final analysis------
 atachWhole <- atachWhole[, grep("^F0", colnames(atachWhole)) := NULL]
 atachWhole <- atachWhole[, grep("^F1", colnames(atachWhole)) := NULL]
 atachWhole <- atachWhole[, grep("^F21", colnames(atachWhole)) := NULL]
@@ -376,6 +373,11 @@ atachWhole <- atachWhole[, grep("^COUN", colnames(atachWhole)) := NULL]
 
 erichWhole <- as.data.table(read_sas("./data/ERICH/erich_case_100217.sas7bdat"))
 
+erichWhole <- erichWhole[, .(ID, age, gender, Race, Ethnicity, D6, D7, D7A, D21, D21A, D22, D25, PHY4, PHY6, PHY6A, PHY9, 1510, 1511, 1512, ED7, ED8, ED36, ED37, ED38, ED39, ED44, ED47_SYS, ED47_DIAS, SH1, SH8, SH9, SH10, SMK1A, MHX1, MHX3, MHX3A, MHX6, MHX8, MHX9, MHX11, MHX12, MHX13, MHX13A, MHX14, MHX16, MHX18, MHX20, MHX21, MHX22, MHX27C, MHX28, SA1, SA4, SA5A, SA6, SE1, SE2, SE3, SE15, LAB5, LAB6, IT5D, detail, ICH_Loc_CT1, ICH_Vol_CT1, IVH_Vol_CT1, ICH_Vol_CT2, ICH_Vol_CT3, ICH_Vol_CT4, IVH_Present, IT5D_DATE, IT5D_TIME, IT5H, IT5H_DATE, IT5H_TIME, IT5I, IT5I_DATE, IT5I_TIME, IT5F, IT5F_SPEC, IT5F_DATE, IT5F_TIME, IT7, IT8, IT8A, IT8B, IT13, IT13A, IT24, IT24A_DATE, IT24A_TIME, IT24A_UNITS, IT29, IT29A, IT29B, IT21, IT22, IT25, IT23, IT28, CC16, CC16_DATE, CC16_TIME, CC17, CC17_DATE, CC17_TIME, CC18, CC18_DATE, CC18_TIME, CX15, CX27, CX30,
[…]

### 2023-01-31 code/data-wrangling.qmd
finished combining data

@@
     location_d180 == "Rehab" ~ "Rehabilitation",
-    location_d180 == "LTCF" ~ "Skilled Nursing Facility",library(tidyverse)
-library(tableone)
-library(data.table)
-library(haven) # Importing SAS Files
-library(janitor)
-library(lubridate)
-library(readxl)
-library(Hmisc)
+    location_d180 == "LTCF" ~ "Skilled Nursing Facility",
     location_d180 == "Hospice" ~ "Hospice",

### 2023-01-31 code/importWrangling.r

@@ -9,6 +9,7 @@ library(lubridate)
 library(readxl)
 library(Hmisc)
 library(stringr)
+getRs("reptools.r") # Loads reptools.r from Github
 
 # loading raw data----
 
@@ -105,7 +106,7 @@ atachWhole[, enrollingCountry := factor(fcase(
   COUNTRY == 4, "South Korea",
   COUNTRY == 5, "Taiwan",
   COUNTRY == 6, "Germany",
-  default = "NA"
+  default = NA
 ))]
 
 ### past medical history-----
@@ -262,7 +263,7 @@ atachCT[, ichLaterality := factor(fcase(
   CTCRQ05 == 4, "Left",
   CTCRQ05 == 5, "Left",
   CTCRQ05 == 6, "Left",
-  default = "NA"
+  default = NA
 ))]
 
 # IVH
@@ -306,19 +307,20 @@ atachWhole[, dnr := factor(fcase(
 
 atachWhole[, comfortCare := factor(fcase(
   F21Q06 == 1, "Yes",
-  F21Q06 == 0, "No"
+  F21Q06 == 0, "No",
+  default = NA
 ))]
 
 # trachesotomy
 atachWhole[, tracheostomy := factor(fcase(
   is.na(trachDay), "No",
   !is.na(trachDay), "Yes",
-  default = "NA"
+  default = NA
 ))]
 
 ### outcome variables-----
 
-#### mRs-----
+#### mRS-----
 
 atachMRS <- as.data.table(atachImportFunc("09"))
 atachMRS <- setnames(atachMRS, "SUBJECT_ID", "id")
@@ -357,7 +359,7 @@ atachWhole[, hospDischargeDisp := factor(fcase(
   F11Q03 == 6, "Assisted Living",
   F11Q03 == 7, "Nursing Home Care",
   F11Q03 == 8, "Dead",
-  default = "NA"
+  default = NA
 ))]
 
 ## filtering only variables need in final analysis------
@@ -372,6 +374,7 @@ atachWhole <- atachWhole[, grep("^TREAT", colnames(atachWhole)) := NULL]
 atachWhole <- atachWhole[, grep("^SITE", colnames(atachWhole)) := NULL]
 atachWhole <- atachWhole[, grep("^RANDO", colnames(atachWhole)) := NULL]
 atachWhole <- atachWhole[, grep("^COUN", colnames(atachWhole)) := NULL]
[…]

### 2023-04-19 code/dataAnalysis/02analysis.qmd
updated code

---
title: "ICH Laterality Analysis"
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    html-math-method: katex
    css: "styles.css"
---

```{r setup and data import}
#| output: false

library(tidyverse)
library(tableone)
library(data.table)
library(haven) # Importing SAS Files
library(janitor)
library(lubridate)
library(readxl)
library(Hmisc)
library(stringr)
library(qs)
getRs("reptools.r") # Loads reptools.r from Github

d <- qread("./nindsCombined.qs")
```

**Causal question**: does the hemispheric location of an intracerebral hemorrhage (ICH) influence death or major disabilty?

```{r}

d ```

### 2023-04-19 manuscript/manuscript.qmd
updated code

---
title: "ICH Laterality Analysis"
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
---

Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue. It is a significant cause of morbidity and mortality, with high rates of disability and death. Hemispheric lateralization, which refers to the specialization of the left and right hemispheres of the brain for specific functions, has been suggested to play a role in the functional outcomes of patients with ICH. The left hemisphere is generally associated with language, verbal memory, and logical reasoning, while the right hemisphere is associated with spatial processing, attention, and emotional regulation.

Several studies have investigated the relationship between hemispheric lateralization and functional outcomes in patients with ICH, with conflicting results. Some studies have found that left hemisphere ICH is associated with worse functional outcomes than right hemisphere ICH, while others have found the opposite. Additionally, the specific functions affected by hemispheric lateralization in patients with ICH remain unclear.

Understanding the relationship between hemispheric lateralization and functional outcomes in patients with ICH is essential for improving patient care and developing targeted rehabilitation strategies. Therefore, this paper aims to review the current literature on hemispheric lateralization and functional outcomes in patients with ICH. We will discuss the potential mechanisms underlying these relationships and highlight the implications for clinical practice.

### 2023-04-19 code/01importCleaningCombining.r

@@ -9,6 +9,7 @@ library(lubridate)
 library(readxl)
 library(Hmisc)
 library(stringr)
+library(qs)
 getRs("reptools.r") # Loads reptools.r from Github
 
 # loading raw data----
@@ -1118,12 +1119,12 @@ mistie3Whole <- mistie3Whole[, .(id, age, tobacco, hld, cocaine, dm2, htn, cad, 
 ## exporting variable list-----
 mistie3VariableList <- names(mistie3Whole[, ])
 
-# combing all datasets
+# combing all datasets-----
 a <- list(atachWhole, erichWhole, clear3Whole, mistie2Whole, mistie3Whole)
 
 nindsCombined <- rbindlist(a, fill = TRUE)
 
-# cleaning up master dataset
+## cleaning up master dataset-----
 setcolorder(nindsCombined, .q(id, study, age, sex, race, ethnicity, gcsBaseline, ichLaterality, ichLocation))
 
 nindsCombined[, .q(arrivalTimeInitial, arrivalTimeStrokeCenter, enrollingCountry, mechVentilation, symptomOnsetDateTime, symptomOnsetEstimatedDateTime, symptomOnsetCombined, edDateTime, nsgyEvacDateTime, nsgyEvacHour, peg, deathDate, anticoagulated, comfortDateTime, comfortCareHour, dniDateTime, dnrDateTime, dnrHour, dniHour, dniDay) := NULL]
@@ -1136,11 +1137,47 @@ nindsCombined[, hour24IVHVolume := ceiling(hour24IVHVolume)]
 nindsCombined[, hour24PHEVolume := ceiling(hour24PHEVolume)]
 
 nindsCombined$ichLaterality <- na_if(nindsCombined$ichLaterality, "NA")
-nindsCombined$ichLaterality <- na_if(nindsCombined$ichLaterality, "NA")
 
 nindsCombined <- nindsCombined %>%
   mutate(across(.q(ichLocation, stroke, chf, afib, htn, pvd, cad, hld, tobacco, dnr, comfortCare), na_if, "NA"))
 
-contents(nindsCombined)
+## Computing ICH Score-----
 
-table(nindsCombined$comfortCare)
+nindsCombined[, ichScoreVolume := fcase(
+  baselineICHVolume >= 30, 1,
+  baselineICHVolume < 30, 0,
+  default = 0
+)]
+
[…]

### 2023-04-19 code/olddatawrangling.r

@@ -1,361 +1,0 @@
-
-library(tidyverse)
-library(tableone)
-library(data.table)
-library(haven) # Importing SAS Files
-library(janitor)
-library(lubridate)
-library(readxl)
-library(Hmisc)
-library(kableExtra)
-
-# Create function to paste together string of FORMxx.txt
-atach.import <- function(y) {
-  dataset <- paste0("atach2",".",y)
-  filepath <- paste0("./data/ATACH2/FORM",y,".txt")
-  dataset <- read.table(filepath, sep = "\t", header = TRUE)
-}
-
-atach00 <- atach.import("00")
-atach01 <- atach.import("01")
-atach02 <- atach.import("02")
-atach03 <- atach.import("03")
-atach04 <- atach.import("04")
-atach05 <- atach.import("05")
-# atach05c <- atach.import("05c") Detailed Blood Pressure Data, Unsure if Needed
-atach06 <- atach.import("06")
-atach07 <- atach.import("07")
-atach08 <- atach.import("08")
-atach09 <- atach.import("09")
-atach10 <- atach.import("10")
-atach11 <- atach.import("11")
-atach13 <- atach.import("13")
-atach14 <- atach.import("14")
-atach15 <- atach.import("15")
-atach16 <- atach.import("16")
-atach17 <- atach.import("17")
-atach18 <- atach.import("18")
-atach20 <- atach.import("20")
-atach21 <- atach.import("21")
-atach23 <- atach.import("23")
-atach33 <- atach.import("33")
-atachct <- read.table("./data/ATACH2/CTCR.txt", sep = "\t", header = TRUE)
-
-# Merging all tidy data
-atach.tidy <- list(atach00, atach01, atach02, atach03, atach04, atach05, atach07, atach08, atach11, atach16, atach17, atach21, atach33)
-atach.whole <- reduce(atach.tidy, full_join, by = "SUBJECT_ID")
-atach.whole <-atach.whole %>% mutate(study = "ATACH2")
-
-# Importing CT Data
-atachct <- atachct %>%
-  mutate(ctnumber = case_when(
-    CTCRMIN < 0 ~ "baseline",
-    CTCRMIN > 0 ~ "24hourpost"
-  ))
-
-# Extracting First CT Data
-atachctbaseline <- atachct %>% 
-  filter(ctnumber == "baseline") %>%
[…]

### 2023-05-17 code/dataAnalysis/02analysis.qmd
updates

@@
 ---
-title: "ICH Laterality Analysis"
+title: "Hemispheric Lateralization in ICH"
+subtitle: "Analysis"
+author: "Nikhil Patel"
+project:
+  type: default
 format:
@@
 library(tidyverse)
+# library(gt) # Fancy tables
 library(tableone)
@@
 library(qs)
-getRs("reptools.r") # Loads reptools.r from Github
+library(glue)
+library(scales) # Nicer labeling functions
+library(ggmosaic) # Mosaic plots with ggplot
+library(ggplot2)
+library(ggpattern) # Pattern fills in ggplot
+library(patchwork) # Combine plots nicely
+library(systemfonts) ## use custom fonts (need to be installed on your OS)
+library(scico) ## scico color palettes(http://www.fabiocrameri.ch/colourmaps.php) in R
+library(ggtext) ## add improved text rendering to ggplot2
+library(ggforce) ## add missing functionality to ggplot2
+library(ggdist) ## add uncertainity visualizations to ggplot2
 
-d <- qread("./nindsCombined.qs")
+d <- qread("../../data/nindsCombined.qs")
+
 ```
 
-**Causal question**: does the hemispheric location of an intracerebral hemorrhage (ICH) influence death or major disabilty?
+## Causal Question
+**Does the hemispheric location of an intracerebral hemorrhage (ICH) influence outcomes such as death, disability, or health-related quality of life?**
+
+Clinical manifestations of brain injury differ between left and right hemispheric lesions because of differences in function, in particular language and spatial awareness.
+
+It is unclear if these clinical differences result in differences in functional disability or health-related quality of life.
+Previous studies have demonstrated conflicting results, with some demonstrating worse levels of disability for patients with left hemispheric lesions as well as the opposite.
+
+We aim to evaluate the question of if patients with left-hemispheric lesions have worse levels of disability or health-related quality of life.
+
+### Data Source
+
+We used individual patient data from 5 NINDS multicenter prospective studies enrolling patients with intracerebral hemorrhage (4 randomized trials: ATACH2, CLEAR3, MISTIE-II, and MISTIE-III and 1 observational study: ERICH).
+
+### Visualizing the Data
 
@@
 
-d ```
+mrsCounts <- d |>
+  filter(!is.na(mrs90)) |>
+  group_by(ichLaterality, mrs90) |>
+  tally(name = "patients") |>
+  group_by(ichLaterality) |>
+  mutate(percentPatients = patients / sum(patients)) |>
+  ungroup() |>
+  mutate(percentPatientsLabel = percent(percentPatients, accuracy = 1))
+
+theme_set(theme_minimal(base_size = 10, base_family = "Figtree"))
+
+myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
+
+ggplot(
+  data = mrsCounts,
+  aes(x = ichLaterality, y = percentPatients, fill = forcats::fct_rev(mrs90))
+) +
+  geom_bar(stat = "identity", position = "fill", width = 0.5) +
+  geom_text(aes(label = percentPatientsLabel),
+    position = position_stack(vjust = 0.5),
+    color = "white",
+    size = 5,
+    family = "FigTree"
+  ) +
+  coord_flip() +
+  scale_y_continuous(
+    name = "Patients (%)",
+    labels = scales::percent_format(accuracy = 1),
+    breaks = c(seq(0, 1, 0.2))
+  ) +
+  scale_fill_manual(values = myColors) +
+  theme(
+    plot.title = element_text(face = "bold", size = 100, lineheight = 0.1),
+    plot.subtitle = element_text(size = 72, lineheight = 0.1),
+    plot.caption = element_text(size = 22),
+    plot.title.position = "plot",
+    plot.caption.position = "plot",
+    legend.title = element_blank(),
+    legend.position = "bottom",
+    legend.text = element_text(size = 20),
+    legend.margin = margin(r = 75),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+    axis.title.x = element_text(size = 20, lineheight = 0.5),
+    axis.title.y = element_text(family = "outfit")
+  )
+
+```
+
+## Draw Assumptions using a Causal Diagram
+
+## Model Assumptions
+
+## Diagnose Models
+
+## Estimate the Causal Effect
+
+## Conduct Sensitivity Analysis on the Effect Estimate
+
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#
+#

### 2023-05-17 code/dataAnalysis/script.r

@@ -1,0 +1,11 @@
+library(tidyverse)
+
+getwd()
+
+d <- qread("./data/nindsCombined.qs")
+
+mrsCounts <- d |>
+  filter(!is.na(mrs90)) |>
+  group_by(ichLaterality, mrs90) |>
+  tally() |>
+  dplyr::mutate(percent = scales::percent(n / sum(n), accuracy = 1, trim = FALSE))

### 2023-05-17 code/dataAnalysis/02analysis.qmd
mrs figures done

@@
     css: "styles.css"
+    fig-width: 12
 ---
@@
 library(ggdist) ## add uncertainity visualizations to ggplot2
+library(viridis)
 
@@
 
+Comparing outcomes between left and right hemisphere location overall. There does not appear to be a big difference.
+
 ```{r}
@@
 
-theme_set(theme_minimal(base_size = 10, base_family = "Figtree"))
+theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
 
@@
   geom_bar(stat = "identity", position = "fill", width = 0.5) +
-  geom_text(aes(label = percentPatientsLabel),
+  geom_text(aes(label = patients),
     position = position_stack(vjust = 0.5),
@@
   scale_y_continuous(
-    name = "Patients (%)",
+    name = NULL,
     labels = scales::percent_format(accuracy = 1),
@@
   ) +
-  scale_fill_manual(values = myColors) +
+  scale_x_discrete(name = NULL) +
+  scale_fill_manual(values = myColors, breaks = c(0, 1, 2, 3, 4, 5, 6)) +
+  labs(
+    title = "Hemispheric Laterality",
+    subtitle = "modified Rankin score at 90 days"
+  ) +
   theme(
-    plot.title = element_text(face = "bold", size = 100, lineheight = 0.1),
-    plot.subtitle = element_text(size = 72, lineheight = 0.1),
-    plot.caption = element_text(size = 22),
+    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.subtitle = element_text(size = 20, lineheight = 0.1),
+    plot.caption = element_text(size = 20),
     plot.title.position = "plot",
@@
     legend.title = element_blank(),
-    legend.position = "bottom",
-    legend.text = element_text(size = 20),
-    legend.margin = margin(r = 75),
+    legend.position = "right",
+    legend.text = element_text(size = 15, lineheight = 0.5),
+    # legend.margin = margin(r = 50),
     panel.grid.minor.x = element_blank(),
@@
     axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
-    axis.title.x = element_text(size = 20, lineheight = 0.5),
-    axis.title.y = element_text(family = "outfit")
+    axis.title.x = element_text(size = 20, lineheight = 0.5)
+  )
+
+```
+
+```{r}
+
+mrsCounts <- d |>
+  filter(!is.na(mrs180)) |>
+  group_by(ichLaterality, mrs180) |>
+  tally(name = "patients") |>
+  group_by(ichLaterality) |>
+  mutate(percentPatients = patients / sum(patients)) |>
+  ungroup() |>
+  mutate(percentPatientsLabel = percent(percentPatients, accuracy = 1))
+
+theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
+
+myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
+
+ggplot(
+  data = mrsCounts,
+  aes(x = ichLaterality, y = percentPatients, fill = forcats::fct_rev(mrs180))
+) +
+  geom_bar(stat = "identity", position = "fill", width = 0.5) +
+  geom_text(aes(label = patients),
+    position = position_stack(vjust = 0.5),
+    color = "white",
+    size = 5,
+    family = "FigTree"
+  ) +
+  coord_flip() +
+  scale_y_continuous(
+    name = NULL,
+    labels = scales::percent_format(accuracy = 1),
+    breaks = c(seq(0, 1, 0.2))
+  ) +
+  scale_x_discrete(name = NULL) +
+  scale_fill_manual(values = myColors, breaks = c(0, 1, 2, 3, 4, 5, 6)) +
+  labs(
+    title = "Hemispheric Laterality",
+    subtitle = "modified Rankin score at 180 days"
+  ) +
+  theme(
+    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.subtitle = element_text(size = 20, lineheight = 0.1),
+    plot.caption = element_text(size = 20),
+    plot.title.position = "plot",
+    plot.caption.position = "plot",
+    legend.title = element_blank(),
+    legend.position = "right",
+    legend.text = element_text(size = 15, lineheight = 0.5),
+    # legend.margin = margin(r = 50),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+    axis.title.x = element_text(size = 20, lineheight = 0.5)
+  )
+
+```
+
+```{r}
+
+mrsCounts <- d |>
+  filter(!is.na(mrs365)) |>
+  group_by(ichLaterality, mrs365) |>
+  tally(name = "patients") |>
+  group_by(ichLaterality) |>
+  mutate(percentPatients = patients / sum(patients)) |>
+  ungroup() |>
+  mutate(percentPatientsLabel = percent(percentPatients, accuracy = 1))
+
+theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
+
+myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
+
+ggplot(
+  data = mrsCounts,
+  aes(x = ichLaterality, y = percentPatients, fill = forcats::fct_rev(mrs365))
+) +
+  geom_bar(stat = "identity", position = "fill", width = 0.5) +
+  geom_text(aes(label = patients),
+    position = position_stack(vjust = 0.5),
+    color = "white",
+    size = 5,
+    family = "FigTree"
+  ) +
+  coord_flip() +
+  scale_y_continuous(
+    name = NULL,
+    labels = scales::percent_format(accuracy = 1),
+    breaks = c(seq(0, 1, 0.2))
+  ) +
+  scale_x_discrete(name = NULL) +
+  scale_fill_manual(values = myColors, breaks = c(0, 1, 2, 3, 4, 5, 6)) +
+  labs(
+    title = "Hemispheric Laterality",
+    subtitle = "modified Rankin score at 365 days"
+  ) +
+  theme(
+    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.subtitle = element_text(size = 20, lineheight = 0.1),
+    plot.caption = element_text(size = 20),
+    plot.title.position = "plot",
+    plot.caption.position = "plot",
+    legend.title = element_blank(),
+    legend.position = "right",
+    legend.text = element_text(size = 15, lineheight = 0.5),
+    # legend.margin = margin(r = 50),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+    axis.title.x = element_text(size = 20, lineheight = 0.5)
   )
@@
 #
+#
+#
+#
+#

### 2023-05-18 code/dataAnalysis/02analysis.qmd
euroQOL

@@
 
+```{r}
+
+euroQOLCounts <- d |>
+  filter(!is.na(euroMobility90)) |>
+  group_by(ichLaterality, euroMobility90) |>
+  tally(name = "patients") |>
+  group_by(ichLaterality) |>
+  mutate(percentPatients = patients / sum(patients)) |>
+  ungroup() |>
+  mutate(percentPatientsLabel = percent(percentPatients, accuracy = 1))
+
+theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
+
+myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
+
+ggplot(
+  data = euroQOLCounts,
+  aes(x = ichLaterality, y = percentPatients, fill = euroMobility90)
+) +
+  geom_bar(stat = "identity", position = "fill", width = 0.5) +
+  geom_text(aes(label = patients),
+    position = position_stack(vjust = 0.5),
+    color = "white",
+    size = 5,
+    family = "FigTree"
+  ) +
+  coord_flip() +
+  scale_y_continuous(
+    name = NULL,
+    labels = scales::percent_format(accuracy = 1),
+    breaks = c(seq(0, 1, 0.2))
+  ) +
+  scale_x_discrete(name = NULL) +
+  # scale_fill_viridus_c() +
+  labs(
+    title = "Hemispheric Laterality",
+    subtitle = "euroMobility90"
+  ) +
+  theme(
+    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.subtitle = element_text(size = 20, lineheight = 0.1),
+    plot.caption = element_text(size = 20),
+    plot.title.position = "plot",
+    plot.caption.position = "plot",
+    legend.title = element_blank(),
+    legend.position = "right",
+    legend.text = element_text(size = 15, lineheight = 0.5),
+    # legend.margin = margin(r = 50),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+    axis.title.x = element_text(size = 20, lineheight = 0.5)
+  )
+
+```
+
 ## Draw Assumptions using a Causal Diagram

### 2023-05-19 code/dataAnalysis/02analysis.qmd
Update 02analysis.qmd

@@
 library(viridis)
+library(dagitty)
 
@@
 
+```{r}
+
+dag {
+bb="0,0,1,1"
+"Disease Severity" [adjusted,pos="0.152,0.791"]
+"Early WLST" [adjusted,pos="0.326,0.215"]
+"Functional Outcome" [outcome,pos="0.916,0.455"]
+"ICH Laterality" [exposure,pos="0.070,0.455"]
+Age [adjusted,pos="0.253,0.858"]
+Comorbidities [adjusted,pos="0.380,0.903"]
+Neurosurgery [adjusted,pos="0.736,0.148"]
+Rehabilitation [latent,pos="0.849,0.703"]
+"Disease Severity" -> "Early WLST"
+"Disease Severity" -> "Functional Outcome"
+"Disease Severity" -> Neurosurgery
+"Disease Severity" -> Rehabilitation
+"Early WLST" -> "Functional Outcome"
+"ICH Laterality" -> "Early WLST"
+"ICH Laterality" -> "Functional Outcome"
+"ICH Laterality" -> Neurosurgery
+Age -> "Functional Outcome"
+Age -> Neurosurgery
+Age -> Rehabilitation
+Comorbidities -> "Functional Outcome"
+Comorbidities -> Neurosurgery
+Comorbidities -> Rehabilitation
+Neurosurgery -> "Functional Outcome"
+Rehabilitation -> "Functional Outcome"
+}
+
+```
+
 ## Model Assumptions

### 2023-05-20 code/dataAnalysis/02analysis.qmd
Update 02analysis.qmd

--- 2023-05-17 code/dataAnalysis/02analysis.qmd
@@
     css: "styles.css"
-    fig-width: 12
 ---
@@
 #| output: false
+#| echo: false
 
@@
 library(Hmisc)
+library(qreport)
 library(stringr)
@@
 library(viridis)
+library(dagitty)
+library(rlang)
+library(ggokabeito)
 
 d <- qread("../../data/nindsCombined.qs")
+
+# ggplot themes and settings
+theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
+
+myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
 
@@
 
-Comparing outcomes between left and right hemisphere location overall. There does not appear to be a big difference.
+#### modified Rankin Scale
+
+Overall, there does not appear to be a big difference in disability as measured by modified Rankin Scale. This is true at any of the time points used.
 
 ```{r}
+#| echo: false
+#| fig.width: 12
+#| fig.height: 15
 
+# mRs 90
 mrsCounts <- d |>
@@
 
-theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
-
-myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
-
-ggplot(
+gmrs90 <- ggplot(
   data = mrsCounts,
@@
   labs(
-    title = "Hemispheric Laterality",
-    subtitle = "modified Rankin score at 90 days"
-  ) +
+    title = "90 days") +
   theme(
-    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.title = element_text(face = "bold", size = 20, lineheight = 0.1),
     plot.subtitle = element_text(size = 20, lineheight = 0.1),
@@
 
-```
-
-```{r}
-
+# mRs 180
 mrsCounts <- d |>
@@
 
-theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
-
-myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
-
-ggplot(
+gmrs180 <- ggplot(
   data = mrsCounts,
@@
   labs(
-    title = "Hemispheric Laterality",
-    subtitle = "modified Rankin score at 180 days"
-  ) +
+    title = "180 days") +
   theme(
-    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.title = element_text(face = "bold", size = 20, lineheight = 0.1),
     plot.subtitle = element_text(size = 20, lineheight = 0.1),
@@
 
-```
-
-```{r}
-
+# mRs 365
 mrsCounts <- d |>
@@
 
-theme_set(theme_minimal(base_size = 20, base_family = "Figtree"))
-
-myColors <- c("#ffbe63", "#f39264", "#d76c6a", "#ae4f6e", "#7c3c6b", "#462d5d", "#0a1e45") # https://learnui.design/tools/data-color-picker.html#palette
-
-ggplot(
+gmrs365 <- ggplot(
   data = mrsCounts,
@@
   labs(
-    title = "Hemispheric Laterality",
-    subtitle = "modified Rankin score at 365 days"
-  ) +
+    title = "365 days") +
   theme(
-    plot.title = element_text(face = "bold", size = 30, lineheight = 0.1),
+    plot.title = element_text(face = "bold", size = 20, lineheight = 0.1),
     plot.subtitle = element_text(size = 20, lineheight = 0.1),
@@
     legend.text = element_text(size = 15, lineheight = 0.5),
-    # legend.margin = margin(r = 50),
     panel.grid.minor.x = element_blank(),
@@
 
+mrsCombined <- 
+(gmrs90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (gmrs180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (gmrs365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
+  plot_annotation(title = 'ICH Laterality and modified Rankin Scale',
+                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))) +
+  plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
+
+mrsCombined
+```
+
+#### Health-Related Quality of Life
+
+Overall, there appears to be a small differences in health related quality of life, especially with mobility and self-care. This is true at any of the time points used.
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig.width: 12
+#| fig.height: 30
+
+myColors <- c("#d76c6a", "#ae4f6e", "#7c3c6b") # https://learnui.design/tools/data-color-picker.html#palette
+
+euroQolPlot <- function(euroSelfCare365, legendLabels, plotTitle) {
+  euroQOLCounts <- d |>
+  filter(!is.na({{ euroSelfCare365 }})) |>
+  group_by(ichLaterality, {{ euroSelfCare365 }}) |>
+  tally(name = "patients") |>
+  group_by(ichLaterality) |>
+  mutate(percentPatients = patients / sum(patients)) |>
+  ungroup() |>
+  mutate(percentPatientsLabel = percent(percentPatients, accuracy = 1))
+
+ggplot(
+  data = euroQOLCounts,
+  aes(x = ichLaterality, y = percentPatients, fill = fct_rev({{ euroSelfCare365 }}))
+) +
+  geom_bar(stat = "identity", position = "fill", width = 0.5) +
+  geom_text(aes(label = percentPatientsLabel),
+    position = position_stack(vjust = 0.5),
+    color = "white",
+    size = 5) +
+  coord_flip() +
+  scale_y_continuous(
+    name = NULL,
+    labels = scales::percent_format(accuracy = 1),
+    breaks = c(seq(0, 1, 0.2))
+  ) +
+  scale_x_discrete(name = NULL) +
+  scale_fill_manual(values = myColors, breaks = legendLabels) +
+  labs(
+    title = plotTitle) +
+  theme(
+    plot.title = element_text(face = "bold", size = 20, lineheight = 0.1),
+    plot.subtitle = element_text(size = 20, lineheight = 0.1),
+    plot.caption = element_text(size = 20),
+    plot.title.position = "plot",
+    plot.caption.position = "plot",
+    legend.title = element_blank(),
+    legend.position = "bottom",
+    legend.text = element_text(size = 15, lineheight = 0.5),
+    panel.grid.minor.x = element_blank(),
+    panel.grid.major.y = element_blank(),
+    axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+    axis.title.x = element_text(size = 20, lineheight = 0.5)
+  )
+}
+legendLabels <- c("No Problems", "Some Problems", "Confined to Bed")
+ggeuroMobility90 <- euroQolPlot(euroMobility90, legendLabels, "EuroQOL Mobility")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")
+ggeuroSelfCare90 <- euroQolPlot(euroSelfCare90, legendLabels, "EuroQOL Self-Care")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")
+ggeuroUsual90 <- euroQolPlot(euroUsual90, legendLabels, "EuroQOL Usual Activities")
+
+legendLabels <- c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")
+ggeuroPain90 <- euroQolPlot(euroPain90, legendLabels, "EuroQOL Pain")
+
+legendLabels <- c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")
+ggeuroAnxiety90 <- euroQolPlot(euroAnxiety90, legendLabels, "EuroQOL Anxiety")
+
+legendLabels <- c("No Problems", "Some Problems", "Confined to Bed")
+ggeuroMobility180 <- euroQolPlot(euroMobility180, legendLabels, "EuroQOL Mobility")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")
+ggeuroSelfCare180 <- euroQolPlot(euroSelfCare180, legendLabels, "EuroQOL Self-Care")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")
+ggeuroUsual180 <- euroQolPlot(euroUsual180, legendLabels, "EuroQOL Usual Activities")
+
+legendLabels <- c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")
+ggeuroPain180 <- euroQolPlot(euroPain180, legendLabels, "EuroQOL Pain")
+
+legendLabels <- c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")
+ggeuroAnxiety180 <- euroQolPlot(euroAnxiety180, legendLabels, "EuroQOL Anxiety")
+
+legendLabels <- c("No Problems", "Some Problems", "Confined to Bed")
+ggeuroMobility365 <- euroQolPlot(euroMobility365, legendLabels, "EuroQOL Mobility")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")
+ggeuroSelfCare365 <- euroQolPlot(euroSelfCare365, legendLabels, "EuroQOL Self-Care")
+
+legendLabels <- c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")
+ggeuroUsual365 <- euroQolPlot(euroUsual365, legendLabels, "EuroQOL Usual Activities")
+
+legendLabels <- c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")
+ggeuroPain365 <- euroQolPlot(euroPain365, legendLabels, "EuroQOL Pain")
+
+legendLabels <- c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")
+ggeuroAnxiety365 <- euroQolPlot(euroAnxiety365, legendLabels, "EuroQOL Anxiety")
+
+euroQolCombined90 <- 
+(ggeuroMobility90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
+(ggeuroPain90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
+(ggeuroAnxiety90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
+  plot_annotation(title = 'ICH Laterality and HRQOL at 90 days',
+                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
+
+euroQolCombined180 <- 
+(ggeuroMobility180 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
+(ggeuroPain180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
+(ggeuroAnxiety180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
+  plot_annotation(title = 'ICH Laterality and HRQOL at 180 days',
+                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
+
+euroQolCombined365 <- 
+(ggeuroMobility365 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
+(ggeuroPain365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
+(ggeuroAnxiety365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
+  plot_annotation(title = 'ICH Laterality and HRQOL at 365 days',
+                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
+
+euroQolCombined90
+euroQolCombined180
+euroQolCombined365
+
+```
+
+#### Health-Related Quality of Life - Visual Analog Scale
+
+```{r}
+#| echo: false
+#| warning: false
+#| fig-width: 15
+#| fig-height: 25
+
+euroVASplot <- function(x, title) {
+  d |>
+  ggplot(aes({{x}}, fill = ichLaterality)) +
+  geom_density(color = NA, alpha = 0.8) +
+  labs(title = {{ title }},
+        x = NULL,
+        y = NULL) +
+  theme(
+  plot.title = element_text(face = "bold", size = 20, lineheight = 0.1),
+  plot.subtitle = element_text(size = 20, lineheight = 0.1),
+  plot.caption = element_text(size = 20),
+  plot.title.position = "plot",
+  plot.caption.position = "plot",
+  legend.title = element_blank(),
+  legend.position = "bottom",
+  legend.text = element_text(size = 15, lineheight = 0.5),
+  panel.grid.minor.x = element_blank(),
+  panel.grid.major.y = element_blank(),
+  axis.text.y = element_text(size = 20, lineheight = 0.5, margin = margin(r = -10)),
+  axis.title.x = element_text(size = 20, lineheight = 0.5)
+  )
+}
+
+ggeurovas90 <- euroVASplot(euroVAS90, "90 Days")
+ggeurovas180 <- euroVASplot(euroVAS180, "180 Days")
+ggeurovas365 <- euroVASplot(euroVAS365, "365 Days")
+
+ggeuroVASCombined <- 
+(ggeurovas90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeurovas180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeurovas365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
+  plot_annotation(title = 'ICH Laterality EuroQOL VAS',
+                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+
+ggeuroVASCombined
+
 ```
@@
 ## Draw Assumptions using a Causal Diagram
+
+```{r}
+#| echo: false
+
+dag <- dagitty('dag {
+bb="0,0,1,1"
+"Disease Severity" [adjusted,pos="0.152,0.791"]
+"Early WLST" [adjusted,pos="0.326,0.215"]
+"Functional Outcome" [outcome,pos="0.916,0.455"]
+"ICH Laterality" [exposure,pos="0.070,0.455"]
+Age [adjusted,pos="0.153,0.958"]
+Comorbidities [adjusted,pos="0.380,0.950"]
+Neurosurgery [adjusted,pos="0.736,0.148"]
+Rehabilitation [latent,pos="0.849,0.703"]
+"Disease Severity" -> "Early WLST"
+"Disease Severity" -> "Functional Outcome"
+"Disease Severity" -> Neurosurgery
+"Disease Severity" -> Rehabilitation
+"Early WLST" -> "Functional Outcome"
+"ICH Laterality" -> "Early WLST"
+"ICH Laterality" -> "Functional Outcome"
+"ICH Laterality" -> Neurosurgery
+Age -> "Functional Outcome"
+Age -> Neurosurgery
+Age -> Rehabilitation
+Age -> Comorbidities
+Comorbidities -> "Functional Outcome"
+Comorbidities -> Neurosurgery
+Comorbidities -> Rehabilitation
+Comorbidities -> "Disease Severity"
+Neurosurgery -> "Functional Outcome"
+Rehabilitation -> "Functional Outcome"
+}')
+
+plot(dag)
+```
+
+**Add Deaths to ATACH data!**
 
@@
 ## Conduct Sensitivity Analysis on the Effect Estimate
-
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#
-#

### 2023-05-20 code/01importCleaningCombining.r

@@ -344,15 +344,41 @@ atachMRS <- unique(atachMRS) # removing duplicate NIHSS rows
 atachWhole <- atachMRS[atachWhole, on = .(id)]
 
 #### euroQOL-----
-levels(atachWhole$euroMobility90) <- c("No Problems", "Some Problems", "Confined to Bed")
 
-levels(atachWhole$euroSelfCare90) <- c("No Problems", "Some Problems", "Unable to Wash or Dress Myself")
+atachWhole[, euroMobility90 := factor(fcase(
+  euroMobility90 == "1", "No Problems",
+  euroMobility90 == "2", "Some Problems",
+  euroMobility90 == "3", "Confined to Bed",
+  default = NA
+))]
 
-levels(atachWhole$euroUsual90) <- c("No Problems", "Some Problems", "Unable to Perform My Usual Activities")
+atachWhole[, euroSelfCare90 := factor(fcase(
+  euroSelfCare90 == "1", "No Problems",
+  euroSelfCare90 == "2", "Some Problems",
+  euroSelfCare90 == "3", "Unable to Wash or Dress Myself",
+  default = NA
+))]
 
-levels(atachWhole$euroPain90) <- c("No Pain/Discomfort", "Some Moderate Pain/Discomfort", "Extreme Pain/Discomfort")
+atachWhole[, euroUsual90 := factor(fcase(
+  euroUsual90 == "1", "No Problems",
+  euroUsual90 == "2", "Some Problems",
+  euroUsual90 == "3", "Unable to Perform My Usual Activities",
+  default = NA
+))]
 
-levels(atachWhole$euroAnxiety90) <- c("Not Anxious/Depressed", "Moderately Anxious/Depressed", "Extremely Anxious/Depressed")
+atachWhole[, euroPain90 := factor(fcase(
+  euroPain90 == "1", "No Pain/Discomfort",
+  euroPain90 == "2", "Some Moderate Pain/Discomfort",
+  euroPain90 == "3", "Extreme Pain/Discomfort",
+  default = NA
+))]
+
+atachWhole[, euroAnxiety90 := factor(fcase(
+  euroAnxiety90 == "1", "Not Anxious/Depressed",
+  euroAnxiety90 == "2", "Moderately Anxious/Depressed",
+  euroAnxiety90 == "3", "Extremely Anxious/Depressed",
+  default = NA
+))]
 
 #### hospital discharge disposition
 
[…]

### 2023-05-20 code/dataAnalysis/script.r

@@ -1,4 +1,8 @@
 library(tidyverse)
+library(qs)
+library(data.table)
+library(Hmisc)
+library(qreport)
 
 getwd()
 
@@ -9,3 +13,11 @@ mrsCounts <- d |>
   group_by(ichLaterality, mrs90) |>
   tally() |>
   dplyr::mutate(percent = scales::percent(n / sum(n), accuracy = 1, trim = FALSE))
+
+dlist <- c(d$mrs30, d$mrs90)
+
+missChk(dlist)
+
+str(d$euroMobility90)
+
+fct_count(d$euroSelfCare365)

### 2023-05-24 code/dataAnalysis/02analysis.qmd
who knows

@@
 library(Hmisc)
-library(qreport)
+# library(qreport)
 library(stringr)
@@
 library(rlang)
-library(ggokabeito)
+# library(ggokabeito)
 

### 2023-06-02 code/dataAnalysis/02analysis.qmd
changes

@@
   labs(
-    title = "90 days") +
+    title = "90 days"
+  ) +
   theme(
@@
   labs(
-    title = "180 days") +
+    title = "180 days"
+  ) +
   theme(
@@
   labs(
-    title = "365 days") +
+    title = "365 days"
+  ) +
   theme(
@@
 mrsCombined <-
-(gmrs90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (gmrs180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (gmrs365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
-  plot_annotation(title = 'ICH Laterality and modified Rankin Scale',
-                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))) +
-  plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
+  (gmrs90 + theme(plot.margin = unit(c(50, 0, 75, 0), "pt"))) / (gmrs180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) / (gmrs365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) +
+    plot_annotation(
+      title = "ICH Laterality and modified Rankin Scale",
+      theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))
+    ) +
+    plot_layout(guides = "collect") & theme(legend.position = "right", legend.direction = "vertical")
 
@@
       color = "white",
-    size = 5) +
+      size = 5
+    ) +
     coord_flip() +
@@
     labs(
-    title = plotTitle) +
+      title = plotTitle
+    ) +
     theme(
@@
 euroQolCombined90 <-
-(ggeuroMobility90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
-(ggeuroPain90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
-(ggeuroAnxiety90 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
-  plot_annotation(title = 'ICH Laterality and HRQOL at 90 days',
-                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  (ggeuroMobility90 + theme(plot.margin = unit(c(50, 0, 75, 0), "pt"))) / (ggeuroSelfCare90 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) / (ggeuroUsual90 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroPain90 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroAnxiety90 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) +
+  plot_annotation(
+    title = "ICH Laterality and HRQOL at 90 days",
+    theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))
+  )
 # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
@@
 euroQolCombined180 <-
-(ggeuroMobility180 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
-(ggeuroPain180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
-(ggeuroAnxiety180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
-  plot_annotation(title = 'ICH Laterality and HRQOL at 180 days',
-                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  (ggeuroMobility180 + theme(plot.margin = unit(c(50, 0, 75, 0), "pt"))) / (ggeuroSelfCare180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) / (ggeuroUsual180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroPain180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroAnxiety180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) +
+  plot_annotation(
+    title = "ICH Laterality and HRQOL at 180 days",
+    theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))
+  )
 # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
@@
 euroQolCombined365 <-
-(ggeuroMobility365 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeuroSelfCare365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeuroUsual365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) /
-(ggeuroPain365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / 
-(ggeuroAnxiety365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
-  plot_annotation(title = 'ICH Laterality and HRQOL at 365 days',
-                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  (ggeuroMobility365 + theme(plot.margin = unit(c(50, 0, 75, 0), "pt"))) / (ggeuroSelfCare365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) / (ggeuroUsual365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroPain365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) /
+  (ggeuroAnxiety365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) +
+  plot_annotation(
+    title = "ICH Laterality and HRQOL at 365 days",
+    theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))
+  )
 # plot_layout(guides = 'collect') & theme(legend.position='right', legend.direction="vertical")
@@
   d |>
-  ggplot(aes({{x}}, fill = ichLaterality)) +
+    ggplot(aes({{ x }}, fill = ichLaterality)) +
     geom_density(color = NA, alpha = 0.8) +
-  labs(title = {{ title }},
+    labs(
+      title = {{ title }},
       x = NULL,
-        y = NULL) +
+      y = NULL
+    ) +
     theme(
@@
 ggeuroVASCombined <-
-(ggeurovas90 + theme(plot.margin = unit(c(50,0,75,0), "pt"))) / (ggeurovas180 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) / (ggeurovas365 + theme(plot.margin = unit(c(0,0,75,0), "pt"))) +
-  plot_annotation(title = 'ICH Laterality EuroQOL VAS',
-                theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1)))
+  (ggeurovas90 + theme(plot.margin = unit(c(50, 0, 75, 0), "pt"))) / (ggeurovas180 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) / (ggeurovas365 + theme(plot.margin = unit(c(0, 0, 75, 0), "pt"))) +
+  plot_annotation(
+    title = "ICH Laterality EuroQOL VAS",
+    theme = theme(plot.title = element_text(face = "bold", size = 35, lineheight = 0.1))
+  )
 

### 2023-06-06 manuscript/irb.qmd
irb

--- 2023-04-19 manuscript/manuscript.qmd
@@
 ---
-title: "ICH Laterality Analysis"
+title: "Hemispheric Lateralization in ICH, Functional Disability, and Health-Related Quality of Life"
 format:
@@
     code-tools: true
-    toc: true
 ---
 
-Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue.
+**Investigators**: Nikhil Patel, Rahul Karamchandani, Andrew Asimos
+
+**Background**: Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue.
 It is a significant cause of morbidity and mortality, with high rates of disability and death.
@@
 Some studies have found that left hemisphere ICH is associated with worse functional outcomes than right hemisphere ICH, while others have found the opposite.
-Additionally, the specific functions affected by hemispheric lateralization in patients with ICH remain unclear.
 
@@
 We will discuss the potential mechanisms underlying these relationships and highlight the implications for clinical practice.
+
+**Objectives**:
+
+- Hello 1
+- Hello 2
+
+**Study Type**: Retrospective analysis of anonymized individual patient data from NINDS observational studies and randomized clinical trials
+
+**Primary outcome**:
+
+**Time period**:
+
+**Variables to select**:
+
+- Age
+- Sex
+- Race
+- Ethnicity
+- Hypertension
+- Diabetes mellitus
+- Atrial fibrillation
+- Previous stroke or ICH
+- Chronic kidney disease
+- Coronary artery disease
+- Congestive heart failure
+- ICH location
+- ICH hemisphere
+- Admission Glasgow Coma Scale
+- Admission ICH volume
+- Admission IVH volume
+- modified Rankin score at 90, 180, and 365 days
+- EuroQOL at 90, 180, and 365 days
+- Withdrawal of life-sustaining therapy
+- Neurosurgery
+
+**Statistical Analysis**:
+
+**Data Security**:
+
+**Study Implications**:

### 2023-06-14 manuscript/irb.qmd
irb submission

--- 2023-04-19 manuscript/manuscript.qmd
@@
     code-tools: true
-    toc: true
+bibliography: bibliography.bib
+csl: jama.csl
 ---
 
-Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue.
+## Protocol Version 1.0
+
+**Investigators**: Nikhil Patel, Rahul Karamchandani, Andrew Asimos
+
+**Sponsor or funding source**: none
+
+**Background, Rationale, and Context**: Intracerebral hemorrhage (ICH) is a devastating type of stroke that results from bleeding within the brain tissue.
 It is a significant cause of morbidity and mortality, with high rates of disability and death.
-Hemispheric lateralization, which refers to the specialization of the left and right hemispheres of the brain for specific functions, has been suggested to play a role in the functional outcomes of patients with ICH.
-The left hemisphere is generally associated with language, verbal memory, and logical reasoning, while the right hemisphere is associated with spatial processing, attention, and emotional regulation.
+Hemispheric lateralization, which refers to the specialization of the left and right hemispheres of the brain for specific functions, has been suggested to play a role in the functional outcomes of patients with brain injury.
+[@leggeNeglectingDifference2006; @finkDoesHemisphericLateralization2008a] The left hemisphere is generally associated with language, verbal memory, and logical reasoning, while the right hemisphere is associated with spatial processing, attention, and emotional regulation.
+[@liStrokeLateralizationLarge2021]
 
@@
 Some studies have found that left hemisphere ICH is associated with worse functional outcomes than right hemisphere ICH, while others have found the opposite.
-Additionally, the specific functions affected by hemispheric lateralization in patients with ICH remain unclear.
+[@sreekrishnanIntracerebralHemorrhageLocation2016; @satoHigherMortalityPatients2015]
 
 Understanding the relationship between hemispheric lateralization and functional outcomes in patients with ICH is essential for improving patient care and developing targeted rehabilitation strategies.
-Therefore, this paper aims to review the current literature on hemispheric lateralization and functional outcomes in patients with ICH.
-We will discuss the potential mechanisms underlying these relationships and highlight the implications for clinical practice.
+Therefore, this paper aims to pool individual patient data from several prospective studies funded by the National Institute of Neurological Disorders and Stroke (NINDS), a division of the National Institute of Health (NIH), to investigate the association of hemispheric lateralization on outcomes.
+This data is publicly available upon submitting on application to the NINDS and contains anonymized individual patient data.
+
+**Objectives**: The primary objective of this research project is to determine any association between hemispheric laterality and functional outcomes in patients with ICH.
+Secondary objectives are to investigate any association between hemispheric laterality and aggressiveness of care as well as any association between hemispheric laterality and health-related quality of life.
+
+**Study Type**: Retrospective analysis of anonymized individual patient data from NINDS observational studies and randomized clinical trials
+
+**Selection criteria**:
+
+- **Inclusion**:
+
+  - Enrolled in the following NINDS studies: ERICH, ATACH-2, iDEF, MISTIE, MISTIE-III, and CLEAR-III
+  - At least 18 years of age
+  - ICH location in the following areas: lobar, basal ganglia, or thalamus
+
+- **Exclusion**:
+
+  - ICH location in the brainstem or cerebellum
+  - Primary intraventricular hemorrhage (IVH)
+  - Midline location of hemorrhage
+
+**Time period**: 2017 - 2013
+
+**Outcomes**:
+
+- **Primary**: modified Rankin Scale at 90 days
+- **Secondary**: 
+  - Percentage of patients undergoing neurosurgical intervention
+  - Percentage of patients with early withdrawal of life-sustaining therapy
+  - Percentage of patients with tracheostomy
+  - EuroQOL Visual Analog Score (VAS) at 90 days
+  - Barthel Index at 90 days
+
+**Definitions**:
+
+- Neurosurgical intervention: any patient undergoing hematoma evacuation, external ventricular drain, or decompressive craniectomy
+- Early withdrawal of life-sustaining therapy: any patient with code status change to DNR or DNI as well as any patient transitioned to comfort-measures only within 72 hours of admission
+
+**Data collection**
+
+- **Baseline characteristics**:
+
+  - Age
+  - Gender
+  - Race
+  - Ethnicity
+  - Admission Glasgow Coma Score (GCS)
+  - ICH Volume (mL)
+  - Presence of IVH
+  - ICH Location
+  - Admission Systolic Blood Pressure
+  - Admission Diastolic Blood Pressure
+  - Comorbidities:
+    - Coronary artery disease
+    - Congestive heart failure
+    - Atrial fibrillation
+    - Chronic kidney disease
+    - Previous stroke
+    - Tobacco use
+    - Alcohol use
+    - Recreational drug use
+
+- **Inpatient care characteristics**:
+
+  - Neurosurgical intervention
+  - Early withdrawal of life-sustaining therapies
+  - Tracheostomy
+  - Gastric tube
+
+- **Outcomes**:
+
+  - modified Rankin scale at 90 days
+  - EuroQOL VAS at 90 days
+  - Barthel Index at 90 days
+
+**Analytical Plan**: Results will be analyzed initially using descriptive statistics with means and standard deviations or counts and percentages.
+Ordinal regression analysis will be utilized for the primary outcome and logistic regression for secondary outcomes.
+Other inferential statistical analysis will be conducted as appropriate.
+Alpha will be set at 0.05 to be considered statistically significant.
+
+**Human Subjects Protection**:
+
+- **Informed Consent**: Written informed consent will not be obtained. The requested data is publicly available and contains anonymized patient data that cannot be traced back to the original patient.
+
+- **Confidentiality and Privacy**: Patient data is already anonymized and cannot be traced back to the original patient.
+All terms agreed upon by signing the NINDS Data Request Agreement will be followed.
+
+- **Data and Safety Monitoring**: The principal investigator will be responsible for the overall monitoring of the data and safety of study participants.
+The principal investigator will be assisted by other members of the study staff.
+All terms agreed upon by signing the NINDS Data Request Agreement will be followed.

### 2023-06-14 manuscript/irb.tex
irb submission

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Hemispheric Lateralization in ICH, Functional Disability, and Health-Related Quality of Life},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Hemispheric Lateralization in ICH, Functional Disability, and
Health-Related Quality of Life}
\author{}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{protocol-version-1.0}{%
\subsection{Protocol Version 1.0}\label{protocol-version-1.0}}

\textbf{Investigators}: Nikhil Patel, Rahul Karamchandani, Andrew Asimos

\textbf{Sponsor or funding source}: none

\textbf{Background, Rationale, and Context}: Intracerebral hemorrhage
(ICH) is a devastating type of stroke that results from bleeding within
the brain tissue. It is a significant cause of morbidity and mortality,
with high rates of disability and death. Hemispheric lateralization,
which refers to the specialization of the left and right hemispheres of
the brain for specific functions, has been suggested to play a role in
the functional outcomes of patients with ICH. The left hemisphere is
generally associated with language, verbal memory, and logical
reasoning, while the right hemisphere is associated with spatial
processing, attention, and emotional regulation.

Several studies have investigated the relationship between hemispheric
lateralization and functional outcomes in patients with ICH, with
conflicting results. Some studies have found that left hemisphere ICH is
associated with worse functional outcomes than right hemisphere ICH,
while others have found the opposite.

Understanding the relationship between hemispheric lateralization and
functional outcomes in patients with ICH is essential for improving
patient care and developing targeted rehabilitation strategies.
Therefore, this paper aims to pool individual patient data from several
prospective studies funded by the National Institute of Neurological
Disorders and Stroke (NINDS), a division of the National Institute of
Health (NIH), to investigate the association of hemispheric
lateralization on outcomes. This data is publicly available upon
submitting on application to the NINDS and contains anonymized
individual patient data.

\textbf{Objectives}: The primary objective of this research project is
to determine any association between hemispheric laterality and
functional outcomes in patients with ICH. Secondary objectives are to
investigate any association between hemispheric laterality and
aggressiveness of care as well as any association between hemispheric
laterality and health-related quality of life.

\textbf{Study Type}: Retrospective analysis of anonymized individual
patient data from NINDS observational studies and randomized clinical
trials

\textbf{Selection criteria}:

\begin{itemize}
\item
  \textbf{Inclusion}:

  \begin{itemize}
  \tightlist
  \item
    Enrolled in the following NINDS studies: ERICH, ATACH-2, iDEF,
    MISTIE, MISTIE-III, and CLEAR-III
  \item
    At least 18 years of age
  \item
    ICH location in the following areas: lobar, basal ganglia, or
    thalamus
  \end{itemize}
\item
  \textbf{Exclusion}:

  \begin{itemize}
  \tightlist
  \item
    ICH location in the brainstem or cerebellum
  \item
    Primary intraventricular hemorrhage (IVH)
  \item
    Midline location of hemorrhage
  \end{itemize}
\end{itemize}

\textbf{Time period}: 2017 - 2013

\textbf{Outcomes}:

\begin{itemize}
\tightlist
\item
  \textbf{Primary}: modified Rankin Scale at 90 days
\item
  \textbf{Secondary}:

  \begin{itemize}
  \tightlist
  \item
    Percentage of patients undergoing neurosurgical intervention
  \item
    Percentage of patients with early withdrawal of life-sustaining
    therapy
  \item
    Percentage of patients with tracheostomy
  \item
    EuroQOL Visual Analog Score (VAS) at 90 days
  \item
    Barthel Index at 90 days
  \end{itemize}
\end{itemize}

\textbf{Definitions}:

\begin{itemize}
\tightlist
\item
  Neurosurgical intervention: any patient undergoing hematoma
  evacuation, external ventricular drain, or decompressive craniectomy
\item
  Early withdrawal of life-sustaining therapy: any patient with code
  status change to DNR or DNI as well as any patient transitioned to
  comfort-measures only within 72 hours of admission
\end{itemize}

\textbf{Data collection}

\begin{itemize}
\item
  \textbf{Baseline characteristics}:

  \begin{itemize}
  \tightlist
  \item
    Age
  \item
    Gender
  \item
    Race
  \item
    Ethnicity
  \item
    Admission Glasgow Coma Score (GCS)
  \item
    ICH Volume (mL)
  \item
    Presence of IVH
  \item
    ICH Location
  \item
    Admission Systolic Blood Pressure
  \item
    Admission Diastolic Blood Pressure
  \item
    Comorbidities:

    \begin{itemize}
    \tightlist
    \item
      Coronary artery disease
    \item
      Congestive heart failure
    \item
      Atrial fibrillation
    \item
      Chronic kidney disease
    \item
      Previous stroke
    \item
      Tobacco use
    \item
      Alcohol use
    \item
      Recreational drug use
    \end{itemize}
  \end{itemize}
\item
  \textbf{Inpatient care characteristics}:

  \begin{itemize}
  \tightlist
  \item
    Neurosurgical intervention
  \item
    Early withdrawal of life-sustaining therapies
  \item
    Tracheostomy
  \item
    Gastric tube
  \end{itemize}
\item
  \textbf{Outcomes}:

  \begin{itemize}
  \tightlist
  \item
    modified Rankin scale at 90 days
  \item
    EuroQOL VAS at 90 days
  \item
    Barthel Index at 90 days
  \end{itemize}
\end{itemize}

\textbf{Analytical Plan}: Results will be analyzed initially using
descriptive statistics with means and standard deviations or counts and
percentages. Ordinal regression analysis will be utilized for the
primary outcome and logistic regression for secondary outcomes. Other
inferential statistical analysis will be conducted as appropriate. Alpha
will be set at 0.05 to be considered statistically significant.

\textbf{Human Subjects Protection}:

\begin{itemize}
\item
  \textbf{Informed Consent}: Written informed consent will not be
  obtained. The requested data is publicly available and contains
  anonymized patient data that cannot be traced back to the original
  patient.
\item
  \textbf{Confidentiality and Privacy}: Patient data is already
  anonymized and cannot be traced back to the original patient. All
  terms agreed upon by signing the NINDS Data Request Agreement will be
  followed.
\item
  \textbf{Data and Safety Monitoring}: The principal investigator will
  be responsible for the overall monitoring of the data and safety of
  study participants. The principal investigator will be assisted by
  other members of the study staff. All terms agreed upon by signing the
  NINDS Data Request Agreement will be followed.
\end{itemize}

\end{document}

### 2023-06-28 code/dataAnalysis/02analysis.qmd
euroVAS regression

@@
 library(rlang)
+library(bayesrules)
+library(rstanarm)
+library(bayesplot)
+library(broom.mixed)
+library(tidybayes)
 # library(ggokabeito)
@@
 
-**Add Deaths to ATACH data!**
+## Model Assumptions
 
-## Model Assumptions
+### EuroQOL Visual Analog Scale
+
+Let's define functional outcome as EuroQOL VAS, an outcome measure from 0-100.
+
+```{r}
+d |>
+  ggplot(aes(x = euroVAS90, fill = ichLaterality)) +
+  geom_density(alpha = 0.5)
+
+eurovas90Model <- stan_glm(
+  euroVAS90 ~ ichLaterality,
+  data = d,
+  family = gaussian,
+  iter = 5000,
+  warmup = 2000,
+  thin = 10,
+  cores = 4,
+  seed = 1234
+)
+
+summary(eurovas90Model)
+
+tidy(eurovas90Model,
+  effects = c("fixed", "aux"),
+  conf.int = TRUE, conf.level = 0.89
+) |>
+  select(-std.error)
+
+# MCMC diagnostics
+
+mcmc_trace(eurovas90Model, size = 0.1)
+mcmc_dens_overlay(eurovas90Model)
+mcmc_acf(eurovas90Model)
+
+## Posterior predictive checks
+
+pp_check(eurovas90Model)
+
+## Posterior models
+
+head(as.data.frame(eurovas90Model))
+
+```
 

### 2023-06-28 code/01importCleaningCombining.r

@@ -75,7 +75,9 @@ renameFunc(
     F21Q07DAY = comfortCareDay,
     F21Q09DAY = dnrDay,
     F11Q02 = icuLOS,
-    F11Q01DAY = hospitalLOS
+    F11Q01DAY = hospitalLOS,
+    F21Q01 = studyEnd,
+    F21Q03DAY = studyEndDate
   )
 )
 
@@ -342,6 +344,24 @@ atachMRS <- unique(atachMRS) # removing duplicate NIHSS rows
 
 # Merge data into bigger dataset
 atachWhole <- atachMRS[atachWhole, on = .(id)]
+
+# Coding mrs 6 as mrsModified
+
+atachWhole <- atachWhole |>
+  mutate(patientDied = ifelse(studyEnd == 4, "Yes", "No")) |>
+  mutate(mrs30Modified = case_when(
+    patientDied == "No" ~ mrs30,
+    patientDied == "Yes" & studyEndDate < 30 ~ 6,
+    TRUE ~ mrs30
+  )) |>
+  mutate(mrs90Modified = case_when(
+    patientDied == "No" ~ mrs90,
+    patientDied == "Yes" & studyEndDate < 90 ~ 6,
+    TRUE ~ mrs90
+  )) |>
+  select(-c(mrs30, mrs90)) |>
+  rename(mrs30 = mrs30Modified) |>
+  rename(mrs90 = mrs90Modified)
 
 #### euroQOL-----
 
@@ -1383,4 +1403,4 @@ qsave(nindsCombined, "./data/nindsCombined.qs")
 
 ## Visualizing Data-----
 
-table(nindsCombined$study, nindsCombined$mrs180)
+table(nindsCombined$study, nindsCombined$mrs90)

### 2023-08-21 code/dataAnalysis/logistic.qmd
who knows

---
title: "Hemispheric Lateralization in ICH"
subtitle: "Logistic Regression"
author: "Nikhil Patel"
project:
  type: default
format:
  html:
    code-fold: true
    code-tools: true
    toc: true
    html-math-method: katex
    css: "styles.css"
---

```{r hemisphere only}
priors <- c(prior())

```

### 2023-09-09 code/dataAnalysis/logistic.qmd
ordinal

@@
   html:
-    code-fold: true
+    code-fold: false
     code-tools: true
@@
 
-```{r hemisphere only}
-priors <- c(prior())
+We are trying to see if there is an association between hemispheric laterality and outcomes.
+The outcomes we have are both functional disability as measured by the mRS and quality of life as measured by the EuroQOL.
+
+```{r setup, warning=FALSE, message=FALSE}
+
+# import libraries
+library(bayesrules)
+library(tidyverse)
+library(brms)
+library(cmdstanr)
+library(rstanarm)
+library(marginaleffects)
+library(broom)
+library(broom.mixed)
+library(tidybayes)
+library(ggdist)
+library(patchwork)
+library(scales)
+library(qs)
+
+# call this once to distribute chains across cpu cores:
+options(mc.cores = parallel::detectCores())
+
+# plot themes
+colors <- MetBrewer::met.brewer("Derain", 6)
+theme_set(theme_bw())
+
+# Tell bayesplot to use the viridis rocket palette for things like pp_check()
+viridisLite::viridis(6, option = "rocket", end = 0.85, direction = -1) |>
+  # Take off the trailing "FF" in the hex codes
+  map_chr(~ str_sub(., 1, 7)) |>
+  bayesplot::color_scheme_set()
+
+# import data
+d <- qread("../../data/nindsCombined.qs")
 

### 2023-09-28 code/01importCleaningCombining.r
file structure

@@ -1,1406 +1,0 @@
-# loading libraries----
-
-library(tidyverse)
-library(dtplyr)
-library(tableone)
-library(data.table)
-library(haven) # Importing SAS Files
-library(janitor)
-library(lubridate)
-library(readxl)
-library(Hmisc)
-library(stringr)
-library(qs)
-getRs("reptools.r") # Loads reptools.r from Github
-
-# loading raw data----
-
-# variables needed from each dataset-----
-# demographics: age, gender, site of enrollment
-# pmh: hypertension, diabetes, stroke, ICH, CAD, CKD
-# meds: anticoagulation, antiplatelet, lipids, bp
-# presentation features: symptom onset, SBP/DBP,GCS, NIHSS
-# presentation features: intubated
-# imaging: ICH volume, location, IVH, midline shift?
-# processes: EVD, neurosurgery, trach, PEG, WLST
-# outcomes: mRS,
-
-
-# atach-2----
-
-## import raw data----
-
-# Create function to paste together string of FORMxx.txt
-atachImportFunc <- function(y) { # nolint: object_name_linter.
-  filepath <- paste0("./data/ATACH2/FORM", y, ".txt")
-  dataset <- read.table(filepath, sep = "\t", header = TRUE) # nolint
-}
-
-## import data files that are tidy-structured
-atachImportVector <- c("00", "01", "02", "03", "04", "05", "07", "08", "11", "16", "17", "21", "33") # nolint # nolint
-
-atachList <- lapply(atachImportVector, FUN = atachImportFunc) # nolint
-
-atachWhole <- as.data.table(reduce(atachList, full_join, by = "SUBJECT_ID")) # nolint
-
-atachWhole[, study := "ATACH2"]
-
-## renaming variables----
-renameFunc <- function(x, n) setnames(x, names(n), n)
-
-renameFunc(
-  atachWhole,
-  .q(
-    SUBJECT_ID = id,
-    F33Q02 = age,
-    F00Q02MIN = symptomOnset,
-    F02Q03MIN = arrivalTimeInitial,
[…]

### 2023-09-28 code/dataAnalysis/script.r

@@ -1,23 +1,0 @@
-library(tidyverse)
-library(qs)
-library(data.table)
-library(Hmisc)
-library(qreport)
-
-getwd()
-
-d <- qread("./data/nindsCombined.qs")
-
-mrsCounts <- d |>
-  filter(!is.na(mrs90)) |>
-  group_by(ichLaterality, mrs90) |>
-  tally() |>
-  dplyr::mutate(percent = scales::percent(n / sum(n), accuracy = 1, trim = FALSE))
-
-dlist <- c(d$mrs30, d$mrs90)
-
-missChk(dlist)
-
-str(d$euroMobility90)
-
-fct_count(d$euroSelfCare365)

### 2023-09-29 analysis/clean.qmd
re-arranging

---
title: "Process and Clean Data"
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(targets)
library(scales)
library(patchwork)
library(DT)
library(naniar)
library(here)

```

*Data has been cleaned and merged elsewhere.*

```{r description of data}

data <- read_csv(here("data", "all.csv"))

```

### 2023-09-29 analysis/clean.qmd
dags

@@
 library(tidyverse)
-library(targets)
-library(scales)
 library(patchwork)
-library(DT)
-library(naniar)
 library(here)
@@
 
+# converting some columns to factors
+factors <- c("sex", "race", "euro_mobility_90", "euro_mobility_180", "euro_mobility_365", "euro_selfcare_90", "euro_selfcare_180", "euro_selfcare_365", "euro_usual_90", "euro_usual_180", "euro_usual_365", "euro_pain_90", "euro_pain_180", "euro_pain_365", "euro_anxiety_90", "euro_anxiety_180", "euro_anxiety_365", "location_90", "location_180", "location_365", "ethnicity", "stroke", "cad", "chf", "afib", "pvd", "hld", "dm2", "tobacco", "htn_meds", "dm2_meds", "ich_location", "ich_laterality", "ivh", "evd", "neurosurgery_evac", "tracheostomy", "peg", "dnr_binary", "dni_binary", "comfort_care_binary", "mrs_90", "mrs_180", "mrs_365", "euroqol_90_assessment_source", "euroqol_180_assessment_source", "euroqol_365_assessment_source", "euro_mobility_365_binary", "euro_selfcare_365_binary", "euro_usual_365_binary", "euro_pain_365_binary", "euro_anxiety_365_binary", "study", "enrolling_country", "treatment_group", "euro_mobility_90_binary", "euro_selfcare_90_binary", "euro_usual_90_binary", "euro_pain_90_binary", "euro_anxiety_90_binary", "cocaine", "anticoagulated", "etoh", "mrs_270", "location_30")
+
+data <- data |>
+  mutate_at(factors, factor)
+
+glimpse(data)
+
 ```

### 2023-09-29 analysis/dag.qmd
dags

---
title: "Causal Assumptions - DAG"
---

```{r setup}

library(tidyverse)
library(dagitty)
library(ggdag)

```

```{r dag}

dag_laterality <- dagify(ich_laterality ~ functional_outcome)

dag_laterality |>
  ggdag()

```

### 2023-09-29 analysis/dag.qmd
dags, bruh
dag updates

@@
 title: "Causal Assumptions - DAG"
+code-fold: true
 ---
 
-```{r setup}
+```{r setup, warning=FALSE, message=FALSE}
 
@@
 library(ggdag)
+library(here)
 
@@
 
+## Specify a Causal Question
+
+Do patients with **left hemispheric** injuries have worse otucomes as compared to patients with **right hemispheric** injuries?
+
+## Exploratory Data Analysis
+
+```{r eda, warning=FALSE, message=FALSE}
+#| fig-width: 10
+
+data <- read_csv(here("data", "all.csv"))
+
+data |>
+  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
+  filter(!is.na(ich_laterality)) |>
+  ggplot(aes(x = ich_laterality, y = mrs_90)) +
+  geom_bar()
+
+data |>
+  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
+  filter(!is.na(ich_laterality)) |>
+  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
+  geom_density(color = NA, alpha = .8)
+
+```
+
+## Model Our Assumptions
+
 ```{r dag}
+#| fig-width: 10
 

### 2023-09-29 analysis/dag.qmd
descriptive analysis

@@
 library(ggdag)
+library(MetBrewer)
 library(here)
@@
   filter(!is.na(ich_laterality)) |>
-  ggplot(aes(x = ich_laterality, y = mrs_90)) +
-  geom_bar()
-
-data |>
-  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
-  filter(!is.na(ich_laterality)) |>
   ggplot(aes(euro_vas_90, fill = ich_laterality)) +
@@
 
-```{r dag}
-#| fig-width: 10
+```{r dag, warning=FALSE, message=FALSE}
+#| fig-width: 12
+#| fig-height: 8
 
-dag_laterality <- dagify(ich_laterality ~ functional_outcome)
+node_details <- tribble(
+  ~name, ~label, ~x, ~y,
+  "functional_disability", "Functional Disability", 6, 1,
+  "ich_laterality", "Hemispheric Laterality", 2, 1,
+  "age", "Age", 2, 2,
+  "ich_location", "ICH Location", 2.5, 3,
+  "ich_volume", "ICH Volume", 3, 1.5,
+  "ivh", "IVH", 3, 4,
+  "gcs_baseline", "Admission GCS", 4, 2.5,
+  "wlst", "WLST", 5, 2,
+  "neurosurgery", "Neurosurgery", 5, 3,
+  "rehab", "Access to Rehab", 4, 5
+)
 
-dag_laterality |>
-  ggdag()
+node_labels <- node_details$label
+names(node_labels) <- node_details$name
+
+dag_laterality <- dagify(
+  functional_disability ~ ich_laterality + age + ich_volume + ivh + ich_location + gcs_baseline + neurosurgery + wlst + rehab,
+  neurosurgery ~ age + ich_location + ich_volume + gcs_baseline,
+  wlst ~ age + ich_volume + ich_location + gcs_baseline,
+  gcs_baseline ~ ich_location + ich_volume + ivh + age,
+  ich_location ~ age,
+  rehab ~ ich_laterality + age + ich_location,
+  exposure = "ich_laterality",
+  outcome = "functional_disability",
+  latent = "rehab",
+  coords = node_details,
+  labels = node_labels
+)
+
+dag_laterality_tidy <- dag_laterality |>
+  tidy_dagitty() |>
+  node_status()
+
+colors <- met.brewer(name = "VanGogh1", n = 7, type = "discrete")
+status_colors <- c(exposure = colors[3], outcome = colors[7], latent = "grey50")
+
+# Fancy DAG
+ggplot(dag_laterality_tidy, aes(x = x, y = y, xend = xend, yend = yend)) +
+  geom_dag_edges() +
+  geom_dag_point(aes(color = status)) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold"
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE) +
+  theme_dag()
+
+ggdag_paths(dag_laterality_tidy)
+ggdag_adjustment_set(dag_laterality_tidy)
 

### 2023-09-29 analysis/descriptive.qmd
descriptive analysis

--- 2023-09-29 analysis/clean.qmd
@@
 ---
-title: "Process and Clean Data"
+title: "Descriptive Analysis"
 ---
@@
 library(tidyverse)
-library(patchwork)
+library(dagitty)
+library(ggdag)
+library(MetBrewer)
 library(here)
-
-```
-
-*Data has been cleaned and merged elsewhere.*
-
-```{r description of data}
+library(tableone)
+library(gt)
+library(gtExtras)
+library(broom)
 
@@
 
-glimpse(data)
+data <- data |>
+  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
+  filter(!is.na(ich_laterality))
 
 ```
+
+```{r tableone}
+#| fig-width: 20
+
+# have to fix time from symptom onset to presentation in ED (units seems to be jacked up)
+
+table1_vars <- c("age", "sex", "race", "ethnicity", "htn", "dm2", "tobacco", "sbp_baseline", "gcs_baseline", "nihss_baseline", "ich_volume_baseline", "ivh", "ich_location", "neurosurgery_evac", "comfort_care_binary", "study")
+
+table1 <- CreateTableOne(vars = table1_vars, data = data, strata = "ich_laterality")
+
+table1
+
+```

### 2023-09-29 analysis/models.qmd
descriptive analysis

---
title: "Models"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(dagitty)
library(ggdag)
library(MetBrewer)
library(here)
library(brms)
library(broom)
library(tableone)

data <- read_csv(here("data", "all.csv"))

```

First let's do a Table 1

```{r interact2 analysis replication}

CreateTableOne(data = data)

```

### 2023-10-02 analysis/descriptive.qmd
more frustration

--- 2023-09-29 analysis/clean.qmd
@@
----
-title: "Process and Clean Data"
----
-
-```{r setup, warning=FALSE, message=FALSE}
+#
+#
+#
+#
+#
 
 library(tidyverse)
-library(targets)
-library(scales)
-library(patchwork)
-library(DT)
-library(naniar)
+library(dagitty)
+library(ggdag)
+library(MetBrewer)
 library(here)
-
-```
-
-*Data has been cleaned and merged elsewhere.*
-
-```{r description of data}
+library(tableone)
+library(gt)
+library(gtExtras)
+library(broom)
 
@@
 
-```
+#
+#
+#
+#
+#
+#| fig-width: 20
+
+# have to fix time from symptom onset to presentation in ED (units seems to be jacked up)
+
+table1_vars <- c("age", "sex", "race", "ethnicity", "htn", "dm2", "tobacco", "sbp_baseline", "gcs_baseline", "nihss_baseline", "ich_volume_baseline", "ivh", "ich_location", "neurosurgery_evac", "comfort_care_binary", "study")
+
+table1 <- CreateTableOne(vars = table1_vars, data = data, strata = "ich_laterality")
+
+table1
+
+#
+#
+#

### 2023-10-02 analysis/models.qmd
more frustration

@@
 library(tidyverse)
-library(dagitty)
-library(ggdag)
-library(MetBrewer)
-library(here)
+library(marginaleffects)
 library(brms)
+library(ggplot2)
+library(ggdist)
 library(broom)
-library(tableone)
 
@@
 
+data <- data |>
+  mutate(mrs_90 = fct_relevel(
+    mrs_90,
+    "0",
+    "1",
+    "2",
+    "3",
+    "4",
+    "5",
+    "6",
+  ))
+
+is.ordered(data$mrs_90)
+
 ```
 
-First let's do a Table 1
+First let's do a replication of the INTERACT re-analysis.
 
@@
 
-CreateTableOne(data = data)
+priors <- c(prior(normal(0, 1.5), class = Intercept))
+
+inits <- list(Intercept = c(-2, -1, 0, 1, 2, 2.5))
+
+mod <- brm(mrs_90 ~ 1,
+  data = data,
+  family = cumulative("logit"),
+  prior = priors,
+  init = rep(list(inits), 4),
+  chains = 4, iter = 5000 * 2, seed = 28203, cores = 4,
+  backend = "cmdstanr", refresh = 0
+)
 

### 2023-10-02 analysis/dag.qmd
ggdag updates

--- 2023-09-29 analysis/dag.qmd
@@
 library(ggdag)
+library(MetBrewer)
 library(here)
+library(broom) # For converting model output to data frames
+
+# set theme of all DAGs to `theme_dag()`
+theme_set(theme_dag(base_family = "Cantarell"))
 
@@
 
-```{r dag}
+```{r dag, warning=FALSE, message=FALSE}
 #| fig-width: 10
+#| fig-height: 7
 
-dag_laterality <- dagify(ich_laterality ~ functional_outcome)
+# original DAG: dagitty.net/mcLgZ8X, pw: CDKM93kbXs
 
-dag_laterality |>
-  ggdag()
+node_details <- tribble(
+  ~name, ~label, ~x, ~y,
+  "functional_disability", "Functional Disability", 4, 1,
+  "ich_laterality", "Hemispheric Laterality", 1, 1,
+  "age", "Age", 2.75, 3,
+  "ich_location", "ICH Location", 2, 2.75,
+  "ich_volume", "ICH Volume", 1, 2.75,
+  "ivh", "IVH", 1.5, 2.25,
+  "gcs_baseline", "Admission GCS", 2, 3.25,
+  "wlst", "Early WLST", 3.75, 2.25,
+  "neurosurgery", "Neurosurgery", 2, 1.5,
+)
+
+node_labels <- node_details$label
+names(node_labels) <- node_details$name
+
+ich_dag <- dagify(
+  functional_disability ~ ich_laterality + age + ich_location + ich_volume + ivh + gcs_baseline + wlst + neurosurgery,
+  neurosurgery ~ ich_laterality + age + ich_location + gcs_baseline,
+  wlst ~ age + ich_location + gcs_baseline + ich_volume + ivh,
+  gcs_baseline ~ age + ich_location + ich_volume + ivh,
+  ich_location ~ age,
+  ivh ~ ich_location,
+  exposure = "ich_laterality",
+  outcome = "functional_disability",
+  coords = node_details,
+  labels = node_labels
+)
+
+ich_dag_tidy <- ich_dag |>
+  tidy_dagitty() |>
+  node_status()
+
+colors <- met.brewer(name = "VanGogh1", n = 7, type = "discrete")
+status_colors <- c(exposure = colors[3], outcome = colors[6], latent = "grey50")
+
+# Fancy DAG
+ggplot(ich_dag_tidy, aes(x = x, y = y, xend = xend, yend = yend)) +
+  geom_dag_edges(aes(color = "grey50")) +
+  geom_dag_point(aes(color = status), size = 20) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold", family = "Cantarell", size = 5
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE)
+
+ggdag_adjustment_set(ich_dag_tidy) +
+  geom_dag_edges(aes(color = "grey50")) +
+  geom_dag_point(aes(color = status), size = 20) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold", family = "Cantarell", size = 5
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE)
+
+ggdag_paths(ich_dag_tidy) +
+  geom_dag_edges(aes(color = "grey50")) +
+  geom_dag_point(aes(color = status), size = 20) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold", family = "Cantarell", size = 5
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE)
+
+adjustment_vars <- c("neurosurgery", "age", "ich_volume", "ivh", "gcs_baseline", "ich_location", "wlst")
+
+control_for(ich_dag_tidy, var = adjustment_vars)
+
+ggdag_adjust(ich_dag_tidy, var = adjustment_vars) +
+  geom_dag_edges(aes(color = "grey50")) +
+  geom_dag_point(aes(color = status), size = 20) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold", family = "Cantarell", size = 5
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE)
+
+ggdag_dseparated(ich_dag_tidy, controlling_for = adjustment_vars) +
+  geom_dag_edges(aes(color = "grey50")) +
+  geom_dag_point(aes(color = status), size = 20) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold", family = "Cantarell", size = 5
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = FALSE, fill = FALSE)
 
 ```
+
+```{r}
+
+```

### 2023-10-02 analysis/descriptive.qmd
ggdag updates

@@
-#
-#
-#
-#
-#
-
-library(tidyverse)
-library(dagitty)
-library(ggdag)
-library(MetBrewer)
-library(here)
-library(tableone)
-library(gt)
-library(gtExtras)
-library(broom)
-
-data <- read_csv(here("data", "all.csv"))
-
-#
-#
-#
-#
-#
-#| fig-width: 20
-
-# have to fix time from symptom onset to presentation in ED (units seems to be jacked up)
-
-table1_vars <- c("age", "sex", "race", "ethnicity", "htn", "dm2", "tobacco", "sbp_baseline", "gcs_baseline", "nihss_baseline", "ich_volume_baseline", "ivh", "ich_location", "neurosurgery_evac", "comfort_care_binary", "study")
-
-table1 <- CreateTableOne(vars = table1_vars, data = data, strata = "ich_laterality")
-
-table1
-
-#
-#
-#
+---
+title: "Descriptive Statistics"
+code-fold: true
+---

### 2023-10-02 analysis/models.qmd
ggdag updates

--- 2023-09-29 analysis/models.qmd
@@
 library(tidyverse)
-library(dagitty)
-library(ggdag)
-library(MetBrewer)
+library(marginaleffects)
+library(brms)
+library(ggplot2)
+library(ggdist)
+library(broom)
 library(here)
-library(brms)
-library(broom)
-library(tableone)
 
@@
 
-First let's do a Table 1
+## Estimate the Causal Effect
+
+First let's do a replication of the INTERACT re-analysis.
 
@@
 
-CreateTableOne(data = data)
+priors <- c(prior(normal(0, 1.5), class = Intercept))
+
+inits <- list(Intercept = c(-2, -1, 0, 1, 2, 2.5))
+
+# mod <- brm(mrs_90 ~ 1,
+#   data = data,
+#   family = cumulative("logit"),
+#   prior = priors,
+#   init = rep(list(inits), 4),
+#   chains = 4, iter = 5000 * 2, seed = 28203, cores = 4,
+#   backend = "cmdstanr", refresh = 0
+# )
 
 ```
+
+## Conduct Sensitivity Analyses

### 2023-10-02 analysis/models.qmd
Update models.qmd

--- 2023-10-02 analysis/models.qmd
@@
 library(broom)
+library(here)
 
@@
 
-data <- data |>
-  mutate(mrs_90 = fct_relevel(
-    mrs_90,
-    "0",
-    "1",
-    "2",
-    "3",
-    "4",
-    "5",
-    "6",
-  ))
+```
 
-is.ordered(data$mrs_90)
-
-```
+## Estimate the Causal Effect
 
@@
 ```
+
+## Conduct Sensitivity Analyses

### 2023-10-03 analysis/models.qmd
bayes

@@
 library(tidyverse)
+library(ordinal)
 library(marginaleffects)
+library(modelsummary)
 library(brms)
+library(tidybayes)
+library(broom.mixed)
 library(ggplot2)
@@
 
-data <- read_csv(here("data", "all.csv"))
+data <- read_rds(here("data", "all.rds"), refhook = NULL)
+
+# plot stuff
+theme_set(theme_minimal(base_family = "Cantarell"))
+
+# convert mrs to ordered factors
+mrs_levels <- c("0", "1", "2", "3", "4", "5", "6")
+
+data <- data |>
+  mutate(
+    mrs_30 = fct_inseq(as.character(mrs_30), ordered = TRUE),
+    mrs_90 = fct_inseq(as.character(mrs_90), ordered = TRUE),
+    mrs_180 = fct_inseq(as.character(mrs_180), ordered = TRUE),
+    mrs_270 = fct_inseq(as.character(mrs_270), ordered = TRUE),
+    mrs_365 = fct_inseq(as.character(mrs_365), ordered = TRUE),
+  ) |>
+  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar")
 
@@
 
-## Estimate the Causal Effect
-
-First let's do a replication of the INTERACT re-analysis.
+## Visualization
 
@@
 
-priors <- c(prior(normal(0, 1.5), class = Intercept))
+mrs_prop <- data |>
+  count(mrs_90) |>
+  mutate(p = n / sum(n), cumsum_p = cumsum(p)) |>
+  na.omit()
 
-inits <- list(Intercept = c(-2, -1, 0, 1, 2, 2.5))
+mrs_prop |>
+  ggplot(aes(x = mrs_90, y = p)) +
+  geom_col(fill = "coral") +
+  scale_y_continuous(labels = scales::percent) +
+  labs(x = "mRS 90", y = "Proportion")
 
-mod <- brm(mrs_90 ~ 1,
-  data = data,
-  family = cumulative("logit"),
-  prior = priors,
-  init = rep(list(inits), 4),
-  chains = 4, iter = 5000 * 2, seed = 28203, cores = 4,
-  backend = "cmdstanr", refresh = 0
-)
+mrs_prop |>
+  ggplot(aes(x = as.integer(mrs_90), y = cumsum_p)) +
+  geom_point(size = 2) +
+  geom_line(linewidth = 1) +
+  labs(x = "mRS 90", y = "Cumulative Proportion")
+
+mrs_prop |>
+  ggplot(aes(x = as.integer(mrs_90), y = log(cumsum_p) - log(1 - cumsum_p))) +
+  geom_point(size = 2) +
+  geom_line(linewidth = 1) +
+  labs(x = "mRS 90", y = "Logit Cumulative Proportion")
 
@@
 
-## Conduct Sensitivity Analyses
+## Frequentist Analysis
+
+### Mediation Analysis with Neurosurgery and ICH Laterality
+
+```{r frequentist ordinal}
+
+mrs_90_ich_lat <- clm(
+  mrs_90 ~ ich_laterality,
+  data = data,
+  link = "logit"
+)
+
+neurosurg_ich_lat <- clm(
+  neurosurgery_evac ~ ich_laterality,
+  data = data,
+  link = "logit"
+)
+
+mrs_90_ich_lat_neurosurg <- clm(
+  mrs_90 ~ ich_laterality + neurosurgery_evac,
+  data = data,
+  link = "logit"
+)
+
+summary(mrs_90_ich_lat)
+summary(neurosurg_ich_lat)
+summary(mrs_90_ich_lat_neurosurg)
+
+```
+
+## Bayesian Analysis
+
+```{r bayesian ordinal}
+
+fit1 <- brm(
+  data = data,
+  family = cumulative(probit),
+  mrs_90 ~ 1,
+)
+
+fit2 <- brm(
+  data = data,
+  family = cumulative(logit),
+  mrs_90 ~ 1,
+)
+
+fit3 <- brm(
+  data = data,
+  family = cumulative(logit),
+  mrs_90 ~ ich_laterality,
+)
+
+print(fit1)
+print(fit2)
+print(fit3)
+
+pred <- predictions(fit3)
+
+```

### 2023-10-03 analysis/models.qmd
finally ran the model

@@
 library(brms)
+library(cmdstanr)
 library(tidybayes)
@@
 library(broom)
+library(gt)
 library(here)
+
+options(mc.cores = parallel::detectCores())
 
@@
   filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar")
+
+data$ich_location <- fct_drop(data$ich_location)
+table(data$ich_location)
 
@@
 
-mrs_90_ich_lat <- clm(
-  mrs_90 ~ ich_laterality,
+fit_freq <- clm(
+  mrs_90 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
   data = data,
@@
 
-neurosurg_ich_lat <- clm(
-  neurosurgery_evac ~ ich_laterality,
-  data = data,
-  link = "logit"
-)
+summary(fit_freq)
 
-mrs_90_ich_lat_neurosurg <- clm(
-  mrs_90 ~ ich_laterality + neurosurgery_evac,
-  data = data,
-  link = "logit"
-)
+predictions(fit_freq, by = "ich_laterality", type = "cum.prob") |>
+  gt()
 
-summary(mrs_90_ich_lat)
-summary(neurosurg_ich_lat)
-summary(mrs_90_ich_lat_neurosurg)
+predictions(fit_freq, by = c("ich_location", "ich_laterality"), type = "cum.prob") |>
+  gt()
+
+plot_predictions(fit_freq, condition = c("ich_location", "ich_laterality"), type = "cum.prob")
+
+tidy(fit_freq)
 
@@
 
-```{r bayesian ordinal}
+```{r bayesian ordinal, warning=FALSE, message=FALSE}
 
-fit1 <- brm(
+fit_brms <- brm(
+  mrs_90 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
   data = data,
-  family = cumulative(probit),
-  mrs_90 ~ 1,
+  family = cumulative(logit),
+  prior = set_prior("normal(0,5)"),
+  chains = 2,
+  cores = 6,
+  backend = "cmdstanr"
 )
 
-fit2 <- brm(
-  data = data,
-  family = cumulative(logit),
-  mrs_90 ~ 1,
-)
+print(fit_brms)
 
-fit3 <- brm(
-  data = data,
-  family = cumulative(logit),
-  mrs_90 ~ ich_laterality,
-)
+tidy(fit_brms) |>
+  select(c(estimate, conf.low, conf.high))
 
-print(fit1)
-print(fit2)
-print(fit3)
+predictions(fit_brms, by = c("ich_location", "ich_laterality"), type = "response") |>
+  gt()
 
-pred <- predictions(fit3)
+plot_predictions(fit_brms, condition = c("ich_location", "ich_laterality"), type = "response")
+
+plot_predictions(fit_brms, condition = c("ich_location", "ich_laterality"), type = "prediction")
 

### 2023-10-03 analysis/logistic.qmd
logistic

---
title: "Logistic Models"
code-fold: true
---

These analyses are intended to build up intuition on using Bayesian logistic regression as a stepping stone to understanding and using Bayesian ordinal regression. We will work with dichotimized variables of the modified Rankin score as well as the EuroQOL.

```{r setup}
#| code-fold: true
#| code-summary: "Setup"
#| warning: false
#| message: false

library(bayesrules)
library(tidyverse)
library(janitor)
library(brms)
library(cmdstanr)
library(rstanarm)
library(marginaleffects)
library(broom)
library(broom.mixed)
library(parameters)
library(tidybayes)
library(ggdist)
library(patchwork)
library(latex2exp)
library(scales)
library(here)

# plot stuff
colors <- MetBrewer::met.brewer("Cross", 6)
theme_set(theme_bw(base_family = "Cantarell"))
bayesplot::color_scheme_set(colors)

# seeds
set.seed(160)
BAYES_SEED <- 160

# data
data <- read_rds(here("data", "all.rds"), refhook = NULL)

data <- data |>
  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
  filter(study == "ERICH" | study == "ATACH2") |>
  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04) |>
  drop_na(ich_laterality)

data$ich_location <- fct_drop(data$ich_location)

```

## General Setup

We are interested in modeling the probability of a "good" outcome for patients with ICH at 90 days as defined by the **modified Rankin Scale**. For this session, we will work with a dichotimized version of the mRS at various cutpoints (0-1 vs 2-6, 0-2 vs 3-6, 0-3 vs 4-6, 0-45 vs 5-6). All of these cutpoints have been used in various randomized trials and observational studies.

We will base our model on these predictors:

- `age`
- `ich_laterality`: the main "predictor" of interest for the main analysis
- `ich_location`: only supratentorial locations are used (lobar, thalamus, basal ganglia)
- `ich_volume_baseline`
- `ivh`
- `neurosurgery_evac`

### Modeling poor outcome given age and ICH Volume

We will use continuous variables first before jumping into categorical predictors.

```{r}
#| fig-width: 10

data |>
  drop_na(mrs_90_03) |>
  ggplot(aes(x = mrs_90_03)) +
  geom_density(fill = colors[3], alpha = 0.2) +
  facet_wrap(~ich_location)

data |>
  drop_na(mrs_90_02) |>
  tabyl(ich_location, mrs_90_02) |>
  adorn_percentages()

```

### 2023-10-03 analysis/models.qmd
logistic

@@
 
-# plot stuff
-theme_set(theme_minimal(base_family = "Cantarell"))
-
-# convert mrs to ordered factors
-mrs_levels <- c("0", "1", "2", "3", "4", "5", "6")
-
 data <- data |>
-  mutate(
-    mrs_30 = fct_inseq(as.character(mrs_30), ordered = TRUE),
-    mrs_90 = fct_inseq(as.character(mrs_90), ordered = TRUE),
-    mrs_180 = fct_inseq(as.character(mrs_180), ordered = TRUE),
-    mrs_270 = fct_inseq(as.character(mrs_270), ordered = TRUE),
-    mrs_365 = fct_inseq(as.character(mrs_365), ordered = TRUE),
-  ) |>
   filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar")
@@
 table(data$ich_location)
+
+# plot stuff
+theme_set(theme_minimal(base_family = "Cantarell"))
 
@@
 
-tidy(fit_brms) |>
-  select(c(estimate, conf.low, conf.high))
-
 predictions(fit_brms, by = c("ich_location", "ich_laterality"), type = "response") |>
@@
 
-plot_predictions(fit_brms, condition = c("ich_location", "ich_laterality"), type = "prediction")
-
 ```

### 2023-10-04 analysis/aggressive_care.qmd
blah

---
title: "Aggressive Care and Hemispheric Laterality"
---

```{r setup}
#| code-fold: true
#| code-summary: "Setup"
#| warning: false
#| message: false

library(tidyverse) # ggplot, dplyr, and friends
library(janitor)
library(glue) # Easier string interpolation
library(gt) # Fancy tables
library(scales) # Nicer labeling functions
library(patchwork) # Combine plots nicely
library(ggmosaic) # Mosaic plots with ggplot
library(parameters) # Extract model parameters as data frames
library(cmdstanr) # Run Stan code from R
library(brms) # Nice frontend for Stan
library(tidybayes) # Manipulate MCMC chains in a tidy way
library(here)

# Set some global Stan options
CHAINS <- 4
ITER <- 2000
WARMUP <- 1000
BAYES_SEED <- 160
options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")

# ggplot theming
theme_nice <- function() {
  theme_minimal(base_family = "Cantarell") +
    theme(
      panel.grid.minor = element_blank(),
      plot.title = element_text(family = "Cantarell", face = "bold"),
      axis.title = element_text(family = "Cantarell Medium"),
      axis.title.x = element_text(hjust = 0),
      axis.title.y = element_text(hjust = 1),
      strip.text = element_text(
        family = "Cantarell", face = "bold",
        size = rel(1), hjust = 0
      ),
      strip.background = element_rect(fill = NA, color = NA)
    )
}

```

I am interested in seeing if hemispheric laterality affects the aggressiveness of care that a patient receives. This mostly includes neurosurgery, but also other processes such as days of mechanical ventilation, EVD placement, rates of tracheostomy/PEG, and if patients were transitioned to comfort care.

I will be using observational data from the ERICH study, as well as the randomized trial data from ATACH-2 to answer these questions.

## Exploratory Data Analysis

Loading the data from ATACH-2 and ERICH, and filtering out the variables needed for the analysis.

```{r loading data}

# load data
data <- read_rds(here("data", "all.rds"), refhook = NULL)

# filter out only ERICH and ATACH2 and keep relevant variables
cleaned_data <- data |>
  mutate(id = 1:n()) |>
  filter(study == "ERICH" | study == "ATACH-2") |>
  select(id, study, ich_laterality, age, sex, race, ethnicity, sbp_baseline, gcs_baseline, ich_volume_baseline, ivh, ich_location, neurosurgery_evac, evd, tracheostomy, peg, comfort_care_binary, comfort_care_day) |>
  drop_na(ich_laterality, neurosurgery_evac)

```

Now we can summarize the preliminary data.

```{r exploratory data analysis}

counts <- cleaned_data |>
  tabyl(ich_laterality, neurosurgery_evac, ich_location) |>
  adorn_percentages("row") |>
  adorn_pct_formatting(rounding = "half up", digits = 1)

counts

counts_tidy <- cleaned_data |>
  group_by(ich_laterality, neurosurgery_evac) |>
  summarize(n = n()) |>
  mutate(
    total = sum(n),
    prop = n / sum(n)
  ) |>
  ungroup()

counts_tidy

ggplot(cleaned_data) +
  geom_mosaic(
    aes(
      x = product(neurosurgery_evac, ich_laterality),
      fill = ich_laterality, alpha = neurosurgery_evac
    ),
    divider = mosaic("v")
  )

```

## The Question

**Do patients with left hemispheric ICH receive neurosurgical evacuations less frequently as compared to patients with right hemispheric ICH?**

We want to determine if there's a substantial difference in the proportion of patients who receive neurosurgical intervention in the right or left hemisphere.

```{r color setup}

# Left/Right hemisphere question colors
clrs_saguaro <- NatParksPalettes::natparks.pals("Saguaro")
clr_left <- clrs_saguaro[1]
clr_right <- clrs_saguaro[6]
clr_diff <- clrs_saguaro[4]

```

## Frequentist Way

```{r frequentist}

matrix <- counts_tidy |>
  filter(neurosurgery_evac == "Yes") |>
  mutate(n_no = total - n) |>
  select(n_yes = n, n_no) |>
  as.matrix()

prop.test(matrix) |>
  model_parameters()

```

```{r binomial}

ggplot() +
  stat_function(fun = ~ dbeta(., 2, 6), geom = "area", fill = "#AC3414") +
  scale_x_continuous(labels = label_percent()) +
  labs(x = "Possible values for π", y = NULL, fill = NULL) +
  theme_nice() +
  theme(axis.text.y = element_blank())

```

### 2023-10-06 analysis/posterior.qmd
making progress

---
title: "Visualizing Posterior Distributions and Predictions"
code-overflow: wrap
---

This is an attempt to teach myself some concepts about the visualizing the posterior distribution in Bayesian models by re-creating Andrew Heiss' blog post [about this](https://doi.org/10.59350/xge39-emt86).

```{r setup}
#| code-fold: true
#| code-summary: "Setup"
#| warning: false
#| message: false

library(tidyverse) # ggplot, dplyr, and friends
library(patchwork) # Combine ggplot plots
library(ggtext) # Fancier text in ggplot plots
library(scales) # Labeling functions
library(brms) # Bayesian modeling through Stan
library(tidybayes) # Manipulate Stan objects in a tidy way
library(marginaleffects) # Calculate marginal effects
library(modelr) # For quick model grids
library(extraDistr) # For dprop() beta distribution with mu/phi
library(distributional) # For plotting distributions with ggdist
library(palmerpenguins) # Penguins!
library(kableExtra) # For nicer tables
library(here)

# plot stuff
colors <- MetBrewer::met.brewer("Cross", 6)
theme_set(theme_minimal(base_family = "Cantarell"))
bayesplot::color_scheme_set(colors)

# seeds
set.seed(160)
BAYES_SEED <- 160

# Bayes
options(
  mc.cores = parallel::detectCores(),
  brms.backend = "cmdstanr"
)

# Add a couple new variables to the penguins data:
#  - is_gentoo: Indicator for whether or not the penguin is a Gentoo
#  - bill_ratio: The ratio of a penguin's bill depth (height) to its bill length
penguins <- penguins |>
  drop_na(sex) |>
  mutate(is_gentoo = species == "Gentoo") |>
  mutate(bill_ratio = bill_depth_mm / bill_length_mm)

# data
data <- read_rds(here("data", "all.rds"), refhook = NULL)

data <- data |>
  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
  filter(study == "ERICH" | study == "ATACH2") |>
  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04) |>
  drop_na(ich_laterality)

data$ich_location <- fct_drop(data$ich_location)

```

## Normal Gaussian model

### Palmer Penguins Example

Andrew first uses the Palmer Penguins dataset to examine the relationship between penguin flipper length and body mass.

```{r eda penguins, warning=FALSE, message=FALSE}

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(size = 1, alpha = 0.7) +
  geom_smooth(method = "lm", color = colors[5], se = FALSE) +
  scale_y_continuous(labels = label_comma()) +
  coord_cartesian(ylim = c(2000, 6000)) +
  labs(x = "Flipper length (mm)", y = "Body mass (g)")

```

This can be written more formally as:

$$
\begin{aligned}
\text{Body mass}_i &\sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta \ \text{Flipper length}_i
\end{aligned}
$$

Or more generally:

$$
\begin{aligned}
y_i &\sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta x_i
\end{aligned}
$$

This states that body mass is normally distributed with an average that is conditional on flipper length.

```{r brms gaussian penguins, output=FALSE, warning=FALSE, message=FALSE}

model_normal_penguins <- brm(
  bf(body_mass_g ~ flipper_length_mm),
  family = gaussian(),
  data = penguins
)

```

```{r brms gaussian penguins results}

broom.mixed::tidy(model_normal_penguins, conf.level = 0.89) |>
  bind_cols(parameter = c("α", "β", "σ")) |>
  select(parameter, term, estimate, std.error, conf.low, conf.high)

```

The intercept (α) is huge and negative because flipper length is far away from 0, so it’s pretty uninterpretable. We could fix that by centering the data, so α would indicate the body mass for an average flipper length.

The coefficient β shows that a one-mm increase in flipper length is associated with a 50 gram increase in body mass. And the overall model standard deviation σ shows that there’s roughly 400 grams of deviation around the mean body mass.

These are all the posterior means but remember that we have complete distributions for these parameters.

However, here we want to examine the posterior distribution of body mass, which is the outcome we are ultimately interested in. Remember this is represented by "mu" and the overall posterior distribution of "y". Both of these represent body mass, but they are really different things. Mu represents the mean body mass, and y represents the total distribution including the standard deviation.

There are three different `brms` functions that we can use, listed below along with their corresponding `tidybayes` wrappers:

- `posterior_predict()`: `predicted_draws`
- `posterior_epred()`: `epred_draws`
- `posterior_linpred()` `linpred_draws`

We have to feed the functions `newdata`, which is a data frame of values to make the predictions. If we feed the original dataset, it would plug each row of the data into the model and generate a bunch of posterior draws (4000). This would result in 333 * 4000 rows of data, which isn't super useful to visualize.

We can first look at the predictions based on a single value of flipper lengths. Andrew uses the average, which is 200.967 mm. You can use the median or whatever number you want.

```{r penguins single value prediction}

# Make a little dataset of just the average flipper length
penguins_avg_flipper <- penguins |>
  summarize(flipper_length_mm = mean(flipper_length_mm))

# Extract different types of posteriors
normal_linpred <- model_normal_penguins |>
  linpred_draws(newdata = penguins_avg_flipper)

normal_epred <- model_normal_penguins |>
  epred_draws(newdata = penguins_avg_flipper)

normal_predicted <- model_normal_penguins |>
  predicted_draws(
    newdata = penguins_avg_flipper,
    seed = 12345
  ) # So that the manual results with rnorm() are the same later

```

Now we can look at each distribution's mean, median, standard deviation, and shapes.

```{r penguins posterior predictions}

summary_normal_linpred <- normal_linpred |>
  ungroup() |>
  summarize(across(.linpred, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_epred <- normal_epred |>
  ungroup() |>
  summarize(across(.epred, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_predicted <- normal_predicted |>
  ungroup() |>
  summarize(across(.prediction, lst(mean, sd, median), .names = "{.fn}"))

tribble(
  ~Function, ~`Model element`,
  "<code>posterior_linpred()</code>", "\\(\\mu\\) in the model",
  "<code>posterior_epred()</code>", "\\(\\operatorname{E(y)}\\) and \\(\\mu\\) in the model",
  "<code>posterior_predict()</code>", "Random draws from posterior \\(\\operatorname{Normal}(\\mu_i, \\sigma)\\)"
) |>
  bind_cols(bind_rows(summary_normal_linpred, summary_normal_epred, summary_normal_predicted)) |>
  kbl(escape = FALSE) |>
  kable_styling()

p1 <- ggplot(normal_linpred, aes(x = .linpred)) +
  stat_halfeye(fill = colors[3]) +
  scale_x_continuous(labels = label_comma()) +
  coord_cartesian(xlim = c(4100, 4300)) +
  labs(
    x = "Body mass (g)", y = NULL,
    title = "**Linear predictor** <span style='font-size: 14px;'>*µ* in the model</span>",
    subtitle = "posterior_linpred(..., tibble(flipper_length_mm = 201))"
  ) +
  theme(plot.title = element_markdown())

p2 <- ggplot(normal_epred, aes(x = .epred)) +
  stat_halfeye(fill = colors[2]) +
  scale_x_continuous(labels = label_comma()) +
  coord_cartesian(xlim = c(4100, 4300)) +
  labs(
    x = "Body mass (g)", y = NULL,
    title = "**Expectation of the posterior** <span style='font-size: 14px;'>E[*y*] and *µ* in the model</span>",
    subtitle = "posterior_epred(..., tibble(flipper_length_mm = 201))"
  )

p3 <- ggplot(normal_predicted, aes(x = .prediction)) +
  stat_halfeye(fill = colors[1]) +
  scale_x_continuous(labels = label_comma()) +
  coord_cartesian(xlim = c(2900, 5500)) +
  labs(
    x = "Body mass (g)", y = NULL,
    title = "**Posterior predictions** <span style='font-size: 14px;'>Random draws from posterior Normal(*µ*, *σ*)</span>",
    subtitle = "posterior_predict(..., tibble(flipper_length_mm = 201))"
  )

(p1 / plot_spacer() / p2 / plot_spacer() / p3) +
  plot_layout(heights = c(0.3, 0.05, 0.3, 0.05, 0.3))

```

The range of predictions is the biggest difference among these. The `posterior_linpred()` and `posterior_epred()` have small spreads (standard errors) while `posterior_predict()` has a larger spread.

This is due to `posterior_linpred()` and `posterior_epred()` corresponding to the "mu" part of the model and not incorporating any standard deviation.

We can re-recreate `posterior_linpred()` by remembering that this is the average body mass as predicted by the model, which is B0 + (B1 * `flipper_length`). So if we plug-in 201 into this, we get the same thing as `posterior_linpred()`.

The results of `posterior_predict()` incorporate the standard deviation as well as the mean, thus resulting in more uncertainty.

For `posterior_epred()`, the results in the **Gaussian regression** are the same as that of the linear predictor `posterior_linpred()`.

We can also look at these different types of posterior predictions across a range of possible flipper lengths. There’s a lot more uncertainty in the full posterior, since it incorporates the uncertainty of both and , while the uncertainty of the linear predictor/expected value of the posterior is much more narrow (and equivalent in this case):

```{r penguins full posterior}

p1 <- penguins |>
  data_grid(flipper_length_mm = seq_range(flipper_length_mm, n = 100)) |>
  add_linpred_draws(model_normal_penguins, ndraws = 100) |>
  ggplot(aes(x = flipper_length_mm)) +
  stat_lineribbon(aes(y = .linpred),
    .width = 0.95,
    alpha = 0.5, color = colors[3], fill = colors[3]
  ) +
  geom_point(data = penguins, aes(y = body_mass_g), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Flipper length (mm)", y = "Body mass (g)",
    title = "**Linear predictor** <span style='font-size: 14px;'>*µ* in the model</span>",
    subtitle = "posterior_linpred()"
  )

p2 <- penguins |>
  data_grid(flipper_length_mm = seq_range(flipper_length_mm, n = 100)) |>
  add_epred_draws(model_normal_penguins, ndraws = 100) |>
  ggplot(aes(x = flipper_length_mm)) +
  stat_lineribbon(aes(y = .epred),
    .width = 0.95,
    alpha = 0.5, color = colors[2], fill = colors[2]
  ) +
  geom_point(data = penguins, aes(y = body_mass_g), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Flipper length (mm)", y = "Body mass (g)",
    title = "**Expectation of the posterior** <span style='font-size: 14px;'>E[*y*] and *µ* in the model</span>",
    subtitle = "posterior_epred()"
  )

p3 <- penguins |>
  data_grid(flipper_length_mm = seq_range(flipper_length_mm, n = 100)) |>
  add_predicted_draws(model_normal_penguins, ndraws = 100) |>
  ggplot(aes(x = flipper_length_mm)) +
  stat_lineribbon(aes(y = .prediction),
    .width = 0.95,
    alpha = 0.5, color = colors[1], fill = colors[1]
  ) +
  geom_point(data = penguins, aes(y = body_mass_g), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Flipper length (mm)", y = "Body mass (g)",
    title = "**Posterior predictions** <span style='font-size: 14px;'>Random draws from posterior Normal(*µ*, *σ*)</span>",
    subtitle = "posterior_predict()"
  )

(p1 / plot_spacer() / p2 / plot_spacer() / p3) +
  plot_layout(heights = c(0.3, 0.05, 0.3, 0.05, 0.3))

```

### ICH Example

Here we will examine the relationship between age and ICH volume on admission. It is not as strong of a relationship, but we're not trying to publish anything here so it'll serve just fine.

```{r eda ich, warning=FALSE, message=FALSE}

ggplot(data, aes(x = age, y = ich_volume_baseline)) +
  geom_point(size = 1, alpha = 0.7) +
  geom_smooth(method = "lm", color = colors[5], se = FALSE) +
  scale_y_continuous(labels = label_comma()) +
  # coord_cartesian(ylim = c(2000, 6000)) +
  labs(x = "Age (years)", y = "ICH Volume (mL)")

```

Again, his can be written more formally as:

$$
\begin{aligned}
\text{ICH Volume}_i &\sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta \ \text{Age}_i
\end{aligned}
$$

Or more generally:

$$
\begin{aligned}
y_i &\sim \operatorname{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta x_i
\end{aligned}
$$

This states that ICH Volume is normally distributed with an average that is conditional on age.

```{r brms gaussian ich, output=FALSE, warning=FALSE, message=FALSE}

model_normal_ich <- brm(
  ich_volume_baseline ~ age,
  family = gaussian(),
  data = data
)

```

```{r brms gaussian ich results}

broom.mixed::tidy(model_normal_ich, conf.level = 0.89) |>
  bind_cols(parameter = c("α", "β", "σ")) |>
  select(parameter, term, estimate, std.error, conf.low, conf.high)

```

The intercept (α) is not very big and because it's impossible for age to be 0, it’s pretty uninterpretable. We could fix that by centering the data, so α would indicate the ICH Volume for an average age.

The coefficient β shows that a one year increase in age is associated with a 0.2 mL increase in ICH Volume. And the overall model standard deviation σ shows that there’s roughly 27 mL of deviation around the mean ICH volume.

Now we will feed the average age and examine the distributions of ICH Volume.

```{r ich single value prediction}

# Make a little dataset of just the average flipper length
ich_avg_age <- data |>
  summarize(age = mean(age))

# Extract different types of posteriors
normal_linpred <- model_normal_ich |>
  linpred_draws(newdata = ich_avg_age)

normal_epred <- model_normal_ich |>
  epred_draws(newdata = ich_avg_age)

normal_predicted <- model_normal_ich |>
  predicted_draws(
    newdata = ich_avg_age,
    seed = 12345
  ) # So that the manual results with rnorm() are the same later

```

Now we can look at each distribution's mean, median, standard deviation, and shapes.

```{r ich posterior predictions}

summary_normal_linpred <- normal_linpred |>
  ungroup() |>
  summarize(across(.linpred, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_epred <- normal_epred |>
  ungroup() |>
  summarize(across(.epred, lst(mean, sd, median), .names = "{.fn}"))

summary_normal_predicted <- normal_predicted |>
  ungroup() |>
  summarize(across(.prediction, lst(mean, sd, median), .names = "{.fn}"))

tribble(
  ~Function, ~`Model element`,
  "<code>posterior_linpred()</code>", "\\(\\mu\\) in the model",
  "<code>posterior_epred()</code>", "\\(\\operatorname{E(y)}\\) and \\(\\mu\\) in the model",
  "<code>posterior_predict()</code>", "Random draws from posterior \\(\\operatorname{Normal}(\\mu_i, \\sigma)\\)"
) |>
  bind_cols(bind_rows(summary_normal_linpred, summary_normal_epred, summary_normal_predicted)) |>
  kbl(escape = FALSE) |>
  kable_styling()

p1 <- ggplot(normal_linpred, aes(x = .linpred)) +
  stat_halfeye(fill = colors[3]) +
  scale_x_continuous(labels = label_comma()) +
  # coord_cartesian(xlim = c(4100, 4300)) +
  labs(
    x = "ICH Volume (mL)", y = NULL,
    title = "**Linear predictor** <span style='font-size: 14px;'>*µ* in the model</span>",
    subtitle = "posterior_linpred(..., tibble(flipper_length_mm = 201))"
  ) +
  theme(plot.title = element_markdown())

p2 <- ggplot(normal_epred, aes(x = .epred)) +
  stat_halfeye(fill = colors[2]) +
  scale_x_continuous(labels = label_comma()) +
  # coord_cartesian(xlim = c(4100, 4300)) +
  labs(
    x = "ICH Volume (mL)", y = NULL,
    title = "**Expectation of the posterior** <span style='font-size: 14px;'>E[*y*] and *µ* in the model</span>",
    subtitle = "posterior_epred(..., tibble(flipper_length_mm = 201))"
  )

p3 <- ggplot(normal_predicted, aes(x = .prediction)) +
  stat_halfeye(fill = colors[1]) +
  scale_x_continuous(labels = label_comma()) +
  # coord_cartesian(xlim = c(2900, 5500)) +
  labs(
    x = "ICH Volume (mL)", y = NULL,
    title = "**Posterior predictions** <span style='font-size: 14px;'>Random draws from posterior Normal(*µ*, *σ*)</span>",
    subtitle = "posterior_predict(..., tibble(flipper_length_mm = 201))"
  )

(p1 / plot_spacer() / p2 / plot_spacer() / p3) +
  plot_layout(heights = c(0.3, 0.05, 0.3, 0.05, 0.3))

```

Now here is the full posterior:

```{r ich full posterior}

p1 <- data |>
  data_grid(age = seq_range(age, n = 100)) |>
  add_linpred_draws(model_normal_ich, ndraws = 100) |>
  ggplot(aes(x = age)) +
  stat_lineribbon(aes(y = .linpred),
    .width = 0.95,
    alpha = 0.5, color = colors[3], fill = colors[3]
  ) +
  geom_point(data = data, aes(y = ich_volume_baseline), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  # coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Age (years)", y = "ICH Volume (mL)",
    title = "**Linear predictor** <span style='font-size: 14px;'>*µ* in the model</span>",
    subtitle = "posterior_linpred()"
  )

p2 <- data |>
  data_grid(age = seq_range(age, n = 100)) |>
  add_epred_draws(model_normal_ich, ndraws = 100) |>
  ggplot(aes(x = age)) +
  stat_lineribbon(aes(y = .epred),
    .width = 0.95,
    alpha = 0.5, color = colors[2], fill = colors[2]
  ) +
  geom_point(data = data, aes(y = ich_volume_baseline), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  # coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Age (years)", y = "ICH Volume (mL)",
    title = "**Expectation of the posterior** <span style='font-size: 14px;'>E[*y*] and *µ* in the model</span>",
    subtitle = "posterior_epred()"
  )

p3 <- data |>
  data_grid(age = seq_range(age, n = 100)) |>
  add_predicted_draws(model_normal_ich, ndraws = 100) |>
  ggplot(aes(x = age)) +
  stat_lineribbon(aes(y = .prediction),
    .width = 0.95,
    alpha = 0.5, color = colors[1], fill = colors[1]
  ) +
  geom_point(data = data, aes(y = ich_volume_baseline), size = 1, alpha = 0.7) +
  scale_y_continuous(labels = label_comma()) +
  # coord_cartesian(ylim = c(2000, 6000)) +
  labs(
    x = "Age (years)", y = "ICH Volume (mL)",
    title = "**Posterior predictions** <span style='font-size: 14px;'>Random draws from posterior Normal(*µ*, *σ*)</span>",
    subtitle = "posterior_predict()"
  )

(p1 / plot_spacer() / p2 / plot_spacer() / p3) +
  plot_layout(heights = c(0.3, 0.05, 0.3, 0.05, 0.3))

```

If we had set better priors, we would not be left with ranges of possible ICH volumes that are negative, but that is for another post.

## Logistic Regression Example

```{r logistic penguins eda}

ggplot(penguins, aes(x = bill_length_mm, y = as.numeric(is_gentoo))) +
  geom_dots(aes(side = ifelse(is_gentoo, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.2
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    color = colors[5], se = FALSE
  ) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Bill length (mm)", y = "Probability of being a Gentoo")

```

```{r logistic ich eda}

ggplot(data, aes(x = ich_volume_baseline, y = as.numeric(mrs_90_04))) +
  geom_dots(aes(side = ifelse(mrs_90_04, "bottom", "top")),
    pch = 19, color = "grey20", scale = 0.2
  ) +
  geom_smooth(
    method = "glm", method.args = list(family = binomial(link = "logit")),
    color = colors[5], se = FALSE
  ) +
  scale_y_continuous(labels = label_percent()) +
  labs(x = "Bill length (mm)", y = "Probability of being a Gentoo")

```

### 2023-10-07 analysis/models.qmd
more progress

--- 2023-10-03 analysis/models.qmd
@@
 library(modelsummary)
+library(parameters)
 library(brms)
+library(cmdstanr)
 library(tidybayes)
@@
 library(broom)
+library(janitor)
+library(gt)
 library(here)
 
+# seeds
+set.seed(160)
+BAYES_SEED <- 160
+
+# Bayes
+options(
+  mc.cores = parallel::detectCores(),
+  brms.backend = "cmdstanr"
+)
+
 data <- read_rds(here("data", "all.rds"), refhook = NULL)
+
+data <- data |>
+  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
+  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90) |>
+  drop_na(ich_laterality)
+
+data$ich_location <- fct_drop(data$ich_location)
+data$ich_laterality <- fct_drop(data$ich_laterality)
+
+data_erich_atach <- data |>
+  filter(study == "ERICH" | study == "ATACH2")
 
@@
 
-# convert mrs to ordered factors
-mrs_levels <- c("0", "1", "2", "3", "4", "5", "6")
+```
 
-data <- data |>
-  mutate(
-    mrs_30 = fct_inseq(as.character(mrs_30), ordered = TRUE),
-    mrs_90 = fct_inseq(as.character(mrs_90), ordered = TRUE),
-    mrs_180 = fct_inseq(as.character(mrs_180), ordered = TRUE),
-    mrs_270 = fct_inseq(as.character(mrs_270), ordered = TRUE),
-    mrs_365 = fct_inseq(as.character(mrs_365), ordered = TRUE),
-  ) |>
-  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar")
+## Dichotimized mRS: 0-3 vs. 4-6
+
+### Priors
+
+Based on previous knowledge, we can say that on average, there is a 60% chance of a good outcome of mRS 0-3.
+On the log-odds scale, that would be `qlogis(0.6) = 0.41`.
+That can range a lot from 20% to 80%, so in log-odds we can say that is `qlogis(0.2) = -1.38` as the variance.
+
+```{r priors log_odds}
 
@@
 
-## Visualization
+```{r eda}
 
-```{r interact2 analysis replication}
+model_mrs03 <- brm(
+  mrs_90_03 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
+  family = bernoulli(link = "logit"),
+  data = data
+)
 
-mrs_prop <- data |>
-  count(mrs_90) |>
-  mutate(p = n / sum(n), cumsum_p = cumsum(p)) |>
-  na.omit()
+print(model_mrs03)
 
-mrs_prop |>
-  ggplot(aes(x = mrs_90, y = p)) +
-  geom_col(fill = "coral") +
-  scale_y_continuous(labels = scales::percent) +
-  labs(x = "mRS 90", y = "Proportion")
+pp_check(model_mrs03, ndraws = 500)
 
-mrs_prop |>
-  ggplot(aes(x = as.integer(mrs_90), y = cumsum_p)) +
-  geom_point(size = 2) +
-  geom_line(linewidth = 1) +
-  labs(x = "mRS 90", y = "Cumulative Proportion")
+model_parameters(model_mrs03,
+  centrality = "mean", dispersion = TRUE,
+  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
+  exponentiate = TRUE
+) |>
+  print(digits = 3, zap_small = TRUE)
 
-mrs_prop |>
-  ggplot(aes(x = as.integer(mrs_90), y = log(cumsum_p) - log(1 - cumsum_p))) +
-  geom_point(size = 2) +
-  geom_line(linewidth = 1) +
-  labs(x = "mRS 90", y = "Logit Cumulative Proportion")
+model_mrs_90 <- brm(
+  mrs_90 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
+  family = cumulative(link = "logit"),
+  data = data_erich_atach
+)
+
+pp_check(model_mrs_90, type = "ecdf_overlay", ndraws = 500)
+
+model_parameters(model_mrs_90,
+  centrality = "mean", dispersion = TRUE,
+  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
+  exponentiate = TRUE
+) |>
+  print(digits = 3, zap_small = TRUE)
+
+model_mrs_90_all <- brm(
+  mrs_90 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
+  family = cumulative(link = "logit"),
+  data = data
+)
+
+pp_check(model_mrs_90_all, type = "ecdf_overlay", ndraws = 500)
+
+model_parameters(model_mrs_90_all,
+  centrality = "mean", dispersion = TRUE,
+  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
+  exponentiate = TRUE
+) |>
+  print(digits = 3, zap_small = TRUE)
 
 ```
-
-## Frequentist Analysis
-
-### Mediation Analysis with Neurosurgery and ICH Laterality
-
-```{r frequentist ordinal}
-
-mrs_90_ich_lat <- clm(
-  mrs_90 ~ ich_laterality,
-  data = data,
-  link = "logit"
-)
-
-neurosurg_ich_lat <- clm(
-  neurosurgery_evac ~ ich_laterality,
-  data = data,
-  link = "logit"
-)
-
-mrs_90_ich_lat_neurosurg <- clm(
-  mrs_90 ~ ich_laterality + neurosurgery_evac,
-  data = data,
-  link = "logit"
-)
-
-summary(mrs_90_ich_lat)
-summary(neurosurg_ich_lat)
-summary(mrs_90_ich_lat_neurosurg)
-
-```
-
-## Bayesian Analysis
-
-```{r bayesian ordinal}
-
-fit1 <- brm(
-  data = data,
-  family = cumulative(probit),
-  mrs_90 ~ 1,
-)
-
-fit2 <- brm(
-  data = data,
-  family = cumulative(logit),
-  mrs_90 ~ 1,
-)
-
-fit3 <- brm(
-  data = data,
-  family = cumulative(logit),
-  mrs_90 ~ ich_laterality,
-)
-
-print(fit1)
-print(fit2)
-print(fit3)
-
-pred <- predictions(fit3)
-
-```

### 2023-10-17 analysis/models.qmd
who knows

@@
 library(broom.mixed)
-library(ggplot2)
 library(ggdist)

### 2023-10-17 analysis/kruscke.qmd
got it

---
title: "Ordinal Regression - Kruscke"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(brms)
library(cmdstanr)
library(here)
library(ordinal)

# seeds
set.seed(160)
BAYES_SEED <- 160

# Bayes
options(
  mc.cores = parallel::detectCores(),
  brms.backend = "cmdstanr"
)

```

## The Case of a Single Group

In this scenario, we will use modified Rankin scores from the ERICH dataset. First we will analyze this without any predictors.

```{r import data, warning=FALSE, message=FALSE}

data <- read_rds(here("data", "all.rds"), refhook = NULL) |>
  drop_na(mrs_90)

erich <- data |>
  filter(study == "ERICH")

erich_plot <- erich |>
  count(mrs_90) |>
  mutate(
    prob = n / nrow(erich),
    cum_prob = cumsum(n / nrow(erich))
  )

fit <- brm(
  data = erich,
  family = cumulative(logit),
  mrs_90 ~ 1
)

fit

draws <- as_draws_df(fit) |>
  mutate(across(starts_with("b_"), inv_logit_scaled)) |>
  pivot_longer(starts_with("b_")) |>
  group_by(name) |>
  summarize(
    mean = mean(value),
    sd = sd(value),
    ll = quantile(value, probs = 0.025),
    up = quantile(value, probs = 0.975)
  )

```

### 2023-10-18 analysis/aggressive_care.qmd
got a model!

@@
 ```
-
-## Frequentist Way
-
-```{r frequentist}
-
-matrix <- counts_tidy |>
-  filter(neurosurgery_evac == "Yes") |>
-  mutate(n_no = total - n) |>
-  select(n_yes = n, n_no) |>
-  as.matrix()
-
-prop.test(matrix) |>
-  model_parameters()
-
-```
-
-```{r binomial}
-
-ggplot() +
-  stat_function(fun = ~ dbeta(., 2, 6), geom = "area", fill = "#AC3414") +
-  scale_x_continuous(labels = label_percent()) +
-  labs(x = "Possible values for π", y = NULL, fill = NULL) +
-  theme_nice() +
-  theme(axis.text.y = element_blank())
-
-```

### 2023-10-18 analysis/dag.qmd
got a model!

@@
 # set theme of all DAGs to `theme_dag()`
-theme_set(theme_dag(base_family = "Cantarell"))
+theme_set(theme_dag(base_family = "Noah"))
 
@@
 
-## Specify a Causal Question
+## Specify Causal Questions
+
+Do patients with **left hemispheric** injuries receive less aggressive care as compared to patients with **right hemispheric** injuries?
+We will look at rates of neurosurgery, evd, tracheostomy, withdrawal of life-sustaining therapies.
 
 Do patients with **left hemispheric** injuries have worse otucomes as compared to patients with **right hemispheric** injuries?
-
-## Exploratory Data Analysis
-
-```{r eda, warning=FALSE, message=FALSE}
-#| fig-width: 10
-
-data <- read_csv(here("data", "all.csv"))
-
-data |>
-  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
-  filter(!is.na(ich_laterality)) |>
-  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
-  geom_density(color = NA, alpha = .8)
-
-```
 
@@
 ```
-
-```{r}
-
-```

### 2023-10-18 analysis/eda.qmd
got a model!

---
title: "Exploratory Data Analysis"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(dagitty)
library(ggdag)
library(MetBrewer)
library(here)
library(broom) # For converting model output to data frames

# set theme of all DAGs to `theme_dag()`
theme_set(theme_dag(base_family = "Noah"))

```

## Outcomes

```{r aggressiveness, warning=FALSE, message=FALSE}
#| fig-width: 10

```

```{r euroqol, warning=FALSE, message=FALSE}
#| fig-width: 10

data <- read_csv(here("data", "all.csv")) |>
  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
  filter(!is.na(ich_laterality))

data |>
  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
  geom_density(color = NA, alpha = .8)

```

### 2023-10-18 analysis/outcomes.qmd
got a model!

---
title: "Outcomes"
code-fold: true
---

```{r setup, warning=FALSE, message=FALSE}

library(tidyverse)
library(marginaleffects)
library(brms)
library(cmdstanr)
library(tidybayes)
library(parameters)
library(broom)
library(broom.mixed)
library(ggdist)
library(here)

# seeds
set.seed(160)
BAYES_SEED <- 160

# Bayes
options(
  mc.cores = parallel::detectCores(),
  brms.backend = "cmdstanr"
)

data <- read_rds(here("data", "all.rds"), refhook = NULL)

data <- data |>
  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90) |>
  drop_na(ich_laterality)

data$ich_location <- fct_drop(data$ich_location)
data$ich_laterality <- fct_drop(data$ich_laterality)

data_erich_atach <- data |>
  filter(study == "ERICH" | study == "ATACH2")

# plot stuff
theme_set(theme_minimal(base_family = "Noah"))

```

We will first run a "dummy" regression so I can interpret the results even before all data is cleaned.

```{r brms priors}

priors <- c(
  prior(normal(0, 10), class = "b")
)

mod_mrs_90 <- brm(
  mrs_90 ~ ich_laterality + age + ich_location + ich_volume_baseline + gcs_baseline + neurosurgery_evac,
  data = data,
  family = cumulative(link = "logit"),
  prior = priors,
  chains = 4,
  iter = 5000 * 2,
  seed = BAYES_SEED,
  refresh = 0
)

pp_check(mod_mrs_90, type = "bars", ndraws = 100)
pp_check(mod_mrs_90, type = "ecdf_overlay", ndraws = 100)

model_parameters(mod_mrs_90,
  centrality = "mean", dispersion = TRUE,
  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
  exponentiate = TRUE
) |>
  print(digits = 3, zap_small = TRUE)

get_variables(mod_mrs_90)

draws <- mod_mrs_90 |>
  spread_draws(b_ich_lateralityRight) |>
  mutate(ich_right_OR = exp(b_ich_lateralityRight))

draws |>
  ggplot(aes(x = ich_right_OR)) +
  stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "green"))

# Find the proportion of posterior draws that are bigger than OR of 1
draws %>%
  summarize(prop_greater_0 = sum(ich_right_OR > 1) / n())

# Find the proportion of posterior draws that are bigger than OR of 1.2
draws %>%
  summarize(prop_greater_0 = sum(ich_right_OR > 1.2) / n())

```

### 2023-10-18 analysis/aggressive_care.qmd
changes
Added setup.R script

@@
 ---
-
-```{r setup}
-#| code-fold: true
-#| code-summary: "Setup"
-#| warning: false
-#| message: false
-
-library(tidyverse) # ggplot, dplyr, and friends
-library(janitor)
-library(glue) # Easier string interpolation
-library(gt) # Fancy tables
-library(scales) # Nicer labeling functions
-library(patchwork) # Combine plots nicely
-library(ggmosaic) # Mosaic plots with ggplot
-library(parameters) # Extract model parameters as data frames
-library(cmdstanr) # Run Stan code from R
-library(brms) # Nice frontend for Stan
-library(tidybayes) # Manipulate MCMC chains in a tidy way
-library(here)
-
-# Set some global Stan options
-CHAINS <- 4
-ITER <- 2000
-WARMUP <- 1000
-BAYES_SEED <- 160
-options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")
-
-# ggplot theming
-theme_nice <- function() {
-  theme_minimal(base_family = "Cantarell") +
-    theme(
-      panel.grid.minor = element_blank(),
-      plot.title = element_text(family = "Cantarell", face = "bold"),
-      axis.title = element_text(family = "Cantarell Medium"),
-      axis.title.x = element_text(hjust = 0),
-      axis.title.y = element_text(hjust = 1),
-      strip.text = element_text(
-        family = "Cantarell", face = "bold",
-        size = rel(1), hjust = 0
-      ),
-      strip.background = element_rect(fill = NA, color = NA)
-    )
-}
-
-```
+{{< include 00-setup.qmd >}}
 

### 2023-10-18 analysis/outcomes.qmd
changes
Added setup.R script

@@
 
-```{r setup, warning=FALSE, message=FALSE}
-
-library(tidyverse)
-library(marginaleffects)
-library(brms)
-library(cmdstanr)
-library(tidybayes)
-library(parameters)
-library(broom)
-library(broom.mixed)
-library(ggdist)
-library(here)
-
-# seeds
-set.seed(160)
-BAYES_SEED <- 160
-
-# Bayes
-options(
-  mc.cores = parallel::detectCores(),
-  brms.backend = "cmdstanr"
-)
-
-data <- read_rds(here("data", "all.rds"), refhook = NULL)
-
-data <- data |>
-  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
-  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90) |>
-  drop_na(ich_laterality)
-
-data$ich_location <- fct_drop(data$ich_location)
-data$ich_laterality <- fct_drop(data$ich_laterality)
-
-data_erich_atach <- data |>
-  filter(study == "ERICH" | study == "ATACH2")
-
-# plot stuff
-theme_set(theme_minimal(base_family = "Noah"))
-
-```
+{{< include 00-setup.qmd >}}
 
@@
 
-```{r brms priors}
+```{r brms priors, warning=FALSE, message=FALSE}
 

### 2023-10-18 analysis/R/setup.R

@@ -1,0 +1,56 @@
+options(
+  tidyverse.quiet = TRUE
+)
+
+# libraries
+library(tidyverse)
+library(marginaleffects)
+library(brms)
+library(cmdstanr)
+library(tidybayes)
+library(parameters)
+library(broom)
+library(broom.mixed)
+library(ggdist)
+library(here)
+library(dagitty)
+library(ggdag)
+library(MetBrewer)
+library(patchwork)
+library(tableone)
+library(gt)
+library(gtExtras)
+library(janitor)
+library(scales) # Nicer labeling functions
+library(ggmosaic) # Mosaic plots with ggplot
+
+# Set some global Stan options
+CHAINS <- 4
+ITER <- 2000
+WARMUP <- 1000
+BAYES_SEED <- 160
+options(mc.cores = parallel::detectCores(), brms.backend = "cmdstanr")
+
+# seeds
+set.seed(160)
+BAYES_SEED <- 160
+
+# plot stuff
+theme_set(theme_minimal(base_family = "Noah"))
+
+# data import
+data <- read_rds(here("data", "all.rds"), refhook = NULL)
+
+data <- data |>
+  filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
+  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90) |>
+  drop_na(ich_laterality)
+
+data$ich_location <- fct_drop(data$ich_location)
+data$ich_laterality <- fct_drop(data$ich_laterality)
+
+data_erich_atach <- data |>
+  filter(study == "ERICH" | study == "ATACH2")
+
+data_surgery_trials <- data |>
+  filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III")

### 2023-10-18 analysis/eda.qmd
changes
Added setup.R script
table 1 and 2

@@
 
-```{r setup, warning=FALSE, message=FALSE}
+{{< include 00-setup.qmd >}}
 
-library(tidyverse)
-library(dagitty)
-library(ggdag)
-library(MetBrewer)
-library(here)
-library(broom) # For converting model output to data frames
+```{r table1, warning=FALSE, message=FALSE}
+#| fig-width: 10
 
-# set theme of all DAGs to `theme_dag()`
-theme_set(theme_dag(base_family = "Noah"))
+table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh_volume_baseline", "study")
+
+table1 <- data |>
+  select(table_one_vars) |>
+  tbl_summary(by = ich_laterality) |>
+  add_p() |>
+  add_overall() |>
+  bold_labels()
+
+table1
+
+```
+
+## Aggressiveness of Care
+
+First, we will explore if patients with **left hemispheric** injuries receive less aggressive care.
+
+### Neurosurgical Intervention
+
+```{r neurosurgery, warning=FALSE, message=FALSE}
+
+table_two_vars <- c("neurosurgery_evac", "evd", "tracheostomy", "peg", "ich_laterality", "comfort_care_binary", "")
+
+table2 <- data |>
+  select(table_two_vars) |>
+  tbl_summary(by = ich_laterality) |>
+  add_p() |>
+  add_overall() |>
+  bold_labels()
+
+table2
 
@@
 
-```{r aggressiveness, warning=FALSE, message=FALSE}
+```{r eda, warning=FALSE, message=FALSE}
 #| fig-width: 10

### 2023-10-18 analysis/R/setup.R

@@ -18,6 +18,7 @@ library(ggdag)
 library(MetBrewer)
 library(patchwork)
 library(tableone)
+library(gtsummary)
 library(gt)
 library(gtExtras)
 library(janitor)
@@ -43,11 +44,16 @@ data <- read_rds(here("data", "all.rds"), refhook = NULL)
 
 data <- data |>
   filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
-  select(study, age, ich_laterality, ich_location, ich_volume_baseline, gcs_baseline, ivh, neurosurgery_evac, sbp_baseline, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90) |>
+  select(study, age, sex, race, ethnicity, time_symptoms_to_ed, sbp_baseline, ich_laterality, ich_location, nihss_baseline, gcs_baseline, ich_volume_baseline, ivh, ivh_volume_baseline, htn, hld, dm2, stroke, afib, tobacco, cad, neurosurgery_evac, neurosurgery_evac_day, evd, tracheostomy, peg, dnr_binary, dnr_day, dni_binary, dni_day, comfort_care_binary, comfort_care_day, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90, hospital_los) |>
   drop_na(ich_laterality)
 
-data$ich_location <- fct_drop(data$ich_location)
-data$ich_laterality <- fct_drop(data$ich_laterality)
+glimpse(data)
+
+# # converting some columns to factors
[…]

### 2023-10-31 analysis/eda.qmd
eda updates

--- 2023-10-18 analysis/eda.qmd
@@
 
-```{r setup, warning=FALSE, message=FALSE}
+{{< include 00-setup.qmd >}}
 
-library(tidyverse)
-library(dagitty)
-library(ggdag)
-library(MetBrewer)
-library(here)
-library(broom) # For converting model output to data frames
+**Things to do**:
 
-# set theme of all DAGs to `theme_dag()`
-theme_set(theme_dag(base_family = "Noah"))
+- Collapse some categorical values
+- IVH as binary instead of volume
+
+```{r table1, warning=FALSE, message=FALSE}
+#| fig-width: 10
+
+table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")
+
+table1 <- data |>
+  select(table_one_vars) |>
+  tbl_summary(
+    by = ich_laterality,
+    missing = "no",
+    label = list(
+      age ~ "Age",
+      sex ~ "Sex",
+      race ~ "Race",
+      ethnicity ~ "Ethnicity",
+      time_symptoms_to_ed ~ "LNW to ED Presentation (mins)",
+      sbp_baseline ~ "Baseline SBP (mm Hg)",
+      nihss_baseline ~ "Baseline NIHSS",
+      gcs_baseline ~ "Baseline GCS",
+      htn ~ "Hypertension",
+      dm2 ~ "Type II Diabetes",
+      stroke ~ "Previous Stroke",
+      tobacco ~ "Tobacco Use",
+      ich_location ~ "ICH Location",
+      ich_volume_baseline ~ "ICH Volume (mL)",
+      ivh ~ "IVH",
+      study ~ "Study"
+    )
+  ) |>
+  add_p() |>
+  add_overall() |>
+  bold_labels()
+
+table1
+
+```
+
+## Aggressiveness of Care
+
+First, we will explore if patients with **left hemispheric** injuries receive less aggressive care.
+
+```{r neurosurgery, warning=FALSE, message=FALSE}
+
+table_two_vars <- c("neurosurgery_evac", "evd", "tracheostomy", "ich_laterality", "comfort_care_binary")
+
+table2 <- data |>
+  select(table_two_vars) |>
+  tbl_summary(
+    by = ich_laterality,
+    missing = "no",
+    label = list(
+      neurosurgery_evac ~ "Neurosurgical Intervention",
+      evd ~ "Ventriculostomy",
+      tracheostomy ~ "Tracheostomy", comfort_care_binary ~ "Comfort Care"
+    )
+  ) |>
+  add_p() |>
+  add_overall() |>
+  bold_labels()
+
+table2
+
+```
+
+### Neurosurgical Intervention
+
+```{r neurosurgical intervention, warning=FALSE, message=FALSE}
+#| fig-width: 10
+
+neurosurgery_glm <- glm(
+  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+  data = data,
+  family = binomial
+)
+
+model_parameters(neurosurgery_glm,
+  centrality = "mean", dispersion = TRUE,
+  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
+  exponentiate = TRUE
+) |>
+  print(digits = 2, zap_small = TRUE)
 
@@
 ## Outcomes
-
-```{r aggressiveness, warning=FALSE, message=FALSE}
-#| fig-width: 10
-
-```
 

### 2023-10-31 analysis/eda.qmd
making progress (still)

@@
 - IVH as binary instead of volume
+
+## Table 1
 
@@
 
-## Aggressiveness of Care
+## Table 2 - Aggressiveness of Care (Unadjusted)
 
@@
 
-table_two_vars <- c("neurosurgery_evac", "evd", "tracheostomy", "ich_laterality", "comfort_care_binary")
+table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")
 
@@
       evd ~ "Ventriculostomy",
-      tracheostomy ~ "Tracheostomy", comfort_care_binary ~ "Comfort Care"
+      days_mechanical_ventilation ~ "Intubation Days",
+      tracheostomy ~ "Tracheostomy",
+      comfort_care_binary ~ "Comfort Care"
     )
@@
 
-### Neurosurgical Intervention
+## Neurosurgical Intervention - Multivariate Regression
 

### 2023-10-31 analysis/R/setup.R

@@ -44,7 +44,7 @@ data <- read_rds(here("data", "all.rds"), refhook = NULL)
 
 data <- data |>
   filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
-  select(study, age, sex, race, ethnicity, time_symptoms_to_ed, sbp_baseline, ich_laterality, ich_location, nihss_baseline, gcs_baseline, ich_volume_baseline, ivh, ivh_volume_baseline, htn, hld, dm2, stroke, afib, tobacco, cad, neurosurgery_evac, neurosurgery_evac_day, evd, tracheostomy, peg, dnr_binary, dnr_day, dni_binary, dni_day, comfort_care_binary, comfort_care_day, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90, hospital_los) |>
+  select(study, age, sex, race, ethnicity, time_symptoms_to_ed, sbp_baseline, ich_laterality, ich_location, nihss_baseline, gcs_baseline, ich_volume_baseline, ivh, ivh_volume_baseline, htn, hld, dm2, stroke, afib, tobacco, cad, neurosurgery_evac, neurosurgery_evac_day, evd, days_mechanical_ventilation, tracheostomy, peg, dnr_binary, dnr_day, dni_binary, dni_day, comfort_care_binary, comfort_care_day, mrs_90_01, mrs_90_02, mrs_90_03, mrs_90_04, mrs_90, hospital_los) |>
   drop_na(ich_laterality)
 
 glimpse(data)

### 2023-11-01 analysis/eda.qmd
poisson

@@
 
-table1 <- data |>
+table1 <- data_erich_atach |>
   select(table_one_vars) |>
@@
 
-table2 <- data |>
+table2 <- data_erich_atach |>
   select(table_two_vars) |>
@@
   neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
-  data = data,
+  data = data_erich_atach,
   family = binomial
@@
 
-model_parameters(neurosurgery_glm,
+neurosurgery_glm |>
+  tbl_regression(
+    exponentiate = TRUE,
+    include = ich_laterality
+  )
+
+left_right_regression <- function(name, var, family) {
+  name <- glm(
+    {{ var }} ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+    data = data_erich_atach,
+    family = family
+  )
+
+  neurosurgery_glm |>
+    tbl_regression(
+      exponentiate = TRUE,
+      include = ich_laterality
+    )
+}
+
+left_right_regression(neurosurgery_glm, neurosurgery_evac, binomial)
+
+vent_glm <- glm(
+  days_mechanical_ventilation ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+  data = data_erich_atach,
+  family = poisson
+)
+
+model_parameters(vent_glm,
   centrality = "mean", dispersion = TRUE,

### 2023-11-01 analysis/R/setup.R

@@ -18,8 +18,8 @@ library(ggdag)
 library(MetBrewer)
 library(patchwork)
 library(tableone)
-library(gtsummary)
 library(gt)
+library(gtsummary)
 library(gtExtras)
 library(janitor)
 library(scales) # Nicer labeling functions
@@ -56,7 +56,7 @@ glimpse(data)
 #   mutate(across(factors, factor))
 
 data_erich_atach <- data |>
-  filter(study == "ERICH" | study == "ATACH2")
+  filter(study == "ERICH" | study == "ATACH-2")
 
 data_surgery_trials <- data |>
   filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III")

### 2023-11-03 analysis/eda.qmd
Update eda.qmd

@@
 
-left_right_regression <- function(name, var, family) {
+left_right_regression <- function(name, x, family) {
   name <- glm(
-    {{ var }} ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+    {{ x }} ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
     data = data_erich_atach,
@@
 
-  neurosurgery_glm |>
+  {{ x }} |>
     tbl_regression(

### 2023-11-20 analysis/aggressive_care.qmd
who knows

@@
 I will be using observational data from the ERICH study, as well as the randomized trial data from ATACH-2 to answer these questions.
-
-## Exploratory Data Analysis
-
-Loading the data from ATACH-2 and ERICH, and filtering out the variables needed for the analysis.
-
-```{r loading data}
-
-# load data
-data <- read_rds(here("data", "all.rds"), refhook = NULL)
-
-# filter out only ERICH and ATACH2 and keep relevant variables
-cleaned_data <- data |>
-  mutate(id = 1:n()) |>
-  filter(study == "ERICH" | study == "ATACH-2") |>
-  select(id, study, ich_laterality, age, sex, race, ethnicity, sbp_baseline, gcs_baseline, ich_volume_baseline, ivh, ich_location, neurosurgery_evac, evd, tracheostomy, peg, comfort_care_binary, comfort_care_day) |>
-  drop_na(ich_laterality, neurosurgery_evac)
-
-```
-
-Now we can summarize the preliminary data.
-
-```{r exploratory data analysis}
-
-counts <- cleaned_data |>
-  tabyl(ich_laterality, neurosurgery_evac, ich_location) |>
-  adorn_percentages("row") |>
-  adorn_pct_formatting(rounding = "half up", digits = 1)
-
-counts
-
-counts_tidy <- cleaned_data |>
-  group_by(ich_laterality, neurosurgery_evac) |>
-  summarize(n = n()) |>
-  mutate(
-    total = sum(n),
-    prop = n / sum(n)
-  ) |>
-  ungroup()
-
-counts_tidy
-
-ggplot(cleaned_data) +
-  geom_mosaic(
-    aes(
-      x = product(neurosurgery_evac, ich_laterality),
-      fill = ich_laterality, alpha = neurosurgery_evac
-    ),
-    divider = mosaic("v")
-  )
-
-```
 
@@
 ```
+
+```{r neurosurgery, warning=FALSE, message=FALSE}
+
+table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")
+
+table2 <- data_erich_atach |>
+  select(table_two_vars) |>
+  tbl_summary(
+    by = ich_laterality,
+    missing = "no",
+    label = list(
+      neurosurgery_evac ~ "Neurosurgical Intervention",
+      evd ~ "Ventriculostomy",
+      days_mechanical_ventilation ~ "Intubation Days",
+      tracheostomy ~ "Tracheostomy",
+      comfort_care_binary ~ "Comfort Care"
+    )
+  ) |>
+  add_p() |>
+  add_overall() |>
+  bold_labels()
+
+table2
+
+```
+
+## Neurosurgical Intervention - Multivariate Regression
+
+```{r neurosurgical intervention, warning=FALSE, message=FALSE}
+#| fig-width: 10
+
+neurosurgery_glm <- glm(
+  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+  data = data_erich_atach,
+  family = binomial
+)
+
+neurosurgery_glm |>
+  tbl_regression(
+    exponentiate = TRUE,
+    include = ich_laterality
+  )
+
+vent_glm <- glm(
+  days_mechanical_ventilation ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+  data = data_erich_atach,
+  family = poisson
+)
+
+vent_glm |>
+  tbl_regression(
+    exponentiate = TRUE,
+    include = ich_laterality
+  )
+
+comfort_glm <- glm(
+  comfort_care_binary ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
+  data = data_erich_atach,
+  family = binomial
+)
+
+comfort_glm |>
+  tbl_regression(
+    exponentiate = TRUE,
+    include = ich_laterality
+  )
+
+```

### 2023-11-20 analysis/eda.qmd
who knows

@@
 - IVH as binary instead of volume
+- Some percentages values are off, likely due to missing data (e.g., Hypertension)
 
@@
 #| fig-width: 10
-
+#|
 table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")
@@
 First, we will explore if patients with **left hemispheric** injuries receive less aggressive care.
-
-```{r neurosurgery, warning=FALSE, message=FALSE}
-
-table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")
-
-table2 <- data_erich_atach |>
-  select(table_two_vars) |>
-  tbl_summary(
-    by = ich_laterality,
-    missing = "no",
-    label = list(
-      neurosurgery_evac ~ "Neurosurgical Intervention",
-      evd ~ "Ventriculostomy",
-      days_mechanical_ventilation ~ "Intubation Days",
-      tracheostomy ~ "Tracheostomy",
-      comfort_care_binary ~ "Comfort Care"
-    )
-  ) |>
-  add_p() |>
-  add_overall() |>
-  bold_labels()
-
-table2
-
-```
-
-## Neurosurgical Intervention - Multivariate Regression
-
-```{r neurosurgical intervention, warning=FALSE, message=FALSE}
-#| fig-width: 10
-
-neurosurgery_glm <- glm(
-  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
-  data = data_erich_atach,
-  family = binomial
-)
-
-neurosurgery_glm |>
-  tbl_regression(
-    exponentiate = TRUE,
-    include = ich_laterality
-  )
-
-left_right_regression <- function(name, x, family) {
-  name <- glm(
-    {{ x }} ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
-    data = data_erich_atach,
-    family = family
-  )
-
-  {{ x }} |>
-    tbl_regression(
-      exponentiate = TRUE,
-      include = ich_laterality
-    )
-}
-
-left_right_regression(neurosurgery_glm, neurosurgery_evac, binomial)
-
-vent_glm <- glm(
-  days_mechanical_ventilation ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
-  data = data_erich_atach,
-  family = poisson
-)
-
-model_parameters(vent_glm,
-  centrality = "mean", dispersion = TRUE,
-  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
-  exponentiate = TRUE
-) |>
-  print(digits = 2, zap_small = TRUE)
-
-```
-
-## Outcomes
-
-```{r euroqol, warning=FALSE, message=FALSE}
-#| fig-width: 10
-
-data <- read_csv(here("data", "all.csv")) |>
-  filter(ich_location == c("Lobar", "Basal Ganglia", "Thalamus")) |>
-  filter(!is.na(ich_laterality))
-
-data |>
-  ggplot(aes(euro_vas_90, fill = ich_laterality)) +
-  geom_density(color = NA, alpha = .8)
-
-```

### 2023-11-20 analysis/R/setup.R

@@ -37,7 +37,7 @@ set.seed(160)
 BAYES_SEED <- 160
 
 # plot stuff
-theme_set(theme_minimal(base_family = "Noah"))
+theme_set(theme_minimal(base_family = "Figtree"))
 
 # data import
 data <- read_rds(here("data", "all.rds"), refhook = NULL)

### 2024-01-11 analysis/aggressive_care.qmd
updates

@@
 
-**Do patients with left hemispheric ICH receive neurosurgical evacuations less frequently as compared to patients with right hemispheric ICH?**
-
-We want to determine if there's a substantial difference in the proportion of patients who receive neurosurgical intervention in the right or left hemisphere.
-
-```{r color setup}
+```{r color setup, warning=FALSE, message=FALSE}
 
@@
 ```
+
+**Do patients with left hemispheric ICH receive neurosurgical evacuations less frequently as compared to patients with right hemispheric ICH?**
+
+We want to determine if there's a substantial difference in the proportion of patients who receive aggressive care (mostly neurosurgical intervention) in the right vs. left hemisphere.
+
+### Table 2 - Aggressiveness of Care (Unadjusted)
+
+Things to clarify:
+
+- How is comfort care defined?
+- Put mean instead of median?
+- Run "unadjusted analyses" and put results of both adjusted and unadjusted into one table
 
@@
 
-## Neurosurgical Intervention - Multivariate Regression
+### Aggressiveness of Care (Adjusted)
+
+#### Neurosurgical Intervention
 
@@
 
+```
+
+#### Days of Mechanical Ventilation
+
+```{r days of ventilation, warning=FALSE, message=FALSE}
+#| fig-width: 10
+
 vent_glm <- glm(
@@
   )
+
+```
+
+#### Comfort Care
+
+```{r comfort care, warning=FALSE, message=FALSE}
+#| fig-width: 10
 

### 2024-01-11 analysis/eda.qmd
updates

@@
 
-- Collapse some categorical values
-- IVH as binary instead of volume
-- Some percentages values are off, likely due to missing data (e.g., Hypertension)
+- Collapse some categorical values especially those that are missing
+- Some percentages values are off, likely due to missing data (e.g., Hypertension). Confirm PMH data.
 

### 2024-01-12 analysis/03-eda.qmd
neurosurgery evac bayes models

---
title: "Exploratory Data Analysis"
code-fold: true
---

{{< include 00-setup.qmd >}}

::: {.callout-warning appearance="simple"}

Things to clarify:

- Collapse some categorical values especially those that are missing
- Some percentages values are off, likely due to missing data (e.g., Hypertension). Confirm PMH data.

:::

## Table 1 - ERICH + ATACH-2

```{r table1 erich and atach2, warning=FALSE, message=FALSE}
#| fig-width: 10
#|
table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")

table1 <- data_erich_atach |>
  select(table_one_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      age ~ "Age",
      sex ~ "Sex",
      race ~ "Race",
      ethnicity ~ "Ethnicity",
      time_symptoms_to_ed ~ "LNW to ED Presentation (mins)",
      sbp_baseline ~ "Baseline SBP (mm Hg)",
      nihss_baseline ~ "Baseline NIHSS",
      gcs_baseline ~ "Baseline GCS",
      htn ~ "Hypertension",
      dm2 ~ "Type II Diabetes",
      stroke ~ "Previous Stroke",
      tobacco ~ "Tobacco Use",
      ich_location ~ "ICH Location",
      ich_volume_baseline ~ "ICH Volume (mL)",
      ivh ~ "IVH",
      study ~ "Study"
    )
  ) |>
  add_p() |>
  add_overall() |>
  bold_labels()

table1

```

## Table 1 - All Studies

```{r table1 all, warning=FALSE, message=FALSE}
#| fig-width: 10
#|
table_one_vars <- c("age", "sex", "race", "ethnicity", "time_symptoms_to_ed", "sbp_baseline", "nihss_baseline", "gcs_baseline", "ich_laterality", "htn", "dm2", "stroke", "tobacco", "ich_location", "ich_volume_baseline", "ivh", "study")

table1_all <- data |>
  select(table_one_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      age ~ "Age",
      sex ~ "Sex",
      race ~ "Race",
      ethnicity ~ "Ethnicity",
      time_symptoms_to_ed ~ "LNW to ED Presentation (mins)",
      sbp_baseline ~ "Baseline SBP (mm Hg)",
      nihss_baseline ~ "Baseline NIHSS",
      gcs_baseline ~ "Baseline GCS",
      htn ~ "Hypertension",
      dm2 ~ "Type II Diabetes",
      stroke ~ "Previous Stroke",
      tobacco ~ "Tobacco Use",
      ich_location ~ "ICH Location",
      ich_volume_baseline ~ "ICH Volume (mL)",
      ivh ~ "IVH",
      study ~ "Study"
    )
  ) |>
  add_p() |>
  add_overall() |>
  bold_labels()

table1_all

```

### 2024-01-12 analysis/04-aggressive_care.qmd
neurosurgery evac bayes models

---
title: "Differences in Care"
code-fold: false
toc-depth: 5
---

{{< include 00-setup.qmd >}}

```{r color setup, warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: Setup

# Left/Right hemisphere colors
clrs_saguaro <- NatParksPalettes::natparks.pals("Saguaro")
clr_left <- clrs_saguaro[1]
clr_right <- clrs_saguaro[6]
clr_diff <- clrs_saguaro[4]

```

**How are patients treated differently if they experience a left hemispheric ICH vs. right hemispheric ICH?**

- Do they receive neurosurgical intervention at a different rate?
- Are they on the ventilator for longer?
- Are they less likely to receive a tracheostomy or PEG?
- Are they more likely to proceed to comfort care?

There are five studies available to analyze: `ERICH`, `ATACH-2`, `MISTIE-II`, `CLEAR-IVH`, and `MISTIE-III`. Since we are interested in the potential differences in rates of neurosurgical intervention, we will exclude studies that randomized patients to an intervention, leaving `ERICH` and `ATACH-2` to help answer these questions.

### Unadjusted Analysis

::: {.callout-warning appearance="simple"}

Things to clarify:

- How should `comfort care` be defined? Below, it is a binary variable, but would it be better to model it based on a time-to-event process?
- Run "unadjusted analyses" and put results of both adjusted and unadjusted into one table?

:::

```{r table two unadjusted, warning=FALSE, message=FALSE}

table_two_vars <- c("neurosurgery_evac", "evd", "days_mechanical_ventilation", "tracheostomy", "ich_laterality", "comfort_care_binary")

table2 <- data_erich_atach |>
  select(table_two_vars) |>
  tbl_summary(
    by = ich_laterality,
    missing = "no",
    label = list(
      neurosurgery_evac ~ "Neurosurgical Intervention",
      evd ~ "Ventriculostomy",
      days_mechanical_ventilation ~ "Mechanical Ventilation Days",
      tracheostomy ~ "Tracheostomy",
      comfort_care_binary ~ "Comfort Care"
    )
  ) |>
  add_overall() |>
  bold_labels()

table2

```

It appears that patients with left hemispheric lesions tend to receive neurosurgical intervention at a lower rate, as well as stay on the ventilator for longer (based on IQR). There does not seem to be a large difference in rates of tracheostomy and comfort care.

### Modeling Differences in Care

#### Neurosurgical Intervention

> How does the probability of neurosurgical intervention (outcome) change if a patient has a left hemispheric ICH vs. right hemispheric ICH?

We will incoporate the following predictors into the model: `ich_laterality`, `age`, `ich_location`, `gcs_baseline`, `ich_volume_baseline`, `ivh`, and `time_symptoms_to_ed`.

##### Priors

We are using default priors as defined in the `brms` package.

::: {.callout-warning appearance="simple"}

Think about defining weakly informative priors and running prior predictive checks to make this analysis more robust.

:::

##### Posterior simulation

```{r neurosurgical intervention, warning=FALSE, message=FALSE}
#| fig-width: 10

neurosurgery_glm <- brm(
  neurosurgery_evac ~ ich_laterality + age + ich_location + gcs_baseline + ich_volume_baseline + ivh + time_symptoms_to_ed,
  data = data_erich_atach,
  family = bernoulli(link = "logit")
)

neurosurgery_glm |>
  gather_draws(`^b_.*`, regex = TRUE) |>
  ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
  geom_line(size = 0.1) +
  scale_color_viridis_d(option = "rocket", end = 0.85) +
  facet_wrap(vars(.variable), scales = "free_y")

pp_check(neurosurgery_glm, ndraws = 500)

model_parameters(neurosurgery_glm,
  centrality = "mean", dispersion = TRUE,
  test = FALSE, verbose = FALSE, ci_method = "hdi", priors = TRUE,
  exponentiate = TRUE
) |>
  print(digits = 2, zap_small = TRUE)

```

Based on this model, there is a **much higher odds (~120% increase)** of a patient with right hemispheric ICH receiving neurosurgical intervention as compared to a patient with left hemispheric ICH **(OR 2.24, 95% CI 1.55 - 3.14)**.

```{r neurosurgical intervention marginaleffects, warning=FALSE, message=FALSE}
#| fig-width: 10

```

### 2024-01-12 analysis/R/setup.R

@@ -5,6 +5,7 @@ options(
 # libraries
 library(tidyverse)
 library(marginaleffects)
+library(modelsummary)
 library(brms)
 library(cmdstanr)
 library(tidybayes)

### 2024-02-09 reports/aggressiveness.qmd
table_one for eda complete

---
title: Differences in care for patients experiencing left vs. right hemisphere ICH
author: Nikhil Patel, MD
format: html
toc: true
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
#| code-fold: false

library(targets)
library(readr)
library(tidyr)
library(janitor)
library(dplyr)
library(forcats)
library(stringr)
library(skimr)
library(gt)
library(gtsummary)
library(marginaleffects)
library(modelsummary)

tar_load(ich)
tar_load(non_surgery_trials)
tar_load(surgery_trials)
tar_load(table_one)

```

## Exploratory Data Analysis

::: {.callout-warning title="Issues to Address"}
`htn` is missing mostly all information. It is also listed as a character variable.

All other co-morbidities are missing information; if this is the case, should default to "No".

`ivh` is also missing some values, which should also default to "No"?

Any "Yes/No" variable should not have any missing information, should just default to "No"

Verify accuracy/validity of `time_symptoms_to_ed`, `sbp_baseline`,

Fix/drop unused levels in `race`, and fix "American Indian"

Fix "unknown" category in `ethnicity` to "Not Hispanic or Latino"

Fix/drop unused levels in `ich_location` and `study`
:::

```{r}

glimpse(ich)
skim(ich)

```

## Table 1

::: {.callout-warning title="Issues to Address"}

Change order of the displayed to frequency

:::

```{r}

table_one

```

### 2024-02-09 R/data_cleaning.R

@@ -50,3 +50,9 @@ filter_variables <- function(x) {
     filter(ich_location == "Basal Ganglia" | ich_location == "Thalamus" | ich_location == "Lobar") |>
     drop_na(ich_laterality)
 }
+
+non_surgery_trial_function <- function(x) {
+  x |>
+    filter(study == "ERICH" | study == "ATACH-2")
+}
+

### 2024-02-09 R/exploratory_data_analysis.R

@@ -1,0 +1,49 @@
+table_1_function <- function(x) {
+  table_one_vars <- c(
+    "age",
+    "sex",
+    "race",
+    "ethnicity",
+    "time_symptoms_to_ed",
+    "sbp_baseline",
+    "nihss_baseline",
+    "gcs_baseline",
+    "ich_laterality",
+    "htn",
+    "dm2",
+    "stroke",
+    "tobacco",
+    "ich_location",
+    "ich_volume_baseline",
+    "ivh",
+    "study"
+  )
+
+  x |>
+    select(all_of(table_one_vars)) |>
+    tbl_summary(
+      by = ich_laterality,
+      missing = "no",
+      label = list(
+        age ~ "Age",
+        sex ~ "Sex",
+        race ~ "Race",
+        ethnicity ~ "Ethnicity",
+        time_symptoms_to_ed ~ "LNW to ED Presentation (mins)",
+        sbp_baseline ~ "Baseline SBP (mm Hg)",
+        nihss_baseline ~ "Baseline NIHSS",
+        gcs_baseline ~ "Baseline GCS",
+        htn ~ "Hypertension",
+        dm2 ~ "Type II Diabetes",
+        stroke ~ "Previous Stroke",
+        tobacco ~ "Tobacco Use",
+        ich_location ~ "ICH Location",
+        ich_volume_baseline ~ "ICH Volume (mL)",
+        ivh ~ "IVH",
+        study ~ "Study"
+      )
+    ) |>
+    add_p() |>
+    add_overall() |>
+    bold_labels()
+}

### 2024-02-09 _targets.R

@@ -32,16 +32,30 @@ set.seed(4641) # From random.org
 
 source("R/packages.R")
 source("R/data_cleaning.R")
+source("R/exploratory_data_analysis.R")
 
 # Pipeline ----
 tar_plan(
+
   ## Define raw data file ----
   tar_file_read(
     imported_data,
     "data/raw_data/all.rds",
     read_rds(!!.x)
   ),
+
   ## Select, filter, and clean data ----
   selected_data = select_variables(imported_data),
-  ich_cleaned = filter_variables(selected_data)
+  ich = filter_variables(selected_data),
+  non_surgery_trials = ich |> filter(study == "ERICH" | study == "ATACH-2"),
+  surgery_trials = ich |> filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III"),
+
+  ## Exploratory data analysis ----
+  table_one = table_1_function(non_surgery_trials),
+
+  ## Render report ----
+  tar_quarto(
+    report,
+    "reports/aggressiveness.qmd"
+  )
 )

### 2024-02-09 reports/aggressiveness.qmd
shit

@@
 library(tidyr)
-library(janitor)
+library(skimr)
 library(dplyr)
-library(forcats)
-library(stringr)
-library(skimr)
 library(gt)
 library(gtsummary)
+library(ggplot2)
 library(marginaleffects)
 library(modelsummary)
+
+theme_set(theme_minimal(base_family = "Figtree"))
 
@@
 tar_load(surgery_trials)
-tar_load(table_one)
+tar_load(table_1)
+tar_load(table_2)
 
@@
 ## Exploratory Data Analysis
+
+### Data Overview
 
@@
 
-glimpse(ich)
-skim(ich)
+glimpse(non_surgery_trials)
+skim(non_surgery_trials)
 
@@
 
-## Table 1
+### Table 1
 
@@
 
-Change order of the displayed to frequency
+Collapse unused levels in `race`, `ethnicity`, `ich_location` and `study`
+
+Change order of the displayed factors to frequency
 
@@
 
-table_one
+table_1
 
 ```
+
+### Relationships between Variables
+
+```{r}
+#| fig-align: center
+
+table_2
+
+```
+
+## Statistical Models
+
+::: {.callout-important title="Other Considerations"}
+
+Add analyses stratifying by each study: `ERICH` and `ATACH2`
+
+Construct DAGs for **each** outcome (e.g., each model)!
+
+Add `DNR` in addition to `comfort_care`?
+
+:::
+
+* `neurosurgery_evac`: logistic
+* `ventriculostomy`: logistic
+* `days_mechanical_ventilation`: poisson (cumulative logit?)
+* `tracheostomy`: logistic
+* `comfort_care`: logistic (days to comfort care, time to event?)
+
+### DAGs

### 2024-02-09 R/dags.R

@@ -1,0 +1,17 @@
+dag_function <- function(x) {
+  node_details <- tribble(
+    ~name, ~label, ~x, ~y,
+    "functional_disability", "Functional Disability", 4, 1,
+    "ich_laterality", "Hemispheric Laterality", 1, 1,
+    "age", "Age", 2.75, 3,
+    "ich_location", "ICH Location", 2, 2.75,
+    "ich_volume", "ICH Volume", 1, 2.75,
+    "ivh", "IVH", 1.5, 2.25,
+    "gcs_baseline", "Admission GCS", 2, 3.25,
+    "wlst", "Early WLST", 3.75, 2.25,
+    "neurosurgery", "Neurosurgery", 2, 1.5,
+  )
+
+  node_labels <- node_details$label
+  names(node_labels) <- node_details$name
+}

### 2024-02-09 R/exploratory_data_analysis.R

@@ -47,3 +47,30 @@ table_1_function <- function(x) {
     add_overall() |>
     bold_labels()
 }
+
+table_2_function <- function(x) {
+  table_two_vars <- c(
+    "neurosurgery_evac",
+    "evd",
+    "days_mechanical_ventilation",
+    "tracheostomy",
+    "ich_laterality",
+    "comfort_care_binary"
+  )
+
+  x |>
+    select(all_of(table_two_vars)) |>
+    tbl_summary(
+      by = ich_laterality,
+      missing = "no",
+      label = list(
+        neurosurgery_evac ~ "Neurosurgical Intervention",
+        evd ~ "Ventriculostomy",
+        days_mechanical_ventilation ~ "Mechanical Ventilation Days",
+        tracheostomy ~ "Tracheostomy",
+        comfort_care_binary ~ "Comfort Care"
+      )
+    ) |>
+    add_overall() |>
+    bold_labels()
+}

### 2024-02-09 _targets.R

@@ -51,7 +51,8 @@ tar_plan(
   surgery_trials = ich |> filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III"),
 
   ## Exploratory data analysis ----
-  table_one = table_1_function(non_surgery_trials),
+  table_1 = table_1_function(non_surgery_trials),
+  table_2 = table_2_function(non_surgery_trials),
 
   ## Render report ----
   tar_quarto(

### 2024-02-13 reports/aggressiveness.qmd
dags

@@
 library(ggplot2)
+library(ggdag)
+library(dagitty)
 library(marginaleffects)
@@
 
-tar_load(ich)
-tar_load(non_surgery_trials)
-tar_load(surgery_trials)
-tar_load(table_1)
-tar_load(table_2)
+# tar_load(ich)
+# tar_load(non_surgery_trials)
+# tar_load(surgery_trials)
+# tar_load(table_1)
+# tar_load(table_2)
+# tar_load(aggressive_dag)
+tar_load_everything()
 
@@
 ### DAGs
+
+```{r}
+#| width: 1200
+#| height: 2000
+#| fig-align: center
+#|
+theme_set(
+  theme_dag() %+replace%
+    # also add some additional styling
+    theme(
+      legend.position = "bottom",
+      strip.text.x = element_text(margin = margin(2, 0, 2, 0, "mm"))
+    )
+)
+
+ggdag_adjustment_set(aggressive_dag, shadow = TRUE, use_labels = "label", text = FALSE)
+
+ggdag_adjustment_set(outcome_dag, shadow = TRUE, use_labels = "label", text = FALSE)
+
+aggressive_dag |>
+  ggdag_paths(shadow = TRUE, text = FALSE, use_labels = "label")
+
+```

### 2024-02-13 R/dags.R

@@ -1,17 +1,57 @@
-dag_function <- function(x) {
-  node_details <- tribble(
+aggressive_dag_function <- function(x) {
+  aggressiveness_node_details <- tribble(
     ~name, ~label, ~x, ~y,
-    "functional_disability", "Functional Disability", 4, 1,
+    "aggressive_care", "Aggressive Care", 5, 1,
     "ich_laterality", "Hemispheric Laterality", 1, 1,
-    "age", "Age", 2.75, 3,
-    "ich_location", "ICH Location", 2, 2.75,
-    "ich_volume", "ICH Volume", 1, 2.75,
-    "ivh", "IVH", 1.5, 2.25,
-    "gcs_baseline", "Admission GCS", 2, 3.25,
-    "wlst", "Early WLST", 3.75, 2.25,
-    "neurosurgery", "Neurosurgery", 2, 1.5,
+    "age", "Age", 3.75, 2.5,
+    "ich_location", "ICH Location", 2.75, 1.75,
+    "ich_volume", "ICH Volume", 1.25, 4,
+    "ivh", "IVH", 1, 3,
+    "gcs_baseline", "Admission GCS", 4.5, 3.75
   )
 
-  node_labels <- node_details$label
-  names(node_labels) <- node_details$name
+  aggressiveness_node_labels <- aggressiveness_node_details$label
+  names(aggressiveness_node_labels) <- aggressiveness_node_details$name
+
+  dagify(
+    aggressive_care ~ ich_laterality + ivh + ich_volume + gcs_baseline + ich_location + age,
+    ivh ~ ich_location,
+    gcs_baseline ~ ich_volume + age + ivh,
+    ich_location ~ age,
+    exposure = "ich_laterality",
+    outcome = "aggressive_care",
+    coords = aggressiveness_node_details,
+    labels = aggressiveness_node_labels
+  )
+}
+
+outcome_dag_function <- function(x) {
+  outcome_node_details <- tribble(
+    ~name, ~label, ~x, ~y,
+    "functional_outcome", "Functional Outcome", 5, 1,
+    "ich_laterality", "Hemispheric Laterality", 1, 1,
+    "age", "Age", 3.75, 2.5,
+    "ich_location", "ICH Location", 2.75, 1.75,
+    "ich_volume", "ICH Volume", 1.25, 4,
+    "ivh", "IVH", 1, 3,
+    "gcs_baseline", "Admission GCS", 4.5, 3.75,
+    "wlst", "Early WLST", 3.75, 2.25,
[…]

### 2024-02-13 _targets.R

@@ -33,6 +33,8 @@ set.seed(4641) # From random.org
 source("R/packages.R")
 source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
+source("R/dags.R")
+
 
 # Pipeline ----
 tar_plan(
@@ -53,6 +55,10 @@ tar_plan(
   ## Exploratory data analysis ----
   table_1 = table_1_function(non_surgery_trials),
   table_2 = table_2_function(non_surgery_trials),
+
+  ## DAGs ----
+  aggressive_dag = aggressive_dag_function(x),
+  outcome_dag = outcome_dag_function(x),
 
   ## Render report ----
   tar_quarto(

### 2024-02-13 reports/aggressiveness.qmd
dag fixes

--- 2024-02-09 reports/aggressiveness.qmd
@@
 author: Nikhil Patel, MD
-format: html
+format:
+  html: 
+    mainfont: Figtree
+    monofont: Source Code Pro
+    fontsize: 0.97rem
+    embed-resources: true
     toc: true
+    toc-location: left
+    toc-depth: 5
+    toc-title: Contents
+    smooth-scroll: true
+    link-external-icon: true
+    link-external-newwindow: true
+    code-overflow: wrap
+    code-tools: true
+    highlight-style: nord
+    lightbox: true
 ---
 
+::: {.callout-note icon="false"}
+## Research question
+
+Do patients presenting with left hemispheric ICH receive less aggressive care (as measured by rates of neurosurgery, ventriculostomy, tracheostomy, PEG, and early withdrawal of life-sustaining therapies) as compared to those with right hemispheric ICH?
+
+:::
+
 ```{r setup, output=FALSE, message=FALSE, warning=FALSE}
-#| code-fold: false
+#| code-fold: true
+#| code-summary: "Loading Libraries and Other Code Setup"
 
 library(targets)
-library(readr)
-library(tidyr)
 library(skimr)
@@
 
+# theme customization
 theme_set(theme_minimal(base_family = "Figtree"))
 
-tar_load(ich)
-tar_load(non_surgery_trials)
-tar_load(surgery_trials)
-tar_load(table_1)
-tar_load(table_2)
+theme_set(
+  theme_dag() %+replace%
+    # also add some additional styling
+    theme(
+      legend.position = "bottom",
+      strip.text.x = element_text(margin = margin(2, 0, 2, 0, "mm"))
+    )
+)
+
+# load all targets objects
+tar_load_everything()
 
@@
 ### DAGs
+
+```{r}
+#| width: 1200
+#| height: 2000
+#| fig-align: center
+#|
+
+ggdag(aggressive_dag, use_labels = "label", text = FALSE)
+
+ggdag_adjustment_set(aggressive_dag, shadow = TRUE, use_labels = "label", text = FALSE)
+
+ggdag_adjustment_set(outcome_dag, shadow = TRUE, use_labels = "label", text = FALSE)
+
+aggressive_dag |>
+  ggdag_paths(shadow = TRUE, text = FALSE, use_labels = "label")
+
+tidy_dagitty(aggressive_dag)
+
+```

### 2024-02-13 R/dags.R

@@ -7,20 +7,24 @@ aggressive_dag_function <- function(x) {
     "ich_location", "ICH Location", 2.75, 1.75,
     "ich_volume", "ICH Volume", 1.25, 4,
     "ivh", "IVH", 1, 3,
-    "gcs_baseline", "Admission GCS", 4.5, 3.75
+    "gcs_baseline", "Admission GCS", 4.5, 3.75,
+    "stroke", "Previous Stroke", 1, 1,
+    "hypertension", "Hypertension", 1, 1,
+    "amyloid", "Amyloid Angiopathy", 1, 1
   )
 
   aggressiveness_node_labels <- aggressiveness_node_details$label
   names(aggressiveness_node_labels) <- aggressiveness_node_details$name
 
   dagify(
-    aggressive_care ~ ich_laterality + ivh + ich_volume + gcs_baseline + ich_location + age,
+    aggressive_care ~ ich_laterality + ivh + ich_volume + gcs_baseline + ich_location + age + stroke + amyloid,
     ivh ~ ich_location,
     gcs_baseline ~ ich_volume + age + ivh,
-    ich_location ~ age,
+    ich_location ~ amyloid + hypertension,
+    amyloid ~ age,
     exposure = "ich_laterality",
     outcome = "aggressive_care",
-    coords = aggressiveness_node_details,
+    latent = "amyloid",
     labels = aggressiveness_node_labels
   )
 }

### 2024-02-13 reports/aggressiveness.qmd
first dag complete?

--- 2024-02-09 reports/aggressiveness.qmd
@@
 author: Nikhil Patel, MD
-format: html
+format:
+  html: 
+    mainfont: Figtree
+    monofont: Source Code Pro
+    fontsize: 0.97rem
+    embed-resources: false
     toc: true
+    toc-location: left
+    toc-depth: 6
+    toc-expand: true
+    toc-title: Contents
+    smooth-scroll: true
+    link-external-icon: true
+    link-external-newwindow: true
+    code-overflow: wrap
+    code-tools: true
+    highlight-style: nord
 ---
 
+::: {.callout-note icon="false"}
+## Research question
+
+Do patients presenting with left hemispheric ICH receive less aggressive care (as measured by rates of neurosurgery, ventriculostomy, tracheostomy, PEG, and early withdrawal of life-sustaining therapies) as compared to those with right hemispheric ICH?
+
+:::
+
 ```{r setup, output=FALSE, message=FALSE, warning=FALSE}
-#| code-fold: false
-
 library(targets)
-library(readr)
-library(tidyr)
 library(skimr)
@@
 library(ggplot2)
+library(ggdag)
+library(dagitty)
 library(marginaleffects)
 library(modelsummary)
+library(MetBrewer)
 
+# theme customization
 theme_set(theme_minimal(base_family = "Figtree"))
 
+theme_set(
+  theme_dag(base_family = "Figtree") %+replace%
+    # also add some additional styling
+    theme(
+      legend.position = "bottom",
+      strip.text.x = element_text(margin = margin(2, 0, 2, 0, "mm"))
+    )
+)
+
+status_colors <- c(exposure = "#CE4441", outcome = "#004F63", latent = "grey80")
+
+# load all targets objects
+tar_load_everything()
 tar_load(ich)
@@
 tar_load(table_2)
+tar_load(aggressive_dag)
 
@@
 ### DAGs
+
+#### Aggressive Care
+
+We aim to estimate the **total causal effect of hemispheric laterality in intracerebral hemorrhage on the probability of receiving aggressive interventions**, as defined by the following:
+
+- neurosurgery
+- ventriculostomy
+- tracheostomy
+- PEG placement
+- early withdrawal of life-sustaining therapies (< 72 hours from presentation)
+
+::: {.callout-warning title="Issues to Address"}
+
+* Make sure other people think it's easy to read
+
+* Add any connections? Ask Nick, Matt, and others
+
+* Add transparency to colors to be able to see arrows
+
+* Fix position of arrows
+
+:::
+
+```{r dags}
+#| fig-align: center
+#| fig.height: 9
+
+aggressive_dag |>
+  tidy_dagitty() |>
+  node_status() |>
+  ggplot(aes(x = x, y = y, xend = xend, yend = yend)) +
+  geom_dag_point(aes(color = status)) +
+  geom_dag_edges_diagonal() +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 2,
+    color = "white", fontface = "bold", size = 4
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20")
+
+```

### 2024-02-13 R/dags.R

@@ -1,61 +1,33 @@
 aggressive_dag_function <- function(x) {
-  aggressiveness_node_details <- tribble(
+  node_details <- tribble(
     ~name, ~label, ~x, ~y,
-    "aggressive_care", "Aggressive Care", 5, 1,
-    "ich_laterality", "Hemispheric Laterality", 1, 1,
-    "age", "Age", 3.75, 2.5,
-    "ich_location", "ICH Location", 2.75, 1.75,
-    "ich_volume", "ICH Volume", 1.25, 4,
-    "ivh", "IVH", 1, 3,
-    "gcs_baseline", "Admission GCS", 4.5, 3.75,
-    "stroke", "Previous Stroke", 1, 1,
-    "hypertension", "Hypertension", 1, 1,
-    "amyloid", "Amyloid Angiopathy", 1, 1
+    "aggressive_care", "Aggressive Care", 4, 5,
+    "ich_laterality", "Hemispheric Laterality", 3, 0,
+    "age", "Age", 1, 3.25,
+    "ich_location", "ICH Location", 3, 4,
+    "ich_volume", "ICH Volume", 3, 5.5,
+    "ivh", "IVH", 3, 2,
+    "gcs_baseline", "Admission GCS", 3, 8,
+    "stroke", "Previous Stroke", 2, 5,
+    "hypertension", "Hypertension", 2, 2.25,
+    "amyloid", "Amyloid Angiopathy", 2, 6.25,
   )
 
-  aggressiveness_node_labels <- aggressiveness_node_details$label
-  names(aggressiveness_node_labels) <- aggressiveness_node_details$name
+  node_labels <- node_details$label
+  names(node_labels) <- node_details$name
 
   dagify(
     aggressive_care ~ ich_laterality + ivh + ich_volume + gcs_baseline + ich_location + age + stroke + amyloid,
     ivh ~ ich_location,
-    gcs_baseline ~ ich_volume + age + ivh,
+    gcs_baseline ~ ich_volume + age + ivh + ich_location,
     ich_location ~ amyloid + hypertension,
     amyloid ~ age,
+    hypertension ~ age,
+    stroke ~ hypertension + age + amyloid,
+    coords = node_details,
     exposure = "ich_laterality",
     outcome = "aggressive_care",
     latent = "amyloid",
-    labels = aggressiveness_node_labels
-  )
-}
-
-outcome_dag_function <- function(x) {
-  outcome_node_details <- tribble(
[…]

### 2024-02-13 _targets.R

@@ -58,7 +58,6 @@ tar_plan(
 
   ## DAGs ----
   aggressive_dag = aggressive_dag_function(x),
-  outcome_dag = outcome_dag_function(x),
 
   ## Render report ----
   tar_quarto(

### 2024-02-13 reports/aggressiveness.qmd
dags done!

@@
     fontsize: 0.97rem
-    embed-resources: false
+    embed-resources: true
     toc: true
@@
     code-tools: true
+    code-fold: true
     highlight-style: nord
@@
 tar_load_everything()
-tar_load(ich)
-tar_load(non_surgery_trials)
-tar_load(surgery_trials)
-tar_load(table_1)
-tar_load(table_2)
-tar_load(aggressive_dag)
 
@@
 
-table_1
+table_1_non_surgery
 
@@
 
+::: {.callout-important title="Other Considerations"}
+
+Need to point out in table that `days_mechanical_ventilation` is only available in ATACH2 data. Have to clarify if `tracheostomy` and `peg` are included in ATACH2 (def included in ERICH)
+
+:::
+
 ```{r}
@@
 
-table_2
+table_2_non_surgery
 
@@
 
-Add `DNR` in addition to `comfort_care`?
+Add `DNR` in addition to `comfort_care` in first analysis?
 
 :::
+
+We aim to estimate the **total causal effect of hemispheric laterality in intracerebral hemorrhage on the probability of receiving aggressive interventions**, as defined by the following:
 
@@
 * `tracheostomy`: logistic
+* `peg`: logistic
 * `comfort_care`: logistic (days to comfort care, time to event?)
 
-### DAGs
-
-#### Aggressive Care
-
-We aim to estimate the **total causal effect of hemispheric laterality in intracerebral hemorrhage on the probability of receiving aggressive interventions**, as defined by the following:
-
-- neurosurgery
-- ventriculostomy
-- tracheostomy
-- PEG placement
-- early withdrawal of life-sustaining therapies (< 72 hours from presentation)
+The directed acyclic graph (DAG) is described below.
 
@@
 
-* Make sure other people think it's easy to read
+* Make it easier to [read](https://github.com/sakeefkarim/intro_ggdag_script/blob/main/code/code_intro_ggdag.R)!
 
@@
 
+* Include `symptom_onset_to_ED` in the DAG (this is known to differ in patients with ischemic stroke)
+
 :::
 
-```{r dags}
+### DAG
+
+```{r aggressive dag}
 #| fig-align: center
@@
   geom_dag_point(aes(color = status)) +
-  geom_dag_edges_diagonal() +
+  geom_dag_edges() +
   geom_dag_label_repel(aes(label = label, fill = status),
-    seed = 2,
-    color = "white", fontface = "bold", size = 4
+    seed = 51,
+    color = "white", fontface = "bold", size = 4.5
   ) +
@@
 ```
+
+#### DAG-Implied Adjustment Sets
+
+Since there is no variable that affects `hemispheric_laterality` in the DAG, the minimally sufficient adjustment set does not include any variables for the estimand of interest (total effect of `hemispheric_laterality`).
+
+The canonical adjustment set includes `age`, `gcs_baseline`, `ich_location`, `ich_volume`, `ivh`, and `stroke`.
+
+We include the competing exposures in the canonical adjustment set (e.g., `age`, `gcs_baseline`, `ich_location`, `ich_volume`, `ivh`, and `stroke`) as covariates in our final model to increase precision, and also run the minimal adjustment set as a sensitivity analysis.
+
+```{r aggressive dag adjustment sets}
+#| code-fold: false
+#| fig-align: center
+#| fig.height: 9
+
+adjustmentSets(
+  aggressive_dag,
+  type = "minimal",
+  effect = "total",
+  max.results = Inf
+)
+
+adjustmentSets(
+  aggressive_dag,
+  type = "canonical",
+  effect = "total",
+  max.results = Inf
+)
+
+```
+
+### Priors
+
+### Posterior Simulation
+
+### Analysis

### 2024-02-13 reports/outcomes.qmd
dags done!

--- 2024-02-13 reports/aggressiveness.qmd
@@
 ---
-title: Differences in care for patients experiencing left vs. right hemisphere ICH
+title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
 author: Nikhil Patel, MD
@@
 
-Do patients presenting with left hemispheric ICH receive less aggressive care (as measured by rates of neurosurgery, ventriculostomy, tracheostomy, PEG, and early withdrawal of life-sustaining therapies) as compared to those with right hemispheric ICH?
+Do patients presenting with left hemispheric ICH have different functional outcomes (as measured by mRs and EuroQoL at 90 days) as compared to those with right hemispheric ICH?
 
@@
 ::: {.callout-warning title="Issues to Address"}
-`htn` is missing mostly all information. It is also listed as a character variable.
 
-All other co-morbidities are missing information; if this is the case, should default to "No".
+?
 
-`ivh` is also missing some values, which should also default to "No"?
-
-Any "Yes/No" variable should not have any missing information, should just default to "No"
-
-Verify accuracy/validity of `time_symptoms_to_ed`, `sbp_baseline`,
-
-Fix/drop unused levels in `race`, and fix "American Indian"
-
-Fix "unknown" category in `ethnicity` to "Not Hispanic or Latino"
-
-Fix/drop unused levels in `ich_location` and `study`
 :::
@@
 
-glimpse(non_surgery_trials)
-skim(non_surgery_trials)
+glimpse(ich)
+skim(ich)
 
@@
 
-Collapse unused levels in `race`, `ethnicity`, `ich_location` and `study`
-
-Change order of the displayed factors to frequency
+?
 
@@
 
-table_1
+table_1_all
 
@@
 
-```{r}
-#| fig-align: center
+::: {.callout-important title="Other Considerations"}
 
-table_2
+?
 
-```
+:::
 
@@
 
-Construct DAGs for **each** outcome (e.g., each model)!
-
-Add `DNR` in addition to `comfort_care`?
+?
 
@@
 
-* `neurosurgery_evac`: logistic
-* `ventriculostomy`: logistic
-* `days_mechanical_ventilation`: poisson (cumulative logit?)
-* `tracheostomy`: logistic
-* `comfort_care`: logistic (days to comfort care, time to event?)
+### Outcomes
 
-### DAGs
+We aim to estimate the total direct effect of hemispheric laterality in intracerebral hemorrhage on the probability of a good functional outcome, as defined by the following:
 
-#### Aggressive Care
+* `mrs_90`: modified Rankin at 90 days: cumulative logit
+* `euro_vas_90`: EuroQOL VAS at 90 days: beta regression
+* `euro_90`: all 5 domains of the EuroQOL at 90 days: cumulative logit
 
-We aim to estimate the **total causal effect of hemispheric laterality in intracerebral hemorrhage on the probability of receiving aggressive interventions**, as defined by the following:
+#### DAG
 
-- neurosurgery
-- ventriculostomy
-- tracheostomy
-- PEG placement
-- early withdrawal of life-sustaining therapies (< 72 hours from presentation)
-
-::: {.callout-warning title="Issues to Address"}
-
-* Make sure other people think it's easy to read
-
-* Add any connections? Ask Nick, Matt, and others
-
-* Add transparency to colors to be able to see arrows
-
-* Fix position of arrows
-
-:::
-
-```{r dags}
+```{r outcomes dag}
 #| fig-align: center
@@
 
-aggressive_dag |>
+outcomes_dag |>
   tidy_dagitty() |>
@@
 ```
+
+#### DAG-Implied Adjustment Sets
+
+Similar to above, there are no
+
+```{r outcomes dag adjustment sets}
+#| code-fold: false
+#| fig-align: center
+#| fig.height: 10
+#| fig.width: 20
+
+adjustmentSets(
+  outcomes_dag,
+  type = "minimal",
+  effect = "direct",
+  max.results = Inf
+)
+
+adjustmentSets(
+  outcomes_dag,
+  type = "canonical",
+  effect = "total",
+  max.results = Inf
+)
+
+```

### 2024-02-13 R/dags.R

@@ -1,16 +1,16 @@
 aggressive_dag_function <- function(x) {
   node_details <- tribble(
     ~name, ~label, ~x, ~y,
-    "aggressive_care", "Aggressive Care", 4, 5,
-    "ich_laterality", "Hemispheric Laterality", 3, 0,
-    "age", "Age", 1, 3.25,
-    "ich_location", "ICH Location", 3, 4,
-    "ich_volume", "ICH Volume", 3, 5.5,
-    "ivh", "IVH", 3, 2,
+    "aggressive_care", "Aggressive Care", 5, 5,
+    "ich_laterality", "Hemispheric Laterality", 3.25, 0,
+    "age", "Age", 1, 3.75,
+    "ich_location", "ICH Location", 3, 3.75,
+    "ich_volume", "ICH Volume", 3.5, 5.5,
+    "ivh", "IVH", 3.25, 2,
     "gcs_baseline", "Admission GCS", 3, 8,
     "stroke", "Previous Stroke", 2, 5,
-    "hypertension", "Hypertension", 2, 2.25,
-    "amyloid", "Amyloid Angiopathy", 2, 6.25,
+    "hypertension", "Hypertension", 2, 2,
+    "amyloid", "Amyloid Angiopathy", 2, 7,
   )
 
   node_labels <- node_details$label
@@ -28,6 +28,46 @@ aggressive_dag_function <- function(x) {
     exposure = "ich_laterality",
     outcome = "aggressive_care",
     latent = "amyloid",
+    labels = node_labels
+  )
+}
+
+outcomes_dag_function <- function(x) {
+  node_details <- tribble(
+    ~name, ~label, ~x, ~y,
+    "functional_outcomes", "Functional Outcomes", 8, 5,
+    "ich_laterality", "Hemispheric Laterality", 3, 0,
+    "age", "Age", 1, 3.75,
+    "ich_location", "ICH Location", 2.75, 3.75,
+    "ich_volume", "ICH Volume", 3, 6,
+    "ivh", "IVH", 3, 2,
+    "gcs_baseline", "Admission GCS", 3, 8,
+    "neurosurgery", "Neurosurgical Intervention", 6, 4,
+    "wlst", "Early WLST", 5, 2,
+    "rehab", "Participation in Rehab", 7, 1,
+    "stroke", "Previous stroke", 2, 5,
+    "hypertension", "Hypertension", 2, 2,
+    "amyloid", "Amyloid Angiopathy", 2, 7
+  )
+
[…]

### 2024-02-13 R/exploratory_data_analysis.R

@@ -48,7 +48,7 @@ table_1_function <- function(x) {
     bold_labels()
 }
 
-table_2_function <- function(x) {
+table_2_aggressive_function <- function(x) {
   table_two_vars <- c(
     "neurosurgery_evac",
     "evd",

### 2024-02-13 _targets.R

@@ -53,15 +53,25 @@ tar_plan(
   surgery_trials = ich |> filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III"),
 
   ## Exploratory data analysis ----
-  table_1 = table_1_function(non_surgery_trials),
-  table_2 = table_2_function(non_surgery_trials),
+  table_1_non_surgery = table_1_function(non_surgery_trials),
+  table_2_non_surgery = table_2_aggressive_function(non_surgery_trials),
+  table_1_all = table_1_function(ich),
 
   ## DAGs ----
   aggressive_dag = aggressive_dag_function(x),
+  outcomes_dag = outcomes_dag_function(x),
+
+  ## Priors ----
+
+
 
   ## Render report ----
   tar_quarto(
-    report,
+    aggressive,
     "reports/aggressiveness.qmd"
+  ),
+  tar_quarto(
+    outcomes,
+    "reports/outcomes.qmd"
   )
 )

### 2024-02-14 reports/aggressiveness.qmd
neurosurgery models done

@@
 library(ggplot2)
+library(ggdist)
+library(gghalves)
 library(ggdag)
+library(scales)
 library(dagitty)
@@
 library(modelsummary)
-library(MetBrewer)
 
@@
 # load all targets objects
-tar_load_everything()
+# tar_load_everything()
+tar_load(non_surgery_trials)
+tar_load(table_1_non_surgery)
+tar_load(table_2_non_surgery)
+tar_load(aggressive_dag)
 
@@
 
-### Priors
+### Neurosurgical Intervention
+
+There are two models we will run here: 1) the minimally-sufficient adjustment model, 2) the canonical adjustment model
+
+#### Model
+
+```{r}
+
+tar_load(neurosurgery_evac_min)
+
+predictions(neurosurgery_evac_min, by = "ich_laterality")
+
+plot_predictions(neurosurgery_evac_min, condition = "ich_laterality") +
+  labs(x = "ICH Laterality", y = "Probability of Neurosurgery")
+
+# Format things as percentage points
+label_pp <- label_number(
+  accuracy = 1, scale = 100,
+  suffix = " pp.", style_negative = "minus"
+)
+
+label_pp_tiny <- label_number(
+  accuracy = 0.01, scale = 100,
+  suffix = " pp.", style_negative = "minus"
+)
+
+theme_clean <- function() {
+  theme_minimal(base_family = "Barlow Semi Condensed") +
+    theme(
+      panel.grid.minor = element_blank(),
+      plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
+      axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
+      strip.text = element_text(
+        family = "BarlowSemiCondensed-Bold",
+        size = rel(1), hjust = 0
+      ),
+      strip.background = element_rect(fill = "grey80", color = NA)
+    )
+}
+
+avg_comparisons(neurosurgery_evac_min, variables = "ich_laterality") |>
+  posterior_draws() |>
+  ggplot(aes(x = draw)) +
+  stat_halfeye(
+    .width = c(0.8, 0.95), point_interval = "median_hdi",
+    fill = "#bc3032"
+  ) +
+  scale_x_continuous(labels = label_pp) +
+  labs(
+    x = "Effect of hemispheric laterality\n on the probability of receiving neurosurgical intervention",
+    y = NULL, caption = "80% and 95% credible intervals shown in black"
+  ) +
+  theme_clean()
+
+```
+
+```{r canonical}
+tar_load(neurosurgery_evac_can)
+
+predictions(neurosurgery_evac_can, by = "ich_laterality")
+
+plot_predictions(neurosurgery_evac_can, condition = "ich_laterality") +
+  labs(x = "ICH Laterality", y = "Probability of Neurosurgery")
+
+# Format things as percentage points
+label_pp <- label_number(
+  accuracy = 1, scale = 100,
+  suffix = " pp.", style_negative = "minus"
+)
+
+label_pp_tiny <- label_number(
+  accuracy = 0.01, scale = 100,
+  suffix = " pp.", style_negative = "minus"
+)
+
+theme_clean <- function() {
+  theme_minimal(base_family = "Barlow Semi Condensed") +
+    theme(
+      panel.grid.minor = element_blank(),
+      plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
+      axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
+      strip.text = element_text(
+        family = "BarlowSemiCondensed-Bold",
+        size = rel(1), hjust = 0
+      ),
+      strip.background = element_rect(fill = "grey80", color = NA)
+    )
+}
+
+avg_comparisons(neurosurgery_evac_can, variables = "ich_laterality") |>
+  posterior_draws() |>
+  ggplot(aes(x = draw)) +
+  stat_halfeye(
+    .width = c(0.8, 0.95), point_interval = "median_hdi",
+    fill = "#bc3032"
+  ) +
+  scale_x_continuous(labels = label_pp) +
+  labs(
+    x = "Effect of hemispheric laterality\n on the probability of receiving neurosurgical intervention",
+    y = NULL, caption = "80% and 95% credible intervals shown in black"
+  ) +
+  theme_clean()
+
+```
 

### 2024-02-14 reports/outcomes.qmd
neurosurgery models done

@@
 # load all targets objects
-tar_load_everything()
+# tar_load_everything()
+tar_load(table_1_all)
+tar_load(outcomes_dag)
+tar_load(ich)
 
@@
   geom_dag_point(aes(color = status)) +
-  geom_dag_edges() +
+  geom_dag_edges_diagonal() +
   geom_dag_label_repel(aes(label = label, fill = status),

### 2024-02-14 R/models.R

@@ -1,0 +1,46 @@
+# neurosurgical intervention
+neurosurgery_evac_func <- function(x) {
+  BAYES_SEED <- 07204
+
+  options(
+    mc.cores = 4,
+    brms.backend = "cmdstanr"
+  )
+
+  priors <- c(
+    set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b")
+  )
+
+  brm(
+    neurosurgery_evac ~ ich_laterality + (1 | study),
+    data = x,
+    family = bernoulli(link = "logit"),
+    prior = priors,
+    init = 0,
+    chains = 4, iter = 4000, seed = BAYES_SEED
+  )
+}
+
+neurosurgery_evac_func_canon <- function(x) {
+  BAYES_SEED <- 07204
+
+  options(
+    mc.cores = 4,
+    brms.backend = "cmdstanr"
+  )
+
+  priors <- c(
+    set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b")
+  )
+
+  brm(
+    neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + stroke + (1 | study),
+    data = x,
+    family = bernoulli(link = "logit"),
+    prior = priors,
+    init = 0,
+    chains = 4, iter = 4000, seed = BAYES_SEED
+  )
+}

### 2024-02-14 R/priors.R

@@ -1,3 +1,22 @@
 # Variables to include: age, gcs_baseline, hypertension, ich_location, ich_volume, ivh, stroke
 # Outcomes: neurosurgery, evd, days_mechanical_ventilation, tracheostomy, peg, comfort_care
 
+# neurosurgical intervention
+neurosurgery_evac_min <- function(x) {
+  BAYES_SEED <- 07204
+
+  priors <- c(
+    set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b")
+  )
+
+  brm(
+    neurosurgery_evac ~ ich_laterality,
+    data = non_surgery_trials,
+    family = bernoulli(link = "logit"),
+    prior = priors,
+    init = 0,
+    chains = 4, iter = 4000, seed = BAYES_SEED,
+    backend = "cmdstanr"
+  )
+}

### 2024-02-14 _targets.R

@@ -34,7 +34,7 @@ source("R/packages.R")
 source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
 source("R/dags.R")
-
+source("R/models.R")
 
 # Pipeline ----
 tar_plan(
@@ -61,11 +61,11 @@ tar_plan(
   aggressive_dag = aggressive_dag_function(x),
   outcomes_dag = outcomes_dag_function(x),
 
-  ## Priors ----
+  ## Models ----
+  neurosurgery_evac_min = neurosurgery_evac_func(non_surgery_trials),
+  neurosurgery_evac_can = neurosurgery_evac_func_canon(non_surgery_trials),
 
-
-
-  ## Render report ----
+  ## Reports ----
   tar_quarto(
     aggressive,
     "reports/aggressiveness.qmd"

### 2024-02-14 reports/aggressiveness.qmd
data cleaning fixes

@@
 ::: {.callout-warning title="Issues to Address"}
-`htn` is missing mostly all information. It is also listed as a character variable.
-
-All other co-morbidities are missing information; if this is the case, should default to "No".
-
-`ivh` is also missing some values, which should also default to "No"?
-
-Any "Yes/No" variable should not have any missing information, should just default to "No"
 
@@
 
-Fix/drop unused levels in `race`, and fix "American Indian"
-
-Fix "unknown" category in `ethnicity` to "Not Hispanic or Latino"
-
-Fix/drop unused levels in `ich_location` and `study`
 :::

### 2024-02-14 R/testing.r

@@ -1,0 +1,15 @@
+priors <- c(
+  set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
+  set_prior("normal(0, 1)", class = "b")
+)
+
+model <- brm(
+  neurosurgery_evac ~ ich_laterality + (1 | study),
+  data = non_surgery_trials,
+  family = bernoulli(link = "logit"),
+  prior = priors,
+  init = 0,
+  chains = 4, iter = 4000, seed = BAYES_SEED
+)
+
+model

### 2024-02-14 reports/aggressiveness.qmd
things work?

@@
 # tar_load_everything()
-tar_load(non_surgery_trials)
-tar_load(table_1_non_surgery)
-tar_load(table_2_non_surgery)
-tar_load(aggressive_dag)
+tar_load(ich_aggressive)
+tar_load(table_1_aggressive)
+tar_load(table_2_aggressive)
+tar_load(dag_aggressive)
 
@@
 
-glimpse(non_surgery_trials)
-skim(non_surgery_trials)
+glimpse(ich_aggressive)
+skim(ich_aggressive)
 
@@
 
-table_1_non_surgery
+table_1_aggressive
 
@@
 
-table_2_non_surgery
+table_2_aggressive
 
@@
 
-aggressive_dag |>
+dag_aggressive |>
   tidy_dagitty() |>

### 2024-02-14 R/packages.R

@@ -14,3 +14,4 @@ library(cmdstanr)
 library(brms)
 library(marginaleffects)
 library(modelsummary)
+library(scales)

### 2024-02-14 R/priors.R

@@ -1,22 +1,17 @@
-# Variables to include: age, gcs_baseline, hypertension, ich_location, ich_volume, ivh, stroke
-# Outcomes: neurosurgery, evd, days_mechanical_ventilation, tracheostomy, peg, comfort_care
-
-# neurosurgical intervention
-neurosurgery_evac_min <- function(x) {
-  BAYES_SEED <- 07204
-
+# Neurosurgical Intervention: Minimal
+neurosurgery_min_prior_f <- function(x) {
   priors <- c(
-    set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
+    set_prior("normal(-2.19, 0.7)", class = "Intercept"),
     set_prior("normal(0, 1)", class = "b")
   )
 
-  brm(
+  mod_prior <- brm(
     neurosurgery_evac ~ ich_laterality,
-    data = non_surgery_trials,
+    data = x,
     family = bernoulli(link = "logit"),
     prior = priors,
-    init = 0,
-    chains = 4, iter = 4000, seed = BAYES_SEED,
-    backend = "cmdstanr"
+    sample_prior = "only"
   )
+
+  return(mod_prior)
 }

### 2024-02-14 R/testing.r

@@ -1,15 +1,0 @@
-priors <- c(
-  set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
-  set_prior("normal(0, 1)", class = "b")
-)
-
-model <- brm(
-  neurosurgery_evac ~ ich_laterality + (1 | study),
-  data = non_surgery_trials,
-  family = bernoulli(link = "logit"),
-  prior = priors,
-  init = 0,
-  chains = 4, iter = 4000, seed = BAYES_SEED
-)
-
-model

### 2024-02-14 _targets.R

@@ -35,6 +35,7 @@ source("R/packages.R")
 source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
 source("R/dags.R")
+source("R/priors.R")
 source("R/models.R")
 
 # Pipeline ----
@@ -49,22 +50,24 @@ tar_plan(
 
   ## Select, filter, and clean data ----
   selected_data = select_variables(imported_data),
-  ich = filter_variables(selected_data),
-  non_surgery_trials = ich |> filter(study == "ERICH" | study == "ATACH-2") |> droplevels(),
-  surgery_trials = ich |> filter(study == "MISTIE2" | study == "MISTIE-3" | study == "CLEAR III") |> droplevels(),
+  ich_all = filter_variables(selected_data),
+  ich_aggressive = ich_all |> filter(study == "ERICH" | study == "ATACH-2") |> droplevels(),
 
   ## Exploratory data analysis ----
-  table_1_non_surgery = table_1_function(non_surgery_trials),
-  table_2_non_surgery = table_2_aggressive_function(non_surgery_trials),
-  table_1_all = table_1_function(ich),
+  table_1_aggressive = table_1_function(ich_aggressive),
+  table_2_aggressive = table_2_aggressive_function(ich_aggressive),
+  table_1_outcomes = table_1_function(ich_all),
 
   ## DAGs ----
-  aggressive_dag = aggressive_dag_function(x),
-  outcomes_dag = outcomes_dag_function(x),
+  dag_aggressive = aggressive_dag_function(x),
+  dag_outcomes = outcomes_dag_function(x),
+
+  ## Priors ----
+  neurosurgery_min_priors = neurosurgery_min_prior_f(ich_aggressive),
 
   ## Models ----
-  neurosurgery_evac_min = neurosurgery_evac_func(non_surgery_trials),
-  neurosurgery_evac_can = neurosurgery_evac_func_canon(non_surgery_trials),
+  neurosurgery_evac_min = neurosurgery_evac_func(ich_aggressive),
+  neurosurgery_evac_can = neurosurgery_evac_func_canon(ich_aggressive),
 
   ## Reports ----
   tar_quarto(

### 2024-02-15 reports/aggressiveness.qmd
trying to set priors

@@
     monofont: Source Code Pro
-    fontsize: 0.97rem
+    fontsize: 1rem
     embed-resources: true
@@
 ```{r setup, output=FALSE, message=FALSE, warning=FALSE}
+
 library(targets)
@@
 library(gtsummary)
+library(tibble)
 library(ggplot2)
@@
 status_colors <- c(exposure = "#CE4441", outcome = "#004F63", latent = "grey80")
+
+theme_gtsummary_compact()
 
@@
 
-::: {.callout-warning title="Issues to Address"}
-
-Collapse unused levels in `race`, `ethnicity`, `ich_location` and `study`
-
-Change order of the displayed factors to frequency
-
-:::
-
 ```{r}
+#| fig-align: center
 
@@
 adjustmentSets(
-  aggressive_dag,
+  dag_aggressive,
   type = "minimal",
@@
 adjustmentSets(
-  aggressive_dag,
+  dag_aggressive,
   type = "canonical",
@@
 ```
-
-### Neurosurgical Intervention
-
-There are two models we will run here: 1) the minimally-sufficient adjustment model, 2) the canonical adjustment model
-
-#### Model
-
-```{r}
-
-tar_load(neurosurgery_evac_min)
-
-predictions(neurosurgery_evac_min, by = "ich_laterality")
-
-plot_predictions(neurosurgery_evac_min, condition = "ich_laterality") +
-  labs(x = "ICH Laterality", y = "Probability of Neurosurgery")
-
-# Format things as percentage points
-label_pp <- label_number(
-  accuracy = 1, scale = 100,
-  suffix = " pp.", style_negative = "minus"
-)
-
-label_pp_tiny <- label_number(
-  accuracy = 0.01, scale = 100,
-  suffix = " pp.", style_negative = "minus"
-)
-
-theme_clean <- function() {
-  theme_minimal(base_family = "Barlow Semi Condensed") +
-    theme(
-      panel.grid.minor = element_blank(),
-      plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
-      axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
-      strip.text = element_text(
-        family = "BarlowSemiCondensed-Bold",
-        size = rel(1), hjust = 0
-      ),
-      strip.background = element_rect(fill = "grey80", color = NA)
-    )
-}
-
-avg_comparisons(neurosurgery_evac_min, variables = "ich_laterality") |>
-  posterior_draws() |>
-  ggplot(aes(x = draw)) +
-  stat_halfeye(
-    .width = c(0.8, 0.95), point_interval = "median_hdi",
-    fill = "#bc3032"
-  ) +
-  scale_x_continuous(labels = label_pp) +
-  labs(
-    x = "Effect of hemispheric laterality\n on the probability of receiving neurosurgical intervention",
-    y = NULL, caption = "80% and 95% credible intervals shown in black"
-  ) +
-  theme_clean()
-
-```
-
-```{r canonical}
-tar_load(neurosurgery_evac_can)
-
-predictions(neurosurgery_evac_can, by = "ich_laterality")
-
-plot_predictions(neurosurgery_evac_can, condition = "ich_laterality") +
-  labs(x = "ICH Laterality", y = "Probability of Neurosurgery")
-
-# Format things as percentage points
-label_pp <- label_number(
-  accuracy = 1, scale = 100,
-  suffix = " pp.", style_negative = "minus"
-)
-
-label_pp_tiny <- label_number(
-  accuracy = 0.01, scale = 100,
-  suffix = " pp.", style_negative = "minus"
-)
-
-theme_clean <- function() {
-  theme_minimal(base_family = "Barlow Semi Condensed") +
-    theme(
-      panel.grid.minor = element_blank(),
-      plot.title = element_text(family = "BarlowSemiCondensed-Bold"),
-      axis.title = element_text(family = "BarlowSemiCondensed-Medium"),
-      strip.text = element_text(
-        family = "BarlowSemiCondensed-Bold",
-        size = rel(1), hjust = 0
-      ),
-      strip.background = element_rect(fill = "grey80", color = NA)
-    )
-}
-
-avg_comparisons(neurosurgery_evac_can, variables = "ich_laterality") |>
-  posterior_draws() |>
-  ggplot(aes(x = draw)) +
-  stat_halfeye(
-    .width = c(0.8, 0.95), point_interval = "median_hdi",
-    fill = "#bc3032"
-  ) +
-  scale_x_continuous(labels = label_pp) +
-  labs(
-    x = "Effect of hemispheric laterality\n on the probability of receiving neurosurgical intervention",
-    y = NULL, caption = "80% and 95% credible intervals shown in black"
-  ) +
-  theme_clean()
-
-```
-
-### Posterior Simulation
-
-### Analysis

### 2024-02-15 reports/outcomes.qmd
things work?
trying to set priors

--- 2024-02-13 reports/outcomes.qmd
@@
 # load all targets objects
-tar_load_everything()
+# tar_load_everything()
+tar_load(table_1_outcomes)
+tar_load(dag_outcomes)
+tar_load(ich_all)
 
@@
 
-glimpse(ich)
-skim(ich)
+glimpse(ich_all)
+skim(ich_all)
 
@@
 
-table_1_all
+table_1_outcomes
 
@@
 
-outcomes_dag |>
+dag_outcomes |>
   tidy_dagitty() |>
@@
 adjustmentSets(
-  outcomes_dag,
+  dag_outcomes,
   type = "minimal",
@@
 adjustmentSets(
-  outcomes_dag,
+  dag_outcomes,
   type = "canonical",

### 2024-02-15 reports/priors.qmd
trying to set priors

---
title: Differences in care for patients experiencing left vs. right hemisphere ICH
subtitle: Prior Predictive Checks
author: Nikhil Patel, MD
format:
  html: 
    mainfont: Figtree
    monofont: Source Code Pro
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 6
    toc-title: Contents
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
    highlight-style: nord
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(tidyverse)
library(glue)
library(marginaleffects)
library(modelsummary)
library(ggdist)
library(gghalves)
library(scales)
library(systemfonts)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

# Format things as percentage points
label_pp <- label_number(
  accuracy = 1, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

label_pp_tiny <- label_number(
  accuracy = 0.01, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

# load targets objects
tar_load(model_parameters)
tar_load(prior_neurosurgery_minimal)
tar_load(prior_neurosurgery_canonical)
tar_load(prior_evd_minimal)
tar_load(prior_evd_canonical)

# prior functions

prior_printing <- function(x, y, var) {
  p1 <- model_parameters$priors[x]

  p2 <- predictions(y, by = "ich_laterality") |>
    as_tibble()

  p3 <- avg_comparisons(y, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    stat_halfeye(
      .width = c(0.8, 0.95), point_interval = "median_hdi",
      fill = "#bc3032"
    ) +
    scale_x_continuous(labels = label_pp) +
    labs(
      title = "Prior Predictive Check",
      x = paste("Effect of hemispheric laterality\n on the probability of", {{ var }}),
      y = NULL, caption = "80% and 95% credible intervals shown in black"
    )

  return(list(
    priors = p1,
    predictions = p2,
    plot = p3
  ))
}

```

## Prior Predictive Checks

### Neurosurgical Intervention

#### Minimal Model

```{r}

prior_printing("neurosurgery_minimal", prior_neurosurgery_minimal, "receiving neurosurgical intervention")

```

#### Canonical Model

```{r}

prior_printing("neurosurgery_canonical", prior_neurosurgery_canonical, "receiving neurosurgical intervention")

```

### EVD

#### Minimal Model

```{r}

prior_printing("evd_minimal", prior_evd_minimal, "receiving an EVD")

```

#### Canonical Model

```{r}

prior_printing("evd_canonical", prior_evd_canonical, "receiving an EVD")

prior_neurosurgery_minimal

```

### 2024-02-15 R/functions_sandbox.R

@@ -1,0 +1,107 @@
+library(purrr)
+library(repurrrsive)
+
+prior_f <- function(formula, priors, data, family) {
+  mod_prior <- brm(formula,
+    data = data,
+    family = family,
+    prior = priors,
+    sample_prior = "only"
+  )
+  return(mod_prior)
+}
+
+prior_arguments <- list(
+  formula = neurosurgery_evac ~ ich_laterality,
+  priors = c(
+    set_prior("normal(-2.2, 0.7)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b")
+  ),
+  data = ich_aggressive,
+  family = bernoulli(link = "logit")
+)
+
+prior_f(prior_arguments$formula, prior_arguments$priors, prior_arguments$data, prior_arguments$family)
+
+wesanderson
+
+
+## purrr tutorial
+
+map(prior_arguments, "formula")
+
+map(got_chars[5:9], 1)
+map(wesanderson, 1)
+map(prior_arguments, 1)
+
+got_chars |>
+  map(3)
+
+names(got_chars)
+str(got_chars)
+
+map_lgl(got_chars, "alive")
+map_chr(got_chars, as.character("id"))
+
+culture <- factor(map_vec(got_chars, "culture"))
+unique(culture)
+
+map(wesanderson$name, print)
+
+tibble <- names(wesanderson)
+
+map_chr(wesanderson, 1)
+
+wes_tibble <- tibble(
+  movie = names(wesanderson),
+  color1 = map_chr(wesanderson, 1),
+  color2 = map_chr(wesanderson, 2)
+)
+
+map(shades, print)
+
+colors
+
+
+models <- list(
+  simple = "lifeexp ~ pop + gdppercap",
+  medium = "lifeexp ~ pop + gdppercap + continent + year",
+  more = "lifeexp ~ pop + gdppercap + country + year",
+  woah = "lifeexp ~ pop + gdppercap + year*country"
+)
+
+model_frame <- data_frame(model = models) %>%
+  mutate(model_name = names(model))
+
+model_frame <- model_frame %>%
+  mutate(model = map(model, as.formula))
+
+model_frame
+
+prior_formulas <- list(
+
+)
+
+
+lhr_setup <- function() {
+  # Settings
+  CHAINS <- 4
+  ITER <- 2000
+  WARMUP <- 1000
+  BAYES_SEED <- 4045 # From random.org
[…]

### 2024-02-15 R/priors.R

@@ -1,17 +1,70 @@
-# Neurosurgical Intervention: Minimal
-neurosurgery_min_prior_f <- function(x) {
-  priors <- c(
-    set_prior("normal(-2.19, 0.7)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b")
+model_parameters_function <- function(ich_aggressive) {
+  formulas <- list(
+    neurosurgery_minimal = neurosurgery_evac ~ ich_laterality,
+    neurosurgery_canonical = neurosurgery_evac ~ ich_laterality + age + htn + stroke + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    evd_minimal = evd ~ ich_laterality,
+    evd_canonical = evd ~ ich_laterality + age + htn + stroke + gcs_baseline + ich_location + ich_volume_baseline + ivh
   )
 
-  mod_prior <- brm(
-    neurosurgery_evac ~ ich_laterality,
-    data = x,
-    family = bernoulli(link = "logit"),
-    prior = priors,
-    sample_prior = "only"
+  priors <- list(
+    neurosurgery_minimal = c(
+      set_prior("normal(-2.2, 0.1)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    neurosurgery_canonical = c(
+      set_prior("normal(-2.2, 1)", class = "Intercept"),
+      set_prior("normal(0, 1.5)", class = "b")
+    ),
+    evd_minimal = c(
+      set_prior("normal(-2.2, 0.7)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    evd_canonical = c(
+      set_prior("normal(-2.2, 1)", class = "Intercept"),
+      set_prior("normal(0, 1.5)", class = "b")
+    )
   )
 
-  return(mod_prior)
+  data <- list(
+    neurosurgery_minimal = ich_aggressive,
+    neurosurgery_canonical = ich_aggressive,
+    evd_minimal = ich_aggressive,
+    evd_canonical = ich_aggressive
+  )
+
+  family <- list(
+    neurosurgery_minimal = bernoulli(link = "logit"),
+    neurosurgery_canonical = bernoulli(link = "logit"),
+    evd_minimal = bernoulli(link = "logit"),
+    evd_canonical = bernoulli(link = "logit")
+  )
+
+  model_parameters <- tibble(
+    names = names(formulas),
+    formula = formulas,
[…]

### 2024-02-15 R/testing.r

@@ -1,0 +1,66 @@
+Steps needed to build a bunch of prior models:
+
+# Aggressive Care
+
+outcomes: neurosurgery, evd, comfort_care, mech_vent, trach, and peg
+
+All but mech_vent are logistic models
+
+1) define formulas (simple and canonical)
+2) define priors (will this be different for each?)
+3) define data (will be the same for all)
+4) define family (will be the same except for mech_vent)
+
+## formulas
+
+formulas_aggressive <- list(
+  neurosurgery_simple = neurosurgery_evac ~ ich_laterality,
+  neurosurgery_canonical = neurosurgery_evac ~ ich_laterality + age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
+
+)
+
+is.atomic(formulas_aggressive[["neurosurgery_simple"]])
+
+prior_f(prior_arguments)
+
+str(prior_arguments)
+### neurosurgery
+
+neurosurgery_evac ~ ich_laterality
+
+neurosurgery_evac ~ ich_laterality + age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
+
+### evd
+
+evd ~ ich_laterality
+
+evd ~ ich_laterality + age + htn + stroke + gcs_baseline + ich_location + ich_volume_baseline + ivh
+
+### trach
+
+tracheostomy ~ ich_laterality
+
+tracheostomy ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
+
+### peg
+
+peg ~ ich_laterality
+
+peg ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
+
+### vent
+
+days_mechanical_ventilation ~ ich_laterality
+
+days_mechanical_ventilation ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
+
+
+prior_arguments <- list(
+  formula = neurosurgery_evac ~ ich_laterality,
+  priors = c(
+    set_prior("normal(-2.2, 0.7)", class = "Intercept"),
+    set_prior("normal(0, 1)", class = "b")
+  ),
+  data = ich_aggressive,
+  family = bernoulli(link = "logit")
+)

### 2024-02-15 _targets.R

@@ -63,11 +63,16 @@ tar_plan(
   dag_outcomes = outcomes_dag_function(x),
 
   ## Priors ----
-  neurosurgery_min_priors = neurosurgery_min_prior_f(ich_aggressive),
+  model_parameters = model_parameters_function(ich_aggressive),
+  prior_neurosurgery_minimal = prior_func_meta(model_parameters, "neurosurgery_minimal"),
+  prior_neurosurgery_canonical = prior_func_meta(model_parameters, "neurosurgery_canonical"),
+  prior_evd_minimal = prior_func_meta(model_parameters, "evd_minimal"),
+  prior_evd_canonical = prior_func_meta(model_parameters, "evd_canonical"),
 
   ## Models ----
   neurosurgery_evac_min = neurosurgery_evac_func(ich_aggressive),
   neurosurgery_evac_can = neurosurgery_evac_func_canon(ich_aggressive),
+
 
   ## Reports ----
   tar_quarto(
@@ -77,5 +82,9 @@ tar_plan(
   tar_quarto(
     outcomes,
     "reports/outcomes.qmd"
+  ),
+  tar_quarto(
+    prior_report,
+    "reports/priors.qmd"
   )
 )

### 2024-02-16 reports/priors.qmd
aggressive priors complete?
still having issues with vent model

@@
 tar_load(prior_evd_canonical)
+tar_load(prior_trach_minimal)
+tar_load(prior_trach_canonical)
+tar_load(prior_vent_minimal)
+tar_load(prior_vent_canonical)
 
@@
 
-  p2 <- predictions(y, by = "ich_laterality") |>
+  p2 <- y
+
+  p3 <- predictions(y, by = "ich_laterality") |>
     as_tibble()
 
-  p3 <- avg_comparisons(y, variables = "ich_laterality") |>
+  p4 <- avg_comparisons(y, variables = "ich_laterality") |>
     posterior_draws() |>
@@
     priors = p1,
-    predictions = p2,
-    plot = p3
+    model = p2,
+    predictions = p3,
+    plot = p4
   ))
@@
 
-prior_neurosurgery_minimal
+```
+
+### Tracheostomy
+
+#### Minimal Model
+
+```{r}
+
+prior_printing("trach_minimal", prior_trach_minimal, "receiving an tracheostomy")
 
 ```
+
+#### Canonical Model
+
+```{r}
+
+prior_printing("trach_canonical", prior_trach_canonical, "receiving a tracheostomy")
+
+```
+
+### Days of Mechanical Ventilation
+
+#### Minimal Model
+
+```{r}
+
+prior_printing("vent_minimal", prior_vent_minimal, "number of days of mechanical ventilation")
+
+```
+
+#### Canonical Model
+
+```{r}
+
+prior_printing("vent_canonical", prior_vent_canonical, "number of days of mechanical ventilation")
+
+```

### 2024-02-16 R/data_cleaning.R

@@ -36,6 +36,7 @@ select_variables <- function(x) {
       dni_day,
       comfort_care_binary,
       comfort_care_day,
+      early_wlst,
       mrs_90,
       mrs_90_01,
       mrs_90_02,
@@ -55,4 +56,3 @@ non_surgery_trial_function <- function(x) {
   x |>
     filter(study == "ERICH" | study == "ATACH-2")
 }
-

### 2024-02-16 R/exploratory_data_analysis.R

@@ -56,7 +56,9 @@ table_2_aggressive_function <- function(x) {
     "days_mechanical_ventilation",
     "tracheostomy",
     "ich_laterality",
-    "comfort_care_binary"
+    "comfort_care_binary",
+    "early_wlst",
+    "dnr_binary"
   )
 
   x |>
@@ -69,7 +71,9 @@ table_2_aggressive_function <- function(x) {
         evd ~ "Ventriculostomy",
         days_mechanical_ventilation ~ "Mechanical Ventilation Days",
         tracheostomy ~ "Tracheostomy",
-        comfort_care_binary ~ "Comfort Care"
+        comfort_care_binary ~ "Withdrawal of Life-Sustaining Therapy",
+        early_wlst ~ "Early Withdrawal of Life-Sustaining Therapy (< 72 hours)",
+        dnr_binary ~ "DNR Order"
       )
     ) |>
     add_overall() |>

### 2024-02-16 R/functions_sandbox.R

@@ -1,107 +1,0 @@
-library(purrr)
-library(repurrrsive)
-
-prior_f <- function(formula, priors, data, family) {
-  mod_prior <- brm(formula,
-    data = data,
-    family = family,
-    prior = priors,
-    sample_prior = "only"
-  )
-  return(mod_prior)
-}
-
-prior_arguments <- list(
-  formula = neurosurgery_evac ~ ich_laterality,
-  priors = c(
-    set_prior("normal(-2.2, 0.7)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b")
-  ),
-  data = ich_aggressive,
-  family = bernoulli(link = "logit")
-)
-
-prior_f(prior_arguments$formula, prior_arguments$priors, prior_arguments$data, prior_arguments$family)
-
-wesanderson
-
-
-## purrr tutorial
-
-map(prior_arguments, "formula")
-
-map(got_chars[5:9], 1)
-map(wesanderson, 1)
-map(prior_arguments, 1)
-
-got_chars |>
-  map(3)
-
-names(got_chars)
-str(got_chars)
-
-map_lgl(got_chars, "alive")
-map_chr(got_chars, as.character("id"))
-
-culture <- factor(map_vec(got_chars, "culture"))
-unique(culture)
-
-map(wesanderson$name, print)
-
-tibble <- names(wesanderson)
-
-map_chr(wesanderson, 1)
-
-wes_tibble <- tibble(
-  movie = names(wesanderson),
-  color1 = map_chr(wesanderson, 1),
-  color2 = map_chr(wesanderson, 2)
-)
-
-map(shades, print)
-
-colors
-
-
-models <- list(
-  simple = "lifeexp ~ pop + gdppercap",
-  medium = "lifeexp ~ pop + gdppercap + continent + year",
-  more = "lifeexp ~ pop + gdppercap + country + year",
-  woah = "lifeexp ~ pop + gdppercap + year*country"
-)
-
-model_frame <- data_frame(model = models) %>%
-  mutate(model_name = names(model))
-
-model_frame <- model_frame %>%
-  mutate(model = map(model, as.formula))
-
-model_frame
-
-prior_formulas <- list(
-
-)
-
-
-lhr_setup <- function() {
-  # Settings
-  CHAINS <- 4
-  ITER <- 2000
-  WARMUP <- 1000
-  BAYES_SEED <- 4045 # From random.org
[…]

### 2024-02-16 R/testing.r

@@ -1,66 +1,0 @@
-Steps needed to build a bunch of prior models:
-
-# Aggressive Care
-
-outcomes: neurosurgery, evd, comfort_care, mech_vent, trach, and peg
-
-All but mech_vent are logistic models
-
-1) define formulas (simple and canonical)
-2) define priors (will this be different for each?)
-3) define data (will be the same for all)
-4) define family (will be the same except for mech_vent)
-
-## formulas
-
-formulas_aggressive <- list(
-  neurosurgery_simple = neurosurgery_evac ~ ich_laterality,
-  neurosurgery_canonical = neurosurgery_evac ~ ich_laterality + age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
-
-)
-
-is.atomic(formulas_aggressive[["neurosurgery_simple"]])
-
-prior_f(prior_arguments)
-
-str(prior_arguments)
-### neurosurgery
-
-neurosurgery_evac ~ ich_laterality
-
-neurosurgery_evac ~ ich_laterality + age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
-
-### evd
-
-evd ~ ich_laterality
-
-evd ~ ich_laterality + age + htn + stroke + gcs_baseline + ich_location + ich_volume_baseline + ivh
-
-### trach
-
-tracheostomy ~ ich_laterality
-
-tracheostomy ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
-
-### peg
-
-peg ~ ich_laterality
-
-peg ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
-
-### vent
-
-days_mechanical_ventilation ~ ich_laterality
-
-days_mechanical_ventilation ~ age + hypertension + stroke + gcs_baseline + ich_location + ich_volume + ivh
-
-
-prior_arguments <- list(
-  formula = neurosurgery_evac ~ ich_laterality,
-  priors = c(
-    set_prior("normal(-2.2, 0.7)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b")
-  ),
-  data = ich_aggressive,
-  family = bernoulli(link = "logit")
-)

### 2024-02-16 _targets.R

@@ -68,6 +68,11 @@ tar_plan(
   prior_neurosurgery_canonical = prior_func_meta(model_parameters, "neurosurgery_canonical"),
   prior_evd_minimal = prior_func_meta(model_parameters, "evd_minimal"),
   prior_evd_canonical = prior_func_meta(model_parameters, "evd_canonical"),
+  prior_trach_minimal = prior_func_meta(model_parameters, "trach_minimal"),
+  prior_trach_canonical = prior_func_meta(model_parameters, "trach_canonical"),
+  prior_vent_minimal = prior_func_meta(model_parameters, "vent_minimal"),
+  prior_vent_canonical = prior_func_meta(model_parameters, "vent_canonical"),
+
 
   ## Models ----
   neurosurgery_evac_min = neurosurgery_evac_func(ich_aggressive),

### 2024-02-16 reports/posteriors.qmd
POSTERIORS

---
title: Differences in care for patients experiencing left vs. right hemisphere ICH
subtitle: Posteriors!
author: Nikhil Patel, MD
format:
  html: 
    mainfont: Figtree
    monofont: Source Code Pro
    embed-resources: true
    toc: true
    toc-location: left
    toc-depth: 6
    toc-title: Contents
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-overflow: wrap
    code-tools: true
    code-fold: true
    highlight-style: nord
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(tidyverse)
library(glue)
library(marginaleffects)
library(modelsummary)
library(ggdist)
library(gghalves)
library(scales)
library(systemfonts)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

# Format things as percentage points
label_pp <- label_number(
  accuracy = 1, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

label_pp_tiny <- label_number(
  accuracy = 0.01, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

# load targets objects
tar_load(model_parameters)
tar_load(posterior_neurosurgery_minimal)
tar_load(posterior_neurosurgery_canonical)
tar_load(posterior_evd_minimal)
tar_load(posterior_evd_canonical)
tar_load(posterior_trach_minimal)
tar_load(posterior_trach_canonical)
tar_load(posterior_vent_minimal)
tar_load(posterior_vent_canonical)
tar_load(posterior_comfort_care_minimal)
tar_load(posterior_comfort_care_canonical)
tar_load(posterior_early_wlst_minimal)
tar_load(posterior_early_wlst_canonical)
tar_load(posterior_dnr_minimal)
tar_load(posterior_dnr_canonical)

# posterior functions

posterior_printing <- function(x, y, var) {
  p1 <- model_parameters$priors[x]

  p2 <- y

  p3 <- predictions(y, by = "ich_laterality") |>
    as_tibble()

  p4 <- avg_comparisons(y, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    stat_halfeye(
      .width = c(0.8, 0.95), point_interval = "median_hdi",
      fill = "#bc3032"
    ) +
    scale_x_continuous(labels = label_pp) +
    labs(
      title = "Posterior Prediction",
      x = paste("Effect of hemispheric laterality\n on the probability of", {{ var }}),
      y = NULL, caption = "80% and 95% credible intervals shown in black"
    )

  return(list(
    priors = p1,
    model = p2,
    predictions = p3,
    plot = p4
  ))
}

```

## Prior Predictive Checks

### Neurosurgical Intervention

#### Minimal Model

```{r}

posterior_printing("neurosurgery_minimal", posterior_neurosurgery_minimal, "receiving neurosurgical intervention")

```

#### Canonical Model

```{r}

posterior_printing("neurosurgery_canonical", posterior_neurosurgery_canonical, "receiving neurosurgical intervention")

```

### EVD

#### Minimal Model

```{r}

posterior_printing("evd_minimal", posterior_evd_minimal, "receiving an EVD")

```

#### Canonical Model

```{r}

posterior_printing("evd_canonical", posterior_evd_canonical, "receiving an EVD")

```

### Tracheostomy

#### Minimal Model

```{r}

posterior_printing("trach_minimal", posterior_trach_minimal, "receiving an tracheostomy")

```

#### Canonical Model

```{r}

posterior_printing("trach_canonical", posterior_trach_canonical, "receiving a tracheostomy")

```

### Days of Mechanical Ventilation

#### Minimal Model

```{r}

posterior_printing("vent_minimal", posterior_vent_minimal, "number of days of mechanical ventilation")

```

#### Canonical Model

```{r}

posterior_printing("vent_canonical", posterior_vent_canonical, "number of days of mechanical ventilation")

```

### Comfort Care

#### Minimal Model

```{r}

posterior_printing("comfort_care_minimal", posterior_comfort_care_minimal, "comfort care")

```

#### Canonical Model

```{r}
posterior_printing("comfort_care_canonical", posterior_comfort_care_canonical, "comfort care")

```

### Early WLST

#### Minimal Model

```{r}
posterior_printing("early_wlst_minimal", posterior_early_wlst_minimal, "early WLST")

```

#### Canonical Model

```{r}
posterior_printing("early_wlst_canonical", posterior_early_wlst_canonical, "early WLST")

```

### DNR

#### Minimal Model

```{r}
posterior_printing("dnr_minimal", posterior_dnr_minimal, "DNR")

```

#### Canonical Model

```{r}
posterior_printing("dnr_canonical", posterior_dnr_canonical, "DNR")

```

### 2024-02-16 _targets.R

@@ -36,7 +36,6 @@ source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
 source("R/dags.R")
 source("R/priors.R")
-source("R/models.R")
 
 # Pipeline ----
 tar_plan(
@@ -72,12 +71,28 @@ tar_plan(
   prior_trach_canonical = prior_func_meta(model_parameters, "trach_canonical"),
   prior_vent_minimal = prior_func_meta(model_parameters, "vent_minimal"),
   prior_vent_canonical = prior_func_meta(model_parameters, "vent_canonical"),
+  prior_comfort_care_minimal = prior_func_meta(model_parameters, "comfort_care_minimal"),
+  prior_comfort_care_canonical = prior_func_meta(model_parameters, "comfort_care_canonical"),
+  prior_early_wlst_minimal = prior_func_meta(model_parameters, "early_wlst_minimal"),
+  prior_early_wlst_canonical = prior_func_meta(model_parameters, "early_wlst_canonical"),
+  prior_dnr_minimal = prior_func_meta(model_parameters, "dnr_minimal"),
+  prior_dnr_canonical = prior_func_meta(model_parameters, "dnr_canonical"),
 
-
-  ## Models ----
-  neurosurgery_evac_min = neurosurgery_evac_func(ich_aggressive),
-  neurosurgery_evac_can = neurosurgery_evac_func_canon(ich_aggressive),
-
+  ## Posterior Simulation ----
+  posterior_neurosurgery_minimal = posterior_func_meta(model_parameters, "neurosurgery_minimal"),
+  posterior_neurosurgery_canonical = posterior_func_meta(model_parameters, "neurosurgery_canonical"),
+  posterior_evd_minimal = posterior_func_meta(model_parameters, "evd_minimal"),
+  posterior_evd_canonical = posterior_func_meta(model_parameters, "evd_canonical"),
+  posterior_trach_minimal = posterior_func_meta(model_parameters, "trach_minimal"),
+  posterior_trach_canonical = posterior_func_meta(model_parameters, "trach_canonical"),
+  posterior_vent_minimal = posterior_func_meta(model_parameters, "vent_minimal"),
+  posterior_vent_canonical = posterior_func_meta(model_parameters, "vent_canonical"),
+  posterior_comfort_care_minimal = posterior_func_meta(model_parameters, "comfort_care_minimal"),
[…]

### 2024-02-20 reports/posteriors.qmd
cumulative logit woes

@@
     code-fold: true
+    code-link: true
     highlight-style: nord
@@
 library(systemfonts)
+library(brms)
+library(tidybayes)
 
@@
 
-posterior_printing <- function(x, y, var) {
+posterior_printing <- function(x, y, z) {
   p1 <- model_parameters$priors[x]
@@
     as_tibble()
+
+  p35 <- avg_comparisons(y, variables = "ich_laterality")
 
@@
       title = "Posterior Prediction",
-      x = paste("Effect of hemispheric laterality\n on the probability of", {{ var }}),
+      x = paste("Effect of hemispheric laterality\n on the probability of", {{ z }}),
       y = NULL, caption = "80% and 95% credible intervals shown in black"
     )
+
+  p5 <- y |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    ggplot(aes(ich_right_or)) +
+    stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "bc3032")) +
+    theme(legend.position = "none")
+
+  p6 <- y |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      or_1 = sum(ich_right_or > 1) / n(),
+      or_1.1 = sum(ich_right_or > 1.1) / n(),
+      or_1.2 = sum(ich_right_or > 1.2) / n()
+    )
+
+  p5 <- pp_check(p2, ndraws = 500)
 
@@
     model = p2,
+    fit = p5,
     predictions = p3,
-    plot = p4
+    comparison = p35,
+    plot = p4,
+    plot2 = p5,
+    probabilities = p6
   ))
@@
 
-## Prior Predictive Checks
+## Posterior Predictions
 
@@
 ```{r}
-
 posterior_printing("neurosurgery_minimal", posterior_neurosurgery_minimal, "receiving neurosurgical intervention")

### 2024-02-20 reports/priors.qmd
cumulative logit woes

@@
 tar_load(prior_vent_canonical)
+tar_load(prior_comfort_care_minimal)
+tar_load(prior_comfort_care_canonical)
+tar_load(prior_early_wlst_minimal)
+tar_load(prior_early_wlst_canonical)
+tar_load(prior_dnr_minimal)
+tar_load(prior_dnr_canonical)
 
@@
 ```{r}
-
 prior_printing("vent_canonical", prior_vent_canonical, "number of days of mechanical ventilation")
@@
 ```
+
+### Comfort Care
+
+#### Minimal Model
+
+```{r}
+
+prior_printing("comfort_care_minimal", prior_comfort_care_minimal, "comfort care")
+
+```
+
+#### Canonical Model
+
+```{r}
+prior_printing("comfort_care_canonical", prior_comfort_care_canonical, "comfort care")
+
+```
+
+### Early WLST
+
+#### Minimal Model
+
+```{r}
+prior_printing("early_wlst_minimal", prior_early_wlst_minimal, "early WLST")
+
+```
+
+#### Canonical Model
+
+```{r}
+prior_printing("early_wlst_canonical", prior_early_wlst_canonical, "early WLST")
+
+```
+
+### DNR
+
+#### Minimal Model
+
+```{r}
+prior_printing("dnr_minimal", prior_dnr_minimal, "DNR")
+
+```
+
+#### Canonical Model
+
+```{r}
+prior_printing("dnr_canonical", prior_dnr_canonical, "DNR")
+
+```

### 2024-02-20 R/priors.R

@@ -42,12 +42,12 @@ model_parameters_function <- function(ich_aggressive) {
       set_prior("normal(0, 1)", class = "b")
     ),
     vent_minimal = c(
-      set_prior("normal(1.5, 0.2)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
+      set_prior("normal(0, 1.5)", class = "Intercept"),
+      set_prior("normal(0, 0.5)", class = "b")
     ),
     vent_canonical = c(
-      set_prior("normal(0.95, 0.5)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
+      set_prior("normal(0, 1.5)", class = "Intercept"),
+      set_prior("normal(0, 0.5)", class = "b")
     ),
     comfort_care_minimal = c(
       set_prior("normal(1.5, 0.2)", class = "Intercept"),
@@ -99,8 +99,8 @@ model_parameters_function <- function(ich_aggressive) {
     evd_canonical = bernoulli(link = "logit"),
     trach_minimal = bernoulli(link = "logit"),
     trach_canonical = bernoulli(link = "logit"),
-    vent_minimal = poisson,
-    vent_canonical = poisson,
+    vent_minimal = cumulative(link = "logit"),
+    vent_canonical = cumulative(link = "logit"),
     comfort_care_minimal = bernoulli(link = "logit"),
     comfort_care_canonical = bernoulli(link = "logit"),
     early_wlst_minimal = bernoulli(link = "logit"),

### 2024-02-20 reports/aggressiveness.qmd
diagnostics and visuals done
need to do OR and ROPE stuff

@@
 ---
-title: Differences in care for patients experiencing left vs. right hemisphere ICH
+title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
 author: Nikhil Patel, MD
 format:
-  html: 
-    mainfont: Figtree
-    monofont: Source Code Pro
-    fontsize: 1rem
-    embed-resources: true
-    toc: true
-    toc-location: left
-    toc-depth: 6
-    toc-expand: true
-    toc-title: Contents
-    smooth-scroll: true
-    link-external-icon: true
-    link-external-newwindow: true
-    code-overflow: wrap
-    code-tools: true
-    code-fold: true
-    highlight-style: nord
+  html
 ---

### 2024-02-20 reports/outcomes.qmd
diagnostics and visuals done
need to do OR and ROPE stuff

@@
 format:
-  html: 
-    mainfont: Figtree
-    monofont: Source Code Pro
-    fontsize: 0.97rem
-    embed-resources: true
-    toc: true
-    toc-location: left
-    toc-depth: 6
-    toc-expand: true
-    toc-title: Contents
-    smooth-scroll: true
-    link-external-icon: true
-    link-external-newwindow: true
-    code-overflow: wrap
-    code-tools: true
-    code-fold: true
-    highlight-style: nord
+  html
 ---
@@
 ```
-
-#### DAG-Implied Adjustment Sets
-
-Similar to above, there are no
-
-```{r outcomes dag adjustment sets}
-#| code-fold: false
-#| fig-align: center
-#| fig.height: 10
-#| fig.width: 20
-
-adjustmentSets(
-  dag_outcomes,
-  type = "minimal",
-  effect = "direct",
-  max.results = Inf
-)
-
-adjustmentSets(
-  dag_outcomes,
-  type = "canonical",
-  effect = "total",
-  max.results = Inf
-)
-
-```

### 2024-02-20 reports/posteriors.qmd
diagnostics and visuals done
need to do OR and ROPE stuff

--- 2024-02-16 reports/posteriors.qmd
@@
 ---
-title: Differences in care for patients experiencing left vs. right hemisphere ICH
-subtitle: Posteriors!
+title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
 author: Nikhil Patel, MD
 format:
-  html: 
-    mainfont: Figtree
-    monofont: Source Code Pro
-    embed-resources: true
-    toc: true
-    toc-location: left
-    toc-depth: 6
-    toc-title: Contents
-    smooth-scroll: true
-    link-external-icon: true
-    link-external-newwindow: true
-    code-overflow: wrap
-    code-tools: true
-    code-fold: true
-    highlight-style: nord
+  html
 ---
@@
 
-# posterior functions
-
-posterior_printing <- function(x, y, var) {
-  p1 <- model_parameters$priors[x]
-
-  p2 <- y
-
-  p3 <- predictions(y, by = "ich_laterality") |>
-    as_tibble()
-
-  p4 <- avg_comparisons(y, variables = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw)) +
-    stat_halfeye(
-      .width = c(0.8, 0.95), point_interval = "median_hdi",
-      fill = "#bc3032"
-    ) +
-    scale_x_continuous(labels = label_pp) +
-    labs(
-      title = "Posterior Prediction",
-      x = paste("Effect of hemispheric laterality\n on the probability of", {{ var }}),
-      y = NULL, caption = "80% and 95% credible intervals shown in black"
-    )
-
-  return(list(
-    priors = p1,
-    model = p2,
-    predictions = p3,
-    plot = p4
-  ))
-}
-
 ```
-
-## Prior Predictive Checks
-
-### Neurosurgical Intervention
-
-#### Minimal Model
-
-```{r}
-
-posterior_printing("neurosurgery_minimal", posterior_neurosurgery_minimal, "receiving neurosurgical intervention")
-
-```
-
-#### Canonical Model
-
-```{r}
-
-posterior_printing("neurosurgery_canonical", posterior_neurosurgery_canonical, "receiving neurosurgical intervention")
-
-```
-
-### EVD
-
-#### Minimal Model
-
-```{r}
-
-posterior_printing("evd_minimal", posterior_evd_minimal, "receiving an EVD")
-
-```
-
-#### Canonical Model
-
-```{r}
-
-posterior_printing("evd_canonical", posterior_evd_canonical, "receiving an EVD")
-
-```
-
-### Tracheostomy
-
-#### Minimal Model
-
-```{r}
-
-posterior_printing("trach_minimal", posterior_trach_minimal, "receiving an tracheostomy")
-
-```
-
-#### Canonical Model
-
-```{r}
-
-posterior_printing("trach_canonical", posterior_trach_canonical, "receiving a tracheostomy")
-
-```
-
-### Days of Mechanical Ventilation
-
-#### Minimal Model
-
-```{r}
-
-posterior_printing("vent_minimal", posterior_vent_minimal, "number of days of mechanical ventilation")
-
-```
-
-#### Canonical Model
-
-```{r}
-
-posterior_printing("vent_canonical", posterior_vent_canonical, "number of days of mechanical ventilation")
-
-```
-
-### Comfort Care
-
-#### Minimal Model
-
-```{r}
-
-posterior_printing("comfort_care_minimal", posterior_comfort_care_minimal, "comfort care")
-
-```
-
-#### Canonical Model
-
-```{r}
-posterior_printing("comfort_care_canonical", posterior_comfort_care_canonical, "comfort care")
-
-```
-
-### Early WLST
-
-#### Minimal Model
-
-```{r}
-posterior_printing("early_wlst_minimal", posterior_early_wlst_minimal, "early WLST")
-
-```
-
-#### Canonical Model
-
-```{r}
-posterior_printing("early_wlst_canonical", posterior_early_wlst_canonical, "early WLST")
-
-```
-
-### DNR
-
-#### Minimal Model
-
-```{r}
-posterior_printing("dnr_minimal", posterior_dnr_minimal, "DNR")
-
-```
-
-#### Canonical Model
-
-```{r}
-posterior_printing("dnr_canonical", posterior_dnr_canonical, "DNR")
-
-```

### 2024-02-20 reports/posteriors_template.qmd
diagnostics and visuals done
need to do OR and ROPE stuff

---
title: Differences in outcomes for patients experiencing left vs. right hemisphere ICH
author: Nikhil Patel, MD
format:
  html
---

```{r setup, output=FALSE, message=FALSE, warning=FALSE}
library(targets)
library(tidyverse)
library(glue)
library(marginaleffects)
library(modelsummary)
library(ggdist)
library(gghalves)
library(scales)
library(systemfonts)
library(brms)
library(tidybayes)

# theme customization
theme_set(theme_minimal(base_family = "Figtree"))

# Format things as percentage points
label_pp <- label_number(
  accuracy = 1, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

label_pp_tiny <- label_number(
  accuracy = 0.01, scale = 100,
  suffix = " pp.", style_negative = "minus"
)

# load targets objects
tar_load(model_parameters)
tar_load(posterior_neurosurgery_minimal)
tar_load(posterior_neurosurgery_canonical)
tar_load(posterior_evd_minimal)
tar_load(posterior_evd_canonical)
tar_load(posterior_trach_minimal)
tar_load(posterior_trach_canonical)
tar_load(posterior_vent_minimal)
tar_load(posterior_vent_canonical)
tar_load(posterior_comfort_care_minimal)
tar_load(posterior_comfort_care_canonical)
tar_load(posterior_early_wlst_minimal)
tar_load(posterior_early_wlst_canonical)
tar_load(posterior_dnr_minimal)
tar_load(posterior_dnr_canonical)

# posterior functions

posterior_printing <- function(x, y, z) {
  p1 <- model_parameters$priors[x]

  p2 <- y

  p3 <- predictions(y, by = "ich_laterality") |>
    as_tibble()

  p35 <- avg_comparisons(y, variables = "ich_laterality")

  p4 <- avg_comparisons(y, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    stat_halfeye(
      .width = c(0.8, 0.95), point_interval = "median_hdi",
      fill = "#bc3032"
    ) +
    scale_x_continuous(labels = label_pp) +
    labs(
      title = "Posterior Prediction",
      x = paste("Effect of hemispheric laterality\n on the probability of", {{ z }}),
      y = NULL, caption = "80% and 95% credible intervals shown in black"
    )

  p5 <- y |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    ggplot(aes(ich_right_or)) +
    stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "bc3032")) +
    theme(legend.position = "none")

  p6 <- y |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize(
      or_1 = sum(ich_right_or > 1) / n(),
      or_1.1 = sum(ich_right_or > 1.1) / n(),
      or_1.2 = sum(ich_right_or > 1.2) / n()
    )

  p5 <- pp_check(p2, ndraws = 500)

  return(list(
    priors = p1,
    model = p2,
    fit = p5,
    predictions = p3,
    comparison = p35,
    plot = p4,
    plot2 = p5,
    probabilities = p6
  ))
}

```

## Posterior Predictions

### Neurosurgical Intervention

#### Minimal Model

```{r}
posterior_printing("neurosurgery_minimal", posterior_neurosurgery_minimal, "receiving neurosurgical intervention")

```

#### Canonical Model

```{r}

posterior_printing("neurosurgery_canonical", posterior_neurosurgery_canonical, "receiving neurosurgical intervention")

```

### EVD

#### Minimal Model

```{r}

posterior_printing("evd_minimal", posterior_evd_minimal, "receiving an EVD")

```

#### Canonical Model

```{r}

posterior_printing("evd_canonical", posterior_evd_canonical, "receiving an EVD")

```

### Tracheostomy

#### Minimal Model

```{r}

posterior_printing("trach_minimal", posterior_trach_minimal, "receiving an tracheostomy")

```

#### Canonical Model

```{r}

posterior_printing("trach_canonical", posterior_trach_canonical, "receiving a tracheostomy")

```

### Days of Mechanical Ventilation

#### Minimal Model

```{r}

posterior_printing("vent_minimal", posterior_vent_minimal, "number of days of mechanical ventilation")

```

#### Canonical Model

```{r}

posterior_printing("vent_canonical", posterior_vent_canonical, "number of days of mechanical ventilation")

```

### Comfort Care

#### Minimal Model

```{r}

posterior_printing("comfort_care_minimal", posterior_comfort_care_minimal, "comfort care")

```

#### Canonical Model

```{r}
posterior_printing("comfort_care_canonical", posterior_comfort_care_canonical, "comfort care")

```

### Early WLST

#### Minimal Model

```{r}
posterior_printing("early_wlst_minimal", posterior_early_wlst_minimal, "early WLST")

```

#### Canonical Model

```{r}
posterior_printing("early_wlst_canonical", posterior_early_wlst_canonical, "early WLST")

```

### DNR

#### Minimal Model

```{r}
posterior_printing("dnr_minimal", posterior_dnr_minimal, "DNR")

```

#### Canonical Model

```{r}
posterior_printing("dnr_canonical", posterior_dnr_canonical, "DNR")

```

### 2024-02-20 R/models.R

@@ -1,54 +1,159 @@
-# neurosurgical intervention
-neurosurgery_evac_func <- function(x) {
-  BAYES_SEED <- 07204
-
-  options(
-    mc.cores = 4,
-    brms.backend = "cmdstanr"
+model_parameters_function <- function(ich_aggressive) {
+  formulas <- list(
+    neurosurgery_minimal = neurosurgery_evac ~ ich_laterality,
+    neurosurgery_canonical = neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    evd_minimal = evd ~ ich_laterality,
+    evd_canonical = evd ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    trach_minimal = tracheostomy ~ ich_laterality,
+    trach_canonical = tracheostomy ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    vent_minimal = days_mechanical_ventilation ~ ich_laterality,
+    vent_canonical = days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    comfort_care_minimal = comfort_care_binary ~ ich_laterality,
+    comfort_care_canonical = comfort_care_binary ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    early_wlst_minimal = early_wlst ~ ich_laterality,
+    early_wlst_canonical = early_wlst ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    dnr_minimal = dnr_binary ~ ich_laterality,
+    dnr_canonical = dnr_binary ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh
   )
 
-  priors <- c(
-    set_prior("student_t(3, 0, 2.5)", class = "Intercept"),
-    set_prior("normal(0, 1)", class = "b")
-  )
-
-  brm(
-    neurosurgery_evac ~ ich_laterality + (1 | study),
-    data = x,
-    family = bernoulli(link = "logit"),
-    prior = priors,
-    init = 0,
-    control = list(
-      adapt_delta = 0.99,
-      max_treedepth = 12
+  priors <- list(
+    neurosurgery_minimal = c(
[…]

### 2024-02-20 R/packages.R

@@ -13,5 +13,12 @@ library(ggdag)
 library(cmdstanr)
 library(brms)
 library(marginaleffects)
-library(modelsummary)
+library(tidybayes)
+library(bayesplot)
 library(scales)
+library(broom)
+library(broom.mixed)
+library(patchwork)
+library(systemfonts)
+library(monochromeR)
+library(glue)

### 2024-02-20 R/posterior_results.R

@@ -1,0 +1,89 @@
+results_visual <- function(model, title) {
+  model <- model
+  subtitle <- title
+
+  draws <- model |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize()
+
+  pred <- predictions(model, by = "ich_laterality") |>
+    tibble()
+
+  theme_ich <- function(base_family = "Inter",
+                        base_size = 14) {
+    theme_minimal(base_size = base_size) +
+      theme(
+        text = element_text(lineheight = 2),
+        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+        axis.text.y = element_text(size = rel(0.8)),
+        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+        axis.text.x = element_text(size = rel(2)),
+        axis.title.x = element_blank(),
+        legend.position = "bottom",
+        legend.justification = 1,
+        panel.grid = element_line(colour = "#F3F4F5"),
+        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
+        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
+      )
+  }
+
+  p1 <- predictions(model, by = "ich_laterality") |>
+    posterior_draws() |>
+    ggplot(aes(x = draw, fill = ich_laterality)) +
+    geom_density(alpha = .7) +
+    scale_fill_manual(values = c("#422f05", "#E69F00")) +
+    scale_x_continuous(labels = scales::percent) +
+    labs(
+      title = "Predicted Probability",
+      fill = "ICH Laterality",
+      y = NULL
+    ) +
+    theme_ich()
+
+  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
+    posterior_draws() |>
+    ggplot(aes(x = draw)) +
+    geom_density(color = "#422f05", fill = "#E69F00") +
+    scale_x_continuous(labels = scales::percent) +
+    labs(
+      title = "Average Difference between Left vs. Right Hemisphere",
+      y = NULL
+    ) +
+    geom_vline(
+      xintercept = 0, linetype = "dashed",
[…]

### 2024-02-20 R/testing.R

@@ -1,6 +1,0 @@
-vent_mod <- brm(
-  days_mechanical_ventilation ~ ich_laterality,
-  data = ich_aggressive,
-  family = cumulative(link = "logit"),
-  prior = c(set_prior("normal(0, 1.5)", class = "Intercept"))
-)

### 2024-02-20 _targets.R

@@ -29,13 +29,18 @@ options(
 set.seed(908)
 BAYES_SEED <- 908
 
+
+
 # R functions ----
 
 source("R/packages.R")
 source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
 source("R/dags.R")
-source("R/priors.R")
+source("R/models.R")
+source("R/diagnostics.R")
+source("R/posterior_results.R")
+
 
 # Pipeline ----
 tar_plan(
@@ -94,21 +99,53 @@ tar_plan(
   posterior_dnr_minimal = posterior_func_meta(model_parameters, "dnr_minimal"),
   posterior_dnr_canonical = posterior_func_meta(model_parameters, "dnr_canonical"),
 
+  ## Diagnostics ----
+  neurosugery_minimal_diagnostics = posterior_diagnostics(posterior_neurosurgery_minimal),
+  neurosugery_canonical_diagnostics = posterior_diagnostics(posterior_neurosurgery_canonical),
+  evd_minimal_diagnostics = posterior_diagnostics(posterior_evd_minimal),
+  evd_canonical_diagnostics = posterior_diagnostics(posterior_evd_canonical),
+  trach_minimal_diagnostics = posterior_diagnostics(posterior_trach_minimal),
+  trach_canonical_diagnostics = posterior_diagnostics(posterior_trach_canonical),
+  vent_minimal_diagnostics = posterior_diagnostics(posterior_vent_minimal),
+  vent_canonical_diagnostics = posterior_diagnostics(posterior_vent_canonical),
+  comfort_care_minimal_diagnostics = posterior_diagnostics(posterior_comfort_care_minimal),
+  comfort_care_canonical_diagnostics = posterior_diagnostics(posterior_comfort_care_canonical),
+  early_wlst_minimal_diagnostics = posterior_diagnostics(posterior_early_wlst_minimal),
+  early_wlst_canonical_diagnostics = posterior_diagnostics(posterior_early_wlst_canonical),
+  dnr_minimal_diagnostics = posterior_diagnostics(posterior_dnr_minimal),
+  dnr_canonical_diagnostics = posterior_diagnostics(posterior_dnr_canonical),
+
+  ## Plots ----
+  neurosugery_minimal_visuals = results_visual(posterior_neurosurgery_minimal, "Neurosurgical Intervention"),
[…]

### 2024-02-21 manuscripts/index.qmd
recoded aggressive models
Got rid of previous functions, they were confusing. Kept the code in models.R

---
title: Differences in care for patients with left vs. right hemisphere intracerebral hemorrhage
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Neurology, Wake Forest School of Medicine
    email: nikhil.patel@atriumhealth.org
    roles: writing
    corresponding: true
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest School of Medicine
    corresponding: false
    roles: writing
bibliography: references.bib
---

## Section

This is a simple placeholder for the manuscript's main document [@audetat_diagnosis_2017-1].

### 2024-02-21 R/aggressive_care_posterior_models.R

@@ -1,0 +1,323 @@
+# Posterior models ----------------------------------------------------------
+
+## Neurosurgery as outcome -----
+
+f_posterior_neurosurgery_minimal <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(neurosurgery_evac ~ ich_laterality,
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed,
+    backend = "cmdstanr"
+  )
+
+  return(model)
+}
+
+f_posterior_neurosurgery_canonical <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed,
+    backend = "cmdstanr"
+  )
+
+  return(model)
+}
+
+## EVD as outcome -----
+
+f_posterior_evd_minimal <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(evd ~ ich_laterality,
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed,
+    backend = "cmdstanr"
+  )
+
+  return(model)
+}
+
+f_posterior_evd_canonical <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(evd ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
[…]

### 2024-02-21 R/aggressive_care_prior_models.R

@@ -1,0 +1,359 @@
+# Settings ----------------------------------------------------------------
+
+model_setup <- function() {
+  # Settings
+  CHAINS <- 4
+  CORES <- 6
+  THREADS <- threading(4)
+  ITER <- 2000
+  WARMUP <- 1000
+  BAYES_SEED <- 908
+
+  # Priors
+  priors_vague <- c(
+    set_prior("normal(0, 10)", class = "Intercept"),
+    set_prior("normal(0, 3)", class = "b")
+  )
+
+  return(list(
+    chains = CHAINS, cores = CORES, threads = THREADS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED, priors_vague = priors_vague
+  ))
+}
+
+# Prior models ----------------------------------------------------------
+
+## Neurosurgery as outcome -----
+
+f_prior_neurosurgery_minimal <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(neurosurgery_evac ~ ich_laterality,
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    sample_prior = "only",
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed,
+    backend = "cmdstanr"
+  )
+
+  return(model)
+}
+
+f_prior_neurosurgery_canonical <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 1)", class = "b")
+    ),
+    sample_prior = "only",
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed,
+    backend = "cmdstanr"
+  )
+
+  return(model)
+}
+
+## EVD as outcome -----
+
+f_prior_evd_minimal <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(evd ~ ich_laterality,
[…]

### 2024-02-21 R/data_cleaning.R

@@ -52,7 +52,4 @@ filter_variables <- function(x) {
     drop_na(ich_laterality)
 }
 
-non_surgery_trial_function <- function(x) {
-  x |>
-    filter(study == "ERICH" | study == "ATACH-2")
-}
+ 

### 2024-02-21 R/outcomes_prior_models.R

@@ -1,0 +1,1 @@
+# Prior models ----------------------------------------------------------

### 2024-02-21 _targets.R

@@ -6,29 +6,11 @@ library(tarchetypes)
 suppressPackageStartupMessages(library(brms))
 
 # Taken from Andrew Heiss (https://github.com/andrewheiss/cautioning-canary/blob/master/_targets.R)
-# Bayes-specific stuff
-# Specify the number of cores and threads
-#
-# If using brms's default rstan backend, set mc.threads to NULL
-#
-#   - If using cmdstanr, you can use within-chain threading
-#   - (https://cran.r-project.org/web/packages/brms/vignettes/brms_threading.html)
-#   - On a four-core computer, use mc.cores = 2 and threads = threading(2)
-#   - On an eight+-core computer, use mc.cores = 4 and threads = threading(2)
-options(
-  mc.cores = 4,
-  mc.threads = 2,
-  brms.backend = "cmdstanr"
-)
 
 options(
   tidyverse.quiet = TRUE,
   dplyr.summarise.inform = FALSE
 )
-
-set.seed(908)
-BAYES_SEED <- 908
-
 
 
 # R functions ----
@@ -37,9 +19,10 @@ source("R/packages.R")
 source("R/data_cleaning.R")
 source("R/exploratory_data_analysis.R")
 source("R/dags.R")
-source("R/models.R")
+source("R/aggressive_care_prior_models.R")
+source("R/aggressive_care_posterior_models.R")
 source("R/diagnostics.R")
-source("R/posterior_results.R")
+source("R/results_visualization.R")
 
 
 # Pipeline ----
@@ -67,69 +50,69 @@ tar_plan(
   dag_outcomes = outcomes_dag_function(x),
 
   ## Priors ----
-  model_parameters = model_parameters_function(ich_aggressive),
-  prior_neurosurgery_minimal = prior_func_meta(model_parameters, "neurosurgery_minimal"),
-  prior_neurosurgery_canonical = prior_func_meta(model_parameters, "neurosurgery_canonical"),
-  prior_evd_minimal = prior_func_meta(model_parameters, "evd_minimal"),
-  prior_evd_canonical = prior_func_meta(model_parameters, "evd_canonical"),
-  prior_trach_minimal = prior_func_meta(model_parameters, "trach_minimal"),
-  prior_trach_canonical = prior_func_meta(model_parameters, "trach_canonical"),
-  prior_vent_minimal = prior_func_meta(model_parameters, "vent_minimal"),
-  prior_vent_canonical = prior_func_meta(model_parameters, "vent_canonical"),
[…]

### 2024-03-06 manuscript 01/_manuscript/index.qmd
manuscript actually started

---
title: Does Hemispheric Lateralization Influence Aggressive Treatment in Patients with Intracerebral Hemorrhage?
subtitle: An Analysis of ATACH-2 and ERICH
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Neurology, Wake Forest University School of Medicine
    roles: data analysis, writing
    corresponding: true
  - name: Matt Jaffa
    affiliation: Department of Neurology, University of Connecticut School of Medicine
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest University School of Medicine
date: last-modified
abstract: |
  Hemispheric laterality blah blah blah
bibliography: references.bib
csl: american-heart-association.csl
mainfont: Figtree

---

```{r}
library(targets)
library(dplyr)
library(targets)
library(withr)
library(here)

with_dir(here(), tar_load(ich_aggressive))

sample_size <- count(ich_aggressive)

```

## Introduction
The clinical presentation of patients with stroke can vary based on the hemisphere affected, in particular with language, spatial orientation, and emotional processing. Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.[@legge_neglecting_2006; @di_legge_impact_2005-1] On the other hand, there is a hesitation among some practitioners to offer decompressive craniectomy in patients with dominant-hemispheric infarcts. [@lin_decompressive_2021-1]

It is unclear how previous research in the ischemic stroke population translates to patients experiencing intracerebral hemorrhage. Not much has been investigated.

We aimed to evaluate the effect of hemispheric lateralization on the probability of receiving aggressive care in patients with ICH.

## Methods

### Study Design and Participants

We conducted a cohort study using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshi_intensive_2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH)[@woo_ethnicracial_2013] study. ATACH-2 was a large, multicenter, international, open-label, blinded-endpoint randomized controlled trial enrolling 1000 patients with ICH while ERICH was a large, multicenter, prospective observational study enrolling patients in the United States admitted with ICH. Study protocols and results of both studies have been previously published and are available elsewhere.

We included all patients enrolled in these two studies with supratentorial location of ICH. We excluded patients with cerebellar or brainstem ICH as well as any patient with primary intraventricular hemorrhage (IVH).

The study was approved by the institutional review board of the Wake Forest School of Medicine.

#### Data Availability Statement

The ATACH-2 (PI: Adnan I. Qureshi, MD, NCT01176565) and ERICH (PI: Daniel Woo, MD, MS, NCT01202864) data and analyses presented in this manuscript are based on the research files downloaded from the Archived Clinical Research web site of the National Institute of Neurologic Disease and Stroke (NINDS). The datasets can be obtained after approval of a written proposal as outlined on the NINDS website.

The analysis code is publicly available as outlined by the NINDS Data Request Agreement at [https://github.com/nikhilpatelmd/ich-laterality](https://github.com/nikhilpatelmd/ich-laterality).

### Exposures and Outcomes

Our exposure was hemispheric laterality (left versus right). Our primary outcome was neurosurgical intervention, defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.

To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), tracheostomy, percutaneous endoscopic gastrostomy (PEG) tube, do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (< 72 from symptom onset).

### Statistical Analysis

#### Overall Analysis Plan

## Results

A total of `r sample_size` patients were included in the analysis.

```{r}

library(gtsummary)

theme_gtsummary_compact(set_theme = TRUE, font_size = 14)
with_dir(here(), tar_read(table_1_aggressive))

```

## Acknowledgments {.appendix}

## References {.appendix}

### 2024-03-06 manuscripts/index.qmd
manuscript actually started

@@
 ---
-title: Differences in care for patients with left vs. right hemisphere intracerebral hemorrhage
-authors:
-  - name: Nikhil Patel
-    orcid: 0000-0002-3193-8590
-    affiliation: Department of Neurology, Wake Forest School of Medicine
-    email: nikhil.patel@atriumhealth.org
-    roles: writing
-    corresponding: true
-  - name: Rahul Karamchandani
-    affiliation: Department of Neurology, Wake Forest School of Medicine
-    corresponding: false
-    roles: writing
-bibliography: references.bib
+title: "ICH Laterality"
+subtitle: "Project Website"
 ---
 
-## Section
-
-This is a simple placeholder for the manuscript's main document [@audetat_diagnosis_2017-1].
+This is the landing page for the website.

### 2024-03-06 manuscripts/manuscript1.qmd
manuscript actually started

---
title: "Differences in Treatment Driven by Hemispheric Laterality"
author: "Nikhil Patel, MD"
---

This is the actual manuscript

```{r}

library(tidyverse)

```

This is more [@noauthor_simvastatin_2023]

### 2024-03-06 R/exploratory_data_analysis.R

@@ -1,4 +1,6 @@
 table_1_function <- function(x) {
+
+
   table_one_vars <- c(
     "age",
     "sex",

### 2024-03-07 manuscript 01/_manuscript/index.qmd
working on table 2

@@
 ---
-title: Does Hemispheric Lateralization Influence Aggressive Treatment in Patients with Intracerebral Hemorrhage?
-subtitle: An Analysis of ATACH-2 and ERICH
+title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
+subtitle: A Bayesian Analysis of ATACH-2 and ERICH
 authors:
@@
     affiliation: Department of Neurology, Wake Forest University School of Medicine
+  - name: Andrew Asimos
+    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
+  - name: Jonathan Parish
+    affiliation: Carolina Neurosurgery & Spine Associates
+  - name: Nicholas Allen Morris
+    affiliation: Department of Neurology, University of Maryland School of Medicine
 date: last-modified
@@
 mainfont: Figtree
-
 ---
@@
 library(targets)
-library(dplyr)
-library(targets)
+library(tidyverse)
 library(withr)
 library(here)
+library(scales)
+library(gt)
+library(gtsummary)
+library(modelsummary)
+library(marginaleffects)
+library(tidybayes)
+library(bayesplot)
+library(ggdist)
+library(patchwork)
+
+# plot stuff
+theme_ich <- function(base_family = "Figtree",
+                      base_size = 14) {
+  theme_minimal(base_size = base_size) +
+    theme(
+      text = element_text(lineheight = 2),
+      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+      axis.text.y = element_text(size = rel(0.8)),
+      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+      axis.text.x = element_text(size = rel(2)),
+      axis.title.x = element_blank(),
+      legend.position = "bottom",
+      legend.justification = 1,
+      panel.grid = element_line(colour = "#F3F4F5"),
+      plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
+      plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
+    )
+}
 
@@
 
+left_hemisphere_count <- ich_aggressive |>
+  filter(ich_laterality == "Left") |>
+  count()
+
 ```
@@
 The clinical presentation of patients with stroke can vary based on the hemisphere affected, in particular with language, spatial orientation, and emotional processing.
-Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.[@legge_neglecting_2006; @di_legge_impact_2005-1] On the other hand, there is a hesitation among some practitioners to offer decompressive craniectomy in patients with dominant-hemispheric infarcts.
+Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.[@legge_neglecting_2006; @di_legge_impact_2005] On the other hand, there is a hesitation among some practitioners to offer decompressive craniectomy in patients with dominant-hemispheric infarcts.
 [@lin_decompressive_2021-1]
@@
 
-A total of `r sample_size` patients were included in the analysis.
+### Baseline Characteristics
 
-```{r}
+A total of `r sample_size` patients were included in the analysis, of which `r left_hemisphere_count` (`r left_hemisphere_count/sample_size`) had left hemispheric ICH.
+They had higher NIHSS scores, higher GCS scores, and lower rates of hypertension as compared to those with right hemispheric ICH (@tbl-1).
 
-library(gtsummary)
+```{r table_1}
+#| label: tbl-1
+#| tbl-cap: "Baseline Characteristics"
 
@@
 
+### Left hemispheric ICH and neurosurgical intervention
+
+::: {.callout-note}
+What am I trying to do here?
+I want to present the results of the logistic regression showing that after controlling for other factors, there is still a higher probability of a right-hemisphere ICH patient receiving neurosurgical intervention.
+
+I have to change the baseline factor in `ich_laterality` to "Left".
+
+I should report the following: point estimate, 95% CIs, and posterior probability that effect is greater than 0, or for minimal clinically important difference.
+:::
+
+```{r neurosurgery}
+models <- list(
+  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neurosurgery_canonical)),
+  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_evd_canonical)),
+  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_trach_canonical)),
+  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
+  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical)),
+  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
+  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical))
+)
+
+table_2 <- with_dir(here(), tar_read(table_2_aggressive))
+
+summary <- modelsummary(models,
+  exponentiate = TRUE,
+  fmt = fmt_significant(3),
+  estimate = "{estimate}",
+  statistic = "{conf.low} - {conf.high}",
+  coef_omit = "^(?!b_ich_lat)",
+  coef_rename = c("b_ich_lateralityRight" = "Right Hemisphere ICH"),
+  gof_map = c("none")
+)
+
+summary
+
+```
+
+### Left hemispheric ICH and other measures of aggressive care
+
+### Sensitivity analysis
+
+## Discussion
+
+Hello
+
 ## Acknowledgments {.appendix}

### 2024-03-07 manuscript 01/index.qmd
working on table 2

--- 2024-03-07 manuscript 01/_manuscript/index.qmd
@@
 library(patchwork)
+library(tinytable)
 
@@
 ```{r neurosurgery}
+neurosurgery_mod <- with_dir(here(), tar_read(m_posterior_neurosurgery_canonical))
+
 models <- list(
@@
 
-summary
+noquote(summary$"Neurosurgical Intervention"[1])
+
+odds_ratios <- list(
+  "Neurosurgical Intervention" = noquote(summary$"Neurosurgical Intervention"[1]),
+  "External Ventricular Drain" = noquote(summary$"External Ventricular Drain"[1]),
+  "Tracheostomy" = noquote(summary$"Tracheostomy"[1]),
+  "Days of Mechanical Ventilation" = noquote(summary$"Days of Mechanical Ventilation"[1]),
+  "DNR Order" = noquote(summary$"DNR Order"[1]),
+  "Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Withdrawal of Life-Sustaining Therapy"[1]),
+  "Early Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Early Withdrawal of Life-Sustaining Therapy"[1])
+)
+
+credible_ints <- list(
+  "Neurosurgical Intervention" = noquote(summary$"Neurosurgical Intervention"[2]),
+  "External Ventricular Drain" = noquote(summary$"External Ventricular Drain"[2]),
+  "Tracheostomy" = noquote(summary$"Tracheostomy"[2]),
+  "Days of Mechanical Ventilation" = noquote(summary$"Days of Mechanical Ventilation"[2]),
+  "DNR Order" = noquote(summary$"DNR Order"[2]),
+  "Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Withdrawal of Life-Sustaining Therapy"[2]),
+  "Early Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Early Withdrawal of Life-Sustaining Therapy"[2])
+)
+
+table_2
 

### 2024-03-08 manuscript 01/index.qmd
making progress on table 2

@@
 
-noquote(summary$"Neurosurgical Intervention"[1])
+# gather odds ratios from modelsummary object
+odds_ratios <- summary |> map(pluck(1))
+odds_ratios[1] <- NULL
 
-odds_ratios <- list(
-  "Neurosurgical Intervention" = noquote(summary$"Neurosurgical Intervention"[1]),
-  "External Ventricular Drain" = noquote(summary$"External Ventricular Drain"[1]),
-  "Tracheostomy" = noquote(summary$"Tracheostomy"[1]),
-  "Days of Mechanical Ventilation" = noquote(summary$"Days of Mechanical Ventilation"[1]),
-  "DNR Order" = noquote(summary$"DNR Order"[1]),
-  "Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Withdrawal of Life-Sustaining Therapy"[1]),
-  "Early Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Early Withdrawal of Life-Sustaining Therapy"[1])
+# gather credible intervals from modelsummary object
+credible_ints <- summary |> map(pluck(2))
+credible_ints[1] <- NULL
+
+# gather posterior probabilities of various magnitudes of difference
+neurosurgery <- models$"Neurosurgical Intervention" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
   )
 
-credible_ints <- list(
-  "Neurosurgical Intervention" = noquote(summary$"Neurosurgical Intervention"[2]),
-  "External Ventricular Drain" = noquote(summary$"External Ventricular Drain"[2]),
-  "Tracheostomy" = noquote(summary$"Tracheostomy"[2]),
-  "Days of Mechanical Ventilation" = noquote(summary$"Days of Mechanical Ventilation"[2]),
-  "DNR Order" = noquote(summary$"DNR Order"[2]),
-  "Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Withdrawal of Life-Sustaining Therapy"[2]),
-  "Early Withdrawal of Life-Sustaining Therapy" = noquote(summary$"Early Withdrawal of Life-Sustaining Therapy"[2])
+evd <- models$"External Ventricular Drain" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
   )
 
-table_2
+trach <- models$"Tracheostomy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+vent <- models$"Days of Mechanical Ventilation" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+dnr <- models$"DNR Order" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+wlst <- models$"Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+early_wlst <- models$"Early Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+# combine probability of OR > 1 (prob of difference)
+prob_difference <- list(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 1),
+  "External Ventricular Drain" = pluck(evd, 1),
+  "Tracheostomy" = pluck(trach, 1),
+  "Days of Mechanical Ventilation" = pluck(vent, 1),
+  "DNR Order" = pluck(dnr, 1),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1)
+)
+
+# combine % of posterior within the ROPE
+rope <- list(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 4),
+  "External Ventricular Drain" = pluck(evd, 4),
+  "Tracheostomy" = pluck(trach, 4),
+  "Days of Mechanical Ventilation" = pluck(vent, 4),
+  "DNR Order" = pluck(dnr, 4),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4)
+)
+
+table1 <- bind_rows(odds_ratios, credible_ints)
+table2 <- bind_rows(prob_difference, rope) |> mutate(across(1:7, as.character))
+
+ult_table <- bind_rows(table1, table2)
 
@@
 
-### Sensitivity analysis
+### Subgroup analyses
+
+### Sensitivity analyses
 

### 2024-03-08 manuscript 01/index.qmd
Update index.qmd

@@
   "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
-  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical)),
   "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
-  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical))
+  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical)),
+  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical))
 )
@@
   "Days of Mechanical Ventilation" = pluck(vent, 1),
-  "DNR Order" = pluck(dnr, 1),
   "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
-  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1)
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1),
+  "DNR Order" = pluck(dnr, 1)
 )
@@
   "Days of Mechanical Ventilation" = pluck(vent, 4),
-  "DNR Order" = pluck(dnr, 4),
   "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
-  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4)
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4),
+  "DNR Order" = pluck(dnr, 4)
 )
@@
 
+ult_table <- t(ult_table)
 ```

### 2024-03-08 manuscript 01/_manuscript/index.qmd
flextable table 2

@@
 library(patchwork)
+library(tinytable)
+library(flextable)
 
@@
 
-summary
+# gather odds ratios from modelsummary object
+odds_ratios <- summary |> map(pluck(1))
+odds_ratios[1] <- NULL
 
+# gather credible intervals from modelsummary object
+credible_ints <- summary |> map(pluck(2))
+credible_ints[1] <- NULL
+
+# gather posterior probabilities of various magnitudes of difference
+neurosurgery <- models$"Neurosurgical Intervention" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+evd <- models$"External Ventricular Drain" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+trach <- models$"Tracheostomy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+vent <- models$"Days of Mechanical Ventilation" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+dnr <- models$"DNR Order" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+wlst <- models$"Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+early_wlst <- models$"Early Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+# combine probability of OR > 1 (prob of difference)
+prob_difference <- list(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 1),
+  "External Ventricular Drain" = pluck(evd, 1),
+  "Tracheostomy" = pluck(trach, 1),
+  "Days of Mechanical Ventilation" = pluck(vent, 1),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1),
+  "DNR Order" = pluck(dnr, 1)
+)
+
+# combine % of posterior within the ROPE
+rope <- list(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 4),
+  "External Ventricular Drain" = pluck(evd, 4),
+  "Tracheostomy" = pluck(trach, 4),
+  "Days of Mechanical Ventilation" = pluck(vent, 4),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4),
+  "DNR Order" = pluck(dnr, 4)
+)
+
+table1 <- bind_rows(odds_ratios, credible_ints)
+table2 <- bind_rows(prob_difference, rope) |> mutate(across(1:7, as.character))
+
+ult_table <- bind_rows(table1, table2)
+
+ult_table <- as_tibble(t(ult_table))
+
+tibble_2 <- as_tibble(table_2) |> select(-1)
+
+tibble_2
+
+column_names <- c(
+  "Neurosurgical Intervention",
+  "External Ventricular Drain",
+  "Tracheostomy",
+  "Days of Mechanical Ventilation",
+  "Withdrawal of Life-Sustaining Therapy",
+  "Early Withdrawal of Life-Sustaining Therapy",
+  "DNR Order"
+)
+
+combined <- bind_cols(column_names, tibble_2, ult_table) |>
+  rename(
+    "Outcome" = ...1,
+    "aOR" = V1,
+    "95% Credible Interval" = V2,
+    "Probability of Difference (aOR >1)" = V3,
+    "% ROPE" = V4
+  )
+
+flextable(combined)
 ```

### 2024-03-08 manuscript 01/index.qmd
flextable table 2

@@
 library(tinytable)
+library(flextable)
 
@@
 ```{r neurosurgery}
-neurosurgery_mod <- with_dir(here(), tar_read(m_posterior_neurosurgery_canonical))
-
 models <- list(
@@
 
-ult_table <- t(ult_table)
+ult_table <- as_tibble(t(ult_table))
+
+tibble_2 <- as_tibble(table_2) |> select(-1)
+
+column_names <- c(
+  "Neurosurgical Intervention",
+  "External Ventricular Drain",
+  "Tracheostomy",
+  "Days of Mechanical Ventilation",
+  "Withdrawal of Life-Sustaining Therapy",
+  "Early Withdrawal of Life-Sustaining Therapy",
+  "DNR Order"
+)
+
+combined <- bind_cols(column_names, tibble_2, ult_table) |>
+  rename(
+    "Outcome" = ...1,
+    "aOR" = V1,
+    "95% Credible Interval" = V2,
+    "Probability of Difference (aOR >1)" = V3,
+    "% ROPE" = V4
+  )
+
+vector <- as.numeric(noquote(unlist(odds_ratios, use.names = FALSE)))
+
+flextable(combined)
+get_flextable_defaults()
+
 ```

### 2024-03-08 manuscript 01/index.qmd
table_2 complete?

--- 2024-03-08 manuscript 01/index.qmd
@@
 library(here)
-library(scales)
+library(formattable)
 library(gt)
@@
 
+# load data and models
 with_dir(here(), tar_load(ich_aggressive))
@@
   count()
+
+models <- list(
+  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neurosurgery_canonical)),
+  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_evd_canonical)),
+  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_trach_canonical)),
+  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
+  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
+  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical)),
+  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical))
+)
 
@@
 
-The analysis code is publicly available as outlined by the NINDS Data Request Agreement at [https://github.com/nikhilpatelmd/ich-laterality](https://github.com/nikhilpatelmd/ich-laterality).
+As outlined in the NINDS Data Request Agreement, the analysis code is publicly available at [https://github.com/nikhilpatelmd/ich-laterality](https://github.com/nikhilpatelmd/ich-laterality).
 
@@
 
-A total of `r sample_size` patients were included in the analysis, of which `r left_hemisphere_count` (`r left_hemisphere_count/sample_size`) had left hemispheric ICH.
+A total of `r sample_size` patients were included in the analysis, of which `r left_hemisphere_count` (`r percent(left_hemisphere_count/sample_size, digits = 1)`) had left hemispheric ICH.
 They had higher NIHSS scores, higher GCS scores, and lower rates of hypertension as compared to those with right hemispheric ICH (@tbl-1).
@@
 
-theme_gtsummary_compact(set_theme = TRUE, font_size = 14)
-with_dir(here(), tar_read(table_1_aggressive))
+with_dir(here(), tar_read(table_1_aggressive)) |>
+  as_flex_table() |>
+  font(part = "all", fontname = "Figtree") |>
+  fontsize(size = 11, part = "header") |>
+  fontsize(size = 10) |>
+  theme_zebra() |>
+  line_spacing(space = 1.25, part = "all") |>
+  padding(padding = 4, part = "all")
 
@@
 
-::: {.callout-note}
-What am I trying to do here?
-I want to present the results of the logistic regression showing that after controlling for other factors, there is still a higher probability of a right-hemisphere ICH patient receiving neurosurgical intervention.
-
-I have to change the baseline factor in `ich_laterality` to "Left".
-
-I should report the following: point estimate, 95% CIs, and posterior probability that effect is greater than 0, or for minimal clinically important difference.
-:::
-
 ```{r neurosurgery}
-neurosurgery_mod <- with_dir(here(), tar_read(m_posterior_neurosurgery_canonical))
-
-models <- list(
-  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neurosurgery_canonical)),
-  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_evd_canonical)),
-  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_trach_canonical)),
-  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
-  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
-  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical)),
-  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical))
-)
-
-table_2 <- with_dir(here(), tar_read(table_2_aggressive))
+#| label: tbl-2
+#| tbl-cap: "Association between hemispheric lateralization and care"#|
 
@@
   exponentiate = TRUE,
-  fmt = fmt_significant(3),
+  fmt = 2,
   estimate = "{estimate}",
@@
 
-ult_table <- bind_rows(table1, table2)
+pre_tab <- as_tibble(t(bind_rows(table1, table2)))
 
-ult_table <- t(ult_table)
+table_2 <- with_dir(here(), tar_read(table_2_aggressive))
+tibble_2 <- as_tibble(table_2) |> select(-1)
+
+column_names <- c(
+  "Neurosurgical Intervention",
+  "External Ventricular Drain",
+  "Tracheostomy",
+  "Days of Mechanical Ventilation",
+  "Withdrawal of Life-Sustaining Therapy",
+  "Early Withdrawal of Life-Sustaining Therapy",
+  "DNR Order"
+)
+
+combined <- bind_cols(column_names, tibble_2, pre_tab) |>
+  rename(
+    "Right Hemisphere" = "**Right**, N = 1,757",
+    "Left Hemisphere" = "**Left**, N = 1,720",
+    "Outcome" = ...1,
+    "aOR" = V1,
+    "95% Credible Interval" = V2,
+    "Probability of Difference (aOR >1)" = V3,
+    "% ROPE" = V4
+  )
+
+flextable(combined) |>
+  font(part = "all", fontname = "Figtree") |>
+  fontsize(size = 10) |>
+  theme_zebra() |>
+  line_spacing(space = 1.25, part = "all") |>
+  padding(padding = 4, part = "all") |>
+  bold(part = "header") |>
+  bold(i = 1) |>
+  padding(padding = 5, part = "all") |>
+  width(width = 50, j = 1) |>
+  width(width = 2, j = 2:3) |>
+  width(width = 4, j = 4:6)
+
 ```

### 2024-05-02 manuscripts/manuscript1.qmd
big moves on table 2

@@
 ---
-title: "Differences in Treatment Driven by Hemispheric Laterality"
+title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
 author: "Nikhil Patel, MD"
+subtitle: A Bayesian Re-Analysis of ATACH-2 and ERICH data
+authors:
+  - name: Nikhil Patel
+    orcid: 0000-0002-3193-8590
+    affiliation: Department of Neurology, Wake Forest University School of Medicine
+    roles: data analysis, writing
+    corresponding: true
+  - name: Matt Jaffa
+    affiliation: Department of Neurology, University of Connecticut School of Medicine
+  - name: Rahul Karamchandani
+    affiliation: Department of Neurology, Wake Forest University School of Medicine
+  - name: Andrew Asimos
+    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
+  - name: Jonathan Parish
+    affiliation: Carolina Neurosurgery & Spine Associates
+  - name: Nicholas Allen Morris
+    affiliation: Department of Neurology, University of Maryland School of Medicine
+date: last-modified
+abstract: |
+  Hemispheric laterality blah blah blah
+table-of-contents: true
+toc-title: Contents
 ---
 
-This is the actual manuscript
+```{r setup}
+#| echo: false
+#| output: false
 
-```{r}
+library(targets)
+library(tidyverse)
+library(withr)
+library(here)
+library(formattable)
+library(gt)
+library(gtsummary)
+library(modelsummary)
+library(marginaleffects)
+library(tidybayes)
+library(bayesplot)
+library(ggdist)
+library(patchwork)
+library(tinytable)
+library(flextable)
 
-library(tidyverse)
+# plot stuff
+theme_ich <- function(base_family = "Figtree",
+                      base_size = 14) {
+  theme_minimal(base_size = base_size) +
+    theme(
+      text = element_text(lineheight = 2),
+      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+      axis.text.y = element_text(size = rel(0.8)),
+      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+      axis.text.x = element_text(size = rel(2)),
+      axis.title.x = element_blank(),
+      legend.position = "bottom",
+      legend.justification = 1,
+      panel.grid = element_line(colour = "#F3F4F5"),
+      plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
+      plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
+    )
+}
+
+# load data and models
+with_dir(here(), tar_load(ich_aggressive))
+
+sample_size <- count(ich_aggressive)
+
+left_hemisphere_count <- ich_aggressive |>
+  filter(ich_laterality == "Left") |>
+  count()
+
+models <- list(
+  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neurosurgery_canonical)),
+  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_evd_canonical)),
+  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_trach_canonical)),
+  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
+  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
+  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical)),
+  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical))
+)
 
@@
 
-This is more [@noauthor_simvastatin_2023]
+## Introduction
+The clinical presentation of patients with stroke can vary based on the hemisphere affected, in particular with language, spatial orientation, and emotional processing.
+Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.[@legge_neglecting_2006; @di_legge_impact_2005] On the other hand, there is a hesitation among some practitioners to offer decompressive craniectomy in patients with dominant-hemispheric infarcts.
+[@lin_decompressive_2021-1]
+
+It is unclear how previous research in the ischemic stroke population translates to patients experiencing intracerebral hemorrhage. Not much has been investigated.
+
+We aimed to evaluate the effect of hemispheric lateralization on the probability of receiving aggressive care in patients with ICH.
+
+## Methods
+
+### Study Design and Participants
+
+We conducted a cohort study using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshi_intensive_2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH)[@woo_ethnicracial_2013] study.
+ATACH-2 was a large, multicenter, international, open-label, blinded-endpoint randomized controlled trial enrolling 1000 patients with ICH while ERICH was a large, multicenter, prospective observational study enrolling patients in the United States admitted with ICH.
+Study protocols and results of both studies have been previously published and are available elsewhere.
+
+We included all patients enrolled in these two studies with supratentorial location of ICH.
+We excluded patients with cerebellar or brainstem ICH as well as any patient with primary intraventricular hemorrhage (IVH).
+
+The study was approved by the institutional review board of the Wake Forest School of Medicine.
+
+#### Data Availability Statement
+
+The ATACH-2 (PI: Adnan I.
+Qureshi, MD, NCT01176565) and ERICH (PI: Daniel Woo, MD, MS, NCT01202864) data and analyses presented in this manuscript are based on the research files downloaded from the Archived Clinical Research web site of the National Institute of Neurologic Disease and Stroke (NINDS).
+The datasets can be obtained after approval of a written proposal as outlined on the NINDS website.
+As outlined in the NINDS Data Request Agreement, the analysis code is publicly available at [https://github.com/nikhilpatelmd/ich-laterality](https://github.com/nikhilpatelmd/ich-laterality).
+
+### Exposures and Outcomes
+
+Our exposure was hemispheric laterality (left versus right).
+Our primary outcome was neurosurgical intervention, defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.
+
+To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), tracheostomy, percutaneous endoscopic gastrostomy (PEG) tube, do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (< 72 from symptom onset).
+
+### Statistical Analysis
+
+#### Overall Analysis Plan
+
+## Results
+
+### Baseline Characteristics
+
+A total of `r sample_size` patients were included in the analysis, of which `r left_hemisphere_count` (`r percent(left_hemisphere_count/sample_size, digits = 1)`) had left hemispheric ICH.
+They had higher NIHSS scores, higher GCS scores, and lower rates of hypertension as compared to those with right hemispheric ICH (@tbl-1).
+
+```{r table1}
+#| label: tbl-1
+#| tbl-cap: "Baseline Characteristics"
+#| echo: false
+
+with_dir(here(), tar_read(table_1_aggressive))
+
+```
+
+### Left hemispheric ICH and neurosurgical intervention
+
+```{r neurosurgery}
+#| label: tbl-2
+#| tbl-cap: "Association between hemispheric lateralization and care"
+#| echo: false
+
+# gather odds ratios and 95% credible intervals for each outcome
+summary <- modelsummary(models,
+  exponentiate = TRUE,
+  fmt = 2,
+  estimate = c("aOR" = "estimate"),
+  statistic = c("95% Credible Interval" = "{conf.low} - {conf.high}"),
+  coef_omit = "^(?!b_ich_lat)",
+  coef_rename = c("b_ich_lateralityRight" = "Right Hemisphere ICH"),
+  gof_map = c("none"),
+  shape = model ~ term + statistic,
+)
+
+aOR_CIs_combined <- as_tibble(summary@table_dataframe) |>
+  rename("Outcome" = " ")
+
+aOR_CIs_combined[1] <- c(
+  "Neurosurgical Intervention",
+  "External Ventricular Drain",
+  "Tracheostomy",
+  "Days of Mechanical Ventilation",
+  "Withdrawal of Life-Sustaining Therapy",
+  "Early Withdrawal of Life-Sustaining Therapy",
+  "DNR Order"
+)
+
+# gather posterior probabilities of various magnitudes of difference for each outcome
+
+neurosurgery <- models$"Neurosurgical Intervention" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+evd <- models$"External Ventricular Drain" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+trach <- models$"Tracheostomy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+vent <- models$"Days of Mechanical Ventilation" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+dnr <- models$"DNR Order" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+wlst <- models$"Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+early_wlst <- models$"Early Withdrawal of Life-Sustaining Therapy" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+# combine probability of OR > 1 (prob of difference)
+prob_difference <- tibble(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 1),
+  "External Ventricular Drain" = pluck(evd, 1),
+  "Tracheostomy" = pluck(trach, 1),
+  "Days of Mechanical Ventilation" = pluck(vent, 1),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1),
+  "DNR Order" = pluck(dnr, 1)
+) |>
+  pivot_longer(
+    cols = everything(),
+    names_to = "Outcome",
+    values_to = "Probabilty of any difference (aOR > 1)"
+  )
+
+# combine % of posterior within the ROPE
+rope <- tibble(
+  "Neurosurgical Intervention" = pluck(neurosurgery, 4),
+  "External Ventricular Drain" = pluck(evd, 4),
+  "Tracheostomy" = pluck(trach, 4),
+  "Days of Mechanical Ventilation" = pluck(vent, 4),
+  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
+  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4),
+  "DNR Order" = pluck(dnr, 4)
+) |>
+  pivot_longer(
+    cols = everything(),
+    names_to = "Outcome",
+    values_to = "% within ROPE"
+  )
+
+table_almost_2 <- left_join(aOR_CIs_combined, prob_difference, by = "Outcome") |>
+  left_join(rope, by = "Outcome")
+
+table_2 <- with_dir(here(), tar_read(table_2_aggressive))
+tibble_2 <- as_tibble(table_2) |> select(-1)
+
+column_names <- c(
+  "Neurosurgical Intervention",
+  "External Ventricular Drain",
+  "Tracheostomy",
+  "Days of Mechanical Ventilation",
+  "Withdrawal of Life-Sustaining Therapy",
+  "Early Withdrawal of Life-Sustaining Therapy",
+  "DNR Order"
+)
+
+tibble_2 <- bind_cols(column_names, tibble_2) |>
+  rename(
+    "Right Hemisphere" = "**Right**, N = 1,757",
+    "Left Hemisphere" = "**Left**, N = 1,720",
+    "Outcome" = ...1
+  ) |>
+  full_join(table_almost_2, by = "Outcome")
+
+tt(tibble_2, digits = 4)
+
+```
+
+## Discussion
+
+## Conclusions

### 2024-06-20 manuscripts/manuscript1.qmd
updated intro

@@
 ## Introduction
-The clinical presentation of patients with stroke can vary based on the hemisphere affected, in particular with language, spatial orientation, and emotional processing.
-Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.[@legge_neglecting_2006; @di_legge_impact_2005] On the other hand, there is a hesitation among some practitioners to offer decompressive craniectomy in patients with dominant-hemispheric infarcts.
-[@lin_decompressive_2021-1]
+Intracerebral hemorrhage (ICH) remains a challenging disease, marked by its sudden onset, morbidity, and mortality.
+While ischemic stroke has garnered considerable attention and therapeutic advancements, the acute management of ICH remains limited to blood pressure control, coagulopathy reversal, and more recently surgical evacuation of hemorrhage.
+Overwhelmingly these efforts have focused on reducing mortality and preventing additional morbidity due to secondary injury.
 
-It is unclear how previous research in the ischemic stroke population translates to patients experiencing intracerebral hemorrhage. Not much has been investigated.
+Clinical presentations for patients with stroke are varied based on the hemisphere affected, where hemispheric dominance dictates cognitive processing, attention, language, and emotional responses.
+Consequently, when confronted with the occurrence of ICH, the specific hemisphere affected can profoundly influence not only the clinical presentation but also therapeutic strategies and prognostic outcomes.
 
-We aimed to evaluate the effect of hemispheric lateralization on the probability of receiving aggressive care in patients with ICH.
+Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.1,2 It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts3 despite growing literature that long-term outcomes are better for patients with left hemispheric injuries in terms of both function and health related quality of life (HRQoL).
+[@deb-chatterjiSideMattersDifferences2022]
+
+It is uncertain how the results of prior work in the ischemic stroke literature translate to patients experiencing acute ICH.
+We aimed to evaluate the effect of hemispheric lateralization of ICH on likelihood of receiving aggressive surgical intervention.
 

### 2024-06-20 R/aggressive_care_posterior_models.R

@@ -17,8 +17,7 @@ f_posterior_neurosurgery_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 }
 
@@ -37,8 +36,7 @@ f_posterior_neurosurgery_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
 }
@@ -60,8 +58,7 @@ f_posterior_evd_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -82,8 +79,7 @@ f_posterior_evd_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -106,8 +102,7 @@ f_posterior_trach_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -128,8 +123,7 @@ f_posterior_trach_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -152,8 +146,7 @@ f_posterior_vent_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -174,8 +167,7 @@ f_posterior_vent_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
[…]

### 2024-06-20 R/aggressive_care_prior_models.R

@@ -29,19 +29,18 @@ f_prior_neurosurgery_minimal <- function(dat) {
 
   model <- brm(neurosurgery_evac ~ ich_laterality,
     family = bernoulli(link = "logit"),
-    data = ich_aggressive,
+    data = dat,
     prior = c(
       set_prior("normal(-2.2, 0.5)", class = "Intercept"),
       set_prior("normal(0, 1)", class = "b")
     ),
     sample_prior = "only",
-    # cores = settings$cores,
-    # chains = settings$chains,
-    # threads = settings$threads,
-    # warmup = settings$warmup,
-    # iter = settings$iter,
-    # seed = settings$seed,
-    # backend = "cmdstanr"
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed
   )
 
   return(model)
@@ -63,11 +62,11 @@ f_prior_neurosurgery_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
+
 }
 
 ## EVD as outcome -----
@@ -88,8 +87,7 @@ f_prior_evd_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -111,8 +109,7 @@ f_prior_evd_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -136,8 +133,7 @@ f_prior_trach_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -159,8 +155,7 @@ f_prior_trach_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
[…]

### 2024-06-20 R/outcomes_prior_models.R

@@ -18,8 +18,7 @@ f_prior_mrs_90_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -41,8 +40,7 @@ f_prior_mrs_90_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "cmdstanr"
+    seed = settings$seed
   )
 
   return(model)
@@ -67,7 +65,7 @@ f_prior_mrs_180_minimal <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -90,7 +88,7 @@ f_prior_mrs_180_canonical <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -115,7 +113,7 @@ f_prior_mrs_365_minimal <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -138,7 +136,7 @@ f_prior_mrs_365_canonical <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -163,7 +161,7 @@ f_prior_euro_mobility_90_minimal <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -186,7 +184,7 @@ f_prior_euro_mobility_90_canonical <- function(dat) {
     warmup = settings$warmup,
     iter = settings$iter,
     seed = settings$seed,
-    backend = "cmdstanr"
+    backend = "rstan"
   )
 
   return(model)
@@ -211,7 +209,7 @@ f_prior_euro_mobility_180_minimal <- function(dat) {
     warmup = settings$warmup,
[…]

### 2024-06-20 _targets.R

@@ -7,6 +7,14 @@ suppressPackageStartupMessages(library(brms))
 
 # Taken from Andrew Heiss (https://github.com/andrewheiss/cautioning-canary/blob/master/_targets.R)
 
+#   - If using cmdstanr, you can use within-chain threading
+#   - (https://cran.r-project.org/web/packages/brms/vignettes/brms_threading.html)
+#   - On a four-core computer, use mc.cores = 2 and threads = threading(2)
+#   - On an eight+-core computer, use mc.cores = 4 and threads = threading(2)
+options(mc.cores = 4,
+        mc.threads = 2,
+        brms.backend = "rstan")
+
 options(
   tidyverse.quiet = TRUE,
   dplyr.summarise.inform = FALSE

### 2024-06-23 manuscripts/manuscript1.qmd
chanages

@@
 abstract: |
-  Hemispheric laterality blah blah blah
+
 table-of-contents: true
@@
 Intracerebral hemorrhage (ICH) remains a challenging disease, marked by its sudden onset, morbidity, and mortality.
-While ischemic stroke has garnered considerable attention and therapeutic advancements, the acute management of ICH remains limited to blood pressure control, coagulopathy reversal, and more recently surgical evacuation of hemorrhage.
+While ischemic stroke has garnered considerable attention and therapeutic advancements, the acute management of ICH remains limited to blood pressure control, coagulopathy reversal, and more recently surgical evacuation of hemorrhage [@pradillaTrialEarlyMinimally2024].
 Overwhelmingly these efforts have focused on reducing mortality and preventing additional morbidity due to secondary injury.
@@
 
-Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.1,2 It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts3 despite growing literature that long-term outcomes are better for patients with left hemispheric injuries in terms of both function and health related quality of life (HRQoL).
-[@deb-chatterjiSideMattersDifferences2022]
+Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.
+[@dileggeImpactLesionSide2005; @foerchDifferenceRecognitionRight2005; @portegiesLeftsidedStrokesAre2015;  @liStrokeLateralizationLarge2021] It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts3 despite literature demonstrating that long-term functional outcomes and health-related quality of life (HRQoL) are equivalent if not better for patients with left hemispheric injuries.
+[@deb-chatterjiSideMattersDifferences2022; @leggeNeglectingDifference2006; @beuscherFunctionalLongTermOutcome2017; @satoHigherMortalityPatients2015]
 
@@
 
-We conducted a cohort study using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshi_intensive_2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH)[@woo_ethnicracial_2013] study.
+We conducted a cohort study using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshiIntensiveBloodPressureLowering2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH) [@wooEthnicRacialVariations2013] study.
 ATACH-2 was a large, multicenter, international, open-label, blinded-endpoint randomized controlled trial enrolling 1000 patients with ICH while ERICH was a large, multicenter, prospective observational study enrolling patients in the United States admitted with ICH.
@@
 Our exposure was hemispheric laterality (left versus right).
-Our primary outcome was neurosurgical intervention, defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.
+Our primary outcome was neurosurgical intervention as a binary variable defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.
 
-To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), tracheostomy, percutaneous endoscopic gastrostomy (PEG) tube, do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (< 72 from symptom onset).
+To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), tracheostomy, do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (defined as < 72 from symptom onset).
+
+### Covariates
+
+We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias.
+Based on our specificiations, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH).
+We present results for both unadjusted and adjusted models.
 
 ### Statistical Analysis
+To estimate the effect of hemispheric laterality on the outcomes of interest, we fitted hierarchal Bayesian regression models.
+This report follows the Bayesian Analysis Reporting Guidelines [@kruschkeBayesianAnalysisReporting2021].
+All analyses were performed in R version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria).
 
-#### Overall Analysis Plan
+#### Priors
+
+#### Posterior distribution
+
+We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median of the posterior distribution; 2) posterior distribution 95% highest density interval (HDI); 3) probability of any difference (PD; the probability that the effect size is on the side of the point estimate); 4) probability of "significance" based on a region of practical equivalence (ROPE) using traditional criteria; 5) probability of the effect size is at least equal or greater than what was considered a minimally important difference (MCID), as defined by the study authors; 6) probability of the effect size is at least 1.5 times greater than the MCID (which we considered a "large" effect).
 
@@
 A total of `r sample_size` patients were included in the analysis, of which `r left_hemisphere_count` (`r percent(left_hemisphere_count/sample_size, digits = 1)`) had left hemispheric ICH.
-They had higher NIHSS scores, higher GCS scores, and lower rates of hypertension as compared to those with right hemispheric ICH (@tbl-1).
+They had higher NIHSS scores, lower GCS scores, and lower rates of hypertension as compared to those with right hemispheric ICH (@tbl-1).
 

### 2024-06-25 manuscripts/manuscript1.qmd
methods

@@
 ### Study Design and Participants
-
-We conducted a cohort study using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshiIntensiveBloodPressureLowering2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH) [@wooEthnicRacialVariations2013] study.
-ATACH-2 was a large, multicenter, international, open-label, blinded-endpoint randomized controlled trial enrolling 1000 patients with ICH while ERICH was a large, multicenter, prospective observational study enrolling patients in the United States admitted with ICH.
-Study protocols and results of both studies have been previously published and are available elsewhere.
+We conducted a post-hoc analysis using data from 2 prospective studies: the Antihypertensive Treatment of Acute Cerebral Hemorrhage II (ATACH-2) trial [@qureshiIntensiveBloodPressureLowering2016] and the Ethnic/Racial Variations of Intracerebral Hemorrhage (ERICH) [@wooEthnicRacialVariations2013] study.
+ATACH-2 was a large, multicenter, international, open-label, blinded-endpoint randomized controlled trial enrolling 1000 patients with ICH and randomizing them to either intesive blood pressure control (110-139 mm Hg) or standard blood pressure control (140-179 mm Hg).
+ERICH was a large, multicenter, prospective observational study enrolling patients in the United States admitted with ICH.
+Study protocols and results of both studies have been previously published.
 
@@
 
-#### Data Availability Statement
+### Exposures and Outcomes
+Our exposure was hemispheric laterality (left versus right).
+Our primary outcome was neurosurgical intervention as a binary variable defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.
 
+To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), receipt of tracheostomy, days of mechanical ventilation, receipt of do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (defined as < 72 from symptom onset).
+
+### Statistical Analysis
+To estimate the effect of hemispheric laterality on the outcomes of interest, we fitted hierarchal Bayesian regression models adjusted for the specified covariates below and study as a random intercept.
+This report follows suggestions from the Bayesian Analysis Reporting Guidelines [@kruschkeBayesianAnalysisReporting2021].
+For binary endpoints, hierarchal Bayesian logistic regression models were used.
+For ordinal endpoints, hierarchal Bayesian ordinal logisitic regression models were used.
+All analyses were performed in R version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria) using the 'brms' and 'tidybayes' packages.
+
+#### Covariates
+We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias.
+Based on our specifications, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH).
+
+#### Priors
+Three sets of priors were used in the analyses: 1)
+
+#### Posterior distribution
+We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median of the posterior distribution; 2) posterior distribution 95% highest density interval (HDI); 3) probability of any difference (PD; the probability that the effect size is on the side of the point estimate); 4) probability of "significance" based on a region of practical equivalence (ROPE) using traditional criteria; 5) probability of the effect size is at least equal or greater than what was considered a minimally important difference (MCID), as defined by the study authors; 6) probability of the effect size is at least 1.5 times greater than the MCID (which we considered a "large" effect).
+
+### Data Availability Statement
 The ATACH-2 (PI: Adnan I.
@@
 As outlined in the NINDS Data Request Agreement, the analysis code is publicly available at [https://github.com/nikhilpatelmd/ich-laterality](https://github.com/nikhilpatelmd/ich-laterality).
-
-### Exposures and Outcomes
-
-Our exposure was hemispheric laterality (left versus right).
-Our primary outcome was neurosurgical intervention as a binary variable defined as receiving a hematoma evacuation, decompressive craniectomy, or other craniotomy.
-
-To explore differences in other aspects of aggressive care, we defined several secondary outcomes: receipt of external ventricular drain (EVD), tracheostomy, do-not-resuscitate (DNR) order at any point in the hospitalization, withdrawal of life-sustaining therapy (WLST) at any point in the hospitalization, and early WLST (defined as < 72 from symptom onset).
-
-### Covariates
-
-We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias.
-Based on our specificiations, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH).
-We present results for both unadjusted and adjusted models.
-
-### Statistical Analysis
-To estimate the effect of hemispheric laterality on the outcomes of interest, we fitted hierarchal Bayesian regression models.
-This report follows the Bayesian Analysis Reporting Guidelines [@kruschkeBayesianAnalysisReporting2021].
-All analyses were performed in R version 4.4.1 (R Foundation for Statistical Computing, Vienna, Austria).
-
-#### Priors
-
-#### Posterior distribution
-
-We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median of the posterior distribution; 2) posterior distribution 95% highest density interval (HDI); 3) probability of any difference (PD; the probability that the effect size is on the side of the point estimate); 4) probability of "significance" based on a region of practical equivalence (ROPE) using traditional criteria; 5) probability of the effect size is at least equal or greater than what was considered a minimally important difference (MCID), as defined by the study authors; 6) probability of the effect size is at least 1.5 times greater than the MCID (which we considered a "large" effect).
 

### 2024-06-25 manuscripts/manuscript1.qmd
updated methods

@@
     orcid: 0000-0002-3193-8590
-    affiliation: Department of Neurology, Wake Forest University School of Medicine
+    affiliation: Department of Internal Medicine, Wake Forest University School of Medicine
     roles: data analysis, writing
@@
 Based on our specifications, we included the following covariates in our models: age, admission Glasgow coma score (GCS), ICH location (categorized as lobar, thalamic, or basal ganglia), baseline ICH volume (measured in millilters), and presence of intraventricular hemorrhage (IVH).
+The DAGs used in the analyses are available in the supplementary material.
 
 #### Priors
-Three sets of priors were used in the analyses: 1)
+Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default vague priors as defined by the 'brms' software package; 2) flat priors; 3) weakly informative priors based on previous literature.
+A description of the reasons for these priors is available in the supplementary material.
 
 #### Posterior distribution
-We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median of the posterior distribution; 2) posterior distribution 95% highest density interval (HDI); 3) probability of any difference (PD; the probability that the effect size is on the side of the point estimate); 4) probability of "significance" based on a region of practical equivalence (ROPE) using traditional criteria; 5) probability of the effect size is at least equal or greater than what was considered a minimally important difference (MCID), as defined by the study authors; 6) probability of the effect size is at least 1.5 times greater than the MCID (which we considered a "large" effect).
+We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05).
+A description of the reasons for these thresholds is available in the supplementary material.
+
+#### Missing value policies
 

### 2024-06-25 manuscripts/manuscript1.qmd
prior update

@@
 author: "Nikhil Patel, MD"
-subtitle: A Bayesian Re-Analysis of ATACH-2 and ERICH data
+subtitle: A Post Hoc Bayesian of ATACH-2 and ERICH data
 authors:
@@
 
-#### Covariates
+#### Covariate Selection
 We constructed directed acyclic graphs (DAGs) to model our causal assumptions and to appropriately select variables to minimize confounding and selection bias.
@@
 
-#### Priors
-Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default vague priors as defined by the 'brms' software package; 2) flat priors; 3) weakly informative priors based on previous literature.
+#### Prior Distribution Specification
+Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature.
 A description of the reasons for these priors is available in the supplementary material.
 
-#### Posterior distribution
+#### Computation and Description of the Posterior Distribution
 We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05).

### 2024-06-25 manuscripts/manuscript1_supplement.qmd
prior update

---
title: Patients with Left Hemispheric Intracerebral Hemorrhage are Less Likely to Receive Neurosurgical Intervention
author: "Nikhil Patel, MD"
subtitle: Supplementary Appendix
authors:
  - name: Nikhil Patel
    orcid: 0000-0002-3193-8590
    affiliation: Department of Internal Medicine, Wake Forest University School of Medicine
    roles: data analysis, writing
    corresponding: true
  - name: Matt Jaffa
    affiliation: Department of Neurology, University of Connecticut School of Medicine
  - name: Rahul Karamchandani
    affiliation: Department of Neurology, Wake Forest University School of Medicine
  - name: Andrew Asimos
    affiliation: Department of Emergency Medicine, Wake Forest University School of Medicine
  - name: Jonathan Parish
    affiliation: Carolina Neurosurgery & Spine Associates
  - name: Nicholas Allen Morris
    affiliation: Department of Neurology, University of Maryland School of Medicine
date: last-modified
abstract: |

table-of-contents: true
toc-title: Contents
---

```{r setup}
#| echo: false
#| output: false

library(targets)
library(tidyverse)
library(withr)
library(here)
library(modelsummary)
library(marginaleffects)
library(brms)
library(tidybayes)
library(bayesplot)
library(glue)
library(ggdist)
library(patchwork)
library(tinytable)
library(flextable)

# plot stuff
theme_ich <- function(base_family = "Figtree",
                      base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      text = element_text(lineheight = 2),
      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
      axis.text.y = element_text(size = rel(0.8)),
      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
      axis.text.x = element_text(size = rel(2)),
      axis.title.x = element_blank(),
      legend.position = "bottom",
      legend.justification = 1,
      panel.grid = element_line(colour = "#F3F4F5"),
      plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
    )
}

# visualization function
results_visual <- function(model, title) {
  model <- model
  subtitle <- title

  draws <- model |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    summarize()

  pred <- predictions(model, by = "ich_laterality") |>
    tibble()

  theme_ich <- function(base_family = "Inter",
                        base_size = 14) {
    theme_minimal(base_size = base_size) +
      theme(
        text = element_text(lineheight = 2),
        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
        axis.text.y = element_text(size = rel(0.8)),
        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
        axis.text.x = element_text(size = rel(2)),
        axis.title.x = element_blank(),
        legend.position = "bottom",
        legend.justification = 1,
        panel.grid = element_line(colour = "#F3F4F5"),
        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
      )
  }

  p1 <- predictions(model, by = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw, fill = ich_laterality)) +
    geom_density(alpha = .7) +
    scale_fill_manual(values = c("#422f05", "#E69F00")) +
    scale_x_continuous(labels = scales::percent) +
    labs(
      title = "Predicted Probability",
      fill = "ICH Laterality",
      y = NULL
    ) +
    theme_ich()

  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
    posterior_draws() |>
    ggplot(aes(x = draw)) +
    geom_density(color = "#422f05", fill = "#E69F00") +
    scale_x_continuous(labels = scales::percent) +
    labs(
      title = "Average Difference between Left vs. Right Hemisphere",
      y = NULL
    ) +
    geom_vline(
      xintercept = 0, linetype = "dashed",
      color = "#999999", linewidth = 1
    ) +
    theme_ich()

  p3 <- model |>
    spread_draws(b_ich_lateralityRight) |>
    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
    ggplot(aes(ich_right_or)) +
    geom_density(color = "#422f05", fill = "#E69F00") +
    # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
    scale_x_continuous(breaks = seq(0, 3, 0.25)) +
    labs(
      title = "Odds Ratio - Right Hemisphere",
      y = NULL,
      fill = NULL
    ) +
    geom_vline(
      xintercept = 1, linetype = "dashed",
      color = "#999999", linewidth = 1
    ) +
    theme_ich() +
    theme(legend.position = "none")

  total <- p1 / p2 / p3

  final <- total +
    plot_annotation(
      title = glue(subtitle),
      theme = theme(plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)))
    )

  return(final)
}

# load data and models
with_dir(here(), tar_load(ich_aggressive))
with_dir(here(), tar_load(dag_aggressive))
with_dir(here(), tar_load(m_prior_default_neurosurgery))
with_dir(here(), tar_load(m_prior_informative_neurosurgery))

```

# Details of Statistical Analysis

## Covariate Selection

```{r dags}

dag_aggressive

```

## Prior Distribution Specification

We show prior predictive checks for 2(?) sets of priors.

### Default Priors

```{r default prior predictive check - neurosurgical intervention}
#| echo: false
#| fig-height: 12

avg_predictions(m_prior_default_neurosurgery, variables = "ich_laterality")

results_visual(m_prior_default_neurosurgery, "Prior Predictive Check - Neurosurgical Intervention")

```

### Weakly Informative Priors

```{r informative prior predictive check - neurosurgical intervention}
#| echo: false
#| fig-height: 12

avg_predictions(m_prior_informative_neurosurgery, variables = "ich_laterality")

results_visual(m_prior_informative_neurosurgery, "Prior Predictive Check - Neurosurgical Intervention")

```

## Posterior Simulation Diagnostics

### MCMC Chain Convergence and Resolution

### Posterior predictive checks

###

## Sensitivity Analyses

### 2024-06-25 R/aggressive_care_prior_models.R

@@ -3,20 +3,14 @@
 model_setup <- function() {
   # Settings
   CHAINS <- 4
-  CORES <- 6
-  THREADS <- threading(4)
+  CORES <- 3
+  THREADS <- threading(2)
   ITER <- 2000
   WARMUP <- 1000
   BAYES_SEED <- 908
 
-  # Priors
-  priors_vague <- c(
-    set_prior("normal(0, 10)", class = "Intercept"),
-    set_prior("normal(0, 3)", class = "b")
-  )
-
   return(list(
-    chains = CHAINS, cores = CORES, threads = THREADS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED, priors_vague = priors_vague
+    chains = CHAINS, cores = CORES, threads = THREADS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED
   ))
 }
 
@@ -24,14 +18,13 @@ model_setup <- function() {
 
 ## Neurosurgery as outcome -----
 
-f_prior_neurosurgery_minimal <- function(dat) {
+f_prior_default_neurosurgery <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(neurosurgery_evac ~ ich_laterality,
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
       set_prior("normal(0, 1)", class = "b")
     ),
     sample_prior = "only",
@@ -44,16 +37,17 @@ f_prior_neurosurgery_minimal <- function(dat) {
   )
 
   return(model)
+
 }
 
-f_prior_neurosurgery_canonical <- function(dat) {
+f_prior_informative_neurosurgery <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
+      set_prior("normal(-20, 0.5)", class = "Intercept"),
[…]

### 2024-06-25 _targets.R

@@ -11,9 +11,9 @@ suppressPackageStartupMessages(library(brms))
 #   - (https://cran.r-project.org/web/packages/brms/vignettes/brms_threading.html)
 #   - On a four-core computer, use mc.cores = 2 and threads = threading(2)
 #   - On an eight+-core computer, use mc.cores = 4 and threads = threading(2)
-options(mc.cores = 4,
+options(mc.cores = 3,
         mc.threads = 2,
-        brms.backend = "rstan")
+        brms.backend = "cmdstanr")
 
 options(
   tidyverse.quiet = TRUE,
@@ -63,8 +63,8 @@ tar_plan(
 
   ### Priors ----
   settings = model_setup(),
-  m_prior_neurosurgery_minimal = f_prior_neurosurgery_minimal(ich_aggressive),
-  m_prior_neurosurgery_canonical = f_prior_neurosurgery_canonical(ich_aggressive),
+  m_prior_default_neurosurgery = f_prior_default_neurosurgery(ich_aggressive),
+  m_prior_informative_neurosurgery = f_prior_informative_neurosurgery(ich_aggressive),
   m_prior_evd_minimal = f_prior_evd_minimal(ich_aggressive),
   m_prior_evd_canonical = f_prior_evd_canonical(ich_aggressive),
   m_prior_trach_minimal = f_prior_trach_minimal(ich_aggressive),

### 2024-06-26 manuscripts/manuscript1_supplement.qmd
prior descriptions
Specified 4 priors, have to run prior predictive checks

@@
 ```{r dags}
+#| echo: false
 
-dag_aggressive
+# dag_aggressive
 
@@
 
-## Prior Distribution Specification
+## Prior Distribution Definitions
 
-We show prior predictive checks for 2(?) sets of priors.
+```{r prior definitions}
+#| echo: false
+#| output: false
 
-### Default Priors
+neutral_mean <- 0
+neutral_sd <- 0.35
+neutral_lower_OR <- exp(neutral_mean - (1.96 * neutral_sd))
+neutral_upper_OR <- exp(neutral_mean + (1.96 * neutral_sd))
 
-```{r default prior predictive check - neurosurgical intervention}
-#| echo: false
-#| fig-height: 12
+left_mean <- -0.22
+left_sd <- 0.175
+left_lower_OR <- exp(left_mean - (1.96 * left_sd))
+left_upper_OR <- exp(left_mean + (1.96 * left_sd))
 
-avg_predictions(m_prior_default_neurosurgery, variables = "ich_laterality")
-
-results_visual(m_prior_default_neurosurgery, "Prior Predictive Check - Neurosurgical Intervention")
+right_mean <- 0.18
+right_sd <- 0.175
+right_lower_OR <- exp(right_mean - (1.96 * right_sd))
+right_upper_OR <- exp(right_mean + (1.96 * right_sd))
 
@@
 
-### Weakly Informative Priors
+In defining prior distributions, we generally followed suggestions from previous literature.
+[@zampieriUsingBayesianMethods2021; @kruschkeBayesianAnalysisReporting2021; @gelmanBayesianWorkflow2020] Specifically, we defined 4 sets of priors for each outcome:
+
+**Neutral prior**: mean of **`r neutral_mean`** and standard deviation of **`r neutral_sd`**, centering around an OR of **`r exp(neutral_mean)`** and 95% of the the probability mass for the OR between **`r round(neutral_lower_OR, digits = 2)`** to **`r round(neutral_upper_OR, digits = 2)`**.
+
+**Left hemisphere prior**: mean of **`r left_mean`** and standard deviation of **`r left_sd`**, centering around an OR of **`r round(exp(left_mean), digits = 2)`** and 95% of the the probability mass for the OR between **`r round(left_lower_OR, digits = 2)`** to **`r round(left_upper_OR, digits = 2)`**.
+
+**Right hemisphere prior**: mean of **`r right_mean`** and standard deviation of **`r right_sd`**, centering around an OR of **`r round(exp(right_mean), digits =2)`** and 95% of the the probability mass for the OR between **`r round(right_lower_OR, digits = 2)`** to **`r round(right_upper_OR, digits =2)`**.
+
+**Flat prior**: most mimics a frequentist analysis
+
+```{r prior graphical summary}
+#| echo: false
+#| fig-height: 6
+
+# Return dnorm(x) for 0 < x < 2, and NA for all other x
+dnorm_limit <- function(x) {
+  y <- dnorm(x)
+  y[x < 0] <- NA
+  return(y)
+}
+
+ggplot(data = tibble(x = -2:2), aes(x = x)) +
+  geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
+  # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
+  geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
+  scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
+  theme_ich()
+
+m_prior_informative_neurosurgery
+
+```
+
+### Neurosurgical Intervention
 
@@
 
-results_visual(m_prior_informative_neurosurgery, "Prior Predictive Check - Neurosurgical Intervention")
+ggplot(mutate(spread_draws(m_prior_informative_neurosurgery, b_ich_lateralityRight),
+  ich_right_or = exp(b_ich_lateralityRight)
+), aes(ich_right_or)) +
+  geom_density(color = "#422f05", fill = "#E69F00") +
+  scale_x_continuous(breaks = seq(-10, 10, 0.25)) +
+  labs(
+    title = "Odds Ratio - Right Hemisphere",
+    y = NULL, fill = NULL
+  ) +
+  geom_vline(
+    xintercept = 1, linetype = "dashed",
+    color = "#999999", linewidth = 1
+  ) +
+  theme_ich() +
+  theme(legend.position = "none")
 

### 2024-06-26 manuscripts/sample.R

@@ -1,0 +1,18 @@
+ggplot(mutate(spread_draws(m_prior_informative_neurosurgery, b_ich_lateralityRight),
+        ich_right_or = exp(b_ich_lateralityRight)
+), aes(ich_right_or)) +
+        geom_density(color = "#422f05", fill = "#E69F00") +
+        scale_x_continuous(breaks = seq(
+                0,
+                3, 0.25
+        )) +
+        labs(
+                title = "Odds Ratio - Right Hemisphere",
+                y = NULL, fill = NULL
+        ) +
+        geom_vline(
+                xintercept = 1, linetype = "dashed",
+                color = "#999999", linewidth = 1
+        ) +
+        theme_ich() +
+        theme(legend.position = "none")

### 2024-06-26 manuscripts/manuscript1_supplement.qmd
updated neurosurgery manuscripts

@@
 with_dir(here(), tar_load(ich_aggressive))
-with_dir(here(), tar_load(dag_aggressive))
-with_dir(here(), tar_load(m_prior_default_neurosurgery))
-with_dir(here(), tar_load(m_prior_informative_neurosurgery))
 
@@
 
-m_prior_informative_neurosurgery
-
 ```
@@
 #| fig-height: 12
+#|
 
-avg_predictions(m_prior_informative_neurosurgery, variables = "ich_laterality")
+with_dir(here(), tar_load(m_prior_neutral_neurosurgery))
+with_dir(here(), tar_load(m_prior_left_neurosurgery))
 
-ggplot(mutate(spread_draws(m_prior_informative_neurosurgery, b_ich_lateralityRight),
-  ich_right_or = exp(b_ich_lateralityRight)
-), aes(ich_right_or)) +
-  geom_density(color = "#422f05", fill = "#E69F00") +
-  scale_x_continuous(breaks = seq(-10, 10, 0.25)) +
-  labs(
-    title = "Odds Ratio - Right Hemisphere",
-    y = NULL, fill = NULL
-  ) +
-  geom_vline(
-    xintercept = 1, linetype = "dashed",
-    color = "#999999", linewidth = 1
-  ) +
-  theme_ich() +
-  theme(legend.position = "none")
+avg_predictions(m_prior_left_neurosurgery, variables = "ich_laterality")
+
+results_visual(m_prior_left_neurosurgery, "Prior Predictive Check")
 

### 2024-06-26 R/aggressive_care_prior_models.R

@@ -23,10 +23,10 @@ f_prior_neutral_neurosurgery <- function(dat) {
 
   model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
     family = bernoulli(link = "logit"),
-    data = dat,e
+    data = dat,
     prior = c(
-      set_prior("normal(0, 0.35)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
+      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(0, 0.35)", class = "b")
     ),
     sample_prior = "only",
     cores = settings$cores,
@@ -40,15 +40,59 @@ f_prior_neutral_neurosurgery <- function(dat) {
   return(model)
 }
 
-f_prior_informative_neurosurgery <- function(dat) {
+f_prior_left_neurosurgery <- function(dat) {
   settings <- model_setup()
 
   model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-20, 0.5)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
+      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-0.22, 0.175)", class = "b")
+    ),
+    sample_prior = "only",
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
+    seed = settings$seed
+  )
+
+  return(model)
+}
+
+f_prior_right_neurosurgery <- function(dat) {
+  settings <- model_setup()
+
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
+    family = bernoulli(link = "logit"),
+    data = dat,
+    prior = c(
+      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(0.18, 0.175)", class = "b")
+    ),
+    sample_prior = "only",
+    cores = settings$cores,
+    chains = settings$chains,
+    threads = settings$threads,
+    warmup = settings$warmup,
+    iter = settings$iter,
[…]

### 2024-06-26 _targets.R

@@ -11,9 +11,11 @@ suppressPackageStartupMessages(library(brms))
 #   - (https://cran.r-project.org/web/packages/brms/vignettes/brms_threading.html)
 #   - On a four-core computer, use mc.cores = 2 and threads = threading(2)
 #   - On an eight+-core computer, use mc.cores = 4 and threads = threading(2)
-options(mc.cores = 3,
-        mc.threads = 2,
-        brms.backend = "cmdstanr")
+options(
+  mc.cores = 3,
+  mc.threads = 2,
+  brms.backend = "cmdstanr"
+)
 
 options(
   tidyverse.quiet = TRUE,
@@ -63,8 +65,10 @@ tar_plan(
 
   ### Priors ----
   settings = model_setup(),
-  m_prior_default_neurosurgery = f_prior_default_neurosurgery(ich_aggressive),
-  m_prior_informative_neurosurgery = f_prior_informative_neurosurgery(ich_aggressive),
+  m_prior_neutral_neurosurgery = f_prior_neutral_neurosurgery(ich_aggressive),
+  m_prior_left_neurosurgery = f_prior_left_neurosurgery(ich_aggressive),
+  m_prior_right_neurosurgery = f_prior_right_neurosurgery(ich_aggressive),
+  m_prior_flat_neurosurgery = f_prior_flat_neurosurgery(ich_aggressive),
   m_prior_evd_minimal = f_prior_evd_minimal(ich_aggressive),
   m_prior_evd_canonical = f_prior_evd_canonical(ich_aggressive),
   m_prior_trach_minimal = f_prior_trach_minimal(ich_aggressive),

### 2024-06-27 manuscripts/manuscript1.qmd
more prior stuff

@@
 author: "Nikhil Patel, MD"
-subtitle: A Post Hoc Bayesian of ATACH-2 and ERICH data
+subtitle: A Post Hoc Bayesian Analysis using data from ATACH-2 and ERICH
 authors:
@@
 #| label: tbl-2
-#| tbl-cap: "Association between hemispheric lateralization and care"
+#| tbl-cap: "Association between hemispheric lateralization and aggressive care"
 #| echo: false
+#| warning: false
 
-# gather odds ratios and 95% credible intervals for each outcome
+# gather odds ratios and 95% credible intervals for each model
 summary <- modelsummary(models,

### 2024-06-27 R/aggressive_care_prior_models.R

@@ -291,7 +291,7 @@ f_prior_flat_tracheostomy <- function(dat) {
 f_prior_neutral_days_mechanical_ventilation <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
+  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
     family = cumulative(link = "logit"),
     data = dat,
     prior = c(
@@ -313,7 +313,7 @@ f_prior_neutral_days_mechanical_ventilation <- function(dat) {
 f_prior_left_days_mechanical_ventilation <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
+  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
     family = cumulative(link = "logit"),
     data = dat,
     prior = c(
@@ -335,7 +335,7 @@ f_prior_left_days_mechanical_ventilation <- function(dat) {
 f_prior_right_days_mechanical_ventilation <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
+  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
     family = cumulative(link = "logit"),
     data = dat,
     prior = c(
@@ -357,7 +357,7 @@ f_prior_right_days_mechanical_ventilation <- function(dat) {
 f_prior_flat_days_mechanical_ventilation <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
+  model <- brm(days_mechanical_ventilation ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
     family = cumulative(link = "logit"),
     data = dat,
     prior = c(

### 2024-06-28 manuscripts/manuscript1_supplement.qmd
prior predictive check function

@@
 
+#| echo: false
+#| fig-height: 12
+
 library(targets)
@@
 library(tinytable)
-library(flextable)
+library(gt)
 
@@
     )
-}
-
-# visualization function
-results_visual <- function(model, title) {
-  model <- model
-  subtitle <- title
-
-  draws <- model |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    summarize()
-
-  pred <- predictions(model, by = "ich_laterality") |>
-    tibble()
-
-  theme_ich <- function(base_family = "Inter",
-                        base_size = 14) {
-    theme_minimal(base_size = base_size) +
-      theme(
-        text = element_text(lineheight = 2),
-        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
-        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
-        axis.text.y = element_text(size = rel(0.8)),
-        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
-        axis.text.x = element_text(size = rel(2)),
-        axis.title.x = element_blank(),
-        legend.position = "bottom",
-        legend.justification = 1,
-        panel.grid = element_line(colour = "#F3F4F5"),
-        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
-        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
-      )
-  }
-
-  p1 <- predictions(model, by = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw, fill = ich_laterality)) +
-    geom_density(alpha = .7) +
-    scale_fill_manual(values = c("#422f05", "#E69F00")) +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Predicted Probability",
-      fill = "ICH Laterality",
-      y = NULL
-    ) +
-    theme_ich()
-
-  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw)) +
-    geom_density(color = "#422f05", fill = "#E69F00") +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Average Difference between Left vs. Right Hemisphere",
-      y = NULL
-    ) +
-    geom_vline(
-      xintercept = 0, linetype = "dashed",
-      color = "#999999", linewidth = 1
-    ) +
-    theme_ich()
-
-  p3 <- model |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    ggplot(aes(ich_right_or)) +
-    geom_density(color = "#422f05", fill = "#E69F00") +
-    # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
-    scale_x_continuous(breaks = seq(0, 3, 0.25)) +
-    labs(
-      title = "Odds Ratio - Right Hemisphere",
-      y = NULL,
-      fill = NULL
-    ) +
-    geom_vline(
-      xintercept = 1, linetype = "dashed",
-      color = "#999999", linewidth = 1
-    ) +
-    theme_ich() +
-    theme(legend.position = "none")
-
-  total <- p1 / p2 / p3
-
-  final <- total +
-    plot_annotation(
-      title = glue(subtitle),
-      theme = theme(plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)))
-    )
-
-  return(final)
 }
@@
 
-# Return dnorm(x) for 0 < x < 2, and NA for all other x
-dnorm_limit <- function(x) {
-  y <- dnorm(x)
-  y[x < 0] <- NA
-  return(y)
+# ggplot(data = tibble(x = -2:2), aes(x = x)) +
+#   geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
+#   # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
+#   geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
+#   scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
+#   theme_ich()
+
+```
+
+```{r prior predictive check - function}
+#| echo: false
+#| output: false
+## | # | | fig-height: 12
+
+f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
+  models <- list(
+    "Neutral Prior" = neutral_model,
+    "Left Hemisphere Prior" = left_model,
+    "Right Hemisphere Prior" = right_model,
+    "Flat Prior" = flat_model
+  )
+
+  # gather summary variables for each model
+  neutral <- models$"Neutral Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  left <- models$"Left Hemisphere Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  right <- models$"Right Hemisphere Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  flat <- models$"Flat Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  prior_summary <- rbind(neutral, left, right, flat) |>
+    mutate(prior = c("Neutral", "Left Hemisphere", "Right Hemisphere", "Flat")) |>
+    select(c(prior, aor, lower_ci, upper_ci, or_1, rope))
+
+  # add prior predictions
+  neutral_predictions <-
+    as_tibble(avg_predictions(models$"Neutral Prior", by = "ich_laterality")) |>
+    mutate("prior" = "Neutral") |>
+    select(c("prior", ich_laterality, estimate)) |>
+    pivot_wider(names_from = ich_laterality, values_from = estimate)
+
+  left_predictions <-
+    as_tibble(avg_predictions(models$"Left Hemisphere Prior", by = "ich_laterality")) |>
+    mutate("prior" = "Left Hemisphere") |>
+    select(c("prior", ich_laterality, estimate)) |>
+    pivot_wider(names_from = ich_laterality, values_from = estimate)
+
+  right_predictions <-
+    as_tibble(avg_predictions(models$"Right Hemisphere Prior", by = "ich_laterality")) |>
+    mutate("prior" = "Right Hemisphere") |>
+    select(c("prior", ich_laterality, estimate)) |>
+    pivot_wider(names_from = ich_laterality, values_from = estimate)
+
+  flat_predictions <-
+    as_tibble(avg_predictions(models$"Flat Prior", by = "ich_laterality")) |>
+    mutate("prior" = "Flat") |>
+    select(c("prior", ich_laterality, estimate)) |>
+    pivot_wider(names_from = ich_laterality, values_from = estimate)
+
+  prediction_summary <- rbind(neutral_predictions, left_predictions, right_predictions, flat_predictions)
+
+  prior_prediction_summary <- left_join(prediction_summary, prior_summary, by = "prior") |>
+    gt(rowname_col = "prior") |>
+    tab_stubhead(label = "Type of Prior") |>
+    fmt_number(columns = everything(), decimals = 2) |>
+    fmt_number(columns = 6, rows = 4, decimals = 0) |>
+    fmt_percent(columns = 2:3, rows = 1:4, decimals = 1) |>
+    tab_header(
+      title = md("**Prior Predictive Checks**")
+    ) |>
+    cols_label(
+      Left = "Left Hemisphere",
+      Right = "Right Hemisphere",
+      aor = "aOR",
+      lower_ci = "Lower 95% Credible Interval",
+      upper_ci = "Upper 95% Credible Interval",
+      or_1 = "Probability of Difference (aOR >1)",
+      rope = "% of Distribution outside of ROPE"
+    ) |>
+    cols_width(
+      1 ~ pct(50)
+    )
+
+  return(prior_prediction_summary)
 }
-
-ggplot(data = tibble(x = -2:2), aes(x = x)) +
-  geom_function(fun = dnorm, args = list(mean = -20, sd = 0.5), linewidth = 1) +
-  # stat_function(fun = dnorm_limit, geom = "area", color = "orange", alpha = 0.2) +
-  geom_vline(xintercept = -0.22, linetype = "dashed", color = "orange", linewidth = 1) +
-  scale_x_continuous(breaks = seq(-2, 2, 0.5)) +
-  theme_ich()
 
@@
 
-```{r informative prior predictive check - neurosurgical intervention}
+```{r prior predictive check - neurosurgery}
 #| echo: false
-#| fig-height: 12
-#|
+#| warning: false
 
-with_dir(here(), tar_load(m_prior_neutral_neurosurgery))
-with_dir(here(), tar_load(m_prior_left_neurosurgery))
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_neurosurgery))
+left_model <- with_dir(here(), tar_read(m_prior_left_neurosurgery))
+right_model <- with_dir(here(), tar_read(m_prior_right_neurosurgery))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))
 
-avg_predictions(m_prior_left_neurosurgery, variables = "ich_laterality")
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
 
-results_visual(m_prior_left_neurosurgery, "Prior Predictive Check")
+```
+
+### Ventriculostomy
+
+```{r prior predictive check - evd}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_evd))
+left_model <- with_dir(here(), tar_read(m_prior_left_evd))
+right_model <- with_dir(here(), tar_read(m_prior_right_evd))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_evd))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+predictions(neutral_model, by = "ich_laterality") |>
+  posterior_draws() |>
+  ggplot(aes(x = draw, fill = ich_laterality)) +
+  geom_density(alpha = .3) +
+  scale_fill_manual(values = c("#422f05", "#E69F00")) +
+  scale_x_continuous(labels = scales::percent) +
+  labs(
+    title = "Predicted Probability",
+    fill = "ICH Laterality",
+    y = NULL
+  )
+
+avg_comparisons(neutral_model, variables = "ich_laterality") |>
+  posterior_draws() |>
+  ggplot(aes(x = draw)) +
+  geom_density(color = "#422f05", fill = "#E69F00") +
+  scale_x_continuous(labels = scales::percent) +
+  labs(
+    title = "Average Difference between Left vs. Right Hemisphere",
+    y = NULL
+  ) +
+  geom_vline(
+    xintercept = 0, linetype = "dashed",
+    color = "#999999", linewidth = 1
+  )
+
+neutral_model |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  ggplot(aes(ich_right_or)) +
+  geom_density(color = "#422f05", fill = "#E69F00") +
+  # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
+  scale_x_continuous(breaks = seq(0, 3, 0.25)) +
+  labs(
+    title = "Odds Ratio - Right Hemisphere",
+    y = NULL,
+    fill = NULL
+  ) +
+  geom_vline(
+    xintercept = 1, linetype = "dashed",
+    color = "#999999", linewidth = 1
+  )
 

### 2024-06-28 manuscripts/function.R

@@ -1,0 +1,153 @@
+library(targets)
+library(tidyverse)
+library(withr)
+library(here)
+library(modelsummary)
+library(marginaleffects)
+library(brms)
+library(tidybayes)
+library(bayesplot)
+library(glue)
+library(ggdist)
+library(patchwork)
+library(tinytable)
+library(gt)
+
+neutral_mean <- 0
+neutral_sd <- 0.35
+neutral_lower_OR <- exp(neutral_mean - (1.96 * neutral_sd))
+neutral_upper_OR <- exp(neutral_mean + (1.96 * neutral_sd))
+
+left_mean <- -0.22
+left_sd <- 0.175
+left_lower_OR <- exp(left_mean - (1.96 * left_sd))
+left_upper_OR <- exp(left_mean + (1.96 * left_sd))
+
+right_mean <- 0.18
+right_sd <- 0.175
+right_lower_OR <- exp(right_mean - (1.96 * right_sd))
+right_upper_OR <- exp(right_mean + (1.96 * right_sd))
+
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_neurosurgery))
+left_model <- with_dir(here(), tar_read(m_prior_left_neurosurgery))
+right_model <- with_dir(here(), tar_read(m_prior_right_neurosurgery))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))
+
+f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
+  models <- list(
+    "Neutral Prior" = neutral_model,
+    "Left Hemisphere Prior" = left_model,
+    "Right Hemisphere Prior" = right_model,
+    "Flat Prior" = flat_model
+  )
+
+  # gather summary variables for each model
+  neutral <- models$"Neutral Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  left <- models$"Left Hemisphere Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
[…]

### 2024-06-29 manuscripts/manuscript1_supplement.qmd
stuff

@@
 
+### Prior Predictive Checks
+
 ```{r prior predictive check - function}
@@
       upper_ci = quantile(ich_right_or, 0.975),
-      or_1 = sum(ich_right_or > 1) / n(),
+      or_1 = sum(ich_right_or < 1) / n(),
       rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
@@
       1 ~ pct(50)
+    ) |>
+    tab_footnote(
+      footnote = "For left hemisphere prior, the p(aOR) < 1 is shown",
+      locations = cells_column_labels(columns = 7)
     )
@@
 
+results_visual <- function(model, title) {
+  model <- model
+  subtitle <- title
+
+  draws <- model |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize()
+
+  pred <- predictions(model, by = "ich_laterality") |>
+    tibble()
+
+  theme_ich <- function(base_family = "Inter",
+                        base_size = 14) {
+    theme_minimal(base_size = base_size) +
+      theme(
+        text = element_text(lineheight = 2),
+        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+        axis.text.y = element_text(size = rel(0.8)),
+        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+        axis.text.x = element_text(size = rel(2)),
+        axis.title.x = element_blank(),
+        legend.position = "bottom",
+        legend.justification = 1,
+        panel.grid = element_line(colour = "#F3F4F5"),
+        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
+        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
+      )
+  }
+
+  p1 <- predictions(model, by = "ich_laterality") |>
+    posterior_draws() |>
+    ggplot(aes(x = draw, fill = ich_laterality)) +
+    geom_density(alpha = .7) +
+    scale_fill_manual(values = c("#054224", "#ff8800")) +
+    scale_x_continuous(labels = scales::percent) +
+    labs(
+      title = "Predicted Probability",
+      fill = "ICH Laterality",
+      y = NULL
+    ) +
+    theme_ich()
+
+  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
+    posterior_draws() |>
+    ggplot(aes(x = draw)) +
+    geom_density(color = "#054224", fill = "#ff8800") +
+    scale_x_continuous(labels = scales::percent) +
+    labs(
+      title = "Average Difference between Left vs. Right Hemisphere",
+      y = NULL
+    ) +
+    geom_vline(
+      xintercept = 0, linetype = "dashed",
+      color = "#999999", linewidth = 1
+    ) +
+    theme_ich()
+
+  p3 <- model |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    ggplot(aes(ich_right_or)) +
+    geom_density(color = "#054224", fill = "#ff8800") +
+    # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
+    scale_x_continuous(breaks = seq(0, 3, 0.25)) +
+    labs(
+      title = "Odds Ratio - Right Hemisphere",
+      y = NULL,
+      fill = NULL
+    ) +
+    geom_vline(
+      xintercept = 1, linetype = "dashed",
+      color = "#999999", linewidth = 1
+    ) +
+    theme_ich() +
+    theme(legend.position = "none")
+
+  total <- p1 / p2 / p3
+
+  final <- total +
+    plot_annotation(
+      title = glue(subtitle),
+      theme = theme(plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)))
+    )
+
+  return(final)
+}
+
 ```
 
-### Neurosurgical Intervention
+#### Neurosurgical Intervention
 
@@
 #| warning: false
+#| fig-height: 12
 
@@
 
-### Ventriculostomy
+#### Ventriculostomy
 
@@
 
-predictions(neutral_model, by = "ich_laterality") |>
-  posterior_draws() |>
-  ggplot(aes(x = draw, fill = ich_laterality)) +
-  geom_density(alpha = .3) +
-  scale_fill_manual(values = c("#422f05", "#E69F00")) +
-  scale_x_continuous(labels = scales::percent) +
-  labs(
-    title = "Predicted Probability",
-    fill = "ICH Laterality",
-    y = NULL
-  )
+```
 
-avg_comparisons(neutral_model, variables = "ich_laterality") |>
-  posterior_draws() |>
-  ggplot(aes(x = draw)) +
-  geom_density(color = "#422f05", fill = "#E69F00") +
-  scale_x_continuous(labels = scales::percent) +
-  labs(
-    title = "Average Difference between Left vs. Right Hemisphere",
-    y = NULL
-  ) +
-  geom_vline(
-    xintercept = 0, linetype = "dashed",
-    color = "#999999", linewidth = 1
-  )
+#### Tracheostomy
 
-neutral_model |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  ggplot(aes(ich_right_or)) +
-  geom_density(color = "#422f05", fill = "#E69F00") +
-  # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
-  scale_x_continuous(breaks = seq(0, 3, 0.25)) +
-  labs(
-    title = "Odds Ratio - Right Hemisphere",
-    y = NULL,
-    fill = NULL
-  ) +
-  geom_vline(
-    xintercept = 1, linetype = "dashed",
-    color = "#999999", linewidth = 1
-  )
+```{r prior predictive check - tracheostomy}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_tracheostomy))
+left_model <- with_dir(here(), tar_read(m_prior_left_tracheostomy))
+right_model <- with_dir(here(), tar_read(m_prior_right_tracheostomy))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_tracheostomy))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+```
+
+#### Days of Mechanical Ventilation
+
+```{r prior predictive check - vent}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_days_mechanical_ventilation))
+left_model <- with_dir(here(), tar_read(m_prior_left_days_mechanical_ventilation))
+right_model <- with_dir(here(), tar_read(m_prior_right_days_mechanical_ventilation))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_days_mechanical_ventilation))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+```
+
+#### Comfort Care
+
+```{r prior predictive check - comfort}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_comfort))
+left_model <- with_dir(here(), tar_read(m_prior_left_comfort))
+right_model <- with_dir(here(), tar_read(m_prior_right_comfort))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_comfort))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+```
+
+#### Early WLST
+
+```{r prior predictive check - early wlst}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_early_wlst))
+left_model <- with_dir(here(), tar_read(m_prior_left_early_wlst))
+right_model <- with_dir(here(), tar_read(m_prior_right_early_wlst))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_early_wlst))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+```
+
+#### DNR
+
+```{r prior predictive check - dnr}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_dnr_binary))
+left_model <- with_dir(here(), tar_read(m_prior_left_dnr_binary))
+right_model <- with_dir(here(), tar_read(m_prior_right_dnr_binary))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_dnr_binary))
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
 

### 2024-06-29 R/results_visualization.R

@@ -33,7 +33,7 @@ results_visual <- function(model, title) {
     posterior_draws() |>
     ggplot(aes(x = draw, fill = ich_laterality)) +
     geom_density(alpha = .7) +
-    scale_fill_manual(values = c("#422f05", "#E69F00")) +
+    scale_fill_manual(values = c("#054224", "#ff8800")) +
     scale_x_continuous(labels = scales::percent) +
     labs(
       title = "Predicted Probability",
@@ -45,7 +45,7 @@ results_visual <- function(model, title) {
   p2 <- avg_comparisons(model, variables = "ich_laterality") |>
     posterior_draws() |>
     ggplot(aes(x = draw)) +
-    geom_density(color = "#422f05", fill = "#E69F00") +
+    geom_density(color = "#054224", fill = "#ff8800") +
     scale_x_continuous(labels = scales::percent) +
     labs(
       title = "Average Difference between Left vs. Right Hemisphere",
@@ -61,7 +61,7 @@ results_visual <- function(model, title) {
     spread_draws(b_ich_lateralityRight) |>
     mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
     ggplot(aes(ich_right_or)) +
-    geom_density(color = "#422f05", fill = "#E69F00") +
+    geom_density(color = "#054224", fill = "#ff8800") +
     # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
     scale_x_continuous(breaks = seq(0, 3, 0.25)) +
     labs(

### 2024-07-08 manuscripts/manuscript1.qmd
figures

@@
 library(tinytable)
-library(flextable)
+library(glue)
 
@@
       axis.text.y = element_text(size = rel(0.8)),
-      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
-      axis.text.x = element_text(size = rel(2)),
+      axis.title.y = element_text(size = rel(0.5), margin = margin(0, 4, 0, 0)),
+      axis.text.x = element_text(size = rel(1)),
       axis.title.x = element_blank(),
@@
       legend.justification = 1,
-      panel.grid = element_line(colour = "#F3F4F5"),
+      panel.grid = element_blank(),
+      panel.background = element_blank(),
       plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
@@
 models <- list(
-  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neurosurgery_canonical)),
-  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_evd_canonical)),
-  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_trach_canonical)),
-  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_vent_canonical)),
-  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_comfort_canonical)),
-  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_early_comfort_canonical)),
-  "DNR Order" = with_dir(here(), tar_read(m_posterior_dnr_canonical))
+  "Neurosurgical Intervention" = with_dir(here(), tar_read(m_posterior_neutral_neurosurgery)),
+  "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_neutral_evd)),
+  "Tracheostomy" = with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)),
+  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)),
+  "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_neutral_comfort)),
+  "Early Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_neutral_early_wlst)),
+  "DNR Order" = with_dir(here(), tar_read(m_posterior_neutral_dnr_binary))
 )
@@
   ) |>
-  full_join(table_almost_2, by = "Outcome")
+  full_join(table_almost_2, by = "Outcome") |>
+  gt()
 
-tt(tibble_2, digits = 4)
+tibble_2 |>
+  fmt_number(columns = 6:7, decimals = 2)
+
+```
+
+```{r figure 1}
+#| echo: false
+#| label: figure 1
+#| fig-width: 10
+
+predictions(models$"Neurosurgical Intervention", by = "ich_laterality") |>
+  posterior_draws() |>
+  ggplot(aes(x = draw, fill = ich_laterality)) +
+  geom_density(alpha = .7) +
+  scale_fill_manual(values = c("#054224", "#ff8800")) +
+  scale_y_continuous(
+    limits = c(0,100)
+  ) +
+  scale_x_continuous(
+    limits = c(0, 0.14),
+    breaks = seq(0, 0.14, 0.02),
+    labels = scales::percent
+  ) +
+  labs(
+    title = "Predicted Probability",
+    fill = "ICH Laterality",
+    y = NULL
+  ) +
+  theme_ich()
+
+```
+
+```{r figure 2}
+#| echo: false
+#| label: figure 2
+#| fig-width: 10
+
+predictions(models$"Neurosurgical Intervention", by = c("ich_laterality", "ich_location")) |>
+  posterior_draws() |>
+  ggplot(aes(x = draw, fill = ich_laterality)) +
+  geom_density(alpha = .7) +
+  facet_grid(~ich_location) +
+  scale_fill_manual(values = c("#054224", "#ff8800")) +
+  scale_y_continuous() +
+  scale_x_continuous(
+    limits = c(0, 0.2),
+    breaks = seq(0, 0.20, 0.05),
+    labels = scales::percent
+  ) +
+  labs(
+    fill = "ICH Laterality",
+    y = NULL
+  ) +
+  theme_ich()
 

### 2024-07-08 manuscripts/manuscript1_supplement.qmd
figures

@@
 theme_ich <- function(base_family = "Figtree",
-                      base_size = 14) {
+                      base_size = 10) {
   theme_minimal(base_size = base_size) +
@@
 
-results_visual <- function(model, title) {
+results_visual <- function(model) {
   model <- model
-  subtitle <- title
 
@@
 
-  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw)) +
-    geom_density(color = "#054224", fill = "#ff8800") +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Average Difference between Left vs. Right Hemisphere",
-      y = NULL
-    ) +
-    geom_vline(
-      xintercept = 0, linetype = "dashed",
-      color = "#999999", linewidth = 1
-    ) +
-    theme_ich()
-
-  p3 <- model |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    ggplot(aes(ich_right_or)) +
-    geom_density(color = "#054224", fill = "#ff8800") +
-    # stat_halfeye(aes(fill_ramp = stat(x > 1), fill = "#E69F00", color = "#422f05")) +
-    scale_x_continuous(breaks = seq(0, 3, 0.25)) +
-    labs(
-      title = "Odds Ratio - Right Hemisphere",
-      y = NULL,
-      fill = NULL
-    ) +
-    geom_vline(
-      xintercept = 1, linetype = "dashed",
-      color = "#999999", linewidth = 1
-    ) +
-    theme_ich() +
-    theme(legend.position = "none")
-
-  total <- p1 / p2 / p3
-
-  final <- total +
-    plot_annotation(
-      title = glue(subtitle),
-      theme = theme(plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)))
-    )
-
-  return(final)
+  return(p1)
 }
@@
 #| warning: false
-#| fig-height: 12
 
@@
 flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))
+
+results_visual(neutral_model)
 
@@
 
+results_visual(neutral_model)
+
 f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
@@
 flat_model <- with_dir(here(), tar_read(m_prior_flat_tracheostomy))
+
+results_visual(neutral_model)
 
@@
 flat_model <- with_dir(here(), tar_read(m_prior_flat_comfort))
+
+results_visual(neutral_model)
 
@@
 
+```{r posterior diagnostics setup}
+#| echo: false
+
+posterior_diagnostics <- function(mod) {
+  # plot code modified from Andrew Heiss (https://bayesf22-notebook.classes.andrewheiss.com/bayes-rules/13-chapter.html#simulating-the-posterior)
+
+  trace_plot <- mod |>
+    gather_draws(`^b_.*`, regex = TRUE) |>
+    ggplot(aes(x = .iteration, y = .value, color = factor(.chain))) +
+    geom_line(linewidth = 0.1) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    facet_wrap(vars(.variable), scales = "free_y")
+
+  trank_plot <- mod |>
+    gather_draws(`^b_.*`, regex = TRUE) |>
+    group_by(.variable) |>
+    mutate(draw_rank = rank(.value)) |>
+    ggplot(aes(x = draw_rank, color = factor(.chain))) +
+    stat_bin(geom = "step", binwidth = 250, position = position_identity(), boundary = 0) +
+    scale_color_viridis_d(option = "rocket", end = 0.85) +
+    facet_wrap(vars(.variable), scales = "free_y") +
+    theme(axis.text.y = element_blank(), axis.title.y = element_blank(), axis.ticks.y = element_blank())
+
+  # pp_check <- pp_check(mod, ndraws = 500)
+  # pp_check_bars <- pp_check(mod, ndraws = 500, type = "bars_grouped", group = "ich_laterality")
+
+  return((trank_plot / trace_plot))
+}
+
+```
+
 ### MCMC Chain Convergence and Resolution
+
+#### Neurosurgical Intervention
+
+```{r posterior diagnostics neurosurgery}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_neurosurgery)))
+
+```
+
+#### Ventriculostomy
+
+```{r posterior diagnostics evd}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_evd)))
+
+```
+
+#### Tracheostomy
+
+```{r posterior diagnostics trach}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)))
+
+```
+
+#### Days of Mechanical Ventilation
+
+```{r posterior diagnostics vent}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)))
+
+```
+
+#### Comfort Care
+
+```{r posterior diagnostics comfort}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_comfort)))
+
+```
+
+#### Early WLST
+
+```{r posterior diagnostics early wlst}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_early_wlst)))
+
+```
+
+#### DNR
+
+```{r posterior diagnostics dnr}
+#| echo: false
+#| fig-height: 10
+
+posterior_diagnostics(with_dir(here(), tar_read(m_posterior_neutral_dnr_binary)))
+
+```
 
@@
 
-###
+```{r posterior prediction function}
+#| echo: false
+
+post_pred <- function(mod) {
+  pp_check <- pp_check(mod, ndraws = 500)
+
+  pp_check_bars <- pp_check(mod, ndraws = 500, type = "bars_grouped", group = "ich_laterality")
+
+  return(pp_check / pp_check_bars)
+}
+
+```
+
+### Neurosurgical Intervention
+
+```{r posterior predictions neurosurgery}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_neurosurgery)))
+
+```
+
+### Ventriculostomy
+
+```{r posterior predictions evd}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_evd)))
+
+```
+
+### Tracheostomy
+
+```{r posterior predictions trach}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)))
+
+```
+
+### Days of Mechanical Ventilation
+
+```{r posterior predictions vent}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)))
+
+```
+
+### Comfort Care
+
+```{r posterior predictions comfort}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_comfort)))
+
+```
+
+### Early WLST
+
+```{r posterior predictions early wlst}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_early_wlst)))
+
+```
+
+### DNR
+
+```{r posterior predictions dnr}
+#| echo: false
+
+post_pred(with_dir(here(), tar_read(m_posterior_neutral_dnr_binary)))
+
+```
 

### 2024-07-08 R/aggressive_care_posterior_models.R

@@ -2,56 +2,15 @@
 
 ## Neurosurgery as outcome -----
 
-f_posterior_neurosurgery_minimal <- function(dat) {
+f_posterior_neutral_neurosurgery <- function(dat) {
   settings <- model_setup()
 
-  model <- brm(neurosurgery_evac ~ ich_laterality,
+  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh + (1 | study),
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
-    ),
-    cores = settings$cores,
-    chains = settings$chains,
-    threads = settings$threads,
-    warmup = settings$warmup,
-    iter = settings$iter,
-    seed = settings$seed
-  )
-}
-
-f_posterior_neurosurgery_canonical <- function(dat) {
-  settings <- model_setup()
-
-  model <- brm(neurosurgery_evac ~ ich_laterality + age + gcs_baseline + ich_location + ich_volume_baseline + ivh,
-    family = bernoulli(link = "logit"),
-    data = dat,
-    prior = c(
-      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
-    ),
-    cores = settings$cores,
-    chains = settings$chains,
-    threads = settings$threads,
-    warmup = settings$warmup,
-    iter = settings$iter,
-    seed = settings$seed
-  )
-
-}
-
-## EVD as outcome -----
-
-f_posterior_evd_minimal <- function(dat) {
-  settings <- model_setup()
-
-  model <- brm(evd ~ ich_laterality,
-    family = bernoulli(link = "logit"),
-    data = dat,
-    prior = c(
-      set_prior("normal(-2.2, 0.5)", class = "Intercept"),
-      set_prior("normal(0, 1)", class = "b")
+      set_prior("normal(-7, 0.35)", class = "Intercept"),
+      set_prior("normal(0, 0.5)", class = "b")
     ),
     cores = settings$cores,
     chains = settings$chains,
@@ -64,15 +23,143 @@ f_posterior_evd_minimal <- function(dat) {
   return(model)
 }
 
-f_posterior_evd_canonical <- function(dat) {
+f_posterior_left_neurosurgery <- function(dat) {
+  settings <- model_setup()
+
[…]

### 2024-07-08 R/aggressive_care_prior_models.R

@@ -385,7 +385,7 @@ f_prior_neutral_comfort <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-15, 0.5)", class = "Intercept"),
       set_prior("normal(0, 0.35)", class = "b")
     ),
     sample_prior = "only",
@@ -407,7 +407,7 @@ f_prior_left_comfort <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-15, 0.5)", class = "Intercept"),
       set_prior("normal(-0.22, 0.175)", class = "b")
     ),
     sample_prior = "only",
@@ -429,7 +429,7 @@ f_prior_right_comfort <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-15, 0.5)", class = "Intercept"),
       set_prior("normal(0.18, 0.175)", class = "b")
     ),
     sample_prior = "only",
@@ -475,7 +475,7 @@ f_prior_neutral_early_wlst <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-20, 0.5)", class = "Intercept"),
       set_prior("normal(0, 0.35)", class = "b")
     ),
     sample_prior = "only",
@@ -497,7 +497,7 @@ f_prior_left_early_wlst <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-20, 0.5)", class = "Intercept"),
       set_prior("normal(-0.22, 0.175)", class = "b")
     ),
     sample_prior = "only",
@@ -519,7 +519,7 @@ f_prior_right_early_wlst <- function(dat) {
     family = bernoulli(link = "logit"),
     data = dat,
     prior = c(
-      set_prior("normal(-5, 0.5)", class = "Intercept"),
+      set_prior("normal(-20, 0.5)", class = "Intercept"),
[…]

### 2024-07-08 _targets.R

@@ -14,7 +14,7 @@ suppressPackageStartupMessages(library(brms))
 options(
   mc.cores = 3,
   mc.threads = 2,
-  brms.backend = "cmdstanr"
+  brms.backend = "rstan"
 )
 
 options(
@@ -77,10 +77,10 @@ tar_plan(
   m_prior_left_tracheostomy = f_prior_left_tracheostomy(ich_aggressive),
   m_prior_right_tracheostomy = f_prior_right_tracheostomy(ich_aggressive),
   m_prior_flat_tracheostomy = f_prior_flat_tracheostomy(ich_aggressive),
-  m_prior_neutral_days_mechanical_ventilation = f_prior_neutral_days_mechanical_ventilation(ich_aggressive), # nolint: line_length_linter.
-  m_prior_left_days_mechanical_ventilation = f_prior_left_days_mechanical_ventilation(ich_aggressive), # nolint: line_length_linter.
-  m_prior_right_days_mechanical_ventilation = f_prior_right_days_mechanical_ventilation(ich_aggressive), # nolint: line_length_linter.
-  m_prior_flat_days_mechanical_ventilation = f_prior_flat_days_mechanical_ventilation(ich_aggressive), # nolint: line_length_linter.
+  m_prior_neutral_days_mechanical_ventilation = f_prior_neutral_days_mechanical_ventilation(ich_aggressive),
+  m_prior_left_days_mechanical_ventilation = f_prior_left_days_mechanical_ventilation(ich_aggressive),
+  m_prior_right_days_mechanical_ventilation = f_prior_right_days_mechanical_ventilation(ich_aggressive),
+  m_prior_flat_days_mechanical_ventilation = f_prior_flat_days_mechanical_ventilation(ich_aggressive),
   m_prior_neutral_comfort = f_prior_neutral_comfort(ich_aggressive),
   m_prior_left_comfort = f_prior_left_comfort(ich_aggressive),
   m_prior_right_comfort = f_prior_right_comfort(ich_aggressive),
@@ -95,128 +95,131 @@ tar_plan(
   m_prior_flat_dnr_binary = f_prior_flat_dnr_binary(ich_aggressive),
 
   ### Posterior Simulation ----
-  m_posterior_neurosurgery_minimal = f_posterior_neurosurgery_minimal(ich_aggressive),
[…]

### 2024-07-10 manuscripts/manuscript1.qmd
update quarto

@@
   scale_y_continuous(
-    limits = c(0,100)
+    limits = c(0, 100)
   ) +

### 2024-07-11 manuscripts/manuscript1.qmd
stuff

@@
 ```{r setup}
-#| echo: false
-#| output: false
-
-library(targets)
-library(tidyverse)
-library(withr)
-library(here)
-library(formattable)
-library(gt)
-library(gtsummary)
-library(modelsummary)
-library(marginaleffects)
-library(tidybayes)
-library(bayesplot)
-library(ggdist)
-library(patchwork)
-library(tinytable)
-library(glue)
-
-# plot stuff
-theme_ich <- function(base_family = "Figtree",
-                      base_size = 14) {
-  theme_minimal(base_size = base_size) +
-    theme(
-      text = element_text(lineheight = 2),
-      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
-      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
-      axis.text.y = element_text(size = rel(0.8)),
+#     axis.text.y = element_text(size = rel(0.8)),
       axis.title.y = element_text(size = rel(0.5), margin = margin(0, 4, 0, 0)),
@@
   "External Ventricular Drain" = with_dir(here(), tar_read(m_posterior_neutral_evd)),
+  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)),
   "Tracheostomy" = with_dir(here(), tar_read(m_posterior_neutral_tracheostomy)),
-  "Days of Mechanical Ventilation" = with_dir(here(), tar_read(m_posterior_neutral_days_mechanical_ventilation)),
   "Withdrawal of Life-Sustaining Therapy" = with_dir(here(), tar_read(m_posterior_neutral_comfort)),
@@
 )
+
+bob <- ich_aggressive |>
+  tabyl(neurosurgery_evac, ich_laterality) |>
+  adorn_percentages("col") |>
+  adorn_pct_formatting(digits = 1) |>
+  adorn_ns(position = "front") |>
+  filter(neurosurgery_evac == "Yes")
+
+ich_aggressive |>
+  tabyl(days_mechanical_ventilation, ich_laterality)
 
@@
 
-```{r neurosurgery}
+```{r table 2}
 #| label: tbl-2
@@
   "External Ventricular Drain",
+  "Days of Mechanical Ventilation",
   "Tracheostomy",
-  "Days of Mechanical Ventilation",
   "Withdrawal of Life-Sustaining Therapy",
@@
 
-trach <- models$"Tracheostomy" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
 vent <- models$"Days of Mechanical Ventilation" |>
@@
 
-dnr <- models$"DNR Order" |>
+trach <- models$"Tracheostomy" |>
   spread_draws(b_ich_lateralityRight) |>
@@
 
+dnr <- models$"DNR Order" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
 # combine probability of OR > 1 (prob of difference)
@@
   "External Ventricular Drain" = pluck(evd, 1),
+  "Days of Mechanical Ventilation" = pluck(vent, 1),
   "Tracheostomy" = pluck(trach, 1),
-  "Days of Mechanical Ventilation" = pluck(vent, 1),
   "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
@@
   "External Ventricular Drain" = pluck(evd, 4),
+  "Days of Mechanical Ventilation" = pluck(vent, 4),
   "Tracheostomy" = pluck(trach, 4),
-  "Days of Mechanical Ventilation" = pluck(vent, 4),
   "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
@@
 tibble_2 |>
-  fmt_number(columns = 6:7, decimals = 2)
+  fmt_number(columns = 6:7, decimals = 2) |>
+  tab_footnote(
+    footnote = "Adjusted for: age, admission GCS, ICH location, IVH, and study (as random intercept)",
+    locations = cells_column_labels(columns = 4)
+  ) |>
+  tab_footnote(
+    footnote = "Region of Practical Equivalence, defined as 0.95 > aOR > 1.05",
+    locations = cells_column_labels(columns = 7)
+  )
 
@@
 
+predictions(models$"Neurosurgical Intervention", by = c("ich_laterality", "study")) |>
+  posterior_draws() |>
+  ggplot(aes(x = draw, fill = ich_laterality)) +
+  geom_density(alpha = .7) +
+  facet_grid(~study) +
+  scale_fill_manual(values = c("#054224", "#ff8800")) +
+  scale_y_continuous() +
+  scale_x_continuous(
+    limits = c(0, 0.2),
+    breaks = seq(0, 0.20, 0.05),
+    labels = scales::percent
+  ) +
+  labs(
+    fill = "ICH Laterality",
+    y = NULL
+  ) +
+  theme_ich()
+
+models$"Neurosurgical Intervention" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  ggplot(aes(ich_right_or)) +
+  # geom_density(color = "#054224", fill = "#ff8800") +
+  stat_halfeye(aes(fill_ramp = after_stat(x > 1.05 | x < 0.95), fill = "#E69F00", color = "#422f05")) +
+  scale_x_continuous(breaks = seq(0, 3, 0.25)) +
+  labs(
+    title = "aOR Neurosurgical Intervention - Right Hemisphere",
+    y = NULL,
+    fill = NULL
+  ) +
+  geom_vline(
+    xintercept = 1, linetype = "dashed",
+    color = "#999999", linewidth = 1
+  ) +
+  geom_vline(
+    xintercept = 1.05, linetype = "dotted",
+    color = "#999999", linewidth = 1
+  ) +
+  geom_vline(
+    xintercept = 0.95, linetype = "dotted",
+    color = "#999999", linewidth = 1
+  ) +
+  theme_ich() +
+  theme(legend.position = "none")
+
+models$"External Ventricular Drain" |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  ggplot(aes(ich_right_or)) +
+  # geom_density(color = "#054224", fill = "#ff8800") +
+  stat_halfeye(aes(fill_ramp = after_stat(x > 1.05 | x < 0.95), fill = "#E69F00", color = "#E69F00")) +
+  scale_x_continuous(
+    breaks = seq(0, 3, 0.25)
+  ) +
+  labs(
+    title = "aOR Ventriculostomy - Right Hemisphere",
+    y = NULL,
+    fill = NULL
+  ) +
+  geom_vline(
+    xintercept = 1, linetype = "dashed",
+    color = "#999999", linewidth = 1
+  ) +
+  geom_vline(
+    xintercept = 1.05, linetype = "dotted",
+    color = "#999999", linewidth = 1
+  ) +
+  geom_vline(
+    xintercept = 0.95, linetype = "dotted",
+    color = "#999999", linewidth = 1
+  ) +
+  theme_ich() +
+  theme(legend.position = "none")
+
 ```

### 2024-07-11 manuscripts/manuscript1_supplement.qmd
stuff

@@
 library(here)
+library(ggdag)
 library(modelsummary)
@@
 #| echo: false
+#| warning: false
+#| fig-height: 8
 
-# dag_aggressive
+with_dir(here(), tar_load(dag_aggressive))
+
+status_colors <- c(exposure = "#0074D9", outcome = "#FF4136", latent = "grey50")
+
+dag_aggressive |>
+  tidy_dagitty() |>
+  node_status() |>
+  ggdag() +
+  geom_dag_edges() +
+  geom_dag_point(aes(color = status)) +
+  geom_dag_label_repel(aes(label = label, fill = status),
+    seed = 1234,
+    color = "white", fontface = "bold"
+  ) +
+  scale_color_manual(values = status_colors, na.value = "grey20") +
+  scale_fill_manual(values = status_colors, na.value = "grey20") +
+  guides(color = "none", fill = "none") +
+  theme_dag()
 
@@
 
+#### Days of Mechanical Ventilation
+
+```{r prior predictive check - vent}
+#| echo: false
+#| warning: false
+
+neutral_model <- with_dir(here(), tar_read(m_prior_neutral_days_mechanical_ventilation))
+left_model <- with_dir(here(), tar_read(m_prior_left_days_mechanical_ventilation))
+right_model <- with_dir(here(), tar_read(m_prior_right_days_mechanical_ventilation))
+flat_model <- with_dir(here(), tar_read(m_prior_flat_days_mechanical_ventilation))
+
+results_visual(neutral_model)
+
+f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
+
+```
+
 #### Tracheostomy
@@
 flat_model <- with_dir(here(), tar_read(m_prior_flat_tracheostomy))
-
-results_visual(neutral_model)
-
-f_prior_predictive_check(neutral_model, left_model, right_model, flat_model)
-
-```
-
-#### Days of Mechanical Ventilation
-
-```{r prior predictive check - vent}
-#| echo: false
-#| warning: false
-
-neutral_model <- with_dir(here(), tar_read(m_prior_neutral_days_mechanical_ventilation))
-left_model <- with_dir(here(), tar_read(m_prior_left_days_mechanical_ventilation))
-right_model <- with_dir(here(), tar_read(m_prior_right_days_mechanical_ventilation))
-flat_model <- with_dir(here(), tar_read(m_prior_flat_days_mechanical_ventilation))
 
@@
 ## Sensitivity Analyses
+
+### Different Priors
+
+### Alternative Adjustment Sets
+
+### Alternative Handling of Missing Data

### 2024-07-11 _targets.R

@@ -1,5 +1,7 @@
 library(targets)
 library(tarchetypes)
+library(quarto)
+
 
 # General pipeline settings ----
 
@@ -124,24 +126,6 @@ tar_plan(
   m_posterior_right_dnr_binary = f_posterior_right_dnr_binary(ich_aggressive),
   m_posterior_flat_dnr_binary = f_posterior_flat_dnr_binary(ich_aggressive),
 
-  #### Subgroups with ERICH ----
-  m_posterior_neutral_neurosurgery_erich = f_posterior_neutral_neurosurgery(erich),
-  m_posterior_neutral_evd_erich = f_posterior_neutral_evd(erich),
-  m_posterior_neutral_tracheostomy_erich = f_posterior_neutral_tracheostomy(erich),
-  m_posterior_neutral_comfort_erich = f_posterior_neutral_comfort(erich),
-  m_posterior_neutral_early_wlst_erich = f_posterior_neutral_early_wlst(erich),
-  m_posterior_neutral_dnr_binary_erich = f_posterior_neutral_dnr_binary(erich),
-
-  #### Subgroups with ATACH ----
-  m_posterior_neutral_neurosurgery_atach = f_posterior_neutral_neurosurgery(atach),
-  m_posterior_neutral_evd_atach = f_posterior_neutral_evd(atach),
-  m_posterior_neutral_tracheostomy_atach = f_posterior_neutral_tracheostomy(atach),
-  m_posterior_neutral_days_mechanical_ventilation_atach = f_posterior_neutral_days_mechanical_ventilation(atach),
-  m_posterior_neutral_comfort_atach = f_posterior_neutral_comfort(atach),
-  m_posterior_neutral_early_wlst_atach = f_posterior_neutral_early_wlst(atach),
-  m_posterior_neutral_dnr_binary_atach = f_posterior_neutral_dnr_binary(atach),
-
-
   ## Outcomes ----
 
   ### Priors ----
@@ -221,21 +205,9 @@ tar_plan(
   # m_posterior_euro_anxiety_365_minimal = f_posterior_euro_anxiety_365_minimal(ich_all),
   # m_posterior_euro_anxiety_365_canonical = f_posterior_euro_anxiety_365_canonical(ich_all),
 
-  ## Reports ----
+  # ## Reports ----
   # tar_quarto(
-  #   aggressive,
-  #   "reports/aggressiveness.qmd"
-  # ),
[…]

### 2024-07-20 manuscripts/manuscript1.qmd
added outcomes data

@@
 ```{r setup}
-#     axis.text.y = element_text(size = rel(0.8)),
-      axis.title.y = element_text(size = rel(0.5), margin = margin(0, 4, 0, 0)),
-      axis.text.x = element_text(size = rel(1)),
+#| echo: false
+#| output: false
+
+#| echo: false
+#| fig-height: 12
+
+library(targets)
+library(tidyverse)
+library(withr)
+library(here)
+library(formattable)
+library(gt)
+library(gtsummary)
+library(modelsummary)
+library(marginaleffects)
+library(tidybayes)
+library(bayesplot)
+library(ggdist)
+library(patchwork)
+library(tinytable)
+library(flextable)
+
+# plot stuff
+theme_ich <- function(base_family = "Figtree",
+                      base_size = 10) {
+  theme_minimal(base_size = base_size) +
+    theme(
+      text = element_text(lineheight = 2),
+      plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+      plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+      axis.text.y = element_text(size = rel(0.8)),
+      axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+      axis.text.x = element_text(size = rel(2)),
       axis.title.x = element_blank(),
@@
       legend.justification = 1,
-      panel.grid = element_blank(),
-      panel.background = element_blank(),
+      panel.grid = element_line(colour = "#F3F4F5"),
       plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
@@
 )
-
-bob <- ich_aggressive |>
-  tabyl(neurosurgery_evac, ich_laterality) |>
-  adorn_percentages("col") |>
-  adorn_pct_formatting(digits = 1) |>
-  adorn_ns(position = "front") |>
-  filter(neurosurgery_evac == "Yes")
-
-ich_aggressive |>
-  tabyl(days_mechanical_ventilation, ich_laterality)
 

### 2024-07-20 R/outcomes_posterior_models.R

@@ -17,8 +17,7 @@ f_posterior_mrs_90_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -39,8 +38,7 @@ f_posterior_mrs_90_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -63,8 +61,7 @@ f_posterior_mrs_180_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -85,8 +82,7 @@ f_posterior_mrs_180_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -109,8 +105,7 @@ f_posterior_mrs_365_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -131,8 +126,7 @@ f_posterior_mrs_365_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -155,8 +149,7 @@ f_posterior_euro_mobility_90_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -177,8 +170,7 @@ f_posterior_euro_mobility_90_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
[…]

### 2024-07-20 R/outcomes_prior_models.R

@@ -64,8 +64,7 @@ f_prior_mrs_180_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -87,8 +86,7 @@ f_prior_mrs_180_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -112,8 +110,7 @@ f_prior_mrs_365_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -135,8 +132,7 @@ f_prior_mrs_365_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -160,8 +156,7 @@ f_prior_euro_mobility_90_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -183,8 +178,7 @@ f_prior_euro_mobility_90_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -208,8 +202,7 @@ f_prior_euro_mobility_180_minimal <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
@@ -231,8 +224,7 @@ f_prior_euro_mobility_180_canonical <- function(dat) {
     threads = settings$threads,
     warmup = settings$warmup,
     iter = settings$iter,
-    seed = settings$seed,
-    backend = "rstan"
+    seed = settings$seed
   )
 
   return(model)
[…]

### 2024-07-20 _targets.R

@@ -129,81 +129,81 @@ tar_plan(
   ## Outcomes ----
 
   ### Priors ----
-  # m_prior_mrs_90_minimal = f_prior_mrs_90_minimal(ich_all),
-  # m_prior_mrs_90_canonical = f_prior_mrs_90_canonical(ich_all),
-  # m_prior_mrs_180_minimal = f_prior_mrs_180_minimal(ich_all),
-  # m_prior_mrs_180_canonical = f_prior_mrs_180_canonical(ich_all),
-  # m_prior_mrs_365_minimal = f_prior_mrs_365_minimal(ich_all),
-  # m_prior_mrs_365_canonical = f_prior_mrs_365_canonical(ich_all),
-  # m_prior_euro_mobility_90_minimal = f_prior_euro_mobility_90_minimal(ich_all),
-  # m_prior_euro_mobility_90_canonical = f_prior_euro_mobility_90_canonical(ich_all),
-  # m_prior_euro_mobility_180_minimal = f_prior_euro_mobility_180_minimal(ich_all),
-  # m_prior_euro_mobility_180_canonical = f_prior_euro_mobility_180_canonical(ich_all),
-  # m_prior_euro_mobility_365_minimal = f_prior_euro_mobility_365_minimal(ich_all),
-  # m_prior_euro_mobility_365_canonical = f_prior_euro_mobility_365_canonical(ich_all),
-  # m_prior_euro_selfcare_90_minimal = f_prior_euro_selfcare_90_minimal(ich_all),
-  # m_prior_euro_selfcare_90_canonical = f_prior_euro_selfcare_90_canonical(ich_all),
-  # m_prior_euro_selfcare_180_minimal = f_prior_euro_selfcare_180_minimal(ich_all),
-  # m_prior_euro_selfcare_180_canonical = f_prior_euro_selfcare_180_canonical(ich_all),
-  # m_prior_euro_selfcare_365_minimal = f_prior_euro_selfcare_365_minimal(ich_all),
-  # m_prior_euro_selfcare_365_canonical = f_prior_euro_selfcare_365_canonical(ich_all),
-  # m_prior_euro_usual_90_minimal = f_prior_euro_usual_90_minimal(ich_all),
-  # m_prior_euro_usual_90_canonical = f_prior_euro_usual_90_canonical(ich_all),
-  # m_prior_euro_usual_180_minimal = f_prior_euro_usual_180_minimal(ich_all),
[…]

### 2024-07-20 manuscripts/manuscript1.qmd
something

@@
 library(formattable)
+library(janitor)
 library(gt)
@@
 
+There
+
 ## Discussion

### 2024-07-20 R/exploratory_data_analysis.R
bob

@@ -48,7 +48,8 @@ table_1_function <- function(x) {
     add_p(pvalue_fun = ~ style_pvalue(.x, digits = 2)) |>
     add_overall() |>
     bold_labels() |>
-    modify_header(label ~ "")
+    modify_header(label ~ "") |>
+    as_gt()
 }
 
 table_2_aggressive_function <- function(x) {
@@ -80,5 +81,6 @@ table_2_aggressive_function <- function(x) {
       )
     ) |>
     bold_labels() |>
-    modify_header(label ~ "Outcome")
+    modify_header(label ~ "Outcome") |>
+    as_gt()
 }

### 2024-07-20 _targets.R

@@ -1,6 +1,7 @@
 library(targets)
 library(tarchetypes)
 library(quarto)
+library(here)
 
 
 # General pipeline settings ----
@@ -57,7 +58,6 @@ tar_plan(
   ## Exploratory data analysis ----
   table_1_aggressive = table_1_function(ich_aggressive),
   table_2_aggressive = table_2_aggressive_function(ich_aggressive),
-  table_1_outcomes = table_1_function(ich_all),
 
   ## DAGs ----
   dag_aggressive = aggressive_dag_function(x),
@@ -207,7 +207,7 @@ tar_plan(
 
   # ## Reports ----
   # tar_quarto(
-  #   aggressive_manuscript,
-  #   "manuscripts/manuscript1.qmd"
+  #   table_1,
+  #   here("manuscripts", "test.qmd")
   # )
 )

### 2024-08-23 manuscripts/manuscript1_supplement.qmd
some mrs updates

@@
 ### Alternative Handling of Missing Data
+
+## mRS Data
+
+```{r mrs90}
+
+with_dir(here(), tar_load(m_posterior_mrs_90_canonical))
+
+mod <- avg_predictions(m_posterior_mrs_90_canonical, by = "ich_laterality")
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_90 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+```
+
+```{r mrs180}
+
+with_dir(here(), tar_load(m_posterior_mrs_180_canonical))
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_180 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+mrs_180
+
+```
+
+```{r mrs365}
+
+with_dir(here(), tar_load(m_posterior_mrs_365_canonical))
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_365 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+mrs_365
+
+```
+
+```{r euro anxiety}
+
+with_dir(here(), tar_load(m_posterior_euro_anxiety_90_canonical))
+
+euro_anxiety <- m_posterior_euro_anxiety_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_anxiety
+
+```
+
+```{r euro mobility}
+
+with_dir(here(), tar_load(m_posterior_euro_mobility_90_canonical))
+
+euro_mobility <- m_posterior_euro_mobility_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_mobility
+
+```
+
+```{r euro pain}
+
+with_dir(here(), tar_load(m_posterior_euro_pain_90_canonical))
+
+euro_pain <- m_posterior_euro_pain_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_pain
+
+```
+
+```{r euro selfcare}
+with_dir(here(), tar_load(m_posterior_euro_selfcare_90_canonical))
+
+euro_selfcare <- m_posterior_euro_selfcare_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_selfcare
+
+```
+
+```{r euro usual}
+with_dir(here(), tar_load(m_posterior_euro_usual_90_canonical))
+
+euro_usual <- m_posterior_euro_usual_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_usual
+
+```

### 2024-09-04 manuscripts/manuscript1.qmd
making more functions

@@
 table_2 <- with_dir(here(), tar_read(table_2_aggressive))
-tibble_2 <- as_tibble(table_2) |> select(-1)
+
+tibble_2 <- table_2[[1]] |>
+  select(c(label, stat_1, stat_2))
 
@@
 
-tibble_2 <- bind_cols(column_names, tibble_2) |>
+tibble_2 <- tibble_2 |>
   rename(
-    "Right Hemisphere" = "**Right**, N = 1,757",
-    "Left Hemisphere" = "**Left**, N = 1,720",
-    "Outcome" = ...1
+    "Left Hemisphere" = "stat_1",
+    "Right Hemisphere" = "stat_2",
+    "Outcome" = "label"
   ) |>
@@
 #| label: figure 1
-#| fig-width: 10
-
-predictions(models$"Neurosurgical Intervention", by = "ich_laterality") |>
-  posterior_draws() |>
-  ggplot(aes(x = draw, fill = ich_laterality)) +
-  geom_density(alpha = .7) +
-  scale_fill_manual(values = c("#054224", "#ff8800")) +
-  scale_y_continuous(
-    limits = c(0, 100)
-  ) +
-  scale_x_continuous(
-    limits = c(0, 0.14),
-    breaks = seq(0, 0.14, 0.02),
-    labels = scales::percent
-  ) +
-  labs(
-    title = "Predicted Probability",
-    fill = "ICH Laterality",
-    y = NULL
-  ) +
-  theme_ich()
-
-```
-
-```{r figure 2}
-#| echo: false
-#| label: figure 2
 #| fig-width: 10

### 2024-09-04 R/exploratory_data_analysis.R

@@ -72,11 +72,11 @@ table_2_aggressive_function <- function(x) {
       missing = "no",
       label = list(
         neurosurgery_evac ~ "Neurosurgical Intervention",
-        evd ~ "Ventriculostomy",
-        days_mechanical_ventilation ~ "Mechanical Ventilation Days",
+        evd ~ "External Ventricular Drain",
+        days_mechanical_ventilation ~ "Days of Mechanical Ventilation",
         tracheostomy ~ "Tracheostomy",
         comfort_care_binary ~ "Withdrawal of Life-Sustaining Therapy",
-        early_wlst ~ "Early Withdrawal of Life-Sustaining Therapy (< 72 hours)",
+        early_wlst ~ "Early Withdrawal of Life-Sustaining Therapy",
         dnr_binary ~ "DNR Order"
       )
     ) |>

### 2024-09-04 R/predictive_checks.R

@@ -1,0 +1,112 @@
+f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
+  models <- list(
+    "Neutral Prior" = neutral_model,
+    "Left Hemisphere Prior" = left_model,
+    "Right Hemisphere Prior" = right_model,
+    "Flat Prior" = flat_model
+  )
+
+  # gather summary variables for each model
+  neutral <- models$"Neutral Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  left <- models$"Left Hemisphere Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or < 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  right <- models$"Right Hemisphere Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
+      upper_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
+    )
+
+  flat <- models$"Flat Prior" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      aor = quantile(ich_right_or, 0.5),
+      lower_ci = quantile(ich_right_or, 0.025),
[…]

### 2024-09-04 R/results_outcomes.R

@@ -1,0 +1,101 @@
+
+
+mod <- avg_predictions(m_posterior_mrs_90_canonical, by = "ich_laterality")
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_90 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_180 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+data <- mod |>
+  select(group, ich_laterality, estimate) |>
+  as_tibble() |>
+  mutate(estimate = estimate * 100)
+
+mrs_365 <- m_posterior_mrs_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
+  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+  summarize(
+    or_1 = sum(ich_right_or > 1) / n(),
+    or_1.1 = sum(ich_right_or > 1.1) / n(),
+    or_1.2 = sum(ich_right_or > 1.2) / n(),
+    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+  )
+
+euro_anxiety <- m_posterior_euro_anxiety_90_canonical |>
+  spread_draws(b_ich_lateralityRight) |>
[…]

### 2024-09-04 _targets.R

@@ -37,6 +37,8 @@ source("R/diagnostics.R")
 source("R/results_visualization.R")
 source("R/outcomes_prior_models.R")
 source("R/outcomes_posterior_models.R")
+source("R/results_outcomes.R")
+
 
 # Pipeline ----
 tar_plan(
@@ -166,8 +168,13 @@ tar_plan(
   m_prior_euro_anxiety_365_minimal = f_prior_euro_anxiety_365_minimal(ich_aggressive),
   m_prior_euro_anxiety_365_canonical = f_prior_euro_anxiety_365_canonical(ich_aggressive),
 
+  ### Prior Predictive Checks ----
 
-  # ### Posterior Simulation ----
+
+
+
+
+  #### Posterior Simulation ----
   m_posterior_mrs_90_minimal = f_posterior_mrs_90_minimal(ich_aggressive),
   m_posterior_mrs_90_canonical = f_posterior_mrs_90_canonical(ich_aggressive),
   m_posterior_mrs_180_minimal = f_posterior_mrs_180_minimal(ich_aggressive),

### 2024-09-04 manuscripts/manuscript1_supplement.qmd
making more functions
changes

--- 2024-07-11 manuscripts/manuscript1_supplement.qmd
@@
   return(prior_prediction_summary)
-}
-
-results_visual <- function(model) {
-  model <- model
-
-  draws <- model |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    summarize()
-
-  pred <- predictions(model, by = "ich_laterality") |>
-    tibble()
-
-  theme_ich <- function(base_family = "Inter",
-                        base_size = 14) {
-    theme_minimal(base_size = base_size) +
-      theme(
-        text = element_text(lineheight = 2),
-        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
-        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
-        axis.text.y = element_text(size = rel(0.8)),
-        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
-        axis.text.x = element_text(size = rel(2)),
-        axis.title.x = element_blank(),
-        legend.position = "bottom",
-        legend.justification = 1,
-        panel.grid = element_line(colour = "#F3F4F5"),
-        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
-        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
-      )
-  }
-
-  p1 <- predictions(model, by = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw, fill = ich_laterality)) +
-    geom_density(alpha = .7) +
-    scale_fill_manual(values = c("#054224", "#ff8800")) +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Predicted Probability",
-      fill = "ICH Laterality",
-      y = NULL
-    ) +
-    theme_ich()
-
-  return(p1)
 }

### 2024-09-04 R/predictive_checks.R

@@ -110,3 +110,49 @@ f_prior_predictive_check <- function(neutral_model, left_model, right_model, fla
 
   return(prior_prediction_summary)
 }
+
+results_visual <- function(model) {
+  model <- model
+
+  draws <- model |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize()
+
+  pred <- predictions(model, by = "ich_laterality") |>
+    tibble()
+
+  theme_ich <- function(base_family = "Inter",
+                        base_size = 14) {
+    theme_minimal(base_size = base_size) +
+      theme(
+        text = element_text(lineheight = 2),
+        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
+        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+        axis.text.y = element_text(size = rel(0.8)),
+        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
+        axis.text.x = element_text(size = rel(2)),
+        axis.title.x = element_blank(),
+        legend.position = "bottom",
+        legend.justification = 1,
+        panel.grid = element_line(colour = "#F3F4F5"),
+        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
+        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
+      )
+  }
+
+  p1 <- predictions(model, by = "ich_laterality") |>
+    posterior_draws() |>
+    ggplot(aes(x = draw, fill = ich_laterality)) +
+    geom_density(alpha = .7) +
+    scale_fill_manual(values = c("#054224", "#ff8800")) +
+    scale_x_continuous(labels = scales::percent) +
+    labs(
+      title = "Predicted Probability",
+      fill = "ICH Laterality",
+      y = NULL
+    ) +
+    theme_ich()
+
+  return(p1)
+}

### 2024-09-04 _targets.R

@@ -37,8 +37,8 @@ source("R/diagnostics.R")
 source("R/results_visualization.R")
 source("R/outcomes_prior_models.R")
 source("R/outcomes_posterior_models.R")
-source("R/results_outcomes.R")
-
+# source("R/results_outcomes.R")
+source("R/predictive_checks.R")
 
 # Pipeline ----
 tar_plan(
@@ -169,6 +169,7 @@ tar_plan(
   m_prior_euro_anxiety_365_canonical = f_prior_euro_anxiety_365_canonical(ich_aggressive),
 
   ### Prior Predictive Checks ----
+  # prior_check_neurosurgery = f_prior_predictive_check(m_prior_neutral_neurosurgery, m_prior_left_neurosurgery, m_prior_right_neurosurgery, m_prior_flat_neurosurgery),
 
 
 

### 2024-09-27 manuscripts/manuscript.qmd
all models run again

--- 2024-09-04 manuscripts/manuscript1.qmd
@@
 Previous research involving patients with ischemic stroke provide evidence that those with right hemispheric injury present later to the hospital and may not be treated as frequently with thrombolytic therapy.
-[@dileggeImpactLesionSide2005; @foerchDifferenceRecognitionRight2005; @portegiesLeftsidedStrokesAre2015;  @liStrokeLateralizationLarge2021] It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts3 despite literature demonstrating that long-term functional outcomes and health-related quality of life (HRQoL) are equivalent if not better for patients with left hemispheric injuries.
+[@dileggeImpactLesionSide2005; @foerchDifferenceRecognitionRight2005; @portegiesLeftsidedStrokesAre2015;  @liStrokeLateralizationLarge2021] It has also been noted that there is a hesitation among some practitioners to offer decompressive craniectomy in patients with left dominant-hemispheric ischemic infarcts despite literature demonstrating that long-term functional outcomes and health-related quality of life (HRQoL) are equivalent if not better for patients with left hemispheric injuries.
 [@deb-chatterjiSideMattersDifferences2022; @leggeNeglectingDifference2006; @beuscherFunctionalLongTermOutcome2017; @satoHigherMortalityPatients2015]
 
-It is uncertain how the results of prior work in the ischemic stroke literature translate to patients experiencing acute ICH.
-We aimed to evaluate the effect of hemispheric lateralization of ICH on likelihood of receiving aggressive surgical intervention.
+Previous work has demonstrated mixed results with regard to hemispheric lateralization and outcomes in ICH, and there hasn't been much research examining if patients are treated differently based on the hemisphere affected.
+We aimed to evaluate the effect of hemispheric lateralization of ICH on likelihood of receiving aggressive surgical intervention as well as it's association with functional outcomes.
 
@@
 
+Our primary functional outcome was modified Rankin score at 90 days, examined via ordinal regression.
+
 ### Statistical Analysis
@@
 #### Prior Distribution Specification
-Two sets of priors were applied for all models, both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature.
+Four sets of priors were applied for all models: both assuming no association of hemispheric laterality on the outcomes of interest: 1) default, weakly informative priors as defined by the 'brms' software package; 2) priors informed from review of previous literature.
 A description of the reasons for these priors is available in the supplementary material.
@@
 #### Computation and Description of the Posterior Distribution
-We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a large difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05).
+We report the following metrics for the effect of hemispheric laterality on the primary and secondary outcomes: 1) median odds ratio (OR) of the posterior distribution with 95% credible interval; 2) probability of any difference (P[OR] > 1); 3) probability of a substantial difference (P[OR] > 1.2) as defined by the study authors; 4) percent of the posterior distribution outside of the region of practical equivalence as defined by the study authors (ROPE; 0.95 > P[OR] > 1.05).
 A description of the reasons for these thresholds is available in the supplementary material.
@@
 
-```{r table1}
-#| label: tbl-1
-#| tbl-cap: "Baseline Characteristics"
-#| echo: false
-
-with_dir(here(), tar_read(table_1_aggressive))
-
-```
-
 ### Left hemispheric ICH and neurosurgical intervention
-
-```{r table 2}
-#| label: tbl-2
-#| tbl-cap: "Association between hemispheric lateralization and aggressive care"
-#| echo: false
-#| warning: false
-
-# gather odds ratios and 95% credible intervals for each model
-summary <- modelsummary(models,
-  exponentiate = TRUE,
-  fmt = 2,
-  estimate = c("aOR" = "estimate"),
-  statistic = c("95% Credible Interval" = "{conf.low} - {conf.high}"),
-  coef_omit = "^(?!b_ich_lat)",
-  coef_rename = c("b_ich_lateralityRight" = "Right Hemisphere ICH"),
-  gof_map = c("none"),
-  shape = model ~ term + statistic,
-)
-
-aOR_CIs_combined <- as_tibble(summary@table_dataframe) |>
-  rename("Outcome" = " ")
-
-aOR_CIs_combined[1] <- c(
-  "Neurosurgical Intervention",
-  "External Ventricular Drain",
-  "Days of Mechanical Ventilation",
-  "Tracheostomy",
-  "Withdrawal of Life-Sustaining Therapy",
-  "Early Withdrawal of Life-Sustaining Therapy",
-  "DNR Order"
-)
-
-# gather posterior probabilities of various magnitudes of difference for each outcome
-
-neurosurgery <- models$"Neurosurgical Intervention" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-evd <- models$"External Ventricular Drain" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-vent <- models$"Days of Mechanical Ventilation" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-trach <- models$"Tracheostomy" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-wlst <- models$"Withdrawal of Life-Sustaining Therapy" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-early_wlst <- models$"Early Withdrawal of Life-Sustaining Therapy" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-dnr <- models$"DNR Order" |>
-  spread_draws(b_ich_lateralityRight) |>
-  mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-  summarize(
-    or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
-    or_1.2 = sum(ich_right_or > 1.2) / n(),
-    rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
-  )
-
-# combine probability of OR > 1 (prob of difference)
-prob_difference <- tibble(
-  "Neurosurgical Intervention" = pluck(neurosurgery, 1),
-  "External Ventricular Drain" = pluck(evd, 1),
-  "Days of Mechanical Ventilation" = pluck(vent, 1),
-  "Tracheostomy" = pluck(trach, 1),
-  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 1),
-  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 1),
-  "DNR Order" = pluck(dnr, 1)
-) |>
-  pivot_longer(
-    cols = everything(),
-    names_to = "Outcome",
-    values_to = "Probabilty of any difference (aOR > 1)"
-  )
-
-# combine % of posterior within the ROPE
-rope <- tibble(
-  "Neurosurgical Intervention" = pluck(neurosurgery, 4),
-  "External Ventricular Drain" = pluck(evd, 4),
-  "Days of Mechanical Ventilation" = pluck(vent, 4),
-  "Tracheostomy" = pluck(trach, 4),
-  "Withdrawal of Life-Sustaining Therapy" = pluck(wlst, 4),
-  "Early Withdrawal of Life-Sustaining Therapy" = pluck(early_wlst, 4),
-  "DNR Order" = pluck(dnr, 4)
-) |>
-  pivot_longer(
-    cols = everything(),
-    names_to = "Outcome",
-    values_to = "% within ROPE"
-  )
-
-table_almost_2 <- left_join(aOR_CIs_combined, prob_difference, by = "Outcome") |>
-  left_join(rope, by = "Outcome")
-
-table_2 <- with_dir(here(), tar_read(table_2_aggressive))
-
-tibble_2 <- table_2[[1]] |>
-  select(c(label, stat_1, stat_2))
-
-column_names <- c(
-  "Neurosurgical Intervention",
-  "External Ventricular Drain",
-  "Days of Mechanical Ventilation",
-  "Tracheostomy",
-  "Withdrawal of Life-Sustaining Therapy",
-  "Early Withdrawal of Life-Sustaining Therapy",
-  "DNR Order"
-)
-
-tibble_2 <- tibble_2 |>
-  rename(
-    "Left Hemisphere" = "stat_1",
-    "Right Hemisphere" = "stat_2",
-    "Outcome" = "label"
-  ) |>
-  full_join(table_almost_2, by = "Outcome") |>
-  gt()
-
-tibble_2 |>
-  fmt_number(columns = 6:7, decimals = 2) |>
-  tab_footnote(
-    footnote = "Adjusted for: age, admission GCS, ICH location, IVH, and study (as random intercept)",
-    locations = cells_column_labels(columns = 4)
-  ) |>
-  tab_footnote(
-    footnote = "Region of Practical Equivalence, defined as 0.95 > aOR > 1.05",
-    locations = cells_column_labels(columns = 7)
-  )
-
-```
 
@@
 
-predictions(models$"Neurosurgical Intervention", by = c("ich_laterality", "ich_location")) |>
+predictions(models$"Neurosurgical Intervention", type = "response", by = c("ich_laterality")) |>
   posterior_draws() |>
@@
   ) +
-  theme_ich()
+  theme_minimal()
+
+predictions(models$"Neurosurgical Intervention", type = "response", by = c("ich_laterality", "ich_location")) |>
+  posterior_draws() |>
+  ggplot(aes(x = draw, fill = ich_laterality)) +
+  geom_density(alpha = .7) +
+  # facet_grid(~ich_location) +
+  scale_fill_manual(values = c("#054224", "#ff8800")) +
+  scale_y_continuous() +
+  scale_x_continuous(
+    limits = c(0, 0.2),
+    breaks = seq(0, 0.20, 0.05),
+    labels = scales::percent
+  ) +
+  labs(
+    fill = "ICH Laterality",
+    y = NULL
+  ) +
+  theme_minimal()
 

### 2024-09-27 R/figures.R

@@ -1,0 +1,228 @@
+subgroup_figure_function <- function(model) {
+  left_fill <- "#f69d75"
+  right_fill <- "#551e4e"
+
+  theme_ich <- function(base_size = 10) {
+    theme_minimal(base_size = base_size) +
+      theme(
+        text = element_text(lineheight = 2),
+        plot.title = element_text(size = rel(2), margin = margin(12, 0, 8, 0)),
+        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
+        axis.text.y = element_blank(),
+        axis.title.y = element_text(
+          size = rel(2),
+          margin = margin(0, 40, 0, 0),
+          vjust = 0.5
+        ),
+        axis.text.x = element_text(size = rel(1.5)),
+        axis.title.x = element_text(size = rel(1.5), margin = margin(10, 0, 0, 0)),
+        strip.text.x = element_text(
+          size = rel(2),
+          margin = margin(0, 0, 20, 0),
+          vjust = 0
+        ),
+        legend.position = "bottom",
+        legend.justification = 1,
+        panel.grid = element_blank(),
+        plot.caption = element_text(size = rel(1), margin = margin(8, 0, 0, 0)),
+        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
+      )
+  }
+
+  total_plot <- predictions(model, type = "response", by = c("ich_laterality")) |>
+    posterior_draws() |>
+    ggplot(aes(x = draw, fill = ich_laterality)) +
+    geom_density(alpha = .5) +
+    # facet_grid(~ich_location) +
+    scale_fill_manual(values = c(left_fill, right_fill)) +
+    scale_x_continuous(
+      limits = c(0, 0.15),
+      breaks = seq(0, 0.15, 0.025),
+      labels = scales::percent
+    ) +
+    labs(
+      title = "Overall Probability",
+      x = NULL,
+      y = NULL,
+      fill = "Hemispheric Laterality"
+    ) +
+    theme_ich()
+
+  ich_location_plot <- predictions(model, type = "response", by = c("ich_laterality", "ich_location")) |>
+    posterior_draws() |>
+    ggplot(aes(x = draw, fill = ich_laterality)) +
+    geom_density(alpha = .5) +
+    facet_grid(~ich_location) +
+    scale_color_manual(values = c(left_fill, right_fill)) +
[…]

### 2024-09-27 R/packages.R

@@ -6,6 +6,7 @@ library(forcats)
 library(stringr)
 library(lubridate)
 library(ggplot2)
+library(ggdist)
 library(gt)
 library(gtsummary)
 library(dagitty)
@@ -22,3 +23,4 @@ library(patchwork)
 library(systemfonts)
 library(monochromeR)
 library(glue)
+library(viridis)

### 2024-09-27 R/results_outcomes.R

@@ -1,26 +1,12 @@
-
-
-mod <- avg_predictions(m_posterior_mrs_90_canonical, by = "ich_laterality")
-
-data <- mod |>
-  select(group, ich_laterality, estimate) |>
-  as_tibble() |>
-  mutate(estimate = estimate * 100)
-
 mrs_90 <- m_posterior_mrs_90_canonical |>
   spread_draws(b_ich_lateralityRight) |>
   mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
   summarize(
     or_1 = sum(ich_right_or > 1) / n(),
-    or_1.1 = sum(ich_right_or > 1.1) / n(),
     or_1.2 = sum(ich_right_or > 1.2) / n(),
     rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
   )
 
-data <- mod |>
-  select(group, ich_laterality, estimate) |>
-  as_tibble() |>
-  mutate(estimate = estimate * 100)
 
 mrs_180 <- m_posterior_mrs_90_canonical |>
   spread_draws(b_ich_lateralityRight) |>
@@ -56,7 +42,6 @@ euro_anxiety <- m_posterior_euro_anxiety_90_canonical |>
     or_1.2 = sum(ich_right_or > 1.2) / n(),
     rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
   )
-
 
 euro_mobility <- m_posterior_euro_mobility_90_canonical |>
   spread_draws(b_ich_lateralityRight) |>

### 2024-09-27 R/results_visualization.R

@@ -1,89 +1,0 @@
-results_visual <- function(model, title) {
-  model <- model
-  subtitle <- title
-
-  draws <- model |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    summarize()
-
-  pred <- predictions(model, by = "ich_laterality") |>
-    tibble()
-
-  theme_ich <- function(base_family = "Inter",
-                        base_size = 14) {
-    theme_minimal(base_size = base_size) +
-      theme(
-        text = element_text(lineheight = 2),
-        plot.title = element_text(size = rel(1.3), margin = margin(12, 0, 8, 0)),
-        plot.subtitle = element_text(size = rel(1.1), margin = margin(4, 0, 0, 0)),
-        axis.text.y = element_text(size = rel(0.8)),
-        axis.title.y = element_text(size = 12, margin = margin(0, 4, 0, 0)),
-        axis.text.x = element_text(size = rel(2)),
-        axis.title.x = element_blank(),
-        legend.position = "bottom",
-        legend.justification = 1,
-        panel.grid = element_line(colour = "#F3F4F5"),
-        plot.caption = element_text(size = rel(0.8), margin = margin(8, 0, 0, 0)),
-        plot.margin = margin(0.25, 0.25, 0.25, 0.25, "cm")
-      )
-  }
-
-  p1 <- predictions(model, by = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw, fill = ich_laterality)) +
-    geom_density(alpha = .7) +
-    scale_fill_manual(values = c("#054224", "#ff8800")) +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Predicted Probability",
-      fill = "ICH Laterality",
-      y = NULL
-    ) +
-    theme_ich()
-
-  p2 <- avg_comparisons(model, variables = "ich_laterality") |>
-    posterior_draws() |>
-    ggplot(aes(x = draw)) +
-    geom_density(color = "#054224", fill = "#ff8800") +
-    scale_x_continuous(labels = scales::percent) +
-    labs(
-      title = "Average Difference between Left vs. Right Hemisphere",
-      y = NULL
-    ) +
-    geom_vline(
-      xintercept = 0, linetype = "dashed",
[…]

### 2024-09-27 R/table1.R

@@ -19,7 +19,7 @@ table_1_function <- function(x) {
     "study"
   )
 
-  x |>
+  table_1 <- x |>
     select(all_of(table_one_vars)) |>
     tbl_summary(
       by = ich_laterality,
@@ -42,45 +42,13 @@ table_1_function <- function(x) {
         ivh ~ "IVH",
         study ~ "Study"
       ),
-      sort = list(everything() ~ "frequency"),
+      sort = all_categorical() ~ "frequency",
       digits = starts_with("gcs") ~ 0,
     ) |>
     add_p(pvalue_fun = ~ style_pvalue(.x, digits = 2)) |>
     add_overall() |>
     bold_labels() |>
-    modify_header(label ~ "") |>
     as_gt()
-}
 
-table_2_aggressive_function <- function(x) {
-  table_two_vars <- c(
-    "neurosurgery_evac",
-    "evd",
-    "days_mechanical_ventilation",
-    "tracheostomy",
-    "ich_laterality",
-    "comfort_care_binary",
-    "early_wlst",
-    "dnr_binary"
-  )
-
-  x |>
-    select(all_of(table_two_vars)) |>
-    mutate(days_mechanical_ventilation = as.numeric(days_mechanical_ventilation)) |>
-    tbl_summary(
-      by = ich_laterality,
-      missing = "no",
-      label = list(
-        neurosurgery_evac ~ "Neurosurgical Intervention",
-        evd ~ "External Ventricular Drain",
-        days_mechanical_ventilation ~ "Days of Mechanical Ventilation",
-        tracheostomy ~ "Tracheostomy",
-        comfort_care_binary ~ "Withdrawal of Life-Sustaining Therapy",
-        early_wlst ~ "Early Withdrawal of Life-Sustaining Therapy",
-        dnr_binary ~ "DNR Order"
-      )
-    ) |>
-    bold_labels() |>
-    modify_header(label ~ "Outcome") |>
-    as_gt()
+  return(table_1)
 }

### 2024-09-27 R/table2.R

@@ -1,4 +1,263 @@
-f_table2 <- function(models) {
+models_list_func <- function(a, b, c, d, e, f, g) {
+  list(
+    "Neurosurgical Intervention" = a,
+    "External Ventricular Drain" = b,
+    "Days of Mechanical Ventilation" = c,
+    "Tracheostomy" = d,
+    "Withdrawal of Life-Sustaining Therapy" = e,
+    "Early Withdrawal of Life-Sustaining Therapy" = f,
+    "DNR Order" = g
+  )
+}
 
+table_2_function <- function(x, models) {
+  models <- models
 
+  # gather raw counts for each outcome
+  neurosurgery_n <- x |>
+    tabyl(ich_laterality, neurosurgery_evac) |>
+    adorn_percentages("row") |>
+    adorn_pct_formatting(digits = 1) |>
+    adorn_ns(position = "front") |>
+    select(!No) |>
+    pivot_wider(names_from = ich_laterality, values_from = Yes)
+
+  evd_n <- x |>
+    tabyl(ich_laterality, evd) |>
+    adorn_percentages("row") |>
+    adorn_pct_formatting(digits = 1) |>
+    adorn_ns(position = "front") |>
+    select(!No) |>
+    pivot_wider(names_from = ich_laterality, values_from = Yes)
+
+  days_mechanical_ventilation_n <- x |>
+    drop_na(days_mechanical_ventilation) |>
+    group_by(ich_laterality) |>
+    summarize(
+      median = median(as.numeric(days_mechanical_ventilation)),
+      lower_25 = quantile(as.numeric(days_mechanical_ventilation), 0.25),
+      upper_75 = round(quantile(as.numeric(days_mechanical_ventilation), 0.75))
+    ) |>
+    mutate(
+      iqr = glue("{median} ({lower_25} - {upper_75})")
+    ) |>
+    select(ich_laterality, iqr) |>
+    pivot_wider(names_from = ich_laterality, values_from = iqr)
+
+  tracheostomy_n <- x |>
+    tabyl(ich_laterality, tracheostomy) |>
+    adorn_percentages("row") |>
+    adorn_pct_formatting(digits = 1) |>
+    adorn_ns(position = "front") |>
+    select(!No) |>
+    pivot_wider(names_from = ich_laterality, values_from = Yes)
+
+  comfort_care_binary_n <- x |>
[…]

### 2024-09-27 R/table3.R

@@ -1,0 +1,126 @@
+outcome_models_list_func <- function(a, b, c, d, e, f) {
+  list(
+    "Modified Rankin Score" = a,
+    "EuroQOL - Mobility" = b,
+    "EuroQOL - Self-Care" = c,
+    "EuroQOL - Usual Activities" = d,
+    "EuroQOL - Pain/Discomfort" = e,
+    "EuroQOL - Anxiety/Depression" = f,
+  )
+}
+
+table_3_function <- function(models) {
+  models <- models
+
+  mrs <- models$"Modified Rankin Score" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      or = median(ich_right_or),
+      lower_95_ci = quantile(ich_right_or, 0.025),
+      upper_95_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      or_1.1 = sum(ich_right_or > 1.1) / n(),
+      or_1.2 = sum(ich_right_or > 1.2) / n(),
+      rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+    ) |>
+    mutate(
+      or_ci = glue("{round(or, digits = 2)} ({round(lower_95_ci, digits = 2)} - {round(upper_95_ci, digits = 2)})")
+    ) |>
+    select(or_ci, or_1, or_1.2, rope)
+
+  euro_anxiety <- models$"EuroQOL - Anxiety/Depression" |>
+    spread_draws(b_ich_lateralityRight) |>
+    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
+    summarize(
+      or = median(ich_right_or),
+      lower_95_ci = quantile(ich_right_or, 0.025),
+      upper_95_ci = quantile(ich_right_or, 0.975),
+      or_1 = sum(ich_right_or > 1) / n(),
+      or_1.1 = sum(ich_right_or > 1.1) / n(),
+      or_1.2 = sum(ich_right_or > 1.2) / n(),
+      rope = sum(ich_right_or < 1.05 & ich_right_or > 0.95) / n()
+    ) |>
+    mutate(
+      or_ci = glue("{round(or, digits = 2)} ({round(lower_95_ci, digits = 2)} - {round(upper_95_ci, digits = 2)})")
+    ) |>
[…]

### 2024-09-27 _targets.R

@@ -29,17 +29,21 @@ options(
 
 source("R/packages.R")
 source("R/data_cleaning.R")
-source("R/exploratory_data_analysis.R")
 source("R/dags.R")
 source("R/aggressive_care_prior_models.R")
 source("R/aggressive_care_posterior_models.R")
 source("R/diagnostics.R")
-source("R/results_visualization.R")
+# source("R/results_visualization.R")
 source("R/outcomes_prior_models.R")
 source("R/outcomes_posterior_models.R")
 # source("R/results_outcomes.R")
 source("R/predictive_checks.R")
 source("R/posterior_diagnostics.R")
+source("R/table1.R")
+source("R/table2.R")
+source("R/table3.R")
+source("R/figures.R")
+
 
 # Pipeline ----
 tar_plan(
@@ -57,10 +61,6 @@ tar_plan(
   ich_aggressive = ich_all |> filter(study == "ERICH" | study == "ATACH-2") |> droplevels(),
   erich = ich_all |> filter(study == "ERICH") |> droplevels(),
   atach = ich_all |> filter(study == "ATACH-2") |> droplevels(),
-
-  ## Exploratory data analysis ----
-  table_1_aggressive = table_1_function(ich_aggressive),
-  table_2_aggressive = table_2_aggressive_function(ich_aggressive),
 
   ## DAGs ----
   dag_aggressive = aggressive_dag_function(x),
@@ -171,94 +171,79 @@ tar_plan(
   ### Priors ----
   m_prior_mrs_90_minimal = f_prior_mrs_90_minimal(ich_aggressive),
   m_prior_neutral_mrs_90_canonical = f_prior_neutral_mrs_90_canonical(ich_aggressive),
-  m_prior_left_mrs_90_canonical = f_prior_neutral_mrs_90_canonical(ich_aggressive),
-  m_prior_right_mrs_90_canonical = f_prior_neutral_mrs_90_canonical(ich_aggressive),
-  m_prior_flat_mrs_90_canonical = f_prior_neutral_mrs_90_canonical(ich_aggressive),
+  m_prior_left_mrs_90_canonical = f_prior_left_mrs_90_canonical(ich_aggressive),
+  m_prior_right_mrs_90_canonical = f_prior_right_mrs_90_canonical(ich_aggressive),
+  m_prior_flat_mrs_90_canonical = f_prior_flat_mrs_90_canonical(ich_aggressive),
[…]

### 2024-09-27 manuscripts/function.R

@@ -1,153 +1,0 @@
-library(targets)
-library(tidyverse)
-library(withr)
-library(here)
-library(modelsummary)
-library(marginaleffects)
-library(brms)
-library(tidybayes)
-library(bayesplot)
-library(glue)
-library(ggdist)
-library(patchwork)
-library(tinytable)
-library(gt)
-
-neutral_mean <- 0
-neutral_sd <- 0.35
-neutral_lower_OR <- exp(neutral_mean - (1.96 * neutral_sd))
-neutral_upper_OR <- exp(neutral_mean + (1.96 * neutral_sd))
-
-left_mean <- -0.22
-left_sd <- 0.175
-left_lower_OR <- exp(left_mean - (1.96 * left_sd))
-left_upper_OR <- exp(left_mean + (1.96 * left_sd))
-
-right_mean <- 0.18
-right_sd <- 0.175
-right_lower_OR <- exp(right_mean - (1.96 * right_sd))
-right_upper_OR <- exp(right_mean + (1.96 * right_sd))
-
-
-neutral_model <- with_dir(here(), tar_read(m_prior_neutral_neurosurgery))
-left_model <- with_dir(here(), tar_read(m_prior_left_neurosurgery))
-right_model <- with_dir(here(), tar_read(m_prior_right_neurosurgery))
-flat_model <- with_dir(here(), tar_read(m_prior_flat_neurosurgery))
-
-f_prior_predictive_check <- function(neutral_model, left_model, right_model, flat_model) {
-  models <- list(
-    "Neutral Prior" = neutral_model,
-    "Left Hemisphere Prior" = left_model,
-    "Right Hemisphere Prior" = right_model,
-    "Flat Prior" = flat_model
-  )
-
-  # gather summary variables for each model
-  neutral <- models$"Neutral Prior" |>
-    spread_draws(b_ich_lateralityRight) |>
-    mutate(ich_right_or = exp(b_ich_lateralityRight)) |>
-    summarize(
-      aor = quantile(ich_right_or, 0.5),
-      lower_ci = quantile(ich_right_or, 0.025),
-      upper_ci = quantile(ich_right_or, 0.975),
-      or_1 = sum(ich_right_or > 1) / n(),
-      rope = sum(ich_right_or > 1.05 | ich_right_or < 0.95) / n()
-    )
-
-  left <- models$"Left Hemisphere Prior" |>
-    spread_draws(b_ich_lateralityRight) |>
[…]

### 2024-09-27 manuscripts/sample.R

@@ -1,18 +1,0 @@
-ggplot(mutate(spread_draws(m_prior_informative_neurosurgery, b_ich_lateralityRight),
-        ich_right_or = exp(b_ich_lateralityRight)
-), aes(ich_right_or)) +
-        geom_density(color = "#422f05", fill = "#E69F00") +
-        scale_x_continuous(breaks = seq(
-                0,
-                3, 0.25
-        )) +
-        labs(
-                title = "Odds Ratio - Right Hemisphere",
-                y = NULL, fill = NULL
-        ) +
-        geom_vline(
-                xintercept = 1, linetype = "dashed",
-                color = "#999999", linewidth = 1
-        ) +
-        theme_ich() +
-        theme(legend.position = "none")
